### 2016-03-28 manuscript.tex
initial commit

%XXX: Aaron, do one last check for new typos in my final series of edits.
\documentclass[preprint,11pt]{aastex} 
\let\captionbox=\undefined
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{mdwlist}
\usepackage{natbib}
\usepackage{natbibspacing}
\usepackage[font={footnotesize}]{caption}
\usepackage{wrapfig}
\usepackage{sidecap}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{url}
\setlist[itemize]{noitemsep, topsep=0pt}
\setlist[enumerate]{noitemsep, topsep=0pt}
\setlength{\bibspacing}{0pt}
\setlength{\parskip}{0pt}
\setlength{\parsep}{0pt}
\setlength{\headsep}{0pt}  
\setlength{\topskip}{0pt}
\setlength{\topmargin}{0pt}
\setlength{\topsep}{6pt}
\setlength{\partopsep}{0pt}
\setlength{\footnotesep}{8pt}
\pagestyle{plain}
\citestyle{aa}
\captionsetup[table]{labelsep=space}
\captionsetup[figure]{labelsep=space}

%%% ADDED FROM 2014 PROPOSAL
%\newcommand{\Mycite}[1]{{\bf \cite{#1}}}
%\newcommand{\Mycitet}[1]{{\bf \citet{#1}}}
%\newcommand{\Mycitep}[1]{{\bf \citep{#1}}}
%\newcommand{\Mycitealt}[1]{{\bf \citealt{#1}}}
\newcommand{\Mycite}[1]{\cite{#1}}
\newcommand{\Mycitet}[1]{\citet{#1}}
\newcommand{\Mycitep}[1]{\citep{#1}}
\newcommand{\Mycitealt}[1]{\citealt{#1}}

\newcommand{\kvec}{{\bf k}}
\newcommand{\bvec}{{\bf b}}
\newcommand{\shat}{{\hat s}}
\newcommand{\kpr}{{k_\perp}}
\newcommand{\kvpr}{{\kvec_\perp}}
\newcommand{\AI}{{\langle\tilde A*\tilde I\rangle}}
\newcommand{\AItau}{{\AI(\tau)}}
\newcommand{\hMpci}{{h~{\rm Mpc}^{-1}}}
\newcommand{\inch}{$^{\prime\prime}$}
\newcommand{\foot}{$^{\prime}$}
\renewcommand{\deg}{^\circ}

\newcommand{\simgt}{\stackrel{>}{_{\sim}}}
\newcommand{\compress}{\vspace{-0.25in}}
\newcommand{\captionbaseline}{\renewcommand{\baselinestretch}{0.99}} 
\newcommand{\Caption}[4]{\vspace{#1}\renewcommand{\baselinestretch}{#2}\caption{#4}\vspace{#3}}
\def\kperp{k_{\bot}}
\def\eppsilon{{$\varepsilon$ppsilon}}
\def\kpar{k_{\|}}
\def\k{{\bf k}}
\def\sky{{\theta}}
\def\HI{{H{\small I }}}
\def\HII{{H{\small II }}}
\def\xHI{{x_{\rm\HI}}}
\newcommand{\startsquarepar}{%
    \par\begingroup \parfillskip 0pt \relax}
\newcommand{\stopsquarepar}{%
    \par\endgroup}

%%% UNCOMMENTED
\usepackage{subfig}
%\usepackage[countmax]{subfloat}
\title{Hydrogen Epoch of Reionization Array (HERA)}
\author{A.U.Thor}

\begin{document}
\maketitle

\begin{center}
\textbf{Abstract}
\end{center}

The Hydrogen Epoch of Reionization Array (HERA http://reionization.org) is a staged experiment that uses the unique properties of the 21-cm line from neutral hydrogen to probe the Epoch of Reionization (EOR). During this epoch, roughly 0.3 - 1 billion years after the Big Bang, the first galaxies and black holes heated and reionized the early Universe. Direct observation of the large scale structure of reionization and its evolution with time will have a profound impact on our understanding of the birth of the first galaxies and black holes, their influence on the intergalactic medium (IGM), and cosmology.  This paper will provide an overview of the project and describe the design of the system.

\section{Introduction}

\noindent The Hydrogen Epoch of Reionization Array (HERA) is a staged experiment to use the redshifted 21\,cm line
of neutral hydrogen to characterize our Cosmic Dawn, from
the formation of the first stars and black holes $\sim0.1~$Gyr after the Big
Bang ($z\sim30$) through the full reionization of the intergalactic medium
(IGM) $\sim1~$Gyr later ($z\sim6$).  By directly observing the large scale
structure of the primordial IGM as it is heated and reionized, HERA complements
other probes, adding transformative capabilities for
understanding the astrophysics and fundamental cosmology of our early universe.
Taking advantage of our new mastery of bright foreground systematics, HERA's
purpose-built radio
interferometer is optimized to deliver high signal-to-noise measurements of
redshifted 21\,cm emission.
With the first generation 21\,cm telescopes approaching the limits of their sensitivity, HERA continues the experiment by bringing significantly more sensitivity to bear
while continuing to optimize its design along the delay-spectrum approach.

\begin{figure}[h!]
	\centering
	\includegraphics[height=1.75in]{plots/hera_render.png}
	\includegraphics[height=1.75in]{plots/hera19H.png}
	\caption{Rendering of the 320-element core (left) of the full HERA-350 array and picture of 19 HERA 14-m, zenith-pointing dishes (with PAPER elements in the background) currently deployed in South Africa (right).} 
	\label{fig:HERApictures}
\end{figure}

HERA brings to bear both the sensitivity and the precision to deliver 
its key science results on the basis of proven techniques.
These science results include
\begin{itemize}[noitemsep,nolistsep]
\item measuring the 21\,cm power spectrum with high significance (\S\ref{sec:EoRPowerSpectra}),
\item characterizing parametrized models of reionization and the first ionizing sources (\S\ref{sec:EoRPowerSpectra}),
\item improving cosmic microwave background (CMB) constraints on primordial density fluctuations and neutrino masses by removing the optical depth degeneracy (\S\ref{sec:tau}), and
\item directly imaging large scale reionization structure (\S\ref{sec:imaging}).
\end{itemize}
At the same time, HERA continues the development of new techniques that extend and improve its capabilities.
Examples include improving foreground subtraction to expand the information content 
in power spectral and imaging measurements and, with support from the Moore Foundation,
extending HERA's performance at lower frequencies to probe IGM heating in the pre-reionization epoch (\S\ref{sec:EoX}).
Beyond these directly supported applications, HERA's public data products and instrumentation platform enable other high-impact science, including 
\begin{itemize}[noitemsep,nolistsep]
\item 21\,cm cross-correlation (e.g. with JWST, WFIRST, CO, CII, and Ly$-\alpha$; \S\ref{sec:crossCorr}),
\item searching for auroral radio bursts from exoplanetary magnetospheres (\S\ref{sec:exoplanets}), and
\item performing triggered, low-frequency follow up and characterization of fast radio bursts (\S\ref{sec:FRBs}).
\end{itemize}

The HERA experiment will comprise 
350 14-m parabolic dishes (320 in a dense core $+$ 30 outriggers) in the South African Karoo Radio
Astronomy Reserve.
HERA has been funded under the US National Science Foundation's Mid-Scale Innovations Program to begin building elements optimized for robust power spectrum detection in a 
staged deployment.  The key is to produce inexpensive collecting area optimized to detect the EOR.  This has resulted in a close-packed array of transit 14-meter dishes (Fig. \ref{fig:HERApictures}), with the first 19-element array completed and construction underway for 37 elements, when the first science observing will begin.
The first phase, extending to August 2016, will deploy 37 elements that will test the HERA element design for performance and manufacturability at location.  These 37 elements will more 
than double the sensitivity of the 128-element PAPER array.  The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal 
paths along with , which should provide a very robust detection of the EOR signal.  Finally, extending into 2020, HERA will build out to 350 elements to further EOR science as a function 
of redshift and spatial scale, potentially producing the first images of the EOR.

HERA is a collaboration of the following Partner institutions:  
Arizona State University (Tempe, AZ USA), 
Brown University (Providence, RI USA), 
University of California Berkeley (Berkeley, CA USA), 
University of California Los Angeles (Los Angeles, CA USA), 
University of Cambridge (Cambridge, UK), 
Massachusetts Institute of Technology (Cambridge, MA USA), 
National Radio Astronomy Observatory (Charlottesville, VA USA), 
University of Pennsylvania (Philadelphia, PA USA), 
Scuola Normale Superiore di Pisa (Pisa, Italy),
SKA-South Africa (Cape Town, South Africa), and
University of Washington (Seattle, WA USA).  
Additional Collaborators are at 
Harvard University (Cambridge, MA USA),  
University of KwaZulu Natal (Durban, South Africa), 
University of Western Cape (Cape Town, South Africa), 
Imperial College London  (London, UK) and 
California State Polytechnic University (Pomona, CA USA).

This paper has the following structure.  Section \ref{sec:science} will give a high-level overview of the science motivating HERA's construction, including primary and secondary science objectives.  Section \ref{sec:eormeas} will present the techniques used to make the measurement, focussing on the delay-spectrum approach.  Section \ref{sec:requirements} will present the resulting high-level requirements and section will present the system description \ref{sec:design} as the bulk of the paper.   Section \ref{sec:status} will provide a status and section \ref{sec:conclusion} will conclude by summarizing and providing additional context.

\section{Scientific Background}
\label{sec:science}

\noindent The {\it Cosmic Dawn} of our universe is one of the last unexplored
frontiers in cosmic history, which is summarized in the cartoon in Fig. \ref{fig:cosmos}. During this period, the 
intergalactic medium (IGM) encodes a panoply of information about astrophysical and cosmological phenomena. 
This period saw the evolution of the universe from a very homogeneous ``gas'' to the formation of the first stars and galaxies.
Ultimately this early population of gravitationally condensed material produced sufficiently energetic flux to reionize the IGM
from its previous neutral state in a period called the Epoch of Reionization
Its evolution depends on the cosmic density field, the relative velocities of
baryons and dark matter, and the sizes and clustering of the first galaxies.
But it also depends on the constituents of those galaxies---so-called Population
III stars (stars formed very early with little to no elements heavier than helium), later generation stars, stellar remnants, X-ray binaries,
and early supermassive black holes.  Bulk properties like
ultraviolet and X-ray luminosities and spectra also affect the thermal and
ionization states of the IGM.  The wealth of unexplored physics during the Cosmic Dawn,
culminating in the Epoch of Reionization, 
led the most recent US National Academies astronomy decadal survey entitled {\em New Worlds, New Horizons} (NWNH) to highlight it as one of the top three ``priority science objectives'' for
the decade.

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.7\textwidth]{plots/cosmicEvo.png}
	\caption{Rendering of cosmic evolution from just after the Big Bang to today (credit CCCCC).  The blue line shows the frequency of the red-shifted hydrogen line (rest frequency 1420 MHz) as a function of redshift (lower horizontal axis) and cosmic time (upper horizontal axis).} 
	\label{fig:cosmos}
\end{figure}

Exploring the interplay of galaxies and large-scale structure during the EOR
requires complementary observational approaches. CMB measurements 
provide initial conditions for structure formation, and the Thomson
scattering of CMB photons constrains the integrated column of 
ionized gas, but even with kinetic Sunyaev-Zel'dovich measurements, 
the detailed evolution of the IGM 
is only loosely constrained \citep{haiman_holder2003,mortonson_hu2008,mesinger_et_al2012}.
Lyman-$\alpha$ absorption features in quasar and $\gamma$-ray burst spectra give 
ionization constraints at the tail end of reionization 
($z\la7$, \citealt{fan_et_al2006, mcgreer_et_al2015}), but these features 
saturate at low neutral fractions
($x_{\rm HI} \la 10^{-4}$). Measurements of galaxy populations in
deep {\it Hubble Space Telescope} observations 
have pinned down the bright end of the galaxy luminosity function at $z \la 8$
\citep{schenker_et_al2013, bouwens_et_al2015} and are pushing deeper
(e.g.~\citealt{mcleod_et_al2015}), but producing a consistent ionization history
requires broad extrapolations to
lower-mass galaxies and ad hoc assumptions about the escape fraction of
ionizing photons and the faint-end cutoff of ionizing galaxies
\citep{robertson_et_al2015, bouwens_et_al2015_reion}. 
Similarly, deducing the ionization state of the IGM from quasar proximity zones
\citep{carilli_et_al2010, bolton_et_al2011, bosman_becker2015} and the
demographics of Ly-$\alpha$ emitting galaxies \citep{fontana_et_al2010,
schenker_et_al2012, treu_et_al2012, dijkstra_et_al2014} is uncertain and
highly model-dependent.

\begin{figure}[h!]
\centering
    \includegraphics[width=0.51\textwidth,clip]{plots/ionHist.pdf}
  \caption{Combining direct constraints on $x_\text{HI}$, the hydrogen neutral fraction, as a function of redshift (black points) with \emph{Planck} priors yields
an inferred $95\%$ confidence region (gray).  HERA constraints with (dark red) and without (pale red) 
a conservative $25\%$ modeling error in the $21\,\textrm{cm}$ power spectrum can dramatically narrow this confidence region.}
	\label{fig:IonHist}
\end{figure}

As shown in Figure~\ref{fig:IonHist},
existing probes are limited in
their ability to constrain reionization, and will be for the foreseeable future.
HERA will use
another complementary probe --- the 21\,cm ``spin-flip" transition of neutral hydrogen
--- to bring transformative new capabilities in this area.  The next sub-sections outline these goals.

\subsection{Precision Constraints on Reionization}
\label{sec:EoRPowerSpectra}

\noindent HERA's top science goal is to transform our understanding of the first stars, galaxies, 
and black holes, and their role in driving reionization. 
Through power-spectral measurements of the 21\,cm line of hydrogen in the primordial IGM,
HERA directly constrains the topology and evolution of reionization, 
opening a unique window into the complex astrophysical interplay between the 
first luminous objects and their environments.
Using cosmological redshift, HERA can associate 
the signal at each observing frequency with emission time (or distance) to determine both the time evolution
and three-dimensional spatial structure of ionization in the IGM.
This 3D structure encodes information about the clustering properties of galaxies,
allowing us to distinguish between models, even if they predict the same ionized fraction. 
With a new telescope optimized for 3D power-spectral measurements and with support for theoretical
modeling efforts, the HERA program will advance our understanding of early galaxy formation and cosmic reionization.

\begin{figure}[h!]
\centering
    \includegraphics[width=0.49\textwidth,clip]{plots/sensitivity_wideband.pdf}
  \caption{ 1$\sigma$ thermal noise errors on $\Delta^2(k)$, the 21\,cm power spectrum, at $k\!=\!0.2$\,$h$\,Mpc$^{-1}$ (the dominant error at that $k$)
with 1080 hours of integration (black)
compared with various heating and reionization models (colored).}
	\label{fig:Sensitivities}
\end{figure}

HERA builds on the advances of first-generation
21\,cm EoR experiments led by HERA team members, particularly 
the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}),
the Murchison Widefield Array (MWA; \citealt{bowman_et_al2012,tingay_et_al2013}),
the MIT EoR experiment (MITEoR; \citealt{zheng_et_al2014}) and the Experiment to Detect the Global EoR Step (EDGES; \citealt{bowman_rogers2010}).  
Efforts by members of the HERA team
have produced the first astrophysically constraining upper limits on the 21\,cm EoR power spectrum, 
providing evidence for significant heating in the IGM prior to reionization \citep{parsons_et_al2014,ali_et_al2015,pober_et_al2015}.
However, current experiments cannot expect more than marginal detections of the EoR signal. 
Figure~\ref{fig:Sensitivities} compares telescope sensitivities as a function of redshift to models of 
the evolving, spherically averaged 21\,cm EoR power spectrum $\Delta^2 (k) \equiv k^3 P(k) / 2 \pi^2$ (discussed later).  In contrast to current experiments,
HERA's optimized design and collecting area make it capable of
high-significance detections of virtually any realistic reionization scenario, even using only currently demonstrated foreground
mitigation strategies (see Table~\ref{tab:signif}). The development of enhanced foreground modeling and subtraction further improve the achievable constraints.

\begin{SCtable}
\caption{\hspace{-1.2mm}: Predicted SNRs of 21\,cm experiments for an EoR model with 50\% ionization at $z=9.5$, with 1080 hours observation, integrated over a $\Delta z$ of $0.8$.
Foreground avoidance represents an analysis comparable to \cite{ali_et_al2015}, whereas foreground modeling allows signifcantly more $k$ modes of the cosmological signal to be recovered.}
\small
 \centering
 \begin{tabular}{c||r||r|r} 
%\begin{deluxetable}{c||r||r|r}
%\tabletypesize{\small}
%\tablecaption{\small
\hline
%\startdata
Instrument & \shortstack{Collecting \\ Area (m$^2$)} & \shortstack{Foreground \\Avoidance} & \shortstack{Foreground \\Modeling} \\
\hline
PAPER & 1,188 & 0.77$\sigma$ & 3.04$\sigma$ \\
MWA & 3,584 & 0.31$\sigma$ & 1.63$\sigma$ \\
LOFAR NL Core & 35,762 & 0.38$\sigma$ & 5.36$\sigma$ \\
%\textbf{HERA-127} & \textbf{19,500} & \textbf{10.88} & \textbf{35.65} \\
\textbf{HERA-350} & \textbf{53,878} & \textbf{23.34$\sigma$} & \textbf{90.97}$\boldsymbol{\sigma}$ \\
SKA1 Low Core & 416,595 & 13.4$\sigma$ & 109.90$\sigma$
%\enddata
\end{tabular}
%\Caption{-0.1in}{0.99}{-0.1in}
\hspace{-0.1in}
\label{tab:signif}
\end{SCtable}

\begin{figure}[h!]
\centering
    \includegraphics[width=0.56\textwidth,clip]{plots/LikelihoodContours_smaller_avoid_All3_no_table}
  \caption{Projected likelihood contours from an MCMC analysis for astrophysical parameters of reionization. Model parameters are $T_\textrm{vir}^\textrm{min}$ (minimum virial temperature of ionizing galaxies); $R_\textrm{mfp}$ (mean free path of ionizing photons in HII regions); and $\zeta_0$ (ionizing efficiency of galaxies).  Also shown are constraints on the derived ionizing escape fraction, $f_\textrm{esc}$. }
	\label{fig:paramConstraints}
\end{figure}

HERA's 21\,cm measurements can be used in conjunction with our semi-analytic models to constrain the ionization history. 
The red band in Fig.~\ref{fig:IonHist} shows the forecasted 95\% confidence region derived from HERA data after marginalizing over astrophysical and cosmological parameters.
Figure~\ref{fig:paramConstraints} shows the results of a Markov Chain Monte
Carlo (MCMC) pipeline for fitting models to 21\,cm power spectrum data \citep{greig_and_mesinger2015}, which we have conservatively limited to the range $8 < z < 10$ (although in practice a much broader bandwidth will be available; see Table \ref{tab:BasicParameters}).
Based on the excursion-set formalism of
\citet{furlanetto_et_al2004} and the 21cmFAST code \citep{mesinger_et_al2011},
this code models the astrophysics of
reionization with three free parameters (see Figure~\ref{fig:paramConstraints} for details). 
While the existing experiment with the most collecting area, 
the LOw Frequency ARray (LOFAR; \citealt{yatawatta_et_al2013}),
faces large
uncertainties, 
HERA's tight constraints on these parameters are comparable
to what could be achieved with the SKA.  Such results would be valuable for
informing the design and science case of the low-frequency SKA, highlighting areas where the SKA has unique capability.
Additionally, HERA's constraints enable principal component parameterizations of the
sky-averaged $21\,\textrm{cm}$ signal measurements pursued by experiments such as EDGES, increasing their signal-to-noise and thus their science return \citep{liu_parsons2015}.

\subsection{Secondary Scientific Objectives}%\\Precision Cosmology, Reionization Imaging, and High-$z$ IGM X-Ray Heating Science}

\startsquarepar\emph{\textbf{Precision Cosmology.}}
\label{sec:tau}
By advancing our understanding of reionization astrophysics, HERA improves CMB constraints on 
fundamental cosmological parameters by
removing the optical depth
$\tau$ as a nuisance parameter. HERA breaks the degeneracy between the
constraints on $\tau$ and the sum of the neutrino masses $\sum m_\nu$, which has
been identified as a potential problem for Stage 4 CMB lensing experiments. A
HERA-informed estimate of $\tau$ enables CMB lensing experiments to achieve a
$0.012\,\textrm{eV}$ error on $\sum m_\nu$ \citep{liu_et_al2015}. This would
represent a $\sim$5$\sigma$ cosmological detection of the neutrino masses even
under the most pessimistic assumptions
% that their sum is the minimum value of $0.058\,\textrm{eV}$ 
still allowed by neutrino oscillation \stopsquarepar

\begin{figure}[h!]
\centering
    \includegraphics[width=0.50\textwidth,clip]{plots/sigmaTau.pdf}
\caption{Likelihood contours ($68\%$ and $95\%$) for $\sigma_8$ and $\tau$ using \emph{Planck} constraints (blue) and 
adding HERA data (red). 
%Light and heavy contours signify $68\%$ and $95\%$ confidence regions, respectively. 
The $21\,\textrm{cm}$ constraints break the CMB degeneracy between the amplitude of density fluctuations and the optical depth, improving constraints on both.}
	\label{fig:sigma8Tau}
\end{figure}

\noindent experiments
\citep{allison_et_al2015}, making HERA key to our understanding of neutrino
physics. HERA's estimate of $\tau$ would also break the degeneracy between
$\tau$ and the amplitude of matter fluctuations (expressed in Fig.
\ref{fig:sigma8Tau} as $\sigma_8$) that arises when using only
CMB data. HERA effectively reduces error bars on $\sigma_8$ by more
than a factor of three \citep{liu_et_al2015}, potentially elucidating
current tensions between cluster cosmology constraints and those from primary
CMB anisotropies.

\emph{\textbf{First Images of the Reionization Epoch.}}
\label{sec:imaging}
In addition to measuring the power spectrum, HERA has sufficient 
sensitivity to directly image the IGM during reionization,
and will observe a 1440 deg$^2$ stripe, comparable to future WFIRST
large area near-IR surveys.
 After 100 hours on a
single field (achievable in 200 nights), HERA reaches a surface brightness 
sensitivity of 50 $\mu$Jy/beam (synthesized beam FWHM $\sim$ 24$'$) compared 
to the brightness temperature fluctuations of up to 400 $\mu$Jy/beam in typical 
reionization models (see Fig.~\ref{fig:LSS}). From the standpoint of sensitivity alone, HERA is capable of 
detecting the brightest structures at $z=8$ with SNR $>$ 10. Additionally, the design 
of HERA places it in a unique position to directly explore calibration techniques
(e.g. redundant calibration, \citealt{liu_et_al2010}), while retaining a high quality 
point spread functions for imaging and identifying foregrounds. Leveraging the expertise gained 
through both precursor strategies, HERA will yield exceptionally high dynamic range
images.

%\subsubsection{The 21\,cm Signal in the Pre-Reionization Epoch}

%\begin{wrapfigure}{R}{0.5\textwidth}
\begin{figure}[b!]
\centering
\vspace{-15pt}
    %\includegraphics[width=.5\textwidth,clip]{plots/imaging/HERA_z8_SNR_annotated_2015.png}
    \includegraphics[width=1.0\textwidth,clip]{plots/imaging/HERA_FoV_w_strips.png}
  \vspace{-20pt}
\caption{\footnotesize 
HERA will observe a 1440 deg$^2$ stripe centered near $\delta = -30^\circ$. HERA can measure the ionization state around galaxies in, e.g., the GOODS-South field that contains a third of all known $z\!>\!8$ galaxies. HERA's primary imaging data product to the community will be deep cubes along the HERA stripe suitable for cross-correlation.
\label{fig:LSS} }
\vspace{-10pt}
\end{figure}
%  \vspace{-10pt}
%\end{wrapfigure}

\emph{\textbf{Pre-Reionization Heating.}}
\label{sec:EoX} 
Prior to reionization, the 21\,cm signal is a sensitive probe of the first
luminous sources and IGM heating mechanisms. First stars are expected to form
at $z \sim 25-30$ and their imprint on the 21\,cm signal is expected to be
sensitive to the halo mass where they are formed \citep{mesinger_et_al2015}.
The IGM is then expected to be heated by first
generation X--ray binaries
\citep{furlanetto_et_al2006_global,pritchard_et_al2007,mesinger_et_al2013} or
by the hot interstellar medium produced by the first supernovae
\citep{pacucci_et_al2014}, although the heating timing and magnitude is still
very much debated \citep{fialkov_et_al2012}. Dark matter annihilation
\citep{evoli_et_al2014} could also inject energy in the IGM, leaving its
imprint in the 21\,cm signal.

With support from the Moore Foundation to develop feeds sensitive down to 50\,MHz, HERA will probe the IGM prior to
reionization and constrain the sources of heating, obtaining, in the case of
X--ray heating, percent-level constraints on the efficiency with which star-forming baryons
produce X-rays \citep{ewall-wice_et_al2015}. While the observational and analytical state of the art of pre-reionization 21\,cm science is not as advanced as for the EoR \citep{ewall-wice_et_al2016-EoXLimits}, the risk of lower-frequency observing is borne by the Moore Foundation.
Lower
frequency observations also test feedback
mechanisms that interact with low-mass halos at high redshifts
\citep{Iliev_et_al2007,Iliev_et_al2012,ahn_et_al2012}.
Such constraints, while interesting in their own right, also reduce
the susceptibility of the aforementioned $21\,\textrm{cm}$-derived $\tau$
constraints to uncertainties in high-redshift physics \citep{liu_et_al2015}, especially if they are 
combined with upcoming or proposed measurements of 
the pre-reionization sky-averaged spectrum \citep{fialkov_and_loeb2016}. Additionally, they may be crucial to a correct
interpretation of kinetic Sunyaev-Zel'dovich effect constraints on reionization
from the CMB \citep{park_et_al2013}. The high redshift probe of
structure afforded by the low frequency $21\,\textrm{cm}$ measurements will
permit some of the most direct observations of hypothesized suppressions of
small-scale structure
\citep{dalal_et_al2010,tseliakhovich_et_al2011,fialkov_et_al2012} arising from
predicted supersonic relative velocities between dark matter and baryonic gas
\citep{tseliakhovich_and_hirata2010}. No other electromagnetic probe can provide direct observations of this
epoch.

\subsection{Other Scientific Objectives}
\label{subsec:broader_science}

\noindent With unprecedented sensitivity at this frequency range, HERA can deliver much more than its core 21\,cm science. 
Some of HERA's broader science impacts derive directly from the public data products HERA delivers.  Other
areas of impact may require additional hardware or software that HERA can host at no extra risk. 
In this respect, HERA serves as a ``platform" for additional science programs with external funding.
The Moore Foundation's support of HERA feed development and data analysis targeting the pre-reionization science described in \S\ref{sec:EoX}
is a prime example of how we envision this functioning.
Below, we list examples of the broader science that HERA can deliver.

\emph{\textbf{Cross-Correlations with Other Reionization Probes.}}
\label{sec:crossCorr}
HERA's public data provide new opportunities for cross-correlation studies.
Cross-correlation between HERA 21\,cm images and other high-redshift probes
(e.g. JWST; WFIRST; CMB maps; CO, CII, and Ly-$\alpha$ intensity mapping) 
can provide an independent confirmation of the 21\,cm power spectrum
\citep[i.e.][]{lidz_et_al2009,dore_et_al2014,silva_et_al2015,vrbanec2016} and enable rich new studies of
the interaction between galaxies and their ionization environment. 
In particular, cross-correlating 21\,cm with galaxy surveys can measure the characteristic bubble size around galaxies of
different luminosities \citep{lidz_et_al2009} and help separate the degeneracy
between the fraction of photons escaping the galaxies and the
total number of ionizing photons produced \citep{zackrisson2013}.

HERA is placed in a unique position to enable such cross-correlation: the GOODS-South field---one of the most panchromatically studied regions of the sky, the site of the Hubble UDF, and home to over a third of all known $z>9$
galaxies---lies in its field of view. HERA's images of the IGM can provide environmental context to galaxy surveys through identification of
ionized bubbles \citep{malloy_lidz2013}, or in a more statistical sense as
described in \cite{beardsley_et_al2015}.

\textbf{\emph{Searching for Exoplanetary Radio Bursts.}}
\label{sec:exoplanets}
HERA could be an powerful tool in the search for bright, highly polarized auroral bursts
from exoplanets %due to the electron cyclotron maser instability 
\citep{treumann2006,hallinan_et_al2015}. These
distinctive, coherent bursts occur periodically at the planetary
rotation rate, which cannot otherwise be measured. 
A burst from a terrestrial planet could point toward
habitability, since it implies a powerful magnetic field
protecting the atmosphere and perhaps the biosphere from energetic stellar wind
particles \citep{tarter_et_al2007}. HERA's sensitivity, large FoV, long campaigns, and precise calibration are all well-suited
for discovering exoplanetary radio bursts from Jupiter-like planets out to 25~pc. Many well studied exoplanetary systems within that distance are in the HERA stripe, including Fomalhaut, Gl 667 C, Gl 433, HD 147513, and Gl 317.

\textbf{\emph{Fast Radio Burst Followup.}}
\label{sec:FRBs}
Fast Radio Bursts (FRB) are millisecond-long radio flashes whose origin has remained a great enigma ever since their discovery \citep{2007Sci...318..777L}. 
If equipped with a suitable backend, HERA could be triggered by nearby, higher-frequency telescopes for FRB followup, saving baseband data and thus full sensitivity to all dispersion measures.
Bursts like that discovered by \citet{masui_et_al2015}, the lowest frequency
FRB detection to date, should be seen hourly by HERA at 5--10$\sigma$. %(including the effects of scattering)
 Observations at HERA frequencies are very sensitive to the physics of
the intervening medium, particularly deviations from $\lambda^2$
dispersion. Detecting deviations would rule out broad classes of models and could indicate whether FRBs are at cosmological distances.

\section{Measuring the EOR}
\label{sec:eormeas}
Due to the expansion of the Universe, we can identify and measure the early Universe via the red-shift of spectral lines.  The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire Universe back almost to the period of recombination.  The bandwidth
therefore equates to a cosmic volume along the site of the telescope, with the frequency determining the cosmic age.

Given the difficulty of the experiment, initial measurements of the EOR strive to measure a statistical power spectrum of the signal since the nature of the reionization process should have a specific spatial signature.  The goal is therefore to measure a range of aggregate spatial scales on the sky, rather than to image the signal directly.  Imaging does remain an ultimate goal to fully understand the process, however we will need a greater understanding of the full measurements to achieve this more difficult goal.  MANY PAPERS, see Nithya and Jonnie recent

Obviously between our present observation point and the Epoch of Reionization approximately 13 billion years ago lies the entire intervening Universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).  Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.  As a first step, areas of the sky with where these signals are minimal (for example, outside the galactic plane and away from strong point sources) are targeted (REF FOR THIS).  More importantly and most fortuitously however is the fact that all of these foreground signals are smooth spectrum sources whereas the expected spectrum of the EOR is expected to be rough since it is made up of non-ionized regions which are randomly distributed over a wide range of redshifts.  This fact allows us to try and isolate foregrounds from the EOR, as will be discussed.

The past three
years have seen deep EoR observations with PAPER, the MWA, 
the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2013}), and LOFAR.
PAPER and the MWA have produced progressively deeper limits
\citep{dillon_et_al2015,parsons_et_al2014,ali_et_al2015}, with PAPER
yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
(Fig. \ref{fig:limits}).
The inherent challenge of simultaneously meeting stringent sensitivity requirements 
while suppressing foregrounds $\sim5$ orders of magnitude
brighter than the 21\,cm signal make this progress all the more remarkable.  
To achieve this, HERA team members have refined and improved 
techniques spanning all aspects of the experimental process, from how we
calibrate our data \citep{zheng_et_al2014, jacobs_et_al2016, barry_et_al2016}, to how we understand foreground contamination
\citep{moore_et_al2013,moore_et_al2016,thyagarajan_et_al2015a,pober_et_al2016},
to how we design the interferometer itself \citep{parsons_et_al2012a,dillon_parsons2016}.
In this section, we discuss how the legacy of these first generation
21\,cm experiments inform critical aspects of HERA's design.

\begin{figure}[t]
	\centering
	\includegraphics[width=.56\textwidth]{plots/current_limits.pdf}
	\includegraphics[width=.43\textwidth]{plots/constraints_proposal.pdf}
	\vspace{-20pt}
	\caption{Left: The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, (solid symbols represent analyses led by HERA collaborators) compared to 21cmFAST-generated models at $k=0.2$\,$h$\,Mpc$^{-1}$. 
Analysis is still underway on PAPER and MWA observations that approach their projected 
full sensitivities (see Fig.~\ref{fig:Sensitivities}); 
HERA can deliver sub-$\text{mK}^2$ sensitivities.
Right: The \citet{ali_et_al2015} limit at $z=8.4$ excludes the entire gray shaded area corresponding to a cold IGM \citep{pober_et_al2015}.  The color scale shows the power spectrum amplitude at $k = 0.25~h{\rm Mpc}^{-1}$ for a given spin temperature and neutral fraction.  Constraints are weaker if the (currently unknown) neutral fraction at $z=8.4$ is very high or low, as the 21\,cm signal is brightest during the middle of reionization.}
	\vspace{-10pt}
	\label{fig:limits}
	\label{fig:IGMtemperatureConstraints}
\end{figure}

Perhaps the most important advance informing HERA's design is a
refined understanding of how smooth-spectrum foregrounds interact
with instrument chromaticity to produce a characteristic ``wedge" of
foreground leakage in Fourier space (see Fig.~\ref{fig:wedge}), 
outside of which the 21\,cm signal dominates.
Through theoretical and observational work
\citep{Datta_2010,morales_et_al2012,parsons_et_al2012b,vedantham_2012,thyagarajan_et_al2013,hazelton_et_al2013,pober_et_al2013b,liu_et_al2014a,liu_et_al2014b},
we have learned how the boundary between the wedge and our ``EoR window" is determined by the separation between antennas,
signal reflections within antennas, and the angular response of the antenna beam.  Deep integrations also show us
that, to the limits of current sensitivity, foreground emission is absent outside of the wedge; it can only 
appear there through instrumental leakage \citep{parsons_et_al2014,ali_et_al2015,moore_et_al2016,kohn_et_al2016}.
Thus, to open the widest possible window for EoR measurements, HERA must use close-packed antennas that
minimize signal reflections and deliver significant forward gain relative to their horizon response.
Tests with prototype HERA antennas (Figs. \ref{fig:orbcommexptandbeammap} and \ref{fig:reflectometry}, discussed in \S\ref{sec:antenna})
indicate that a moderately large parabolic dish with a short focal height can meet these requirements
\citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016}.

PAPER adopted a foreground mitigation strategy based largely
on filtering out the wedge in frequency domain (so-called ``delay filtering"; \citealt{parsons_et_al2012b}).  
By reducing the need
for image-domain foreground modeling, this approach allowed PAPER to switch to a grid-based antenna layout that
enhanced sensitivity toward fewer Fourier modes and facilitated calibration
based on the ``redundancy" of different antenna pairs measuring the same sky modes \citep{zheng_et_al2014}.
Combined, redundancy and delay filtering provide a robust, inexpensive, and demonstrably successful solution 
to the foreground problem.  
Yet PAPER's lack of imaging support and its uneven $uv$ sampling 
leave it with limited diagnostic capability, particularly for direction-dependent systematics
such as polarization leakage from Faraday-rotated emission \Mycitep{moore_et_al2013}.
While concern over such effects has decreased markedly since discovering that
intrinsic point source polarization is lower than previously thought
\citep{asad_et_al2015} and that variable Faraday rotation through the
ionosphere averages down the polarized signal over long observing seasons
\Mycitep{moore_et_al2016}, direction-dependent beam effects remain an area of interest.
The MWA's image-based calibration and foreground subtraction strategy provides complementary
capabilities. Imaging with subtraction, while still under development as a viable foreground strategy,
can increase sensitivity by recovering more modes of the cosmological signal (see
Table~\ref{tab:signif}) and help address systematic errors rooted
in the image domain.  HERA's antenna configuration---shown in Figure \ref{fig:arrayConfig} and discussed in \S\ref{sec:arrayConfig}---emphasizes
the proven approaches of redundant calibration and delay filtering, while simultaneously
increasing the extent and density of $uv$ sampling for high-fidelity imaging.

Despite progress, the fact remains that the 21\,cm EoR signal is intrinsically very faint; making a detection requires a
large instrumental collecting area and a long, dedicated observing campaign.
Although published PAPER and MWA results (Fig. \ref{fig:limits}) do not yet include observations 
at full sensitivity that are still being analyzed (e.g. Fig. \ref{fig:Sensitivities}), 
it is clear already that these instruments lack the sensitivity
to make a conclusive detection (see Table~\ref{tab:signif}). 
HERA addresses this shortcoming with a dish element that delivers a much larger collecting area 
while retaining the necessary characteristics for both proven and forward-looking foreground removal strategies.

Though the desired measurement is over a volume, it is instructive to split it into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kvpr$ is determined by the antenna baseline and $\kpar$ by the frequency.  This is useful since it allows us to split out the chromatic response of the interferometer and isolate a phase space where foreground sources ({\em i.e.} everything {\em not} the EOR) contaminate the signal of interest from where they don't.  This contaminated phase space may be shown [REF] to be a wedge-shaped region in $\kpr$:$\kpar$ space:
\begin{equation}
|\kpar| \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
\end{equation}
where $c$ is the speed of light and $S$ is a parameter accounting for offsets related to the combined spectral smoothness of the foregrounds and the antenna response (see Fig. \ref{fig:wedgeWithBG}).  
A key issue to measuring the EOR power spectrum is to understand and minimize such effects.  For an analytical derivation, see \cite{2012ApJ...756..165P,vedantham_et_al2012,liu_et_al2014b},  in simulation see \cite{datta_et_al2010,hazelton_et_al2013} and for observations
\cite{2013ApJ...768L..36P,2015arXiv150601026P,2014ApJ...788..106P,2015arXiv150206016A}.

\begin{figure}[t]
\includegraphics[height=3in]{plots/wedgeWithBG.png}
\caption{Foregrounds are a primary challenge facing 21\,cm cosmology experiments. 
HERA leverages a ``wedge'' in the spherically-averaged ($\mathbf{k}$ is broken into $k_\|$ and $k_\perp$) power spectrum (center panel; \citealt{dillon_et_al2015}). Smooth-spectrum foregrounds (right panel) $\sim$5 orders of magnitude brighter than fiducial EoR models (left panel; \citealt{mesinger_et_al2011}) create the ``wedge'' when they interact with the interferometer's chromatic response. By avoiding foregrounds, PAPER has placed limits within an order
of magnitude (in mK) of these models \citep{ali_et_al2015} and show the ``EoR Window'' to be foreground-free.
HERA's dish and configuration optimize wedge/window isolation and direct sensitivity to low-$k_\perp$ modes where EoR is brightest.
%v2
Cartoon of the wedge.  The gray rectangle represents the smooth-spectrum foregrounds as they would be if the instrument had a perfect achromatic response.  The pink wedge area shows the chromatic response of the interferometer, where additional power is thrown up from the gray region to higher k-values with increasing baseline.  The additional horizontal pink line illustrates the affect that other systematics, such as reflections on cables, can have in contaminating the measurement.  The white region shown the 'eor window', where the measurement is not contaminated by foregrounds or systematics.}
\label{fig:wedge}
\end{figure}

The techniques to measure the power spectrum may be broken down into two principle techniques (delay-space and map-making), plus additional hydrid methods.  These are briefly summarized below, with an emphasis on the delay-spectrum techniqe as the initial approach taken by HERA.  Finally, we provide a short discussion on calibration in the concept of this design.

\subsection{Delay-Spectrum Approach}
\label{sec:delayapproach}
Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement.  There are now numerous papers in the literature on the technique of measuring the power spectrum using an interferometer and the attendant issues [[[list of citations]]].  Below, a high-level overview based on  \cite{2012ApJ...756..165P} will be presented. The so-called delay-spectrum approach leverages this response to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.  Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum.

We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment:
\begin{equation}
{\hat P}(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} \right]
\label{eq:Pk_baseline}
\end{equation}
where $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes, $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.  The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.  
Note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
Rather than ${\hat P}(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.

Since the thermal noise per visibility baseline may be expressed as
\begin{equation}
V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
\label{eq:sensitivity_per_baseline}
\end{equation}
where $\tau$ is the integration time (see {\em e.g.} TMS),
to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq. \ref{eq:Pk_baseline}.  Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
\begin{equation}
\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
\label{eq:sensitivity}
\end{equation}
where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement 
(see \cite{2012ApJ...756..165P}).  Fig. \ref{fig:boost} plots this ``boost'' factor for various telescopes under nominal observing campaigns using the delay-spectrum technique.  The large improvement for PAPER and HERA is the fact that the baselines are redundantly placed on appropriate spatial scales, so the same baselines gets many samples.

\begin{figure}[h!]
\includegraphics[width=0.5\textwidth]{plots/boost.png}
\caption{Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) for representative observing campaigns for various instruments.}
\label{fig:boost}
\end{figure}

The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by the antenna baselines and bandwidths.  Note that largest bandwidth (which determines the smallest $\kpar$) is about 10 MHz, for larger bandwidths cosmic evolution begins to impact the result.  Figure \ref{fig:kperf} shows the allowed k-values based on the beamwidth/baseline (gray plus dark-blue overlap regions, the black lines are the limits) and a bandwidth of 10 MHz with 1024 channels (the blue and dark-blue overlap regions, the blue line is the lower limit).  The red-dashed line is the allowed smallest $k$-value attainable with this technique.  Note that at these red-shifts it is essentially set just by the bandwidth.

\begin{figure}[h!]
\centerline{
\includegraphics[width=\textwidth]{plots/kperf.png} 
}
\caption{\small k
\label{fig:kperf}}
\end{figure}

\subsection{Map-Making Approach}
\label{sec:mapapproach}
In contrast to the delay-spectrum approach, map-making approaches combine baselines to build up information before squaring and calculating the power spectrum.  The conceptually easiest approach is that one makes an image of the sky and takes a 2-D Fourier transform to determine the power spectrum and so this approach is often called the imaging approach.  Advantages are that you actually image the reionized bubbles and can determine the magnitude of the smaller $k$-modes on the sky.  The principle disadvantage is that you have to compute and calibrate the effects of systematics and the propagation of the very bright foregrounds through your system and subtract them incredibly accurately.  In contrast, the delay-spectrum approach localizes those effects to a parameter space where they can much more easily be filtered and not used in subsequent analysis.

\subsection{Hybrid Approaches}
\label{sec:hybridapproach}

\subsection{Calibration}
\label{sec:calibration}

\section{High-Level Requirements} 
\label{sec:requirements}
From the science and techniques discussed above, we may derive some high-level requirements.  This section describes frequency and sensitivity requirements that largely set the overall design concept.

\subsection{Frequency and Bandwidth}
Figure \ref{fig:cosmos} indicates the frequency range requirement to probe the expected timescale of the epoch of reionization using the 21cm line of hydrogen as the probe.  
These limits are derived from to-date complementary probes of reionization which include measurements of
the optical depth to last scattering in the CMB, QSO spectra, Ly-$\alpha$
absorption in the spectra of quasars and gamma ray bursts and the demographics
of Ly-$\alpha$ emitting galaxies
(Figure~\ref{fig:IonHist}). 
Constraints from these probes are still weak:
Ly-$\alpha$ absorption saturates at very small neutral fractions; galaxy
surveys directly constrain only the bright end of the luminosity function and
depend on an unknown escape fraction of ionizing photons to constrain
reionization; CMB measurements probe an integral quantity subject
to large degeneracies. Even when these observations are
combined into a single 95\% confidence region, the bounds remain weak.
For example, $x_{HI}$ spans almost the entire allowable range of [0,1]
at $z=8$. 21\,cm reionization experiments place much tighter constraints on ionization, with the
red band showing the forecasted 95\% confidence region derived from HERA data,
after marginalizing over astrophysical and cosmological parameters.

From these measurements and models, the Phase 1 and 2 minimum spans of redshifts range from $z=$6 to 13, corresponding to a frequency range of 100 - 200 MHz.  Extending the science to include the Dark Ages remains a goal so that efforts are on-going to design a feed to increase the lower limit down to about 50 MHz without compromising the performance in the specification range.  The fall-back position for the low-frequencies is a serial deployment of a scaled version of the HERA feed or to potentially build additional elements specifically for the low-frequencies.  The 100 MHz bandwidth is also the minimum fully digitized and correlated bandwidth.

The scientific requirement on channel bandwidth is to allow access to k-modes of greater than 1 h/Mpc [ref], which requires about 256 channels over the 100 MHz bandwidth.  However, in order to handle radio frequency interference as well as to allow the bandpass to be characterized, a specification of 1024 channels has been chosen.  This yields a channel bandwidth of 97.7 kHz.

As mentioned earlier (see Fig. \ref{fig:kperf}) we should be able to process at least 10 MHz

\begin{figure}[t]
\centerline{
\includegraphics[width=7.5cm]{plots/costfigNew.png} 
}
\caption{\small Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).  The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.  SEE TEXT...
\label{fig:costfig}}
\end{figure}

\subsection{Sensitivity Optimization}
\label{sec:cost}
As an instrument to characterize the power spectrum over its evolution, the specification on sensitivity is to make at least a nominal detection over the redshift region of support, or $z$=6-12, along with a very robust detection at the peak.  As seen in Fig. \ref{fig:Sensitivities} and Table \ref{tab:signif}.
To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, where the total number was constrained to keep constant sensitivity for this measurement, which is shown in Figure \ref{fig:costfig}.  This included the full system, except the post-processing.  By assuming canonical values of 331 14-meter antennas, the number of elements needed for a given diameter to match the sensitivity is
\begin{equation}
N_a = 4634/D_a
\end{equation}
where $N_a$ and $D_a$ are the number of antennas and their diameter respectively.  Note that the number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} Pober) over a range of values for the HERA configuration.  This dependency differs from Mellema, mostly likely due to the assumed configuration.
Figure \ref{fig:costfig} shows the resulting normalized system cost/performance-curve as the horizontal shading relative to the canonical HERA design 
which becomes very flat (colored purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.

\section{System Design} \label{sec:design}

\noindent As described in the previous section, we have applied 
critical insights from first generation 21\,cm EoR experiments to define
the requirements for HERA---an instrument that ensures foregrounds remain bounded
within the wedge while \emph{delivering the sensitivity for
high-significance detections of the 21\,cm reionization power spectrum with
established foreground filtering techniques} 
\citep{pober_et_al2014,greig_and_mesinger2015}.
In this section, we summarize key features of the HERA design (see Table~\ref{tab:BasicParameters}) 
and system architecture (see Fig.~\ref{fig:overallBlockDiagram}).
This architecture
directly inherits from the PAPER and MWA experiments; HERA begins by reusing the
analog, digital, and real-time processing systems deployed for PAPER-128.  As HERA develops, this
architecture is incrementally upgraded to improve performance and add features while simultaneously
addressing issues of modularity and scalability.  As with PAPER, HERA proceeds in stages of development,
with annual observing campaigns driving a cycle of development, testing, system integration, calibration, and analysis.
This cycle ensures that HERA's instrument is always growing, that systematics are being found and eliminated at the
earliest build-out stages, that data analysis pipelines are tested and debugged while data volumes are smaller,
and that HERA is always producing high quality science.

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{l | l}
\multicolumn{1}{c}{\emph{\textbf{Instrument Design Specification}}} & \multicolumn{1}{c}{\emph{\textbf{Observational Performance}}}\\
\hline
\textbf{Element Diameter:} 14\,m & \textbf{Field of View:} 9\arcdeg \\
\textbf{Minimium Baseline:} 14.6\,m & \textbf{Largest Scale:} 7.8\arcdeg\\
\textbf{Maximum Core Baseline:} 292\,m & \textbf{Core Synthesized Beam:} 25\arcmin\\
\textbf{Maximum Outrigger Baseline:} 876\,m & \textbf{Outrigger Synthesized Beam:} 11\arcmin\\
\textbf{Frequency Range:} 50--250\,MHz & \textbf{Redshift Range: $4.7 < z < 27.4$} \\
\textbf{Frequency Resolution:} 97.8\,kHz & \textbf{LoS Comoving Resolution:} $1.7$\,Mpc (at $z=8.5$)\\
\textbf{Survey Area:} $\sim 1440$ deg$^2$ & \textbf{Comoving Survey Volume:} $\sim 150$\,Gpc$^3$ \\
$\mathbf{T_\textbf{sys}}$: $100 + 120 (\nu/\rm{150~MHz})^{-2.55}$ K & \textbf{Sensitivity after 100\,hrs:} 50 $\mu \rm{Jy}~\rm{beam}^{-1}$  \\
\hline
\end{tabular}
\Caption{-0.1in}{0.99}{-0.4in}{HERA-350 design parameters and their observational consequences. Angular scales computed at 150\,MHz.}
\label{tab:BasicParameters}
\vspace{.2in}
\end{center}
\end{table}

\begin{figure}[h]
	\centering
	\includegraphics[width=1\textwidth]{plots/HERA_high_level_block_diagram.pdf}
	\caption{HERA's signal path.  Front-end amplifiers at the antenna feed drive signals on short coaxial cables to 
field nodes.  Nodes contain post-amplifiers and Smart Network ADC Processor (SNAP) boards that digitize, channelize,
and packetize data for optical transmission in 10 Gb Ethernet format.  Optical fibers are aggregated in a field container
onto a 10 km fiber bundle connecting to the Karoo Array Processing Building, where signals are cross-multiplied
in the X processor.  After correlation, visibilities are stored by the Librarian, compressed and redundantly calibrated
by the Real-Time Processor, and transmitted over the network to clusters for storage and analysis.  Final products are
hosted on public-facing NRAO servers, with a web interface for selecting and downloading data.}
	\label{fig:overallBlockDiagram}
	\vspace{-5pt}
\end{figure}

\subsection{Antenna Element}
\label{sec:antenna}
\vspace{-5pt}

As discussed in the previous section, the design of the design principles are three-fold:
\begin{itemize}
\item optimize for the delay-spectrum technique of measuring the EOR power spectrum,
\item minimize costs, and
\item the experiment has a limited lifetime of about five years.
\end{itemize}
The first item primarily means that chromatic effects corresponding to delays appropriate for the measurement described above must be below the expected signal level, which essentially determines the focal length.  The second item constrains the diameter and element count, as well as the focal length over diameter ratio ($f/D$), based on a cost function and maximizing sensitivity per element.  And the third items constrains the construction materials and methods and the operational model.  This led to a fixed transit element with a diameter of 14 meters.

The construction materials are readily available standard construction materials, making for a very cost-effective design.  The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.  Given the close-packed design, each pole (except for the perimeter) are shared between three antennas.  This both reduces the number of poles, but also makes for a balanced load.  Note that the perimeter poles can be stayed with guy lines if needed.  A close-up of one antenna is shown in Fig \ref{fig:elementpic}.  The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.

The center of the antenna is a cast concrete donut-shaped hub with 12 angled sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.  The twelve spars are made of 60mm PVC pipe, which are stressed along three support points to approximate a parabolic shape.  Note that an ideal moment-loaded beam actually attains a true parabola.  Note that the three support points (one at the hub, one about 3 meters out radially and one at the rim-end) are at the correct height and angle for the underlying parabola.

The shape between the spars is actually a one-dimensional parabola rather than a two-dimensional paraboloid, so the dish surface may more properly be called a ``facetted parabola''.  If only twelve spars were used for the entire surface, simple Ruze losses would be quite high (about 10\% at 150 MHz).  Therefore, at about 2.1 meters another horizontal member is placed, which launches another parabolic spar.  The simple Ruze loss is then less than 1\%, which is much more acceptable.  Fig \ref{fig:elementplan} shows this surface scheme and the panelization (left) and the offsets (in mm) of that scheme from an ideal parabola.  The panel that is ``left out'' is the location of a door to allow access into the hub (via a small removable bridge).  The additional intermediate spars also make the panel sizes much more manageable and better secured.  The panel arrangement is also set by a standard width of mesh roll of 1.22m.

\begin{figure}[h!]
\centerline{
\includegraphics[height=6cm]{plots/antPlan.png}
\includegraphics[height=6cm]{plots/offset.png} 
}
\caption{\small Element.
\label{fig:elementplan}}
\end{figure}

The feed is to be deployed along the two separate phases.  Initially, the feed reuses the dipole portion of the PAPER element, however in a new optimized cylindrical feed-cage configuration.  The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on dish), standing waves (blockage, {\em e.g.} Baars), and frequency response.  The design parameters were size of backplane, height of mast, height of skirt, and cone versus cylinder.  Cones were ruled out since it was found they introduce additional frequency structure with no real improvement in performance and a range of cylinders were evaluated.  The phase 1 design is documented in a memo \cite{feedmemo}, while additional studies are on-going at NRAO, University of Cambridge and Stellenbosch University.  The chosen feed design is shown in Figure X.

Additionally, to lessen cross-coupling, a screen fence around the perimeter is also under investigation.  Rather than being vertical, it will have a slight angle and potentially some edge treatment to try and lessen diffraction.  Tests of the antennas are discussed in a number of papers CITE.  Figure Y shows beam patterns.

As mentioned, the feed is supported by lines from the three poles via a rigid support to which the feed cage and dipole attach.  Another line goes down centrally to the hub to hold the feed rigidly in place at the correct height.

The low noise amplifier/balun integrates directly to the base of the feed dipole in the same configuration as per PAPER (REF).  The 150m length of 75 $\Omega$ cable also runs down centrally and then over to the post-amplifier module, which is housed near the RFI-shielded container containing the digital electronics.

\noindent The novel design of HERA's antenna element (Figs. \ref{fig:HERApictures} and \ref{fig:orbcommexptandbeammap}), now extensively field-tested with MSIP seed funding, 
is one of the critical advances that enables HERA to achieve its science
goals cost-effectively.  The 14-m fixed zenith-pointing parabolic dish strikes an optimal
balance between sensitivity and systematics.  The large collecting area of a
HERA element yields nearly 5 times the sensitivity of an MWA tile and more
than 20 times that of a PAPER element, but it does so without substantially
degrading our ability to isolate and remove foreground emission on the basis of
spectral smoothness (Fig. \ref{fig:reflectometry}, right panel).  As shown in
\Mycitet{parsons_et_al2012b}, the amplitude and timescale of signal reflections
relates directly to the leakage of smooth-spectrum foregrounds into regions of Fourier space 
used to measure reionization.
To facilitate foreground filtering, HERA's antenna element is designed to suppress reflections at long time delays.

\begin{figure}[h!]
	\begin{tabular}{ll}
	\begin{minipage}{4.2in}
\includegraphics[width=2.1in]{plots/ref_dipole_and_hera_dish.jpg}
\includegraphics[width=2.1in]{plots/orbcomm_dish_beam_map_530cm_feed.pdf}
	\end{minipage} & 
	\begin{minipage}{2.05in}
	\caption{Left: The first of three prototype dishes at NRAO--Green Bank, used for measuring beam frequency structure with reflectometry and the beam pattern at 137\,MHz by comparing satellite signals to the reference dipole in the foreground \citep{neben_et_al2016}. Right: The measured EW power pattern plotted with dashed lines marking zenith angles of 20$^\circ$, 40$^\circ$, 60$^\circ$, 80$^\circ$.} 
	\label{fig:orbcommexptandbeammap}
	\end{minipage}
	\end{tabular}
	\vspace{-15pt}
\end{figure}

\begin{figure}[h!]
	\centering
    \includegraphics[width=0.49\textwidth]{plots/s11_compare_msip.pdf}
	%\includegraphics[width=0.49\textwidth]{plots/compare_kernels_msip_subband.pdf}
\includegraphics[width=0.5\textwidth]{plots/ps1d_with_delay_kernel.pdf}
	\vspace{-25pt}
	\caption{Left: Electromagnetic simulations (black; \citealt{ewallwice_et_al2016}) of the response of the HERA dish as a function of time delay, $\tau$,
       agree with field measurements (grey; \citealt{patra_et_al2016}). 
Right: Modeled foreground power spectra with (green) and without (blue) the antenna chromaticism over a 10\,MHz subband centered at $z=7.5$. The spectral smoothness of the dish enables foreground isolation below the level of a fiducial reionization signal (black) for most $k_{\parallel}$ modes, yielding percent-level constraints on model parameters.}
	\label{fig:reflectometry}
	\vspace{-10pt}
\end{figure}

As shown in Fig.~\ref{fig:reflectometry}, HERA's element meets this requirement in
reflectometry tests, performing comparably to the PAPER element.  The
relatively short (4.5\,m) focal height of the parabolic dish 
minimizes the timescale of feed-dish reflections while maintaining the feed's
illumination of the dish.  HERA's feed design is based on the PAPER sleeved
dipole element, inverted above a backplane, with a skirting cylinder to
improve symmetry in the polarized beam response patterns.  Figure
\ref{fig:orbcommexptandbeammap} shows the beam pattern measured at 137 MHz with a beam mapping
system using the ORBCOMM satellite network \Mycitep{neben_et_al2016}.  Results
indicate an effective per-element collecting area of 93\,m$^2$, compared with the
theoretical maximum Airy response of 155 m$^2$.  The measured primary beam is
consistent with simulations at the 0.1\% to 0.5\% level, with a full width at
half maximum of $\sim$$10^\circ$, and a first sidelobe at -20 dB \citep{ewallwice_et_al2016,neben_et_al2016,patra_et_al2016,thyagarajan_et_al2016}.
Feed-to-feed coupling between adjacent antennas has been measured to be below -50 dB, indicating
that mutual coupling will not be problematic.

With observing wavelengths of 1.5 to 3 m, meeting the required accuracy
in placing the components of HERA's dish is straightforward.
Dish centers are fixed
by a concrete hub placed with final surveyed accuracy of $\sim$10 cm.  These
hubs constrain radial PVC spars, tensioned into approximate parabolas against a
rim supported by utility poles.  The resulting faceted paraboloid has
a measured RMS surface accuracy of 3 cm, which is well within the 0.1$\lambda$ tolerance for
coherent phasing.  Feed placement relative to the dish is the most stringent requirement, with
a tolerance of 5 cm.  This placement is ensured by spring-tensioned Kevlar lines attached to
three utility poles at the dish rim and a 165-lb tensioning down to the concrete hub with
a fixed length line that ensures a correct focal height.  With 19 dishes constructed in South Africa
and 2 in Green Bank, we are confident in the ability of our trained field teams to construct HERA
elements to specification.

\subsection{Array Configuration}
\label{sec:arrayConfig}

\noindent HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments 
to cover the $uv$-plane with sub-element sampling density (see Fig.~\ref{fig:arrayConfig}). The dense core maximizes 
sensitivity on the short baselines for which PAPER's proven foreground filtering strategy works best.
The core is supplemented by 30 additional outrigger elements 
to tile the $uv$-plane with instantaneously complete sub-aperture 
sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz). This sampling 
strategy suppresses grating lobes in the synthesized beam and provides information for calibrating and correcting 
direction-dependent antenna responses \citep{dillon_parsons2016}.
Even with the sub-aperture dithering and long baselines, all 350 HERA elements can be robustly calibrated by taking advantage of 
its highly redundant configuration \citep{liu_et_al2010,zheng_et_al2014}. 
Resulting calibration errors range from $\sim5\%$ (in the core) to $\sim10\%$ (for outriggers)
of the residual fractional noise per antenna after averaging.

\begin{figure}[h]
	\centering
	\vspace{-5pt}
	\includegraphics[width=1\textwidth,clip]{plots/HERA_Array_Config.eps}
	\vspace{-25pt}
	\caption{HERA's elements are divided between a 320-element, hexagonally-packed core and 30 outriggers (left). This produces instantaneous
$uv$ coverage at triple the element spacing out to 250$\lambda$ at 150\,MHz (middle). All 350 elements can be redundantly calibrated 
using the \citet{liu_et_al2010} technique, yielding calibration errors that are a small fraction of the residual noise per antenna (right).}
		\label{fig:arrayConfig}
			\vspace{-10pt}
\end{figure}

\subsection{Analog Signal Path}

\noindent HERA's analog signal path, which operates from 50--250 MHz, emphasizes spectral smoothness and robustness.
With receiver temperatures of 50--100 K readily achievable with ambient temperature electronics, 
sky noise dominates the system temperature of telescopes operating below 300 MHz.  As a result,
HERA's low-noise amplifiers (LNAs) are inexpensive, passively cooled components integrated into the
antenna feed.  
HERA's signal path emphasizes careful impedance matching between each component to avoid signal reflections that can induce foreground leakage.
In particular, impedance matching between free space (377$\Omega$) and coaxial cable (50$\Omega$) is undertaken
in the feed/balun at the front end of the signal chain, where the time constants of inevitable reflections are
minimized.  For a similar reason, HERA houses analog-to-digital converters (ADCs) in nodes near the antenna elements
to limit the total analog path length.

While building to 128 elements during the first two project years, HERA reuses PAPER's tested and characterized 
LNAs, cables, and post-amplifier modules, operating
from 100--200 MHz.  In the third project year, these components are replaced by the 50--250 MHz, length-constrained signal
path described above, supporting HERA's pre-reionization science and extending the lever-arm for modeling and suppressing smooth-spectrum foreground emission.

\subsection{Digital Signal Path}
\label{sec:digital}

\noindent Although correlator development has historically been one of the most 
expensive aspects of building a large radio interferometer, this is no longer the case.
CASPER \Mycitep{parsons_et_al2006}
open-sourced the development of digital signal processing engines for astronomy and
now has world-wide participation,
with over 500 members at 73 institutions, and 
five generations of hardware.
On a modest budget, PAPER applied CASPER technology to develop and deploy new correlators
annually for five years running, each quadrupling the computational capacity of its predecessor.
Led at UC Berkeley's Radio Astronomy Lab (RAL),
HERA efforts continue this incremental development cycle, using a packet-switched
correlator architecture \Mycitep{parsons_et_al2008} that both PAPER and LEDA have
applied to systems employing FPGAs and GPUs \citep{clark_et_al2011}.

In order to meet the 35\,m specification for a maximum analog signal path, along with a growing need for system scalability, HERA beyond the first 128 elements will adopt a node-based architecture for amplification, digitization, channelization, and digital
transmission in the field, building on MWA heritage. This design is merged with PAPER's clean 
architecture for real-sampling and channelizing the entire analog passband at once, packetizing the data into
10 Gb Ethernet format, and relying on commercial switches to perform the frequency/antenna corner-turn required for FX correlators

{\bf Node}. HERA-350 employs RFI-tight node enclosures that each contain the final gain and digitization stages for
signals from 12 antennas, along with power supplies, cooling, sensors, and a small server for monitor/control.  
As part of previous HERA support,
a new Smart Network ADC Processor (SNAP; Fig.~\ref{fig:hardware}) board has been incorporated 
into the Collaboration for Astronomy Signal Processing and Electronics Research (CASPER) suite of hardware and firmware. This inexpensive board was co-designed by UC Berkeley and NRAO to be
both the digitizer and F-engine in HERA's FX correlator architecture.
Each SNAP board 
digitizes and channelizes a 0--250 MHz band for 6 input signals (3 antennas, dual-polarization).
A 200-MHz band of selectable channels is transmitted over optical fiber
to a central container (see below), and on to the correlator.  Laboratory
tests have successfully demonstrated robust transmission from the SNAP board, through 10 km of fiber optic cable, 
into a 10 GbE switch using inexpensive, commercially available optical transceivers.
Activities under this proposal include 
testing and extending FPGA firmware,
integrating and testing all components in the node subsystem, and providing a monitor/control
access interface.

\begin{figure}[h]
\centering
%\includegraphics[width=2.2in]{plots/ROACH2.png}
\includegraphics[width=1.0\textwidth]{plots/casper_boards.png}
%\includegraphics[height=1.7in]{plots/snap_board.png}
\caption{
Six generations of CASPER digital signal processing (left to right) culminating in the SNAP board (right, along with the long-haul fiber-link test setup), which was developed and tested under the previous HERA MSIP award.
By preserving its toolflow, signal processing libraries, and interface code between hardware generations,
CASPER combines with a scalable correlator architecture to enable the PAPER correlator to be 
easily upgraded for HERA \Mycitep{parsons_et_al2006,parsons_et_al2008}.
%
}\label{fig:hardware}
\end{figure}

{\bf Central Container}.
HERA's central container houses two significant subsystems adjacent to the array.  The first is a timing subsystem
that maintains a GPS-disciplined oscillator and distributes timing
signals (the sampling clock and 1 pulse-per-second synchronization) to the nodes.  The second
subsystem is a passive fiber optic patch panel that couples
the optical network from the nodes into the 192-filament optical fiber bundle 
that connects to the KAPB.

{\bf Karoo Array Processing Building (KAPB)}.
The KAPB houses the switch and CPUs
that 
complete the HERA correlator system.  A fiber optic bundle enters the KAPB and patches
into local fiber optic cables. These cables terminate in optical transceivers which plug into a 240-port 10 GbE switch.
Such switches, while large, are readily available commercially today.  Also connected to
this switch are 30 servers, each hosting two dual-GPU graphics cards and two dual
10 GbE network interface cards, which implement the cross-multiplication (X-Engine) component
of the correlator during observations.  This estimate for the number of X-Engine servers
is extrapolated from current GPU servers deployed on PAPER, assuming no improvement in bus
speeds for transferring data into the GPU cores, but assuming that the computational
capacity of GPU cores doubles once according to Moore's Law prior to purchasing
these servers in year three of the project.
Output data from the correlator are written to the data storage system described
in the following section.

\subsection{Data Management}
\label{sec:data}
\noindent The HERA correlator generates $\sim4$~TB of raw data per 12-hour observing day.
HERA's data management system archives this on-site in a 1.5 PB storage array, calibrates it in real-time
on the basis of redundancy, compresses it by a factor of 20, and transfers it to our computing facility at NRAO.
This real-time pipeline, based on a similar system for PAPER, is mature, computationally manageable, and 
relatively agnostic with respect to observing parameters, making it a safe and robust first step in data processing
\citep{zheng_et_al2014,parsons_et_al2014,ali_et_al2015}.
The Internet link between the Karoo and data centers in the U.S.\ has been extensively tested, with assistance from SKA-SA, TENET
(South African to London link provider) and GEANT (London to U.S.\ provider).   
Transfer speeds are sufficient for streaming compressed data to the U.S.\ via the Internet. Portable storage is also
provisioned as a fallback transfer method.

With support from this grant, NRAO hosts the archiving, indexing, and processing of HERA data products. 
Project data and public data products are served via the existing 
NRAO archive at the Domenici Science Operations Center %(DSOC) 
in Socorro, New Mexico, where it will be accessible through computing accounts and public web-based
archive searches.
NRAO also supports
an attached 30-node processing cluster 
coupled to 
scratch storage
for routine data inspection and lightweight analysis tasks.
For the bulk application of processing-intensive data pipelines to HERA data,
NRAO manages the dynamic acquisition of computing from
external HPC and cloud-computing facilities, including the Amazon spot market and XSEDE. 
Combined with legacy computing resources at UPenn and MIT, this plan leverages NRAO's
expertise in managing and acquiring computing resources while providing distributed
permanent resources for HERA's basic processing needs.

%This effort leverages existing infrastructure at NRAO, with support 
%for the expansion of the data storage and for upgrading to a 30-node computing cluster.  This cluster supports the bulk of the analysis by HERA collaborators (see \S\ref{sec:analysis}) that requires access to the full set of HERA observations.

\subsection{Analysis Pipelines} 
\label{sec:software}

\noindent HERA builds on the rich legacy of PAPER and MWA software and database systems 
developed for field operations, data analysis, and simulation.  Examples range from
the strictly versioned and unit-tested packages for field-deployed systems (e.g. the correlator, real-time processing, and
monitor/control systems) to loose collections of scripts written for exploratory analysis.
Software packages that support HERA analysis are open source, publicly 
hosted\footnote{Including \url{github.com/AaronParsons/aipy}, \url{github.com/JeffZhen/omnical}, 
\url{github.com/MiguelFMorales/FHD}, \url{github.com/MiguelFMorales/eppsilon}, \url{github.com/jpober/21cmSense},
and \url{github.com/Nithyanandan/PRISim}.}, revision controlled, and unit-tested.  
These include Omnical, a complete package for redundant baseline calibration;
Astronomical Interferometry in Python (AIPY), a set of
tools and file-format interfaces for reading visibilities, calibrating,
rephasing, imaging, and deconvolution; Fast Holographic Deconvolution
(FHD), a purpose-built tool for imaging, calibration, and foreground
forward-modeling and subtraction \citep{sullivan_et_al2012}; Precision Radio Interferometry Simulator 
(PRISim), a package for accurately simulating wide-field interferometric observations;
21cmFAST, a fast, semi-numerical 21~cm signal simulator,
and 21cmSense, a tool for forecasting power spectrum sensitivity.
Other project code is aggregated and revision
controlled in a public repository with separate sandboxes for each developer to ensure that
HERA members have up-to-date copies of all project code to facilitate sharing and debugging.
Such code includes the PAPER pipeline for foreground filtering and estimating power spectra from
visibility data,
the MWA power spectrum analysis codes---\eppsilon\ \citep{hazelton_et_al2016} and the 
empirical covariance technique of \citet{dillon_et_al2015}---as well as
machine-learning-based source finding, verification, and removal tools \citep{caroll_et_al2016,jacobs_et_al2016,beardsley_et_al2016}.

New software development will focus on integrating and improving the MWA and PAPER power spectrum and foreground removal pipelines,
developing a monitor and control software interface and database for recording instrument metadata,
extending, with support from Scuola Normale Superiore, semi-analytic and numerical models 
of the 21~cm signal for robust parameter estimation, and developing
machine learning interpolations of simulations for joint Monte Carlo estimation of cosmological and astrophysical parameters.

\section{Schedule and Status}
\label{sec:status}

\noindent On-site construction will proceed in stages, starting from the
initial 37 elements that will be in place by project start.  These will be used
in conjunction with the thoroughly characterized extant PAPER signal path and
processing hardware for a very low risk initiation of scientific observations
from the very beginning.  Additionally, the 19 HERA elements currently in
place will also have had an extensive period of commissioning
and characterization.  As elements 38 to 128 are installed in year 1, they can
immediately be placed within the array and be used for observing.

During this period, infrastructure for the new node-based system is installed
and tested with the first elements beyond 128.  After the HERA-128 observing
season, the full array is transitioned to use HERA's new hardware infrastructure.  In this same time frame, the existing PAPER processing
container is moved to the edge of the array to house the timing sub-system
and fiber optic splice cabinet while the correlator will move to the KAPB.
Also, the data processing activity at Penn is transitioned to NRAO.
Summarizing, the hardware deployment phases are in four categories:
\begin{enumerate}
\item Element construction:  linear process (years 1, 2 \& 3)
\item Infrastructure:  (a) node power and fiber reticulation (years 1 \& 2), (b) container relocation (year 1), (c) node installation (years 2 \& 3)
\item Signal path:  convert to node-based system after 128 elements (years 2 \& 3)
\item Processing:  (a) data processor/server to NRAO (year 1), (b) X-engine \& real-time processor/storage to KAPB (year 2)
\end{enumerate}

\begin{figure}[h]
	\vspace{-7pt}
	\begin{tabular}{ll}
	\begin{minipage}{5in}
		\includegraphics[width=4.9in]{plots/timeline_short.pdf}
		\end{minipage} & \hspace{-.15in}
	\begin{minipage}{1.35in}
\captionsetup{justification=raggedright,
singlelinecheck=false
}
	\caption{Timeline of HERA construction, analysis development, observation, and scientific output.}
	\label{fig:timeline}
	\end{minipage}
	\end{tabular}
	\vspace{-8pt}
\end{figure}

The optimal EoR observing window is from September to April, with power spectrum limit results appearing about one year later.  Concurrent technique development and deployment will be on-going. Additional information is in the project management plan.
\begin{itemize}[leftmargin=0.7in]
\item[Year 1:] Observe with H37.  Real-time data pipeline, delay-space power spectrum (DSPS) pipeline, FHD pipeline.  21\,cm framework for incorporating with other probes. Construct H128.
\item[Year 2:] Observe with H128.  Real-time calibration pipeline.  DSPS/FHD/global sky model integration. Snapshot imaging pipeline.  EoR parameter estimation development.  H37 results.  Construct H240.  {\em Data products:  power spectrum, Stokes I maps.}
\item[Year 3:] Observe with H240.  Foreground-filtered imaging pipeline.  EoR parameter estimation development.   H128 results. Construct H350.  {\em Data products:  power spectrum, Stokes IQUV maps, foreground image cube.}
\item[Year 4:] Observe with H350.  EoR parameter estimates.  H240 results.  {\em Data products:  power spectrum, global sky model IQUV, snapshots, foreground-filtered image cube.}
%\item[Year 5:] Observe with H350.  H350 results.  {\em Data products:  power spectrum, global sky model IQUV, snapshots, foreground-filtered image cube.}
\end{itemize}

\section{Conclusion}
\label{sec:conclusion}

\noindent With first-generation instruments, the HERA team has led a revolution in our understanding of how to best perform 21\,cm cosmology measurements.
In the past three years, we have developed the EoR window paradigm for isolating foreground systematics, implemented novel
calibration and power spectrum analysis pipelines, made precision measurements of astrophysical foregrounds, and published deep power 
spectrum limits that constrain heating in the early universe.  HERA's design, now validated in the field with MSIP seed funding, 
is the result of these advances.

HERA is ``the right instrument at the right time.'' With first generation experiments approaching the limits of their sensitivity, it is now critical to build upon our new knowledge and experience to advance beyond them. HERA's large collecting area and optimized design enable precise constraints of EoR astrophysics and a broad range of high-impact secondary science, all before SKA-low becomes operational.
With CHAMP, HERA expands CAMPARE's proven methods and recruitment network to bring more students from underrepresented groups into cutting-edge astronomy research and support them with year-round mentoring.
By bringing the U.S.\ teams from PAPER, the MWA, MITEoR, and EDGES together with strategic international partners,
the HERA collaboration has consolidated theoretical and experimental leadership in the field, 
and is equipped to deliver the proposed science. By building HERA, we can provide the community with unprecedented insights into this key new area of cosmology.

\clearpage
\setcounter{page}{1}
\thispagestyle{empty}
%\bibliographystyle{apj}
%\bibliographystyle{hapj}
\nocite{Beardsley:thesis}
\nocite{kolopanis_et_al2016}
\bibliographystyle{jponew}
%\bibliographystyle{unsrt}
{\small \bibliography{biblio}}

\end{document}

### 2016-03-30 manuscript.tex
number of changes throughout, but mainly updated kperf figure

@@
 	\centering
-	\includegraphics[width=.56\textwidth]{plots/current_limits.pdf}
-	\includegraphics[width=.43\textwidth]{plots/constraints_proposal.pdf}
-	\vspace{-20pt}
-	\caption{Left: The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, (solid symbols represent analyses led by HERA collaborators) compared to 21cmFAST-generated models at $k=0.2$\,$h$\,Mpc$^{-1}$.
+	\includegraphics[width=.8\textwidth]{plots/current_limits.pdf}
+	\caption{The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, (solid symbols represent analyses led by HERA collaborators) compared to 21cmFAST-generated models at $k=0.2$\,$h$\,Mpc$^{-1}$.
 Analysis is still underway on PAPER and MWA observations that approach their projected 
 full sensitivities (see Fig.~\ref{fig:Sensitivities}); 
-HERA can deliver sub-$\text{mK}^2$ sensitivities.
-Right: The \citet{ali_et_al2015} limit at $z=8.4$ excludes the entire gray shaded area corresponding to a cold IGM \citep{pober_et_al2015}.
-The color scale shows the power spectrum amplitude at $k = 0.25~h{\rm Mpc}^{-1}$ for a given spin temperature and neutral fraction.
-Constraints are weaker if the (currently unknown) neutral fraction at $z=8.4$ is very high or low, as the 21\,cm signal is brightest during the middle of reionization.}
+HERA can deliver sub-$\text{mK}^2$ sensitivities.}
 	\vspace{-10pt}
@@
 \cite{2013ApJ...768L..36P,2015arXiv150601026P,2014ApJ...788..106P,2015arXiv150206016A}.
+Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).
+In order to normalize for updated measurements of Hubble's constant, it is further normalized by $h$, where $H_o=100h$, such that they are expressed as $h$/Mpc.
 
@@
 \centerline{
-\includegraphics[width=\textwidth]{plots/kperf.png} 
+\includegraphics[width=.4\textwidth]{plots/kperf.png} 
 }
@@
 
-From these measurements and models, the Phase 1 and 2 minimum spans of redshifts range from $z=$6 to 13, corresponding to a frequency range of 100 - 200 MHz.
-Extending the science to include the Dark Ages remains a goal so that efforts are on-going to design a feed to increase the lower limit down to about 50 MHz without compromising the performance in the specification range.
+From these measurements and models, HERA is required to cover a redshift range of $z=$6 to 13, corresponding to a frequency range of 101 - 203 MHz.
+We adopt 100-200 MHz for full performance as the requirement.
+Extending the science to include the Dark Ages remains a goal so that efforts are on-going to design a feed to increase the lower limit down to at least 70 MHz without compromising the performance in the above requirement range.
 The fall-back position for the low-frequencies is a serial deployment of a scaled version of the HERA feed or to potentially build additional elements specifically for the low-frequencies.

### 2016-03-30 manuscript.tex
some reordering of manuscript

@@
 have produced the first astrophysically constraining upper limits on the 21\,cm EoR power spectrum, 
-providing evidence for significant heating in the IGM prior to reionization \citep{parsons_et_al2014,ali_et_al2015,pober_et_al2015}.
+providing evidence for significant heating in the IGM prior to reionization \citep{parsons_etal2014,ali_et_al2015,pober_et_al2015}.
 However, current experiments cannot expect more than marginal detections of the EoR signal. 
@@
 \label{sec:eormeas}
-Due to the expansion of the Universe, we can identify and measure the early Universe via the red-shift of spectral lines.
+Due to the expansion of the Universe over cosmological time, we can identify and measure the early Universe via the red-shift of spectral lines.
 The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire Universe back almost to the period of recombination.
@@
 
+This section will provide a brief overview of existing measurements and the theoretical underpinning of how foregrounds contaminate the measurement (the so-called ``wedge''), discuss the various techniques used to make the measurement and finally provide a brief discussion on calibration issues and techniques.
+
+\subsection{The ``Wedge''}
 The past three
@@
 PAPER and the MWA have produced progressively deeper limits
-\citep{dillon_et_al2015,parsons_et_al2014,ali_et_al2015}, with PAPER
+\citep{dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
 yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
@@
 
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.8\textwidth]{plots/current_limits.pdf}
-	\caption{The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, (solid symbols represent analyses led by HERA collaborators) compared to 21cmFAST-generated models at $k=0.2$\,$h$\,Mpc$^{-1}$.
-Analysis is still underway on PAPER and MWA observations that approach their projected 
-full sensitivities (see Fig.~\ref{fig:Sensitivities}); 
-HERA can deliver sub-$\text{mK}^2$ sensitivities.}
-	\vspace{-10pt}
-	\label{fig:limits}
-	\label{fig:IGMtemperatureConstraints}
-\end{figure}
-
 Perhaps the most important advance informing HERA's design is a
@@
 Through theoretical and observational work
-\citep{Datta_2010,morales_et_al2012,parsons_et_al2012b,vedantham_2012,thyagarajan_et_al2013,hazelton_et_al2013,pober_et_al2013b,liu_et_al2014a,liu_et_al2014b},
+\citep{datta_etal2010,morales_et_al2012,parsons_et_al2012b,vedantham_2012,thyagarajan_et_al2013,hazelton_et_al2013,pober_etal2013b,liu_et_al2014a,liu_et_al2014b},
 we have learned how the boundary between the wedge and our ``EoR window" is determined by the separation between antennas,
@@
 that, to the limits of current sensitivity, foreground emission is absent outside of the wedge; it can only 
-appear there through instrumental leakage \citep{parsons_et_al2014,ali_et_al2015,moore_et_al2016,kohn_et_al2016}.
-Thus, to open the widest possible window for EoR measurements, HERA must use close-packed antennas that
-minimize signal reflections and deliver significant forward gain relative to their horizon response.
-Tests with prototype HERA antennas (Figs. \ref{fig:orbcommexptandbeammap} and \ref{fig:reflectometry}, discussed in \S\ref{sec:antenna})
-indicate that a moderately large parabolic dish with a short focal height can meet these requirements
-\citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016}.
-
+appear there through instrumental leakage \citep{parsons_etal2014,ali_et_al2015,moore_et_al2016,kohn_et_al2016}.
+
+Though the desired measurement is over a volume, it is instructive to split it into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kvpr$ is determined by the antenna baseline and $\kpar$ by the frequency.
+This is useful since it allows us to split out the chromatic response of the interferometer and isolate a phase space where foreground sources ({\em i.e.} everything {\em not} the EOR) contaminate the signal of interest from where they don't.
+This contaminated phase space may be shown [REF] to be a wedge-shaped region in $\kpr$:$\kpar$ space:
+\begin{equation}
+|\kpar| \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
+\end{equation}
+where $c$ is the speed of light and $S$ is a parameter accounting for offsets related to the combined spectral smoothness of the foregrounds and the antenna response (see Fig. \ref{fig:wedge}).  
+A key issue to measuring the EOR power spectrum is to understand and minimize such effects.
+For an analytical derivation, see \cite{zahn_etal2012,vedantham_2012,liu_et_al2014b},  in simulation see \cite{datta_etal2010,hazelton_et_al2013} and for observations
+\cite{pober_etal2013b,2015arXiv150601026P,parsons_etal2014,2015arXiv150206016A}.
+Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).
+In order to normalize for updated measurements of Hubble's constant, it is further normalized by $h$, where $H_o=100h$, such that they are expressed as $h$/Mpc.
+
+\begin{figure}[t]
+\includegraphics[height=3in]{plots/wedgeWithBG.png}
+\caption{Foregrounds are a primary challenge facing 21\,cm cosmology experiments. 
+HERA leverages a ``wedge'' in the spherically-averaged ($\mathbf{k}$ is broken into $k_\|$ and $k_\perp$) power spectrum (center panel; \citealt{dillon_et_al2015}).
+Smooth-spectrum foregrounds (right panel) $\sim$5 orders of magnitude brighter than fiducial EoR models (left panel; \citealt{mesinger_et_al2011}) create the ``wedge'' when they interact with the interferometer's chromatic response.
+By avoiding foregrounds, PAPER has placed limits within an order
+of magnitude (in mK) of these models \citep{ali_et_al2015} and show the ``EoR Window'' to be foreground-free.
+HERA's dish and configuration optimize wedge/window isolation and direct sensitivity to low-$k_\perp$ modes where EoR is brightest.
+%v2
+Cartoon of the wedge.
+The gray rectangle represents the smooth-spectrum foregrounds as they would be if the instrument had a perfect achromatic response.
+The pink wedge area shows the chromatic response of the interferometer, where additional power is thrown up from the gray region to higher k-values with increasing baseline.
+The additional horizontal pink line illustrates the affect that other systematics, such as reflections on cables, can have in contaminating the measurement.
+The white region shown the 'eor window', where the measurement is not contaminated by foregrounds or systematics.}
+\label{fig:wedge}
+\end{figure}
+
+The techniques to measure the power spectrum may be broken down into two principle techniques (delay-space and map-making), plus additional hydrid methods.
+These are briefly summarized below, with an emphasis on the delay-spectrum technique as the initial approach taken by HERA.
+Finally, we provide a short discussion on calibration in the concept of this design.
+
+\subsection{Delay-Spectrum Approach}
+\label{sec:delayapproach}
+Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement.
+There are now numerous papers in the literature on the technique of measuring the power spectrum using an interferometer and the attendant issues [[[list of citations]]].
+Below, a high-level overview based on  \cite{parsons_et_al2012b} will be presented.
+The so-called delay-spectrum approach leverages this response to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
+Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum.
+
+We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment:
+\begin{equation}
+{\hat P}(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} \right]
+\label{eq:Pk_baseline}
+\end{equation}
+where $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes, $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.
+The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.
+Note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
+Rather than ${\hat P}(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.
+
+Since the thermal noise per visibility baseline may be expressed as
+\begin{equation}
+V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
+\label{eq:sensitivity_per_baseline}
+\end{equation}
+where $\tau$ is the integration time (see {\em e.g.} TMS),
+to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq.
+\ref{eq:Pk_baseline}.
+Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
+\begin{equation}
+\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
+\label{eq:sensitivity}
+\end{equation}
+where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement 
+(see \cite{parsons_et_al2012b}).
+Fig.
+\ref{fig:boost} plots this ``boost'' factor for various telescopes under nominal observing campaigns using the delay-spectrum technique.
+The large improvement for PAPER and HERA is the fact that the baselines are redundantly placed on appropriate spatial scales, so the same baselines gets many samples.
+
+\begin{figure}[h!]
+\includegraphics[width=0.5\textwidth]{plots/boost.png}
+\caption{Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) for representative observing campaigns for various instruments.}
+\label{fig:boost}
+\end{figure}
+
+The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by the antenna baselines and bandwidths.
+Note that largest bandwidth (which determines the smallest $\kpar$) is about 10 MHz, for larger bandwidths cosmic evolution begins to impact the result.
+Figure \ref{fig:kperf} shows the allowed k-values based on the beamwidth/baseline (gray plus dark-blue overlap regions, the black lines are the limits) and a bandwidth of 10 MHz with 1024 channels (the blue and dark-blue overlap regions, the blue line is the lower limit).
+The red-dashed line is the allowed smallest $k$-value attainable with this technique.
+Note that at these red-shifts it is essentially set just by the bandwidth.
+
+\begin{figure}[h!]
+\centerline{
+\includegraphics[width=.4\textwidth]{plots/kperf.png} 
+}
+\caption{\small ADD NEW HERA CONFIGURATION AND LOFAR AND PUT IN 'HOW TO MAKE THIS PLOT'}
+\label{fig:kperf}
+\end{figure}
+
+\subsection{Map-Making Approach}
+\label{sec:mapapproach}
+In contrast to the delay-spectrum approach, map-making approaches combine baselines to build up information before squaring and calculating the power spectrum.
+The conceptually easiest approach is that one makes an image of the sky and takes a 2-D Fourier transform to determine the power spectrum and so this approach is often called the imaging approach.
+Advantages are that you actually image the reionized bubbles and can determine the magnitude of the smaller $k$-modes on the sky.
+The principle disadvantage is that you have to compute and calibrate the effects of systematics and the propagation of the very bright foregrounds through your system and subtract them incredibly accurately.
+In contrast, the delay-spectrum approach localizes those effects to a parameter space where they can much more easily be filtered and not used in subsequent analysis.
+
+\subsection{Hybrid Approaches}
+\label{sec:hybridapproach}
+
+\subsection{Calibration}
+\label{sec:calibration}
+
+\subsection{Existing Measurements}
+\label{sec:existing}
 PAPER adopted a foreground mitigation strategy based largely
@@
 
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.8\textwidth]{plots/current_limits.pdf}
+	\caption{The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, (solid symbols represent analyses led by HERA collaborators) compared to 21cmFAST-generated models at $k=0.2$\,$h$\,Mpc$^{-1}$.
+Analysis is still underway on PAPER and MWA observations that approach their projected 
+full sensitivities (see Fig.~\ref{fig:Sensitivities}); 
+HERA can deliver sub-$\text{mK}^2$ sensitivities.}
+	\vspace{-10pt}
+	\label{fig:limits}
+	\label{fig:IGMtemperatureConstraints}
+\end{figure}
+
 Despite progress, the fact remains that the 21\,cm EoR signal is intrinsically very faint; making a detection requires a
@@
 
-Though the desired measurement is over a volume, it is instructive to split it into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kvpr$ is determined by the antenna baseline and $\kpar$ by the frequency.
-This is useful since it allows us to split out the chromatic response of the interferometer and isolate a phase space where foreground sources ({\em i.e.} everything {\em not} the EOR) contaminate the signal of interest from where they don't.
-This contaminated phase space may be shown [REF] to be a wedge-shaped region in $\kpr$:$\kpar$ space:
-\begin{equation}
-|\kpar| \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
-\end{equation}
-where $c$ is the speed of light and $S$ is a parameter accounting for offsets related to the combined spectral smoothness of the foregrounds and the antenna response (see Fig.
-\ref{fig:wedgeWithBG}).
-A key issue to measuring the EOR power spectrum is to understand and minimize such effects.
-For an analytical derivation, see \cite{2012ApJ...756..165P,vedantham_et_al2012,liu_et_al2014b},  in simulation see \cite{datta_et_al2010,hazelton_et_al2013} and for observations
-\cite{2013ApJ...768L..36P,2015arXiv150601026P,2014ApJ...788..106P,2015arXiv150206016A}.
-Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).
-In order to normalize for updated measurements of Hubble's constant, it is further normalized by $h$, where $H_o=100h$, such that they are expressed as $h$/Mpc.
-
-\begin{figure}[t]
-\includegraphics[height=3in]{plots/wedgeWithBG.png}
-\caption{Foregrounds are a primary challenge facing 21\,cm cosmology experiments. 
-HERA leverages a ``wedge'' in the spherically-averaged ($\mathbf{k}$ is broken into $k_\|$ and $k_\perp$) power spectrum (center panel; \citealt{dillon_et_al2015}).
-Smooth-spectrum foregrounds (right panel) $\sim$5 orders of magnitude brighter than fiducial EoR models (left panel; \citealt{mesinger_et_al2011}) create the ``wedge'' when they interact with the interferometer's chromatic response.
-By avoiding foregrounds, PAPER has placed limits within an order
-of magnitude (in mK) of these models \citep{ali_et_al2015} and show the ``EoR Window'' to be foreground-free.
-HERA's dish and configuration optimize wedge/window isolation and direct sensitivity to low-$k_\perp$ modes where EoR is brightest.
-%v2
-Cartoon of the wedge.
-The gray rectangle represents the smooth-spectrum foregrounds as they would be if the instrument had a perfect achromatic response.
-The pink wedge area shows the chromatic response of the interferometer, where additional power is thrown up from the gray region to higher k-values with increasing baseline.
-The additional horizontal pink line illustrates the affect that other systematics, such as reflections on cables, can have in contaminating the measurement.
-The white region shown the 'eor window', where the measurement is not contaminated by foregrounds or systematics.}
-\label{fig:wedge}
-\end{figure}
-
-The techniques to measure the power spectrum may be broken down into two principle techniques (delay-space and map-making), plus additional hydrid methods.
-These are briefly summarized below, with an emphasis on the delay-spectrum techniqe as the initial approach taken by HERA.
-Finally, we provide a short discussion on calibration in the concept of this design.
-
-\subsection{Delay-Spectrum Approach}
-\label{sec:delayapproach}
-Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement.
-There are now numerous papers in the literature on the technique of measuring the power spectrum using an interferometer and the attendant issues [[[list of citations]]].
-Below, a high-level overview based on  \cite{2012ApJ...756..165P} will be presented.
-The so-called delay-spectrum approach leverages this response to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
-Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum.
-
-We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment:
-\begin{equation}
-{\hat P}(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} \right]
-\label{eq:Pk_baseline}
-\end{equation}
-where $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes, $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.
-The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.
-Note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
-Rather than ${\hat P}(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.
-
-Since the thermal noise per visibility baseline may be expressed as
-\begin{equation}
-V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
-\label{eq:sensitivity_per_baseline}
-\end{equation}
-where $\tau$ is the integration time (see {\em e.g.} TMS),
-to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq.
-\ref{eq:Pk_baseline}.
-Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
-\begin{equation}
-\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
-\label{eq:sensitivity}
-\end{equation}
-where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement 
-(see \cite{2012ApJ...756..165P}).
-Fig.
-\ref{fig:boost} plots this ``boost'' factor for various telescopes under nominal observing campaigns using the delay-spectrum technique.
-The large improvement for PAPER and HERA is the fact that the baselines are redundantly placed on appropriate spatial scales, so the same baselines gets many samples.
-
-\begin{figure}[h!]
-\includegraphics[width=0.5\textwidth]{plots/boost.png}
-\caption{Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) for representative observing campaigns for various instruments.}
-\label{fig:boost}
-\end{figure}
-
-The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by the antenna baselines and bandwidths.
-Note that largest bandwidth (which determines the smallest $\kpar$) is about 10 MHz, for larger bandwidths cosmic evolution begins to impact the result.
-Figure \ref{fig:kperf} shows the allowed k-values based on the beamwidth/baseline (gray plus dark-blue overlap regions, the black lines are the limits) and a bandwidth of 10 MHz with 1024 channels (the blue and dark-blue overlap regions, the blue line is the lower limit).
-The red-dashed line is the allowed smallest $k$-value attainable with this technique.
-Note that at these red-shifts it is essentially set just by the bandwidth.
-
-\begin{figure}[h!]
-\centerline{
-\includegraphics[width=.4\textwidth]{plots/kperf.png} 
-}
-\caption{\small k
-\label{fig:kperf}}
-\end{figure}
-
-\subsection{Map-Making Approach}
-\label{sec:mapapproach}
-In contrast to the delay-spectrum approach, map-making approaches combine baselines to build up information before squaring and calculating the power spectrum.
-The conceptually easiest approach is that one makes an image of the sky and takes a 2-D Fourier transform to determine the power spectrum and so this approach is often called the imaging approach.
-Advantages are that you actually image the reionized bubbles and can determine the magnitude of the smaller $k$-modes on the sky.
-The principle disadvantage is that you have to compute and calibrate the effects of systematics and the propagation of the very bright foregrounds through your system and subtract them incredibly accurately.
-In contrast, the delay-spectrum approach localizes those effects to a parameter space where they can much more easily be filtered and not used in subsequent analysis.
-
-\subsection{Hybrid Approaches}
-\label{sec:hybridapproach}
-
-\subsection{Calibration}
-\label{sec:calibration}
-
 \section{High-Level Requirements} 
@@
 From the science and techniques discussed above, we may derive some high-level requirements.  This section describes frequency and sensitivity requirements that largely set the overall design concept.
+
+Thus, to open the widest possible window for EoR measurements, HERA must use close-packed antennas that
+minimize signal reflections and deliver significant forward gain relative to their horizon response.
+Tests with prototype HERA antennas (Figs. \ref{fig:orbcommexptandbeammap} and \ref{fig:reflectometry}, discussed in \S\ref{sec:antenna})
+indicate that a moderately large parabolic dish with a short focal height can meet these requirements
+\citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016}.
 
@@
 relatively agnostic with respect to observing parameters, making it a safe and robust first step in data processing
-\citep{zheng_et_al2014,parsons_et_al2014,ali_et_al2015}.
+\citep{zheng_et_al2014,parsons_etal2014,ali_et_al2015}.
 The Internet link between the Karoo and data centers in the U.S.\ has been extensively tested, with assistance from SKA-SA, TENET

### 2016-04-04 manuscript.tex
update kperf figure and various throughout manuscript

@@
 
-\subsection{Secondary Scientific Objectives}%\\Precision Cosmology, Reionization Imaging, and High-$z$ IGM X-Ray Heating Science}
-
-\startsquarepar\emph{\textbf{Precision Cosmology.}}
-\label{sec:tau}
-By advancing our understanding of reionization astrophysics, HERA improves CMB constraints on 
-fundamental cosmological parameters by
-removing the optical depth
-$\tau$ as a nuisance parameter. HERA breaks the degeneracy between the
-constraints on $\tau$ and the sum of the neutrino masses $\sum m_\nu$, which has
-been identified as a potential problem for Stage 4 CMB lensing experiments. A
-HERA-informed estimate of $\tau$ enables CMB lensing experiments to achieve a
-$0.012\,\textrm{eV}$ error on $\sum m_\nu$ \citep{liu_et_al2015}. This would
-represent a $\sim$5$\sigma$ cosmological detection of the neutrino masses even
-under the most pessimistic assumptions
-% that their sum is the minimum value of $0.058\,\textrm{eV}$ 
-still allowed by neutrino oscillation \stopsquarepar
-
-\begin{figure}[h!]
-\centering
-    \includegraphics[width=0.50\textwidth,clip]{plots/sigmaTau.pdf}
-\caption{Likelihood contours ($68\%$ and $95\%$) for $\sigma_8$ and $\tau$ using \emph{Planck} constraints (blue) and 
-adding HERA data (red). 
-%Light and heavy contours signify $68\%$ and $95\%$ confidence regions, respectively. 
-The $21\,\textrm{cm}$ constraints break the CMB degeneracy between the amplitude of density fluctuations and the optical depth, improving constraints on both.}
-	\label{fig:sigma8Tau}
-\end{figure}
-
-\noindent experiments
-\citep{allison_et_al2015}, making HERA key to our understanding of neutrino
-physics. HERA's estimate of $\tau$ would also break the degeneracy between
-$\tau$ and the amplitude of matter fluctuations (expressed in Fig.
-\ref{fig:sigma8Tau} as $\sigma_8$) that arises when using only
-CMB data. HERA effectively reduces error bars on $\sigma_8$ by more
-than a factor of three \citep{liu_et_al2015}, potentially elucidating
-current tensions between cluster cosmology constraints and those from primary
-CMB anisotropies.
-
-\emph{\textbf{First Images of the Reionization Epoch.}}
-\label{sec:imaging}
-In addition to measuring the power spectrum, HERA has sufficient 
-sensitivity to directly image the IGM during reionization,
-and will observe a 1440 deg$^2$ stripe, comparable to future WFIRST
-large area near-IR surveys.
- After 100 hours on a
-single field (achievable in 200 nights), HERA reaches a surface brightness 
-sensitivity of 50 $\mu$Jy/beam (synthesized beam FWHM $\sim$ 24$'$) compared 
-to the brightness temperature fluctuations of up to 400 $\mu$Jy/beam in typical 
-reionization models (see Fig.~\ref{fig:LSS}). From the standpoint of sensitivity alone, HERA is capable of 
-detecting the brightest structures at $z=8$ with SNR $>$ 10. Additionally, the design 
-of HERA places it in a unique position to directly explore calibration techniques
-(e.g. redundant calibration, \citealt{liu_et_al2010}), while retaining a high quality 
-point spread functions for imaging and identifying foregrounds. Leveraging the expertise gained 
-through both precursor strategies, HERA will yield exceptionally high dynamic range
-images.
-
-%\subsubsection{The 21\,cm Signal in the Pre-Reionization Epoch}
-
-%\begin{wrapfigure}{R}{0.5\textwidth}
-\begin{figure}[b!]
-\centering
-\vspace{-15pt}
-    %\includegraphics[width=.5\textwidth,clip]{plots/imaging/HERA_z8_SNR_annotated_2015.png}
-    \includegraphics[width=1.0\textwidth,clip]{plots/imaging/HERA_FoV_w_strips.png}
-  \vspace{-20pt}
-\caption{\footnotesize 
-HERA will observe a 1440 deg$^2$ stripe centered near $\delta = -30^\circ$.
-HERA can measure the ionization state around galaxies in, e.g., the GOODS-South field that contains a third of all known $z\!>\!8$ galaxies.
-HERA's primary imaging data product to the community will be deep cubes along the HERA stripe suitable for cross-correlation.
-\label{fig:LSS} }
-\vspace{-10pt}
-\end{figure}
-%  \vspace{-10pt}
-%\end{wrapfigure}
-
-\emph{\textbf{Pre-Reionization Heating.}}
-\label{sec:EoX} 
-Prior to reionization, the 21\,cm signal is a sensitive probe of the first
-luminous sources and IGM heating mechanisms. First stars are expected to form
-at $z \sim 25-30$ and their imprint on the 21\,cm signal is expected to be
-sensitive to the halo mass where they are formed \citep{mesinger_et_al2015}.
-The IGM is then expected to be heated by first
-generation X--ray binaries
-\citep{furlanetto_et_al2006_global,pritchard_et_al2007,mesinger_et_al2013} or
-by the hot interstellar medium produced by the first supernovae
-\citep{pacucci_et_al2014}, although the heating timing and magnitude is still
-very much debated \citep{fialkov_et_al2012}. Dark matter annihilation
-\citep{evoli_et_al2014} could also inject energy in the IGM, leaving its
-imprint in the 21\,cm signal.
-
-With support from the Moore Foundation to develop feeds sensitive down to 50\,MHz, HERA will probe the IGM prior to
-reionization and constrain the sources of heating, obtaining, in the case of
-X--ray heating, percent-level constraints on the efficiency with which star-forming baryons
-produce X-rays \citep{ewall-wice_et_al2015}.
-While the observational and analytical state of the art of pre-reionization 21\,cm science is not as advanced as for the EoR \citep{ewall-wice_et_al2016-EoXLimits}, the risk of lower-frequency observing is borne by the Moore Foundation.
-Lower
-frequency observations also test feedback
-mechanisms that interact with low-mass halos at high redshifts
-\citep{Iliev_et_al2007,Iliev_et_al2012,ahn_et_al2012}.
-Such constraints, while interesting in their own right, also reduce
-the susceptibility of the aforementioned $21\,\textrm{cm}$-derived $\tau$
-constraints to uncertainties in high-redshift physics \citep{liu_et_al2015}, especially if they are 
-combined with upcoming or proposed measurements of 
-the pre-reionization sky-averaged spectrum \citep{fialkov_and_loeb2016}. Additionally, they may be crucial to a correct
-interpretation of kinetic Sunyaev-Zel'dovich effect constraints on reionization
-from the CMB \citep{park_et_al2013}. The high redshift probe of
-structure afforded by the low frequency $21\,\textrm{cm}$ measurements will
-permit some of the most direct observations of hypothesized suppressions of
-small-scale structure
-\citep{dalal_et_al2010,tseliakhovich_et_al2011,fialkov_et_al2012} arising from
-predicted supersonic relative velocities between dark matter and baryonic gas
-\citep{tseliakhovich_and_hirata2010}. No other electromagnetic probe can provide direct observations of this
-epoch.
-
-\subsection{Other Scientific Objectives}
-\label{subsec:broader_science}
-
-\noindent With unprecedented sensitivity at this frequency range, HERA can deliver much more than its core 21\,cm science. 
-Some of HERA's broader science impacts derive directly from the public data products HERA delivers.  Other
-areas of impact may require additional hardware or software that HERA can host at no extra risk. 
-In this respect, HERA serves as a ``platform" for additional science programs with external funding.
-The Moore Foundation's support of HERA feed development and data analysis targeting the pre-reionization science described in \S\ref{sec:EoX}
-is a prime example of how we envision this functioning.
-Below, we list examples of the broader science that HERA can deliver.
-
-\emph{\textbf{Cross-Correlations with Other Reionization Probes.}}
-\label{sec:crossCorr}
-HERA's public data provide new opportunities for cross-correlation studies.
-Cross-correlation between HERA 21\,cm images and other high-redshift probes
-(e.g. JWST; WFIRST; CMB maps; CO, CII, and Ly-$\alpha$ intensity mapping) 
-can provide an independent confirmation of the 21\,cm power spectrum
-\citep[i.e.][]{lidz_et_al2009,dore_et_al2014,silva_et_al2015,vrbanec2016} and enable rich new studies of
-the interaction between galaxies and their ionization environment. 
-In particular, cross-correlating 21\,cm with galaxy surveys can measure the characteristic bubble size around galaxies of
-different luminosities \citep{lidz_et_al2009} and help separate the degeneracy
-between the fraction of photons escaping the galaxies and the
-total number of ionizing photons produced \citep{zackrisson2013}.
-
-HERA is placed in a unique position to enable such cross-correlation: the GOODS-South field---one of the most panchromatically studied regions of the sky, the site of the Hubble UDF, and home to over a third of all known $z>9$
-galaxies---lies in its field of view. HERA's images of the IGM can provide environmental context to galaxy surveys through identification of
-ionized bubbles \citep{malloy_lidz2013}, or in a more statistical sense as
-described in \cite{beardsley_et_al2015}.
-
-\textbf{\emph{Searching for Exoplanetary Radio Bursts.}}
-\label{sec:exoplanets}
-HERA could be an powerful tool in the search for bright, highly polarized auroral bursts
-from exoplanets %due to the electron cyclotron maser instability 
-\citep{treumann2006,hallinan_et_al2015}. These
-distinctive, coherent bursts occur periodically at the planetary
-rotation rate, which cannot otherwise be measured. 
-A burst from a terrestrial planet could point toward
-habitability, since it implies a powerful magnetic field
-protecting the atmosphere and perhaps the biosphere from energetic stellar wind
-particles \citep{tarter_et_al2007}. HERA's sensitivity, large FoV, long campaigns, and precise calibration are all well-suited
-for discovering exoplanetary radio bursts from Jupiter-like planets out to 25~pc.
-Many well studied exoplanetary systems within that distance are in the HERA stripe, including Fomalhaut, Gl 667 C, Gl 433, HD 147513, and Gl 317.
-
-\textbf{\emph{Fast Radio Burst Followup.}}
-\label{sec:FRBs}
-Fast Radio Bursts (FRB) are millisecond-long radio flashes whose origin has remained a great enigma ever since their discovery \citep{2007Sci...318..777L}. 
-If equipped with a suitable backend, HERA could be triggered by nearby, higher-frequency telescopes for FRB followup, saving baseband data and thus full sensitivity to all dispersion measures.
-Bursts like that discovered by \citet{masui_et_al2015}, the lowest frequency
-FRB detection to date, should be seen hourly by HERA at 5--10$\sigma$. %(including the effects of scattering)
- Observations at HERA frequencies are very sensitive to the physics of
-the intervening medium, particularly deviations from $\lambda^2$
-dispersion. Detecting deviations would rule out broad classes of models and could indicate whether FRBs are at cosmological distances.
-
-\section{Measuring the EOR}
-\label{sec:eormeas}
-Due to the expansion of the Universe over cosmological time, we can identify and measure the early Universe via the red-shift of spectral lines.
-The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire Universe back almost to the period of recombination.
-The bandwidth
-therefore equates to a cosmic volume along the site of the telescope, with the frequency determining the cosmic age.
-
-Given the difficulty of the experiment, initial measurements of the EOR strive to measure a statistical power spectrum of the signal since the nature of the reionization process should have a specific spatial signature.
-The goal is therefore to measure a range of aggregate spatial scales on the sky, rather than to image the signal directly.
-Imaging does remain an ultimate goal to fully understand the process, however we will need a greater understanding of the full measurements to achieve this more difficult goal.
-MANY PAPERS, see Nithya and Jonnie recent
-
-Obviously between our present observation point and the Epoch of Reionization approximately 13 billion years ago lies the entire intervening Universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).
-Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.
-As a first step, areas of the sky with where these signals are minimal (for example, outside the galactic plane and away from strong point sources) are targeted (REF FOR THIS).
-More importantly and most fortuitously however is the fact that all of these foreground signals are smooth spectrum sources whereas the expected spectrum of the EOR is expected to be rough since it is made up of non-ionized regions which are randomly distributed over a wide range of redshifts.
-This fact allows us to try and isolate foregrounds from the EOR, as will be discussed.
-
-This section will provide a brief overview of existing measurements and the theoretical underpinning of how foregrounds contaminate the measurement (the so-called ``wedge''), discuss the various techniques used to make the measurement and finally provide a brief discussion on calibration issues and techniques.
-
-\subsection{The ``Wedge''}
-The past three
-years have seen deep EoR observations with PAPER, the MWA, 
-the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2013}), and LOFAR.
-PAPER and the MWA have produced progressively deeper limits
-\citep{dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
-yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
-(Fig. \ref{fig:limits}).
-The inherent challenge of simultaneously meeting stringent sensitivity requirements 
-while suppressing foregrounds $\sim5$ orders of magnitude
-brighter than the 21\,cm signal make this progress all the more remarkable.  
-To achieve this, HERA team members have refined and improved 
-techniques spanning all aspects of the experimental process, from how we
-calibrate our data \citep{zheng_et_al2014, jacobs_et_al2016, barry_et_al2016}, to how we understand foreground contamination
-\citep{moore_et_al2013,moore_et_al2016,thyagarajan_et_al2015a,pober_et_al2016},
-to how we design the interferometer itself \citep{parsons_et_al2012a,dillon_parsons2016}.
-In this section, we discuss how the legacy of these first generation
-21\,cm experiments inform critical aspects of HERA's design.
-
-Perhaps the most important advance informing HERA's design is a
-refined understanding of how smooth-spectrum foregrounds interact
-with instrument chromaticity to produce a characteristic ``wedge" of
-foreground leakage in Fourier space (see Fig.~\ref{fig:wedge}), 
-outside of which the 21\,cm signal dominates.
-Through theoretical and observational work
-\citep{datta_etal2010,morales_et_al2012,parsons_et_al2012b,vedantham_2012,thyagarajan_et_al2013,hazelton_et_al2013,pober_etal2013b,liu_et_al2014a,liu_et_al2014b},
-we have learned how the boundary between the wedge and our ``EoR window" is determined by the separation between antennas,
-signal reflections within antennas, and the angular response of the antenna beam.  Deep integrations also show us
-that, to the limits of current sensitivity, foreground emission is absent outside of the wedge; it can only 
-appear there through instrumental leakage \citep{parsons_etal2014,ali_et_al2015,moore_et_al2016,kohn_et_al2016}.
-
-Though the desired measurement is over a volume, it is instructive to split it into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kvpr$ is determined by the antenna baseline and $\kpar$ by the frequency.
-This is useful since it allows us to split out the chromatic response of the interferometer and isolate a phase space where foreground sources ({\em i.e.} everything {\em not} the EOR) contaminate the signal of interest from where they don't.
-This contaminated phase space may be shown [REF] to be a wedge-shaped region in $\kpr$:$\kpar$ space:
-\begin{equation}
-|\kpar| \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
-\end{equation}
-where $c$ is the speed of light and $S$ is a parameter accounting for offsets related to the combined spectral smoothness of the foregrounds and the antenna response (see Fig. \ref{fig:wedge}).  
-A key issue to measuring the EOR power spectrum is to understand and minimize such effects.
-For an analytical derivation, see \cite{zahn_etal2012,vedantham_2012,liu_et_al2014b},  in simulation see \cite{datta_etal2010,hazelton_et_al2013} and for observations
-\cite{pober_etal2013b,2015arXiv150601026P,parsons_etal2014,2015arXiv150206016A}.
-Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).
-In order to normalize for updated measurements of Hubble's constant, it is further normalized by $h$, where $H_o=100h$, such that they are expressed as $h$/Mpc.
-
-\begin{figure}[t]
-\includegraphics[height=3in]{plots/wedgeWithBG.png}
-\caption{Foregrounds are a primary challenge facing 21\,cm cosmology experiments. 
-HERA leverages a ``wedge'' in the spherically-averaged ($\mathbf{k}$ is broken into $k_\|$ and $k_\perp$) power spectrum (center panel; \citealt{dillon_et_al2015}).
-Smooth-spectrum foregrounds (right panel) $\sim$5 orders of magnitude brighter than fiducial EoR models (left panel; \citealt{mesinger_et_al2011}) create the ``wedge'' when they interact with the interferometer's chromatic response.
-By avoiding foregrounds, PAPER has placed limits within an order
-of magnitude (in mK) of these models \citep{ali_et_al2015} and show the ``EoR Window'' to be foreground-free.
-HERA's dish and configuration optimize wedge/window isolation and direct sensitivity to low-$k_\perp$ modes where EoR is brightest.
-%v2
-Cartoon of the wedge.
-The gray rectangle represents the smooth-spectrum foregrounds as they would be if the instrument had a perfect achromatic response.
-The pink wedge area shows the chromatic response of the interferometer, where additional power is thrown up from the gray region to higher k-values with increasing baseline.
-The additional horizontal pink line illustrates the affect that other systematics, such as reflections on cables, can have in contaminating the measurement.
-The white region shown the 'eor window', where the measurement is not contaminated by foregrounds or systematics.}
-\label{fig:wedge}
-\end{figure}
-
-The techniques to measure the power spectrum may be broken down into two principle techniques (delay-space and map-making), plus additional hydrid methods.
-These are briefly summarized below, with an emphasis on the delay-spectrum technique as the initial approach taken by HERA.
-Finally, we provide a short discussion on calibration in the concept of this design.
-
-\subsection{Delay-Spectrum Approach}
-\label{sec:delayapproach}
-Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement.
-There are now numerous papers in the literature on the technique of measuring the power spectrum using an interferometer and the attendant issues [[[list of citations]]].
-Below, a high-level overview based on  \cite{parsons_et_al2012b} will be presented.
-The so-called delay-spectrum approach leverages this response to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
-Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum.
-
-We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment:
-\begin{equation}
-{\hat P}(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} \right]
-\label{eq:Pk_baseline}
-\end{equation}
-where $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes, $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.
-The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.
-Note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
-Rather than ${\hat P}(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.
-
-Since the thermal noise per visibility baseline may be expressed as
-\begin{equation}
-V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
-\label{eq:sensitivity_per_baseline}
-\end{equation}
-where $\tau$ is the integration time (see {\em e.g.} TMS),
-to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq.
-\ref{eq:Pk_baseline}.
-Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
-\begin{equation}
-\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
-\label{eq:sensitivity}
-\end{equation}
-where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement 
-(see \cite{parsons_et_al2012b}).
-Fig.
-\ref{fig:boost} plots this ``boost'' factor for various telescopes under nominal observing campaigns using the delay-spectrum technique.
-The large improvement for PAPER and HERA is the fact that the baselines are redundantly placed on appropriate spatial scales, so the same baselines gets many samples.
-
-\begin{figure}[h!]
-\includegraphics[width=0.5\textwidth]{plots/boost.png}
-\caption{Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) for representative observing campaigns for various instruments.}
-\label{fig:boost}
-\end{figure}
-
-The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by the antenna baselines and bandwidths.
-Note that largest bandwidth (which determines the smallest $\kpar$) is about 10 MHz, for larger bandwidths cosmic evolution begins to impact the result.
-Figure \ref{fig:kperf} shows the allowed k-values based on the beamwidth/baseline (gray plus dark-blue overlap regions, the black lines are the limits) and a bandwidth of 10 MHz with 1024 channels (the blue and dark-blue overlap regions, the blue line is the lower limit).
-The red-dashed line is the allowed smallest $k$-value attainable with this technique.
-Note that at these red-shifts it is essentially set just by the bandwidth.
-
-\begin{figure}[h!]
-\centerline{
-\includegraphics[width=.4\textwidth]{plots/kperf.png} 
-}
-\caption{\small ADD NEW HERA CONFIGURATION AND LOFAR AND PUT IN 'HOW TO MAKE THIS PLOT'}
-\label{fig:kperf}
-\end{figure}
-
-\subsection{Map-Making Approach}
-\label{sec:mapapproach}
-In contrast to the delay-spectrum approach, map-making approaches combine baselines to build up information before squaring and calculating the power spectrum.
-The conceptually easiest approach is that one makes an image of the sky and takes a 2-D Fourier transform to determine the power spectrum and so this approach is often called the imaging approach.
-Advantages are that you actually image the reionized bubbles and can determine the magnitude of the smaller $k$-modes on the sky.
-The principle disadvantage is that you have to compute and calibrate the effects of systematics and the propagation of the very bright foregrounds through your system and subtract them incredibly accurately.
-In contrast, the delay-spectrum approach localizes those effects to a parameter space where they can much more easily be filtered and not used in subsequent analysis.
-
-\subsection{Hybrid Approaches}
-\label{sec:hybridapproach}
-
-\subsection{Calibration}
-\label{sec:calibration}
-
 \subsection{Existing Measurements}
@@
 
+\subsection{Secondary Scientific Objectives}%\\Precision Cosmology, Reionization Imaging, and High-$z$ IGM X-Ray Heating Science}
+
+\startsquarepar\emph{\textbf{Precision Cosmology.}}
+\label{sec:tau}
+By advancing our understanding of reionization astrophysics, HERA improves CMB constraints on 
+fundamental cosmological parameters by
+removing the optical depth
+$\tau$ as a nuisance parameter. HERA breaks the degeneracy between the
+constraints on $\tau$ and the sum of the neutrino masses $\sum m_\nu$, which has
+been identified as a potential problem for Stage 4 CMB lensing experiments. A
+HERA-informed estimate of $\tau$ enables CMB lensing experiments to achieve a
+$0.012\,\textrm{eV}$ error on $\sum m_\nu$ \citep{liu_et_al2015}. This would
+represent a $\sim$5$\sigma$ cosmological detection of the neutrino masses even
+under the most pessimistic assumptions
+% that their sum is the minimum value of $0.058\,\textrm{eV}$ 
+still allowed by neutrino oscillation \stopsquarepar
+
+\begin{figure}[h!]
+\centering
+    \includegraphics[width=0.50\textwidth,clip]{plots/sigmaTau.pdf}
+\caption{Likelihood contours ($68\%$ and $95\%$) for $\sigma_8$ and $\tau$ using \emph{Planck} constraints (blue) and 
+adding HERA data (red). 
+%Light and heavy contours signify $68\%$ and $95\%$ confidence regions, respectively. 
+The $21\,\textrm{cm}$ constraints break the CMB degeneracy between the amplitude of density fluctuations and the optical depth, improving constraints on both.}
+	\label{fig:sigma8Tau}
+\end{figure}
+
+\noindent experiments
+\citep{allison_et_al2015}, making HERA key to our understanding of neutrino
+physics. HERA's estimate of $\tau$ would also break the degeneracy between
+$\tau$ and the amplitude of matter fluctuations (expressed in Fig.
+\ref{fig:sigma8Tau} as $\sigma_8$) that arises when using only
+CMB data. HERA effectively reduces error bars on $\sigma_8$ by more
+than a factor of three \citep{liu_et_al2015}, potentially elucidating
+current tensions between cluster cosmology constraints and those from primary
+CMB anisotropies.
+
+\emph{\textbf{First Images of the Reionization Epoch.}}
+\label{sec:imaging}
+In addition to measuring the power spectrum, HERA has sufficient 
+sensitivity to directly image the IGM during reionization,
+and will observe a 1440 deg$^2$ stripe, comparable to future WFIRST
+large area near-IR surveys.
+ After 100 hours on a
+single field (achievable in 200 nights), HERA reaches a surface brightness 
+sensitivity of 50 $\mu$Jy/beam (synthesized beam FWHM $\sim$ 24$'$) compared 
+to the brightness temperature fluctuations of up to 400 $\mu$Jy/beam in typical 
+reionization models (see Fig.~\ref{fig:LSS}). From the standpoint of sensitivity alone, HERA is capable of 
+detecting the brightest structures at $z=8$ with SNR $>$ 10. Additionally, the design 
+of HERA places it in a unique position to directly explore calibration techniques
+(e.g. redundant calibration, \citealt{liu_et_al2010}), while retaining a high quality 
+point spread functions for imaging and identifying foregrounds. Leveraging the expertise gained 
+through both precursor strategies, HERA will yield exceptionally high dynamic range
+images.
+
+%\subsubsection{The 21\,cm Signal in the Pre-Reionization Epoch}
+
+%\begin{wrapfigure}{R}{0.5\textwidth}
+\begin{figure}[b!]
+\centering
+\vspace{-15pt}
+    %\includegraphics[width=.5\textwidth,clip]{plots/imaging/HERA_z8_SNR_annotated_2015.png}
+    \includegraphics[width=1.0\textwidth,clip]{plots/imaging/HERA_FoV_w_strips.png}
+  \vspace{-20pt}
+\caption{\footnotesize 
+HERA will observe a 1440 deg$^2$ stripe centered near $\delta = -30^\circ$.
+HERA can measure the ionization state around galaxies in, e.g., the GOODS-South field that contains a third of all known $z\!>\!8$ galaxies.
+HERA's primary imaging data product to the community will be deep cubes along the HERA stripe suitable for cross-correlation.
+\label{fig:LSS} }
+\vspace{-10pt}
+\end{figure}
+%  \vspace{-10pt}
+%\end{wrapfigure}
+
+\emph{\textbf{Pre-Reionization Heating.}}
+\label{sec:EoX} 
+Prior to reionization, the 21\,cm signal is a sensitive probe of the first
+luminous sources and IGM heating mechanisms. First stars are expected to form
+at $z \sim 25-30$ and their imprint on the 21\,cm signal is expected to be
+sensitive to the halo mass where they are formed \citep{mesinger_et_al2015}.
+The IGM is then expected to be heated by first
+generation X--ray binaries
+\citep{furlanetto_et_al2006_global,pritchard_et_al2007,mesinger_et_al2013} or
+by the hot interstellar medium produced by the first supernovae
+\citep{pacucci_et_al2014}, although the heating timing and magnitude is still
+very much debated \citep{fialkov_et_al2012}. Dark matter annihilation
+\citep{evoli_et_al2014} could also inject energy in the IGM, leaving its
+imprint in the 21\,cm signal.
+
+With support from the Moore Foundation to develop feeds sensitive down to 50\,MHz, HERA will probe the IGM prior to
+reionization and constrain the sources of heating, obtaining, in the case of
+X--ray heating, percent-level constraints on the efficiency with which star-forming baryons
+produce X-rays \citep{ewall-wice_et_al2015}.
+While the observational and analytical state of the art of pre-reionization 21\,cm science is not as advanced as for the EoR \citep{ewall-wice_et_al2016-EoXLimits}, the risk of lower-frequency observing is borne by the Moore Foundation.
+Lower
+frequency observations also test feedback
+mechanisms that interact with low-mass halos at high redshifts
+\citep{Iliev_et_al2007,Iliev_et_al2012,ahn_et_al2012}.
+Such constraints, while interesting in their own right, also reduce
+the susceptibility of the aforementioned $21\,\textrm{cm}$-derived $\tau$
+constraints to uncertainties in high-redshift physics \citep{liu_et_al2015}, especially if they are 
+combined with upcoming or proposed measurements of 
+the pre-reionization sky-averaged spectrum \citep{fialkov_and_loeb2016}. Additionally, they may be crucial to a correct
+interpretation of kinetic Sunyaev-Zel'dovich effect constraints on reionization
+from the CMB \citep{park_et_al2013}. The high redshift probe of
+structure afforded by the low frequency $21\,\textrm{cm}$ measurements will
+permit some of the most direct observations of hypothesized suppressions of
+small-scale structure
+\citep{dalal_et_al2010,tseliakhovich_et_al2011,fialkov_et_al2012} arising from
+predicted supersonic relative velocities between dark matter and baryonic gas
+\citep{tseliakhovich_and_hirata2010}. No other electromagnetic probe can provide direct observations of this
+epoch.
+
+\subsection{Other Scientific Objectives}
+\label{subsec:broader_science}
+
+\noindent With unprecedented sensitivity at this frequency range, HERA can deliver much more than its core 21\,cm science. 
+Some of HERA's broader science impacts derive directly from the public data products HERA delivers.  Other
+areas of impact may require additional hardware or software that HERA can host at no extra risk. 
+In this respect, HERA serves as a ``platform" for additional science programs with external funding.
+The Moore Foundation's support of HERA feed development and data analysis targeting the pre-reionization science described in \S\ref{sec:EoX}
+is a prime example of how we envision this functioning.
+Below, we list examples of the broader science that HERA can deliver.
+
+\emph{\textbf{Cross-Correlations with Other Reionization Probes.}}
+\label{sec:crossCorr}
+HERA's public data provide new opportunities for cross-correlation studies.
+Cross-correlation between HERA 21\,cm images and other high-redshift probes
+(e.g. JWST; WFIRST; CMB maps; CO, CII, and Ly-$\alpha$ intensity mapping) 
+can provide an independent confirmation of the 21\,cm power spectrum
+\citep[i.e.][]{lidz_et_al2009,dore_et_al2014,silva_et_al2015,vrbanec2016} and enable rich new studies of
+the interaction between galaxies and their ionization environment. 
+In particular, cross-correlating 21\,cm with galaxy surveys can measure the characteristic bubble size around galaxies of
+different luminosities \citep{lidz_et_al2009} and help separate the degeneracy
+between the fraction of photons escaping the galaxies and the
+total number of ionizing photons produced \citep{zackrisson2013}.
+
+HERA is placed in a unique position to enable such cross-correlation: the GOODS-South field---one of the most panchromatically studied regions of the sky, the site of the Hubble UDF, and home to over a third of all known $z>9$
+galaxies---lies in its field of view. HERA's images of the IGM can provide environmental context to galaxy surveys through identification of
+ionized bubbles \citep{malloy_lidz2013}, or in a more statistical sense as
+described in \cite{beardsley_et_al2015}.
+
+\textbf{\emph{Searching for Exoplanetary Radio Bursts.}}
+\label{sec:exoplanets}
+HERA could be an powerful tool in the search for bright, highly polarized auroral bursts
+from exoplanets %due to the electron cyclotron maser instability 
+\citep{treumann2006,hallinan_et_al2015}. These
+distinctive, coherent bursts occur periodically at the planetary
+rotation rate, which cannot otherwise be measured. 
+A burst from a terrestrial planet could point toward
+habitability, since it implies a powerful magnetic field
+protecting the atmosphere and perhaps the biosphere from energetic stellar wind
+particles \citep{tarter_et_al2007}. HERA's sensitivity, large FoV, long campaigns, and precise calibration are all well-suited
+for discovering exoplanetary radio bursts from Jupiter-like planets out to 25~pc.
+Many well studied exoplanetary systems within that distance are in the HERA stripe, including Fomalhaut, Gl 667 C, Gl 433, HD 147513, and Gl 317.
+
+\textbf{\emph{Fast Radio Burst Followup.}}
+\label{sec:FRBs}
+Fast Radio Bursts (FRB) are millisecond-long radio flashes whose origin has remained a great enigma ever since their discovery \citep{2007Sci...318..777L}. 
+If equipped with a suitable backend, HERA could be triggered by nearby, higher-frequency telescopes for FRB followup, saving baseband data and thus full sensitivity to all dispersion measures.
+Bursts like that discovered by \citet{masui_et_al2015}, the lowest frequency
+FRB detection to date, should be seen hourly by HERA at 5--10$\sigma$. %(including the effects of scattering)
+ Observations at HERA frequencies are very sensitive to the physics of
+the intervening medium, particularly deviations from $\lambda^2$
+dispersion. Detecting deviations would rule out broad classes of models and could indicate whether FRBs are at cosmological distances.
+
+\section{Measuring the EOR}
+\label{sec:eormeas}
+Due to the expansion of the Universe over cosmological time, we can identify and measure the early Universe via the red-shift of spectral lines.
+The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire Universe back almost to the period of recombination.
+The bandwidth
+therefore equates to a cosmic volume along the site of the telescope, with the frequency determining the cosmic age.
+
+Given the difficulty of the experiment, initial measurements of the EOR strive to measure a statistical power spectrum of the signal since the nature of the reionization process should have a specific spatial signature.
+The goal is therefore to measure a range of aggregate spatial scales on the sky, rather than to image the signal directly.
+Imaging does remain an ultimate goal to fully understand the process, however we will need a greater understanding of the full measurements to achieve this more difficult goal.
+MANY PAPERS, see Nithya and Jonnie recent
+
+Obviously between our present observation point and the Epoch of Reionization approximately 13 billion years ago lies the entire intervening Universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).
+Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.
+As a first step, areas of the sky with where these signals are minimal (for example, outside the galactic plane and away from strong point sources) are targeted (REF FOR THIS).
+More importantly and most fortuitously however is the fact that all of these foreground signals are smooth spectrum sources whereas the expected spectrum of the EOR is expected to be rough since it is made up of non-ionized regions which are randomly distributed over a wide range of redshifts.
+This fact allows us to try and isolate foregrounds from the EOR, as will be discussed.
+
+This section will provide a brief overview of existing measurements and the theoretical underpinning of how foregrounds contaminate the measurement (the so-called ``wedge''), discuss the various techniques used to make the measurement and finally provide a brief discussion on calibration issues and techniques.
+
+\subsection{The ``Wedge''}
+The past three years have seen deep EoR observations with PAPER, the MWA, 
+the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2013}), and LOFAR.
+PAPER and the MWA have produced progressively deeper limits
+\citep{dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
+yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
+(Fig. \ref{fig:limits}).
+The inherent challenge is to simultaneously meet stringent sensitivity requirements 
+while suppressing foregrounds $\sim5$ orders of magnitude
+brighter than the 21\,cm signal, which has been addressed along a number of approaches.  
+To address this challenge,  all aspects of the experimental process must be refined and improved, spanning 
+calibration ({\em e.g.} \cite{zheng_et_al2014, jacobs_et_al2016, barry_et_al2016}), handling foreground contamination
+({\em e.g.} \cite{moore_et_al2013,moore_et_al2016,thyagarajan_et_al2015a,pober_et_al2016}),
+and the interferometer design itself ({\em e.g.} \cite{parsons_et_al2012a,dillon_parsons2016}).
+
+Perhaps the most important advance informing HERA's design is a
+refined understanding of how smooth-spectrum foregrounds interact
+with instrument chromaticity to produce a characteristic ``wedge" of
+foreground leakage in Fourier space (see Fig.~\ref{fig:wedge}), 
+outside of which the 21\,cm signal dominates.
+Through theoretical and observational work
+\citep{datta_etal2010,morales_et_al2012,parsons_et_al2012b,vedantham_2012,thyagarajan_et_al2013,hazelton_et_al2013,pober_etal2013b,liu_et_al2014a,liu_et_al2014b},
+we have learned how the boundary between the wedge and our ``EoR window" is determined by the separation between antennas,
+signal reflections within antennas, and the angular response of the antenna beam.  Deep integrations also show us
+that, to the limits of current sensitivity, foreground emission is absent outside of the wedge; it can only 
+appear there through instrumental leakage \citep{parsons_etal2014,ali_et_al2015,moore_et_al2016,kohn_et_al2016}.
+
+Though the power spectrum measurement is over a volume, it is instructive to split the spatial wavenumbers into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kvpr$ is determined by the antenna baseline and $\kpar$ by the frequency.
+This is useful since it allows us to split out the chromatic response of the interferometer and isolate a phase space where foreground sources ({\em i.e.} everything {\em not} the EOR) contaminate the signal of interest from where they don't.
+This contaminated phase space may be shown [REF] to be a wedge-shaped region in $\kpr$:$\kpar$ space:
+\begin{equation}
+|\kpar| \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
+\end{equation}
+where $c$ is the speed of light and $S$ is a parameter accounting for offsets related to the combined spectral smoothness of the foregrounds and the antenna response (see Fig. \ref{fig:wedge}).  
+A key issue to measuring the EOR power spectrum is to understand and minimize such effects.
+For an analytical derivation, see \cite{zahn_etal2012,vedantham_2012,liu_et_al2014b},  in simulation see \cite{datta_etal2010,hazelton_et_al2013} and for observations
+\cite{pober_etal2013b,2015arXiv150601026P,parsons_etal2014,2015arXiv150206016A}.
+Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).
+In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_o=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.
+
+\begin{figure}[t]
+\includegraphics[height=3in]{plots/wedgeWithBG.png}
+\caption{Foregrounds are a primary challenge facing 21\,cm cosmology experiments. 
+HERA leverages a ``wedge'' in the spherically-averaged ($\mathbf{k}$ is broken into $k_\|$ and $k_\perp$) power spectrum (center panel; \citealt{dillon_et_al2015}).
+Smooth-spectrum foregrounds (right panel) $\sim$5 orders of magnitude brighter than fiducial EoR models (left panel; \citealt{mesinger_et_al2011}) create the ``wedge'' when they interact with the interferometer's chromatic response.
+By avoiding foregrounds, PAPER has placed limits within an order
+of magnitude (in mK) of these models \citep{ali_et_al2015} and show the ``EoR Window'' to be foreground-free.
+HERA's dish and configuration optimize wedge/window isolation and direct sensitivity to low-$k_\perp$ modes where EoR is brightest.
+%v2
+Cartoon of the wedge.
+The gray rectangle represents the smooth-spectrum foregrounds as they would be if the instrument had a perfect achromatic response.
+The pink wedge area shows the chromatic response of the interferometer, where additional power is thrown up from the gray region to higher k-values with increasing baseline.
+The additional horizontal pink line illustrates the affect that other systematics, such as reflections on cables, can have in contaminating the measurement.
+The white region shown the 'eor window', where the measurement is not contaminated by foregrounds or systematics.}
+\label{fig:wedge}
+\end{figure}
+
+The techniques to measure the power spectrum may be broken down into two principle techniques (delay-space and map-making), plus additional hydrid methods.
+These are briefly summarized below, with an emphasis on the delay-spectrum technique as the initial approach taken by HERA.
+Finally, we provide a short discussion on calibration in the concept of this design.
+
+\subsection{Delay-Spectrum Approach}
+\label{sec:delayapproach}
+Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement.
+There are now numerous papers in the literature on the technique of measuring the power spectrum using an interferometer and the attendant issues [[[list of citations]]].
+Below, a high-level overview based on  \cite{parsons_et_al2012b} will be presented.
+The so-called delay-spectrum approach leverages the interferometer measurement to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
+Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum.
+
+We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment:
+\begin{equation}
+{\hat P}(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} \right]
+\label{eq:Pk_baseline}
+\end{equation}
+where $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes (so, relating wavenumber to physical volume at a given redshift), $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.
+The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.
+Note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
+Rather than ${\hat P}(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.
+
+Since the thermal noise per visibility baseline may be expressed as
+\begin{equation}
+V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
+\label{eq:sensitivity_per_baseline}
+\end{equation}
+where $\tau$ is the integration time (see {\em e.g.} TMS),
+to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq.
+\ref{eq:Pk_baseline}.
+Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
+\begin{equation}
+\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
+\label{eq:sensitivity}
+\end{equation}
+where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement 
+(see \cite{parsons_et_al2012b}).
+Fig.
+\ref{fig:boost} plots this ``boost'' factor for various telescopes under nominal observing campaigns using the delay-spectrum technique.
+The large improvement for PAPER and HERA is the fact that the baselines are redundantly placed on appropriate spatial scales, so that a more limited number of unique baselines get sampled many times.
+
+\begin{figure}[h!]
+\includegraphics[width=0.5\textwidth]{plots/boost.png}
+\caption{Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) for representative observing campaigns for various instruments.
+ADD NEW HERA CONFIGURATION AND LOFAR AND PUT IN 'HOW TO MAKE THIS PLOT'}
+\label{fig:boost}
+\end{figure}
+
+The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by the antenna baselines and bandwidths.
+Note that largest bandwidth (which determines the smallest $\kpar$) is about 10 MHz, for larger bandwidths cosmic evolution begins to impact the result.
+Figure \ref{fig:kperf} shows the allowed k-values based on the beamwidth/baseline (gray plus dark-blue overlap regions, the black lines are the limits) and a bandwidth of 10 MHz with 1024 channels (the blue and dark-blue overlap regions, the blue line is the lower limit).
+The red-dashed line is the allowed smallest $k$-value attainable with this technique.
+Note that at these red-shifts it is dominated by the bandwidth.
+
+\begin{figure}[h!]
+\centerline{
+\includegraphics[width=.4\textwidth]{plots/kperf.png} 
+}
+\caption{\small kperf}
+\label{fig:kperf}
+\end{figure}
+
+\subsection{Map-Making Approach}
+\label{sec:mapapproach}
+In contrast to the delay-spectrum approach, map-making approaches combine baselines to build up information before squaring and calculating the power spectrum.
+The conceptually easiest approach is that one makes an image of the sky and takes a 2-D Fourier transform to determine the power spectrum and so this approach is often called the imaging approach.
+Advantages are that you actually image the reionized bubbles and can determine the magnitude of the smaller $k$-modes on the sky.
+The principle disadvantage is that you have to compute and calibrate the effects of systematics and the propagation of the very bright foregrounds through your system and subtract them incredibly accurately.
+In contrast, the delay-spectrum approach localizes those effects to a parameter space where they can much more easily be filtered and not used in subsequent analysis.
+
+\subsection{Hybrid Approaches}
+\label{sec:hybridapproach}
+
+\subsection{Calibration}
+\label{sec:calibration}
+
 \section{High-Level Requirements} 
 \label{sec:requirements}
-From the science and techniques discussed above, we may derive some high-level requirements.  This section describes frequency and sensitivity requirements that largely set the overall design concept.
-
+From the science and techniques discussed above, we may derive some high-level requirements, predicating on optimizing for the delay-spectrum approach for detecting the EOR while not unduly limiting other approaches or other science goals.
+The fact that HERA is an experiment and not meant to be a long-term general use facility greatly facilitates this optimization.
+This section describes the element, configuration, frequency and sensitivity requirements that largely set the overall design concept.
+
+\subsection{Antenna Design and Configuration}
 Thus, to open the widest possible window for EoR measurements, HERA must use close-packed antennas that

### 2016-04-04 plots/kperf.py

@@ -1,0 +1,118 @@
+import sys
+sys.path.append('/Users/ddeboer/Documents/ubase/Code/cosmo')
+import nedclass
+import matplotlib.pyplot as plt
+import math
+import numpy as np
+
+n = nedclass.ned(H0=100.0)
+
+def mkPlot(restFreq,BW,b,coeff):
+    X = []  # X Mpc/rad
+    Xb = [] # scaled version
+    Xinstr = []  # observed extent in Mpc
+    Y = []  # Y Mpc/GHz
+    DA = [] # angular scale term
+    E = []  # Hubble parameter evolution
+    #Below are just used to get a power-law fit and then plot it post facto
+    apX = []    # approximated X (post-fit numbers put back in)
+    apY = []    # approximated Y ( " )
+    z2fit = []  # range of z's to fit for approximation
+    X2fit = []  #     "    X's   "
+    Y2fit = []  #     "    Y's   "
+
+    # Loop over red-shift
+    z = np.arange(0.1,25,.1)
+    for zi in z:
+        DAi = n.calcUniverse(zi)
+        DA.append(DAi)
+        Xi = DAi*(1.+zi)
+        obsFreq = restFreq/(1.+zi)
+        obsWavelength = 0.3/obsFreq
+        Xbi = Xi/((180.0/math.pi)*1.0)  #convert to degrees
+        X.append(Xi)
+        Xb.append(Xbi)
+        Xbi = Xi*coeff*obsWavelength/b
+        Xinstr.append(Xbi)
+        apX.append(4670.*pow(1.+zi,0.312))
+        E.append(n.Ez)
+        Yi = ( n.Ynu )/restFreq
+        Y.append(Yi)
+        apY.append(64.*pow(1.+zi,0.56))
+        if zi > 6.0 and zi < 9.0:
+            z2fit.append(zi)
+            X2fit.append(Xi)
+            Y2fit.append(Yi)
+    Y = np.array(Y)
+    Xinstr = np.array(Xinstr)
+    kperp = 2.0*np.pi/Xinstr
+    kpar = 2.0*np.pi/(BW*Y)
+    ktot = np.sqrt(2.0*kperp**2 + kpar**2)
[…]

### 2016-04-04 manuscript.tex
Edited mapmaking section and added hybrid approaches section

@@
     %\includegraphics[width=.5\textwidth,clip]{plots/imaging/HERA_z8_SNR_annotated_2015.png}
-    \includegraphics[width=1.0\textwidth,clip]{plots/imaging/HERA_FoV_w_strips.png}
+    \includegraphics[width=1.0\textwidth,clip]{plots/HERA_FoV_w_strips.png}
   \vspace{-20pt}
@@
 In contrast to the delay-spectrum approach, map-making approaches combine baselines to build up information before squaring and calculating the power spectrum.
-The conceptually easiest approach is that one makes an image of the sky and takes a 2-D Fourier transform to determine the power spectrum and so this approach is often called the imaging approach.
-Advantages are that you actually image the reionized bubbles and can determine the magnitude of the smaller $k$-modes on the sky.
-The principle disadvantage is that you have to compute and calibrate the effects of systematics and the propagation of the very bright foregrounds through your system and subtract them incredibly accurately.
-In contrast, the delay-spectrum approach localizes those effects to a parameter space where they can much more easily be filtered and not used in subsequent analysis.
+The conceptually easiest approach is to first make an image cube of the sky as a function of angular position and frequency, and then to Fourier transform, square, and bin the cube to determine the power spectrum.
+One advantage of such an approach is that sky images can be interesting in their own right, both for dealing with measurement systematics and for accessing the non-Gaussian observational signatures (such as ionized bubble structures) that are missed by variance statistics such as the power spectrum.
+The principal disadvantage of a map-making approach is its stringent calibration and forward-modeling requirements for propagating models of bright foregrounds through instrument systematics.
+In contrast, the delay-spectrum allows one to stay considerably closer to the format of raw visibility measurements, thus allowing systematic effects to be sequestered to localized regions of parameter space where they can much more easily be filtered out.
 
@@
 \label{sec:hybridapproach}
+
+While they were presented separately above, the delay-spectrum and map-making approaches need not be viewed as mutually exclusive.
+For example, mapmaking need not be limited to the real space image basis; pipelines that combine multiple baselines into estimates of various Fourier amplitudes of the sky are formally also mapmaking pipelines.
+Like the delay-spectrum approach, such pipelines (see, e.g., \citealt{trott_et_al2016}) avoid the image domain entirely.
+This helps to prevent artifacts that may introduced by imaging algorithms.
+
+More generally, it is possible to express a map-making algorithm as a linear operator that acts on visibility data to produce a compressed dataset.
+Squaring this to obtain an estimate of the power spectrum then results in a weighted quadratic combination of visibilities, where one multiplies the visibility from every baseline with that from every other baseline.
+These visibility product pairs are subsequently normalized to form their own individual estimates of the power spectrum before all the normalized pairs are summed together to form a final power spectrum.
+Measuring a power spectrum in this way combines the best aspects of the mapmaking and delay-spectrum approaches.
+Power spectra are formed directly by the cross-multiplication of visibility data, preserving the delay-spectrum approach's strategy of staying close to quantities measured by an interferometer; on the other hand, by multiplying together the data from every possible combination of baselines, one retains all the information captured by mapmaking.
+Indeed, it can be shown \citep{liu_et_al2014a} that in the limit of infinitely fine Fourier space bins, the hybrid method that we have just described is mathematically equivalent to the mapmaking approach.
+In practice, the hybrid method comes with the added benefit of allowing particular baseline pairs that are suspected of being affected by instrumental systematics to be more surgically downweighted.
 

### 2016-04-04 manuscript.tex
update to kperf and small manuscript

--- 2016-04-04 manuscript.tex
@@
 
-The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by the antenna baselines and bandwidths.
-Note that largest bandwidth (which determines the smallest $\kpar$) is about 10 MHz, for larger bandwidths cosmic evolution begins to impact the result.
-Figure \ref{fig:kperf} shows the allowed k-values based on the beamwidth/baseline (gray plus dark-blue overlap regions, the black lines are the limits) and a bandwidth of 10 MHz with 1024 channels (the blue and dark-blue overlap regions, the blue line is the lower limit).
-The red-dashed line is the allowed smallest $k$-value attainable with this technique.
-Note that at these red-shifts it is dominated by the bandwidth.
+The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by the antenna baselines and bandwidths, as discussed above.
+Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.
+For HERA, wavenumbers are dominated by the bandwidth, not the baseline.
+
+Figure \ref{fig:kperf} is a rather busy plot showing many of the dependencies on wavenumber of bandwidth, configuration and the cosmos.
+It shows the perpendicular, parallel and total resolution wavenumber as a function of redshift for various bandwidths and baselines.
+The term {\em resolution} wavenumber is used, since it is a lower limit for that parameter since larger wavenumbers may be obtained at smaller bandwidths.
+Blue lines denote $\kpar$ values, black lines denote $\kvpr$, and green lines denote total magnitude ($|{\bf k}|$) and the lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber.
+Obviously nearly any values may be used, but here a small sampling of values appropriate for HERA are used and the redshift range is appropriate for the extended frequency range that is the goal for HERA.
+Also, by the $\kpar$ lines the additional ranges indicated by the horizontal lines centered at z=6,12, and 18 are the corresponding redshift extent for that bandwidth at those redshifts.
+Note that for 100 MHz, the concept of a power spectrum at a given redshift is not defined, so some evolutionary model would have to be assumed (this is why the line is dashed and only the z=12 range is shown).
+Also in Fig.
+\ref{fig:kperf}, the lime-green shaded region indicates a plausible region in which HERA could make the measurement and the lighter square within that region is the area that is being targeted in the first instance.
 

### 2016-04-04 plots/kperf.py

@@ -7,6 +7,9 @@ import numpy as np
 
 n = nedclass.ned(H0=100.0)
 
+def get_dz(z,B,rest=1420.):
+    return B*(1+z)**2/rest
+
 def mkPlot(restFreq,BW,b,coeff):
     X = []  # X Mpc/rad
     Xb = [] # scaled version
@@ -22,7 +25,7 @@ def mkPlot(restFreq,BW,b,coeff):
     Y2fit = []  #     "    Y's   "
 
     # Loop over red-shift
-    z = np.arange(0.1,25,.1)
+    z = np.arange(0.1,25.0,.1)
     for zi in z:
         DAi = n.calcUniverse(zi)
         DA.append(DAi)
@@ -109,6 +112,25 @@ for i,BW in enumerate(BW_list):
     plt.semilogy(z,kpar,lsc[i],label=lbl,linewidth=3)
     plt.text(txtloc[i][0],txtloc[i][1],lbl,rotation=txtloc[i][2],color='b',fontsize=16)
 #plt.fill_between(z,p[0],p[1],color='b',alpha='0.25')
+
+#fourth part (redshift bins)
+z_list = [6.,12.,18.]
+for z in z_list:
+    for BW in BW_list:
+        zback,kperp,kpar,ktot = mkPlot(restFreq,BW,D,coeff)
+        y = 0.9*kpar[int(np.where(zback>0.99*z)[0][0])]
+        dz = get_dz(z,BW,restFreq)
+        if BW<0.02:
+	    plotit=True
+	elif z>11. and z<13.:
+	    y = 0.9*y
+	    plotit = True
+	else:
+	    plotit = False
+	if plotit:
+            plt.plot([z-dz/2.,z+dz/2.],[y,y],'b',lw=2)
+            plt.plot([z-dz/2.,z-dz/2.],[0.9*y,1.1*y],'b',lw=2)
+            plt.plot([z+dz/2.,z+dz/2.],[0.9*y,1.1*y],'b',lw=2)
 
 
 plt.xlabel('Redshift [z]',fontsize=16)

### 2016-04-04 plots/wedge.py

@@ -1,0 +1,50 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+kperp = np.logspace(-3,-1,100)
+kperp = np.linspace(0,.1,100)
+kzero = np.zeros(len(kperp))
+kflat = np.ones(len(kperp))*0.06
+#plt.plot(kperp,kflat)
+plt.fill_between(kperp,kzero,kflat,color='k',alpha='0.5')
+
+khoriz = 7.5*kperp
+g = np.where(khoriz < max(kflat))
+khoriz[g] = max(kflat)
+#plt.plot(kperp,khoriz)
+plt.fill_between(kperp,kflat,khoriz,color='b',alpha='0.5')
+#1ksys = khoriz + 0.05 + 1.1e-3/(kperp+1e-4)
+#1ksys[g] = ksys[g[0][-1]+1]*np.ones(len(g[0]))
+#2ksys = 4.1*(kperp+0.00) + 0.15
+x = [0.0,0.01,0.09]
+y = [0.15, 0.18, 0.50]
+#x = [-2.,1.,2.]
+#y = [7.,10.,27.]
+num = (y[2]-y[0]) - (y[1]-y[0])*(x[2]**2-x[0]**2)/(x[1]**2-x[0]**2)
+den = (x[2]-x[0]) - (x[1]-x[0])*(x[2]**2-x[0]**2)/(x[1]**2-x[0]**2)
+B = num/den
+A = (y[1]-y[0] - B*(x[1]-x[0]))/(x[1]**2-x[0]**2)
+C = y[0] - A*x[0]**2 - B*x[0]
+print 'A,B,C'
+print A,B,C
+ksys = A*kperp**2 + B*kperp + C  ###Curved
+ksys = 8.5*kperp + 0.08 ###Straight
+#plt.plot(kperp,ksys,'r--')
+plt.fill_between(kperp,khoriz,ksys,color='r',alpha='0.2')
+plt.axis([0.0,0.1,.0,1])#.55])
+plt.xlabel(r'k$_\perp [h$/Mpc]',fontsize=14)
+plt.ylabel(r'k$_{||} [h$/Mpc]',fontsize=14)
+
+useCables = False
+if useCables:
+    kperp_cables = [0,0.085]
+    kpam = 0.45
+    kpar_cables1 = [kpam,kpam]
+    plt.plot(kperp_cables,kpar_cables1,'r--')
+    kpam = 0.465
[…]

### 2016-04-04 manuscript.tex
manuscript edits primarily to kperf figure discussion

--- 2016-04-04 manuscript.tex
@@
 Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.
-As a first step, areas of the sky with where these signals are minimal (for example, outside the galactic plane and away from strong point sources) are targeted (REF FOR THIS).
+As a first step, areas of the sky where these signals are minimal (for example, outside the galactic plane and away from strong point sources) are targeted (REF FOR THIS).
 More importantly and most fortuitously however is the fact that all of these foreground signals are smooth spectrum sources whereas the expected spectrum of the EOR is expected to be rough since it is made up of non-ionized regions which are randomly distributed over a wide range of redshifts.
@@
 
-This section will provide a brief overview of existing measurements and the theoretical underpinning of how foregrounds contaminate the measurement (the so-called ``wedge''), discuss the various techniques used to make the measurement and finally provide a brief discussion on calibration issues and techniques.
+This section will provide a brief overview of  the theoretical underpinning of how foregrounds contaminate the measurement (the so-called ``wedge''), discuss the various techniques used to make the measurement and finally provide a brief discussion on calibration issues and techniques.
 
@@
 foreground leakage in Fourier space (see Fig.~\ref{fig:wedge}), 
-outside of which the 21\,cm signal dominates.
+outside of which the 21\,cm signal dominates (the ``EoR window'').  The coordinates of the wedge are discussed below.
 Through theoretical and observational work
 \citep{datta_etal2010,morales_et_al2012,parsons_et_al2012b,vedantham_2012,thyagarajan_et_al2013,hazelton_et_al2013,pober_etal2013b,liu_et_al2014a,liu_et_al2014b},
-we have learned how the boundary between the wedge and our ``EoR window" is determined by the separation between antennas,
+we have learned how the boundary between the wedge and  the EoR window is determined by the separation between antennas,
 signal reflections within antennas, and the angular response of the antenna beam.  Deep integrations also show us
@@
 
-Though the power spectrum measurement is over a volume, it is instructive to split the spatial wavenumbers into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kvpr$ is determined by the antenna baseline and $\kpar$ by the frequency.
+Though the power spectrum measurement is over a volume, it is instructive to split the spatial wavenumbers into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kvpr$ is determined by the antenna baseline and $\kpar$ by the bandwidth.
 This is useful since it allows us to split out the chromatic response of the interferometer and isolate a phase space where foreground sources ({\em i.e.} everything {\em not} the EOR) contaminate the signal of interest from where they don't.
-This contaminated phase space may be shown [REF] to be a wedge-shaped region in $\kpr$:$\kpar$ space:
+As mentioned above, this contaminated phase space may be shown [REF] to be a wedge-shaped region in $\kpr$:$\kpar$ space:
 \begin{equation}
@@
 \label{sec:delayapproach}
-Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement.
+Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement of the EOR spatial power spectrum.
 There are now numerous papers in the literature on the technique of measuring the power spectrum using an interferometer and the attendant issues [[[list of citations]]].
@@
 The so-called delay-spectrum approach leverages the interferometer measurement to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
-Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum.
+Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum, which contrasts to the map-making techniques briefly described below.
 
@@
 
-The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by the antenna baselines and bandwidths.
-Note that largest bandwidth (which determines the smallest $\kpar$) is about 10 MHz, for larger bandwidths cosmic evolution begins to impact the result.
-Figure \ref{fig:kperf} shows the allowed k-values based on the beamwidth/baseline (gray plus dark-blue overlap regions, the black lines are the limits) and a bandwidth of 10 MHz with 1024 channels (the blue and dark-blue overlap regions, the blue line is the lower limit).
-The red-dashed line is the allowed smallest $k$-value attainable with this technique.
-Note that at these red-shifts it is dominated by the bandwidth.
+The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by both the antenna baseline and bandwidth, as discussed above.
+The $\kvpr$ components are proportional to the visibilities and $\kpar$ are proportional to the Fourier transform of the frequency response.
+The Fourier transform of a frequency spectrum is a delay, hence the name ``delay-spectrum'' for this technique.
+Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.
+For HERA, wavenumbers are dominated by the bandwidth, not the baseline.
+
+As an aside, note the parallels to imaging with an interferometer.
+For imaging, one takes a 2-D spatial Fourier transform of the visibilities, and produces images as a function of frequency.
+Here, one is taking a 1-D Fourier transform along the other axis (the bandpass) and produces a 3-D spatial power spectrum.
+In imaging, point sources are transformed to a single pixel, whereas for the delay-spectrum, smooth sources are transformed to a single channel (but then get forced into the wedge due to the inherent chromatic response of an interferometer).
+For imaging, we are concerned with ripples in the passband due to standing waves, whereas for the delay-spectrum we are concerned with delayed versions of the signal due to multiple reflections.
+
+Figure \ref{fig:kperf} is a rather busy plot showing many of the dependencies on wavenumber of bandwidth, configuration and the cosmos.
+It shows the perpendicular, parallel and total resolution wavenumber as a function of redshift for various bandwidths and baselines.
+The term {\em resolution} wavenumber is used, since it is a lower limit for that parameter since larger wavenumbers may be obtained by channelizing.
+The redshift range is appropriate for the extended frequency range that is the goal for HERA, however $z$=6-13 is the primary region for EOR studies.
+
+In Fig.
+\ref{fig:kperf}, blue lines denote $\kpar$ values, black lines denote $|\kvpr|$, and green lines denote total magnitude ($|{\kvec}|$).
+The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.
+Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.
+Also shown just below the blue $\kpar$ curves are redshift ranges for that bandwidth centered at $z$=6,12, and 18 (they appear as horizontal ``error'' bars).
+Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer really appropriate, so some evolutionary model would have to be assumed (this is why the line is dashed and only the $z$=12 range is shown).
+Also in Fig.
+\ref{fig:kperf}, the lime-green shaded region indicates a plausible region in which HERA could operate and the lighter square within that region is the area that is being targeted in the first instance.
 
@@
 In contrast, the delay-spectrum allows one to stay considerably closer to the format of raw visibility measurements, thus allowing systematic effects to be sequestered to localized regions of parameter space where they can much more easily be filtered out.
+
+This approach is being pursued primarily via FHD.
 

### 2016-04-04 manuscript.tex
in high-level requirements and system design

@@
 \begin{figure}[t]
+\centerline{
 \includegraphics[height=3in]{plots/wedgeWithBG.png}
+}
 \caption{Foregrounds are a primary challenge facing 21\,cm cosmology experiments. 
@@
 \begin{figure}[h!]
+\centerline{
 \includegraphics[width=0.5\textwidth]{plots/boost.png}
+}
 \caption{Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) for representative observing campaigns for various instruments.
@@
 }
-\caption{\small kperf}
+\caption{\small $\kvec, \kpar, \kperp$ resolution wavenumbers with various bandwidths and baselines as a function of redshift.
+See text for discussion.
+The full redshift range is the HERA extended goal.
+The initial HERA delay-spectrum power spectrum goal is the light square region in the upper right.}
 \label{fig:kperf}
@@
 \label{sec:requirements}
-From the science and techniques discussed above, we may derive some high-level requirements, predicating on optimizing for the delay-spectrum approach for detecting the EOR while not unduly limiting other approaches or other science goals.
+From the science and techniques discussed above, we may derive some high-level requirements predicated on optimizing for the delay-spectrum approach of detecting the EOR power spectrum while not unduly limiting other approaches or other science goals.
 The fact that HERA is an experiment and not meant to be a long-term general use facility greatly facilitates this optimization.
@@
 \subsection{Antenna Design and Configuration}
-Thus, to open the widest possible window for EoR measurements, HERA must use close-packed antennas that
-minimize signal reflections and deliver significant forward gain relative to their horizon response.
+The primary new feature of HERA over previous generation experiments (in this context, primarily PAPER) is the use of large elements to increase the sensitivity.
+The cost-performance optimization will
+be discussed below, however we see that to address the central challenges of the delay-spectrum approach in the context of the wedge, to open the widest possible window for EoR measurements, 
+HERA must use close-packed antennas that
+minimize signal reflections over long delays and deliver significant forward gain relative to their horizon response.
 Tests with prototype HERA antennas (Figs. \ref{fig:orbcommexptandbeammap} and \ref{fig:reflectometry}, discussed in \S\ref{sec:antenna})
 indicate that a moderately large parabolic dish with a short focal height can meet these requirements
-\citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016}.
+\citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016}.  Note that to pursue map-making and hybrid techniques, additional antenna outriggers are valuable additions, so the design
+should be able to accommodate ~1km baselines.  In referring to the EOR sensitivity of the array however, the number of antennas just in the core is sometimes used.
 
@@
 
-The scientific requirement on channel bandwidth is to allow access to k-modes of greater than 1 h/Mpc [ref], which requires about 256 channels over the 100 MHz bandwidth.
+The scientific requirement on channel bandwidth is to allow access to k-modes of 1 h/Mpc [ref], which requires about 256 channels over the 100 MHz bandwidth.
 However, in order to handle radio frequency interference as well as to allow the bandpass to be characterized, a specification of 1024 channels has been chosen.
@@
 
-As mentioned earlier (see Fig. \ref{fig:kperf}) we should be able to process at least 10 MHz
-
+The total simultaneously processed bandwidth is the full 100 MHz, since we wish to efficiently probe the entire redshift range.
+The main impact from this requirement is the bandwidth from the digitizer back to the correlator, however this bandwidth is easily handled with current generation digital back-ends.
+
+\subsection{Sensitivity Optimization}
+\label{sec:cost}
+As an instrument to characterize the power spectrum over its evolution, the specification on sensitivity is to make at least a nominal detection over the redshift region of support, or $z$=6-13, along with a very robust detection at the peak.
+As seen in Fig.
+\ref{fig:Sensitivities} and Table \ref{tab:signif}, an array comprising 320 14-m core antennas (note, labelled HERA-350, which includes the outriggers) has sensitivity to a
+fiducial reionization model across that redshift range.  We therefore adopt the delay-spectrum power spectrum sensitivity of 320 14-m close-packed antennas as the minimum sensitivity needed.  
+For this sensitivity, the number of elements needed for a given diameter is
+\begin{equation}
+N_{\text core} = (320*14)/D
+\end{equation}
+where $N_{\text core}$ and $D$ are the number of antennas in the core and their diameter respectively.
+Note that the number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} Pober) over a range of values for the HERA configuration.
+This dependency differs from Mellema and analytical epressions, mostly likely due to the assumed configuration.
+Optimization is 
+therefore done for this sensitivity against cost for various diameters and total element numbers, while trying to minimize multiple reflections on long delays.
+Figure \ref{fig:costfig} summarizes this optimization, which
+is discussed below.
+
+To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, where the total number was constrained to keep constant sensitivity at the minimum discussed in the previous paragraph.
+This included the full system, except the post-processing (that is, everything after the archived correlator output).
+Figure \ref{fig:costfig} shows the resulting normalized system cost/performance-curve as the horizontal shading relative to the canonical HERA design 
+which becomes very flat (colored purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.
 \begin{figure}[t]
@@
 \end{figure}
-
-\subsection{Sensitivity Optimization}
-\label{sec:cost}
-As an instrument to characterize the power spectrum over its evolution, the specification on sensitivity is to make at least a nominal detection over the redshift region of support, or $z$=6-12, along with a very robust detection at the peak.
-As seen in Fig.
-\ref{fig:Sensitivities} and Table \ref{tab:signif}.
-To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, where the total number was constrained to keep constant sensitivity for this measurement, which is shown in Figure \ref{fig:costfig}.
-This included the full system, except the post-processing.
-By assuming canonical values of 331 14-meter antennas, the number of elements needed for a given diameter to match the sensitivity is
-\begin{equation}
-N_a = 4634/D_a
-\end{equation}
-where $N_a$ and $D_a$ are the number of antennas and their diameter respectively.
-Note that the number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} Pober) over a range of values for the HERA configuration.
-This dependency differs from Mellema, mostly likely due to the assumed configuration.
-Figure \ref{fig:costfig} shows the resulting normalized system cost/performance-curve as the horizontal shading relative to the canonical HERA design 
-which becomes very flat (colored purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.
 

### 2016-04-06 manuscript.tex
mainly worked on sensitivity figure + some wordsmithing

@@
 have produced the first astrophysically constraining upper limits on the 21\,cm EoR power spectrum, 
-providing evidence for significant heating in the IGM prior to reionization \citep{parsons_etal2014,ali_et_al2015,pober_et_al2015}.
+providing evidence for significant heating in the IGM prior to reionization \citep{parsons_etal2014, 2015ApJ...809...61A,pober_et_al2015}.
 However, current experiments cannot expect more than marginal detections of the EoR signal. 
@@
 \caption{\hspace{-1.2mm}: Predicted SNRs of 21\,cm experiments for an EoR model with 50\% ionization at $z=9.5$, with 1080 hours observation, integrated over a $\Delta z$ of $0.8$.
-Foreground avoidance represents an analysis comparable to \cite{ali_et_al2015}, whereas foreground modeling allows signifcantly more $k$ modes of the cosmological signal to be recovered.}
+Foreground avoidance represents an analysis comparable to \cite{2015ApJ...809...61A}, whereas foreground modeling allows signifcantly more $k$ modes of the cosmological signal to be recovered.}
 \small
@@
 While the existing experiment with the most collecting area, 
-the LOw Frequency ARray (LOFAR; \citealt{yatawatta_et_al2013}),
+the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V, yatawatta_et_al2013}),
 faces large
@@
 The bandwidth
-therefore equates to a cosmic volume along the site of the telescope, with the frequency determining the cosmic age.
-
-Given the difficulty of the experiment, initial measurements of the EOR strive to measure a statistical power spectrum of the signal since the nature of the reionization process should have a specific spatial signature.
+therefore equates to a cosmic distance along the site of the telescope, with the frequency determining the cosmic age.
+An observation of an area of sky over a given bandwidth is therefore providing an average
+over a cosmic volume.
+
+As the distance to the EOR is great (about 13 billion light-years) the signal is quite weak.
+However, the signal also subtends the entire sky, so initial measurements of the EOR strive to measure a statistical power spectrum of the signal over the sky since the nature of the reionization process should have a specific spatial signature.
 The goal is therefore to measure a range of aggregate spatial scales on the sky, rather than to image the signal directly.
-Imaging does remain an ultimate goal to fully understand the process, however we will need a greater understanding of the full measurements to achieve this more difficult goal.
+Imaging does remain an ultimate goal to fully understand the process, however we will likely need a greater understanding of the signal characteristics and the systematics of our telescopes to achieve this more difficult goal.
 MANY PAPERS, see Nithya and Jonnie recent
 
-Obviously between our present observation point and the Epoch of Reionization approximately 13 billion years ago lies the entire intervening Universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).
+Obviously between our present observation point and the Epoch of Reionization lies the entire intervening Universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).
 Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.
@@
 PAPER and the MWA have produced progressively deeper limits
-\citep{dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
+\citep{dillon_et_al2015,parsons_etal2014,2015ApJ...809...61A}, with PAPER
 yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
@@
 that, to the limits of current sensitivity, foreground emission is absent outside of the wedge; it can only 
-appear there through instrumental leakage \citep{parsons_etal2014,ali_et_al2015,moore_et_al2016,kohn_et_al2016}.
+appear there through instrumental leakage \citep{parsons_etal2014,2015,ApJ...809...61A,moore_et_al2016,kohn_et_al2016}.
 
@@
 By avoiding foregrounds, PAPER has placed limits within an order
-of magnitude (in mK) of these models \citep{ali_et_al2015} and show the ``EoR Window'' to be foreground-free.
+of magnitude (in mK) of these models \citep{2015ApJ...809...61A} and show the ``EoR Window'' to be foreground-free.
 HERA's dish and configuration optimize wedge/window isolation and direct sensitivity to low-$k_\perp$ modes where EoR is brightest.
@@
 The so-called delay-spectrum approach leverages the interferometer measurement to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
-Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum, which contrasts to the map-making techniques briefly described below.
+Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum, which contrasts to the map-making techniques briefly described in the following section.
 
@@
 relatively agnostic with respect to observing parameters, making it a safe and robust first step in data processing
-\citep{zheng_et_al2014,parsons_etal2014,ali_et_al2015}.
+\citep{zheng_et_al2014,parsons_etal2014,2015ApJ...809...61A}.
 The Internet link between the Karoo and data centers in the U.S.\ has been extensively tested, with assistance from SKA-SA, TENET

### 2016-04-06 manuscript.tex
moved existing and started calibration

@@
 
-\subsection{Existing Measurements}
+\subsection{Secondary Scientific Objectives}%\\Precision Cosmology, Reionization Imaging, and High-$z$ IGM X-Ray Heating Science}
+
+\startsquarepar\emph{\textbf{Precision Cosmology.}}
+\label{sec:tau}
+By advancing our understanding of reionization astrophysics, HERA improves CMB constraints on 
+fundamental cosmological parameters by
+removing the optical depth
+$\tau$ as a nuisance parameter. HERA breaks the degeneracy between the
+constraints on $\tau$ and the sum of the neutrino masses $\sum m_\nu$, which has
+been identified as a potential problem for Stage 4 CMB lensing experiments. A
+HERA-informed estimate of $\tau$ enables CMB lensing experiments to achieve a
+$0.012\,\textrm{eV}$ error on $\sum m_\nu$ \citep{liu_et_al2015}. This would
+represent a $\sim$5$\sigma$ cosmological detection of the neutrino masses even
+under the most pessimistic assumptions
+% that their sum is the minimum value of $0.058\,\textrm{eV}$ 
+still allowed by neutrino oscillation \stopsquarepar
+
+\begin{figure}[h!]
+\centering
+    \includegraphics[width=0.50\textwidth,clip]{plots/sigmaTau.pdf}
+\caption{Likelihood contours ($68\%$ and $95\%$) for $\sigma_8$ and $\tau$ using \emph{Planck} constraints (blue) and 
+adding HERA data (red). 
+%Light and heavy contours signify $68\%$ and $95\%$ confidence regions, respectively. 
+The $21\,\textrm{cm}$ constraints break the CMB degeneracy between the amplitude of density fluctuations and the optical depth, improving constraints on both.}
+	\label{fig:sigma8Tau}
+\end{figure}
+
+\noindent experiments
+\citep{allison_et_al2015}, making HERA key to our understanding of neutrino
+physics. HERA's estimate of $\tau$ would also break the degeneracy between
+$\tau$ and the amplitude of matter fluctuations (expressed in Fig.
+\ref{fig:sigma8Tau} as $\sigma_8$) that arises when using only
+CMB data. HERA effectively reduces error bars on $\sigma_8$ by more
+than a factor of three \citep{liu_et_al2015}, potentially elucidating
+current tensions between cluster cosmology constraints and those from primary
+CMB anisotropies.
+
+\emph{\textbf{First Images of the Reionization Epoch.}}
+\label{sec:imaging}
+In addition to measuring the power spectrum, HERA has sufficient 
+sensitivity to directly image the IGM during reionization,
+and will observe a 1440 deg$^2$ stripe, comparable to future WFIRST
+large area near-IR surveys.
+ After 100 hours on a
+single field (achievable in 200 nights), HERA reaches a surface brightness 
+sensitivity of 50 $\mu$Jy/beam (synthesized beam FWHM $\sim$ 24$'$) compared 
+to the brightness temperature fluctuations of up to 400 $\mu$Jy/beam in typical 
+reionization models (see Fig.~\ref{fig:LSS}). From the standpoint of sensitivity alone, HERA is capable of 
+detecting the brightest structures at $z=8$ with SNR $>$ 10. Additionally, the design 
+of HERA places it in a unique position to directly explore calibration techniques
+(e.g. redundant calibration, \citealt{liu_et_al2010}), while retaining a high quality 
+point spread functions for imaging and identifying foregrounds. Leveraging the expertise gained 
+through both precursor strategies, HERA will yield exceptionally high dynamic range
+images.
+
+%\subsubsection{The 21\,cm Signal in the Pre-Reionization Epoch}
+
+%\begin{wrapfigure}{R}{0.5\textwidth}
+\begin{figure}[b!]
+\centering
+\vspace{-15pt}
+    %\includegraphics[width=.5\textwidth,clip]{plots/imaging/HERA_z8_SNR_annotated_2015.png}
+    \includegraphics[width=1.0\textwidth,clip]{plots/HERA_FoV_w_strips.png}
+  \vspace{-20pt}
+\caption{\footnotesize 
+HERA will observe a 1440 deg$^2$ stripe centered near $\delta = -30^\circ$.
+HERA can measure the ionization state around galaxies in, e.g., the GOODS-South field that contains a third of all known $z\!>\!8$ galaxies.
+HERA's primary imaging data product to the community will be deep cubes along the HERA stripe suitable for cross-correlation.
+\label{fig:LSS} }
+\vspace{-10pt}
+\end{figure}
+%  \vspace{-10pt}
+%\end{wrapfigure}
+
+\emph{\textbf{Pre-Reionization Heating.}}
+\label{sec:EoX} 
+Prior to reionization, the 21\,cm signal is a sensitive probe of the first
+luminous sources and IGM heating mechanisms. First stars are expected to form
+at $z \sim 25-30$ and their imprint on the 21\,cm signal is expected to be
+sensitive to the halo mass where they are formed \citep{mesinger_et_al2015}.
+The IGM is then expected to be heated by first
+generation X--ray binaries
+\citep{furlanetto_et_al2006_global,pritchard_et_al2007,mesinger_et_al2013} or
+by the hot interstellar medium produced by the first supernovae
+\citep{pacucci_et_al2014}, although the heating timing and magnitude is still
+very much debated \citep{fialkov_et_al2012}. Dark matter annihilation
+\citep{evoli_et_al2014} could also inject energy in the IGM, leaving its
+imprint in the 21\,cm signal.
+
+With support from the Moore Foundation to develop feeds sensitive down to 50\,MHz, HERA will probe the IGM prior to
+reionization and constrain the sources of heating, obtaining, in the case of
+X--ray heating, percent-level constraints on the efficiency with which star-forming baryons
+produce X-rays \citep{ewall-wice_et_al2015}.
+While the observational and analytical state of the art of pre-reionization 21\,cm science is not as advanced as for the EoR \citep{ewall-wice_et_al2016-EoXLimits}, the risk of lower-frequency observing is borne by the Moore Foundation.
+Lower
+frequency observations also test feedback
+mechanisms that interact with low-mass halos at high redshifts
+\citep{Iliev_et_al2007,Iliev_et_al2012,ahn_et_al2012}.
+Such constraints, while interesting in their own right, also reduce
+the susceptibility of the aforementioned $21\,\textrm{cm}$-derived $\tau$
+constraints to uncertainties in high-redshift physics \citep{liu_et_al2015}, especially if they are 
+combined with upcoming or proposed measurements of 
+the pre-reionization sky-averaged spectrum \citep{fialkov_and_loeb2016}. Additionally, they may be crucial to a correct
+interpretation of kinetic Sunyaev-Zel'dovich effect constraints on reionization
+from the CMB \citep{park_et_al2013}. The high redshift probe of
+structure afforded by the low frequency $21\,\textrm{cm}$ measurements will
+permit some of the most direct observations of hypothesized suppressions of
+small-scale structure
+\citep{dalal_et_al2010,tseliakhovich_et_al2011,fialkov_et_al2012} arising from
+predicted supersonic relative velocities between dark matter and baryonic gas
+\citep{tseliakhovich_and_hirata2010}. No other electromagnetic probe can provide direct observations of this
+epoch.
+
+\subsection{Other Scientific Objectives}
+\label{subsec:broader_science}
+
+\noindent With unprecedented sensitivity at this frequency range, HERA can deliver much more than its core 21\,cm science. 
+Some of HERA's broader science impacts derive directly from the public data products HERA delivers.  Other
+areas of impact may require additional hardware or software that HERA can host at no extra risk. 
+In this respect, HERA serves as a ``platform" for additional science programs with external funding.
+The Moore Foundation's support of HERA feed development and data analysis targeting the pre-reionization science described in \S\ref{sec:EoX}
+is a prime example of how we envision this functioning.
+Below, we list examples of the broader science that HERA can deliver.
+
+\emph{\textbf{Cross-Correlations with Other Reionization Probes.}}
+\label{sec:crossCorr}
+HERA's public data provide new opportunities for cross-correlation studies.
+Cross-correlation between HERA 21\,cm images and other high-redshift probes
+(e.g. JWST; WFIRST; CMB maps; CO, CII, and Ly-$\alpha$ intensity mapping) 
+can provide an independent confirmation of the 21\,cm power spectrum
+\citep[i.e.][]{lidz_et_al2009,dore_et_al2014,silva_et_al2015,vrbanec2016} and enable rich new studies of
+the interaction between galaxies and their ionization environment. 
+In particular, cross-correlating 21\,cm with galaxy surveys can measure the characteristic bubble size around galaxies of
+different luminosities \citep{lidz_et_al2009} and help separate the degeneracy
+between the fraction of photons escaping the galaxies and the
+total number of ionizing photons produced \citep{zackrisson2013}.
+
+HERA is placed in a unique position to enable such cross-correlation: the GOODS-South field---one of the most panchromatically studied regions of the sky, the site of the Hubble UDF, and home to over a third of all known $z>9$
+galaxies---lies in its field of view. HERA's images of the IGM can provide environmental context to galaxy surveys through identification of
+ionized bubbles \citep{malloy_lidz2013}, or in a more statistical sense as
+described in \cite{beardsley_et_al2015}.
+
+\textbf{\emph{Searching for Exoplanetary Radio Bursts.}}
+\label{sec:exoplanets}
+HERA could be an powerful tool in the search for bright, highly polarized auroral bursts
+from exoplanets %due to the electron cyclotron maser instability 
+\citep{treumann2006,hallinan_et_al2015}. These
+distinctive, coherent bursts occur periodically at the planetary
+rotation rate, which cannot otherwise be measured. 
+A burst from a terrestrial planet could point toward
+habitability, since it implies a powerful magnetic field
+protecting the atmosphere and perhaps the biosphere from energetic stellar wind
+particles \citep{tarter_et_al2007}. HERA's sensitivity, large FoV, long campaigns, and precise calibration are all well-suited
+for discovering exoplanetary radio bursts from Jupiter-like planets out to 25~pc.
+Many well studied exoplanetary systems within that distance are in the HERA stripe, including Fomalhaut, Gl 667 C, Gl 433, HD 147513, and Gl 317.
+
+\textbf{\emph{Fast Radio Burst Followup.}}
+\label{sec:FRBs}
+Fast Radio Bursts (FRB) are millisecond-long radio flashes whose origin has remained a great enigma ever since their discovery \citep{2007Sci...318..777L}. 
+If equipped with a suitable backend, HERA could be triggered by nearby, higher-frequency telescopes for FRB followup, saving baseband data and thus full sensitivity to all dispersion measures.
+Bursts like that discovered by \citet{masui_et_al2015}, the lowest frequency
+FRB detection to date, should be seen hourly by HERA at 5--10$\sigma$. %(including the effects of scattering)
+ Observations at HERA frequencies are very sensitive to the physics of
+the intervening medium, particularly deviations from $\lambda^2$
+dispersion. Detecting deviations would rule out broad classes of models and could indicate whether FRBs are at cosmological distances.
+
+\section{Measuring the EOR}
+\label{sec:eormeas}
+Due to the expansion of the Universe over cosmological time, we can identify and measure the early Universe via the red-shift of spectral lines.
+The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire Universe back almost to the period of recombination.
+The bandwidth
+therefore equates to a cosmic distance along the site of the telescope, with the frequency determining the cosmic age.
+An observation of an area of sky over a given bandwidth is therefore providing an average
+over a cosmic volume.
+
+As the distance to the EOR is great (about 13 billion light-years) the signal is quite weak.
+However, the signal also subtends the entire sky, so initial measurements of the EOR strive to measure a statistical power spectrum of the signal over the sky since the nature of the reionization process should have a specific spatial signature.
+The goal is therefore to measure a range of aggregate spatial scales on the sky, rather than to image the signal directly.
+Imaging does remain an ultimate goal to fully understand the process, however we will likely need a greater understanding of the signal characteristics and the systematics of our telescopes to achieve this more difficult goal.
+MANY PAPERS, see Nithya and Jonnie recent
+
+Obviously between our present observation point and the Epoch of Reionization lies the entire intervening Universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).
+Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.
+As a first step, areas of the sky where these signals are minimal (for example, outside the galactic plane and away from strong point sources) are targeted (REF FOR THIS).
+More importantly and most fortuitously however is the fact that all of these foreground signals are smooth spectrum sources whereas the expected spectrum of the EOR is expected to be rough since it is made up of non-ionized regions which are randomly distributed over a wide range of redshifts.
+This fact allows us to try and isolate foregrounds from the EOR, as will be discussed.
+
+This section will provide a brief overview of  the theoretical underpinning of how foregrounds contaminate the measurement (the so-called ``wedge''), discuss the various techniques used to make the measurement and finally provide a brief discussion on calibration issues and techniques.
+
+\subsection{The ``Wedge''}
+The past three years have seen deep EoR observations with PAPER, the MWA, 
+the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2013}), and LOFAR.
+PAPER and the MWA have produced progressively deeper limits
+\citep{dillon_et_al2015,parsons_etal2014,2015ApJ...809...61A}, with PAPER
+yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
+(Fig. \ref{fig:limits}).
+The inherent challenge is to simultaneously meet stringent sensitivity requirements 
+while suppressing foregrounds $\sim5$ orders of magnitude
+brighter than the 21\,cm signal, which has been addressed along a number of approaches.  
+To address this challenge,  all aspects of the experimental process must be refined and improved, spanning 
+calibration ({\em e.g.} \cite{zheng_et_al2014, jacobs_et_al2016, barry_et_al2016}), handling foreground contamination
+({\em e.g.} \cite{moore_et_al2013,moore_et_al2016,thyagarajan_et_al2015a,pober_et_al2016}),
+and the interferometer design itself ({\em e.g.} \cite{parsons_et_al2012a,dillon_parsons2016}).
+
+Perhaps the most important advance informing HERA's design is a
+refined understanding of how smooth-spectrum foregrounds interact
+with instrument chromaticity to produce a characteristic ``wedge" of
+foreground leakage in Fourier space (see Fig.~\ref{fig:wedge}), 
+outside of which the 21\,cm signal dominates (the ``EoR window'').  The coordinates of the wedge are discussed below.
+Through theoretical and observational work
+\citep{datta_etal2010,morales_et_al2012,parsons_et_al2012b,vedantham_2012,thyagarajan_et_al2013,hazelton_et_al2013,pober_etal2013b,liu_et_al2014a,liu_et_al2014b},
+we have learned how the boundary between the wedge and  the EoR window is determined by the separation between antennas,
+signal reflections within antennas, and the angular response of the antenna beam.  Deep integrations also show us
+that, to the limits of current sensitivity, foreground emission is absent outside of the wedge; it can only 
+appear there through instrumental leakage \citep{parsons_etal2014,2015,ApJ...809...61A,moore_et_al2016,kohn_et_al2016}.
+
+Though the power spectrum measurement is over a volume, it is instructive to split the spatial wavenumbers into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kvpr$ is determined by the antenna baseline and $\kpar$ by the bandwidth.
+This is useful since it allows us to split out the chromatic response of the interferometer and isolate a phase space where foreground sources ({\em i.e.} everything {\em not} the EOR) contaminate the signal of interest from where they don't.
+As mentioned above, this contaminated phase space may be shown [REF] to be a wedge-shaped region in $\kpr$:$\kpar$ space:
+\begin{equation}
+|\kpar| \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
+\end{equation}
+where $c$ is the speed of light and $S$ is a parameter accounting for offsets related to the combined spectral smoothness of the foregrounds and the antenna response (see Fig. \ref{fig:wedge}).  
+A key issue to measuring the EOR power spectrum is to understand and minimize such effects.
+For an analytical derivation, see \cite{zahn_etal2012,vedantham_2012,liu_et_al2014b},  in simulation see \cite{datta_etal2010,hazelton_et_al2013} and for observations
+\cite{pober_etal2013b,2015arXiv150601026P,parsons_etal2014,2015arXiv150206016A}.
+Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).
+In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_o=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.
+
+\begin{figure}[t]
+\centerline{
+\includegraphics[height=3in]{plots/wedgeWithBG.png}
+}
+\caption{Foregrounds are a primary challenge facing 21\,cm cosmology experiments. 
+HERA leverages a ``wedge'' in the spherically-averaged ($\mathbf{k}$ is broken into $k_\|$ and $k_\perp$) power spectrum (center panel; \citealt{dillon_et_al2015}).
+Smooth-spectrum foregrounds (right panel) $\sim$5 orders of magnitude brighter than fiducial EoR models (left panel; \citealt{mesinger_et_al2011}) create the ``wedge'' when they interact with the interferometer's chromatic response.
+By avoiding foregrounds, PAPER has placed limits within an order
+of magnitude (in mK) of these models \citep{2015ApJ...809...61A} and show the ``EoR Window'' to be foreground-free.
+HERA's dish and configuration optimize wedge/window isolation and direct sensitivity to low-$k_\perp$ modes where EoR is brightest.
+%v2
+Cartoon of the wedge.
+The gray rectangle represents the smooth-spectrum foregrounds as they would be if the instrument had a perfect achromatic response.
+The pink wedge area shows the chromatic response of the interferometer, where additional power is thrown up from the gray region to higher k-values with increasing baseline.
+The additional horizontal pink line illustrates the affect that other systematics, such as reflections on cables, can have in contaminating the measurement.
+The white region shown the 'eor window', where the measurement is not contaminated by foregrounds or systematics.}
+\label{fig:wedge}
+\end{figure}
+
+The techniques to measure the power spectrum may be broken down into two principle techniques (delay-space and map-making), plus additional hydrid methods.
+These are briefly summarized below, with an emphasis on the delay-spectrum technique as the initial approach taken by HERA.
+Finally, we provide a short discussion on calibration in the concept of this design.
+
+\subsection{Delay-Spectrum Approach}
+\label{sec:delayapproach}
+Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement of the EOR spatial power spectrum.
+There are now numerous papers in the literature on the technique of measuring the power spectrum using an interferometer and the attendant issues [[[list of citations]]].
+Below, a high-level overview based on  \cite{parsons_et_al2012b} will be presented.
+The so-called delay-spectrum approach leverages the interferometer measurement to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
+Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum, which contrasts to the map-making techniques briefly described in the following section.
+
+We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment:
+\begin{equation}
+{\hat P}(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} \right]
+\label{eq:Pk_baseline}
+\end{equation}
+where $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes (so, relating wavenumber to physical volume at a given redshift), $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.
+The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.
+Note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
+Rather than ${\hat P}(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.
+
+Since the thermal noise per visibility baseline may be expressed as
+\begin{equation}
+V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
+\label{eq:sensitivity_per_baseline}
+\end{equation}
+where $\tau$ is the integration time (see {\em e.g.} TMS),
+to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq.
+\ref{eq:Pk_baseline}.
+Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
+\begin{equation}
+\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
+\label{eq:sensitivity}
+\end{equation}
+where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement 
+(see \cite{parsons_et_al2012b}).
+Fig.
+\ref{fig:boost} plots this ``boost'' factor for various telescopes under nominal observing campaigns using the delay-spectrum technique.
+The large improvement for PAPER and HERA is the fact that the baselines are redundantly placed on appropriate spatial scales, so that a more limited number of unique baselines get sampled many times.
+
+\begin{figure}[h!]
+\centerline{
+\includegraphics[width=0.5\textwidth]{plots/boost.png}
+}
+\caption{Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) for representative observing campaigns for various instruments.
+ADD NEW HERA CONFIGURATION AND LOFAR AND PUT IN 'HOW TO MAKE THIS PLOT'}
+\label{fig:boost}
+\end{figure}
+
+The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by both the antenna baseline and bandwidth, as discussed above.
+The $\kvpr$ components are proportional to the visibilities and $\kpar$ are proportional to the Fourier transform of the frequency response.
+The Fourier transform of a frequency spectrum is a delay, hence the name ``delay-spectrum'' for this technique.
+Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.
+For HERA, wavenumbers are dominated by the bandwidth, not the baseline.
+
+As an aside, note the parallels to imaging with an interferometer.
+For imaging, one takes a 2-D spatial Fourier transform of the visibilities, and produces images as a function of frequency.
+Here, one is taking a 1-D Fourier transform along the other axis (the bandpass) and produces a 3-D spatial power spectrum.
+In imaging, point sources are transformed to a single pixel, whereas for the delay-spectrum, smooth sources are transformed to a single channel (but then get forced into the wedge due to the inherent chromatic response of an interferometer).
+For imaging, we are concerned with ripples in the passband due to standing waves, whereas for the delay-spectrum we are concerned with delayed versions of the signal due to multiple reflections.
+
+Figure \ref{fig:kperf} is a rather busy plot showing many of the dependencies on wavenumber of bandwidth, configuration and the cosmos.
+It shows the perpendicular, parallel and total resolution wavenumber as a function of redshift for various bandwidths and baselines.
+The term {\em resolution} wavenumber is used, since it is a lower limit for that parameter since larger wavenumbers may be obtained by channelizing.
+The redshift range is appropriate for the extended frequency range that is the goal for HERA, however $z$=6-13 is the primary region for EOR studies.
+
+In Fig.
+\ref{fig:kperf}, blue lines denote $\kpar$ values, black lines denote $|\kvpr|$, and green lines denote total magnitude ($|{\kvec}|$).
+The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.
+Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.
+Also shown just below the blue $\kpar$ curves are redshift ranges for that bandwidth centered at $z$=6,12, and 18 (they appear as horizontal ``error'' bars).
+Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer really appropriate, so some evolutionary model would have to be assumed (this is why the line is dashed and only the $z$=12 range is shown).
+Also in Fig.
+\ref{fig:kperf}, the lime-green shaded region indicates a plausible region in which HERA could operate and the lighter square within that region is the area that is being targeted in the first instance.
+
+\begin{figure}[h!]
+\centerline{
+\includegraphics[width=.4\textwidth]{plots/kperf.png} 
+}
+\caption{\small $\kvec, \kpar, \kperp$ resolution wavenumbers with various bandwidths and baselines as a function of redshift.
+See text for discussion.
+The full redshift range is the HERA extended goal.
+The initial HERA delay-spectrum power spectrum goal is the light square region in the upper right.}
+\label{fig:kperf}
+\end{figure}
+
+\subsection{Map-Making Approach}
+\label{sec:mapapproach}
+In contrast to the delay-spectrum approach, map-making approaches combine baselines to build up information before squaring and calculating the power spectrum.
+The conceptually easiest approach is to first make an image cube of the sky as a function of angular position and frequency, and then to Fourier transform, square, and bin the cube to determine the power spectrum.
+One advantage of such an approach is that sky images can be interesting in their own right, both for dealing with measurement systematics and for accessing the non-Gaussian observational signatures (such as ionized bubble structures) that are missed by variance statistics such as the power spectrum.
+The principal disadvantage of a map-making approach is its stringent calibration and forward-modeling requirements for propagating models of bright foregrounds through instrument systematics.
+In contrast, the delay-spectrum allows one to stay considerably closer to the format of raw visibility measurements, thus allowing systematic effects to be sequestered to localized regions of parameter space where they can much more easily be filtered out.
+
+This approach is being pursued primarily via FHD.
+
+\subsection{Hybrid Approaches}
+\label{sec:hybridapproach}
+
+While they were presented separately above, the delay-spectrum and map-making approaches need not be viewed as mutually exclusive.
+For example, mapmaking need not be limited to the real space image basis; pipelines that combine multiple baselines into estimates of various Fourier amplitudes of the sky are formally also mapmaking pipelines.
+Like the delay-spectrum approach, such pipelines (see, e.g., \citealt{trott_et_al2016}) avoid the image domain entirely.
+This helps to prevent artifacts that may introduced by imaging algorithms.
+
+More generally, it is possible to express a map-making algorithm as a linear operator that acts on visibility data to produce a compressed dataset.
+Squaring this to obtain an estimate of the power spectrum then results in a weighted quadratic combination of visibilities, where one multiplies the visibility from every baseline with that from every other baseline.
+These visibility product pairs are subsequently normalized to form their own individual estimates of the power spectrum before all the normalized pairs are summed together to form a final power spectrum.
+Measuring a power spectrum in this way combines the best aspects of the mapmaking and delay-spectrum approaches.
+Power spectra are formed directly by the cross-multiplication of visibility data, preserving the delay-spectrum approach's strategy of staying close to quantities measured by an interferometer; on the other hand, by multiplying together the data from every possible combination of baselines, one retains all the information captured by mapmaking.
+Indeed, it can be shown \citep{liu_et_al2014a} that in the limit of infinitely fine Fourier space bins, the hybrid method that we have just described is mathematically equivalent to the mapmaking approach.
+In practice, the hybrid method comes with the added benefit of allowing particular baseline pairs that are suspected of being affected by instrumental systematics to be more surgically downweighted.
+
+\subsection{Calibration}
+\label{sec:calibration}
+As examined above, for maximize sensitivity to a limited number of spatial modes using the delay-spectrum technique, a redundant grid greatly is advantageous.
+This redundancy may also be exploited to calibrate the array as well, if the baselines are sufficiently redundant.
+This technique has been pioneered with the MIT EoR experiment, and has resulted in a package called omnical \citep{zheng_et_al2014}.
+
+\section{Existing Measurements}
 \label{sec:existing}
+As mentioned earlier, a number of programs are underway to make the first detection of the EoR power spectrum using the above techniques.  
 PAPER adopted a foreground mitigation strategy based largely
@@
 	\caption{The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, (solid symbols represent analyses led by HERA collaborators) compared to 21cmFAST-generated models at $k=0.2$\,$h$\,Mpc$^{-1}$.
-Analysis is still underway on PAPER and MWA observations that approach their projected 
-full sensitivities (see Fig.~\ref{fig:Sensitivities}); 
-HERA can deliver sub-$\text{mK}^2$ sensitivities.}
+Analysis is still underway on PAPER and MWA observations that approach their projected full sensitivities;  HERA can deliver sub-$\text{mK}^2$ sensitivities.}
 	\vspace{-10pt}
@@
 while retaining the necessary characteristics for both proven and forward-looking foreground removal strategies.
-
-\subsection{Secondary Scientific Objectives}%\\Precision Cosmology, Reionization Imaging, and High-$z$ IGM X-Ray Heating Science}
-
-\startsquarepar\emph{\textbf{Precision Cosmology.}}
-\label{sec:tau}
-By advancing our understanding of reionization astrophysics, HERA improves CMB constraints on 
-fundamental cosmological parameters by
-removing the optical depth
-$\tau$ as a nuisance parameter. HERA breaks the degeneracy between the
-constraints on $\tau$ and the sum of the neutrino masses $\sum m_\nu$, which has
-been identified as a potential problem for Stage 4 CMB lensing experiments. A
-HERA-informed estimate of $\tau$ enables CMB lensing experiments to achieve a
-$0.012\,\textrm{eV}$ error on $\sum m_\nu$ \citep{liu_et_al2015}. This would
-represent a $\sim$5$\sigma$ cosmological detection of the neutrino masses even
-under the most pessimistic assumptions
-% that their sum is the minimum value of $0.058\,\textrm{eV}$ 
-still allowed by neutrino oscillation \stopsquarepar
-
-\begin{figure}[h!]
-\centering
-    \includegraphics[width=0.50\textwidth,clip]{plots/sigmaTau.pdf}
-\caption{Likelihood contours ($68\%$ and $95\%$) for $\sigma_8$ and $\tau$ using \emph{Planck} constraints (blue) and 
-adding HERA data (red). 
-%Light and heavy contours signify $68\%$ and $95\%$ confidence regions, respectively. 
-The $21\,\textrm{cm}$ constraints break the CMB degeneracy between the amplitude of density fluctuations and the optical depth, improving constraints on both.}
-	\label{fig:sigma8Tau}
-\end{figure}
-
-\noindent experiments
-\citep{allison_et_al2015}, making HERA key to our understanding of neutrino
-physics. HERA's estimate of $\tau$ would also break the degeneracy between
-$\tau$ and the amplitude of matter fluctuations (expressed in Fig.
-\ref{fig:sigma8Tau} as $\sigma_8$) that arises when using only
-CMB data. HERA effectively reduces error bars on $\sigma_8$ by more
-than a factor of three \citep{liu_et_al2015}, potentially elucidating
-current tensions between cluster cosmology constraints and those from primary
-CMB anisotropies.
-
-\emph{\textbf{First Images of the Reionization Epoch.}}
-\label{sec:imaging}
-In addition to measuring the power spectrum, HERA has sufficient 
-sensitivity to directly image the IGM during reionization,
-and will observe a 1440 deg$^2$ stripe, comparable to future WFIRST
-large area near-IR surveys.
- After 100 hours on a
-single field (achievable in 200 nights), HERA reaches a surface brightness 
-sensitivity of 50 $\mu$Jy/beam (synthesized beam FWHM $\sim$ 24$'$) compared 
-to the brightness temperature fluctuations of up to 400 $\mu$Jy/beam in typical 
-reionization models (see Fig.~\ref{fig:LSS}). From the standpoint of sensitivity alone, HERA is capable of 
-detecting the brightest structures at $z=8$ with SNR $>$ 10. Additionally, the design 
-of HERA places it in a unique position to directly explore calibration techniques
-(e.g. redundant calibration, \citealt{liu_et_al2010}), while retaining a high quality 
-point spread functions for imaging and identifying foregrounds. Leveraging the expertise gained 
-through both precursor strategies, HERA will yield exceptionally high dynamic range
-images.
-
-%\subsubsection{The 21\,cm Signal in the Pre-Reionization Epoch}
-
-%\begin{wrapfigure}{R}{0.5\textwidth}
-\begin{figure}[b!]
-\centering
-\vspace{-15pt}
-    %\includegraphics[width=.5\textwidth,clip]{plots/imaging/HERA_z8_SNR_annotated_2015.png}
-    \includegraphics[width=1.0\textwidth,clip]{plots/HERA_FoV_w_strips.png}
-  \vspace{-20pt}
-\caption{\footnotesize 
-HERA will observe a 1440 deg$^2$ stripe centered near $\delta = -30^\circ$.
-HERA can measure the ionization state around galaxies in, e.g., the GOODS-South field that contains a third of all known $z\!>\!8$ galaxies.
-HERA's primary imaging data product to the community will be deep cubes along the HERA stripe suitable for cross-correlation.
-\label{fig:LSS} }
-\vspace{-10pt}
-\end{figure}
-%  \vspace{-10pt}
-%\end{wrapfigure}
-
-\emph{\textbf{Pre-Reionization Heating.}}
-\label{sec:EoX} 
-Prior to reionization, the 21\,cm signal is a sensitive probe of the first
-luminous sources and IGM heating mechanisms. First stars are expected to form
-at $z \sim 25-30$ and their imprint on the 21\,cm signal is expected to be
-sensitive to the halo mass where they are formed \citep{mesinger_et_al2015}.
-The IGM is then expected to be heated by first
-generation X--ray binaries
-\citep{furlanetto_et_al2006_global,pritchard_et_al2007,mesinger_et_al2013} or
-by the hot interstellar medium produced by the first supernovae
-\citep{pacucci_et_al2014}, although the heating timing and magnitude is still
-very much debated \citep{fialkov_et_al2012}. Dark matter annihilation
-\citep{evoli_et_al2014} could also inject energy in the IGM, leaving its
-imprint in the 21\,cm signal.
-
-With support from the Moore Foundation to develop feeds sensitive down to 50\,MHz, HERA will probe the IGM prior to
-reionization and constrain the sources of heating, obtaining, in the case of
-X--ray heating, percent-level constraints on the efficiency with which star-forming baryons
-produce X-rays \citep{ewall-wice_et_al2015}.
-While the observational and analytical state of the art of pre-reionization 21\,cm science is not as advanced as for the EoR \citep{ewall-wice_et_al2016-EoXLimits}, the risk of lower-frequency observing is borne by the Moore Foundation.
-Lower
-frequency observations also test feedback
-mechanisms that interact with low-mass halos at high redshifts
-\citep{Iliev_et_al2007,Iliev_et_al2012,ahn_et_al2012}.
-Such constraints, while interesting in their own right, also reduce
-the susceptibility of the aforementioned $21\,\textrm{cm}$-derived $\tau$
-constraints to uncertainties in high-redshift physics \citep{liu_et_al2015}, especially if they are 
-combined with upcoming or proposed measurements of 
-the pre-reionization sky-averaged spectrum \citep{fialkov_and_loeb2016}. Additionally, they may be crucial to a correct
-interpretation of kinetic Sunyaev-Zel'dovich effect constraints on reionization
-from the CMB \citep{park_et_al2013}. The high redshift probe of
-structure afforded by the low frequency $21\,\textrm{cm}$ measurements will
-permit some of the most direct observations of hypothesized suppressions of
-small-scale structure
-\citep{dalal_et_al2010,tseliakhovich_et_al2011,fialkov_et_al2012} arising from
-predicted supersonic relative velocities between dark matter and baryonic gas
-\citep{tseliakhovich_and_hirata2010}. No other electromagnetic probe can provide direct observations of this
-epoch.
-
-\subsection{Other Scientific Objectives}
-\label{subsec:broader_science}
-
-\noindent With unprecedented sensitivity at this frequency range, HERA can deliver much more than its core 21\,cm science. 
-Some of HERA's broader science impacts derive directly from the public data products HERA delivers.  Other
-areas of impact may require additional hardware or software that HERA can host at no extra risk. 
-In this respect, HERA serves as a ``platform" for additional science programs with external funding.
-The Moore Foundation's support of HERA feed development and data analysis targeting the pre-reionization science described in \S\ref{sec:EoX}
-is a prime example of how we envision this functioning.
-Below, we list examples of the broader science that HERA can deliver.
-
-\emph{\textbf{Cross-Correlations with Other Reionization Probes.}}
-\label{sec:crossCorr}
-HERA's public data provide new opportunities for cross-correlation studies.
-Cross-correlation between HERA 21\,cm images and other high-redshift probes
-(e.g. JWST; WFIRST; CMB maps; CO, CII, and Ly-$\alpha$ intensity mapping) 
-can provide an independent confirmation of the 21\,cm power spectrum
-\citep[i.e.][]{lidz_et_al2009,dore_et_al2014,silva_et_al2015,vrbanec2016} and enable rich new studies of
-the interaction between galaxies and their ionization environment. 
-In particular, cross-correlating 21\,cm with galaxy surveys can measure the characteristic bubble size around galaxies of
-different luminosities \citep{lidz_et_al2009} and help separate the degeneracy
-between the fraction of photons escaping the galaxies and the
-total number of ionizing photons produced \citep{zackrisson2013}.
-
-HERA is placed in a unique position to enable such cross-correlation: the GOODS-South field---one of the most panchromatically studied regions of the sky, the site of the Hubble UDF, and home to over a third of all known $z>9$
-galaxies---lies in its field of view. HERA's images of the IGM can provide environmental context to galaxy surveys through identification of
-ionized bubbles \citep{malloy_lidz2013}, or in a more statistical sense as
-described in \cite{beardsley_et_al2015}.
-
-\textbf{\emph{Searching for Exoplanetary Radio Bursts.}}
-\label{sec:exoplanets}
-HERA could be an powerful tool in the search for bright, highly polarized auroral bursts
-from exoplanets %due to the electron cyclotron maser instability 
-\citep{treumann2006,hallinan_et_al2015}. These
-distinctive, coherent bursts occur periodically at the planetary
-rotation rate, which cannot otherwise be measured. 
-A burst from a terrestrial planet could point toward
-habitability, since it implies a powerful magnetic field
-protecting the atmosphere and perhaps the biosphere from energetic stellar wind
-particles \citep{tarter_et_al2007}. HERA's sensitivity, large FoV, long campaigns, and precise calibration are all well-suited
-for discovering exoplanetary radio bursts from Jupiter-like planets out to 25~pc.
-Many well studied exoplanetary systems within that distance are in the HERA stripe, including Fomalhaut, Gl 667 C, Gl 433, HD 147513, and Gl 317.
-
-\textbf{\emph{Fast Radio Burst Followup.}}
-\label{sec:FRBs}
-Fast Radio Bursts (FRB) are millisecond-long radio flashes whose origin has remained a great enigma ever since their discovery \citep{2007Sci...318..777L}. 
-If equipped with a suitable backend, HERA could be triggered by nearby, higher-frequency telescopes for FRB followup, saving baseband data and thus full sensitivity to all dispersion measures.
-Bursts like that discovered by \citet{masui_et_al2015}, the lowest frequency
-FRB detection to date, should be seen hourly by HERA at 5--10$\sigma$. %(including the effects of scattering)
- Observations at HERA frequencies are very sensitive to the physics of
-the intervening medium, particularly deviations from $\lambda^2$
-dispersion. Detecting deviations would rule out broad classes of models and could indicate whether FRBs are at cosmological distances.
-
-\section{Measuring the EOR}
-\label{sec:eormeas}
-Due to the expansion of the Universe over cosmological time, we can identify and measure the early Universe via the red-shift of spectral lines.
-The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire Universe back almost to the period of recombination.
-The bandwidth
-therefore equates to a cosmic distance along the site of the telescope, with the frequency determining the cosmic age.
-An observation of an area of sky over a given bandwidth is therefore providing an average
-over a cosmic volume.
-
-As the distance to the EOR is great (about 13 billion light-years) the signal is quite weak.
-However, the signal also subtends the entire sky, so initial measurements of the EOR strive to measure a statistical power spectrum of the signal over the sky since the nature of the reionization process should have a specific spatial signature.
-The goal is therefore to measure a range of aggregate spatial scales on the sky, rather than to image the signal directly.
-Imaging does remain an ultimate goal to fully understand the process, however we will likely need a greater understanding of the signal characteristics and the systematics of our telescopes to achieve this more difficult goal.
-MANY PAPERS, see Nithya and Jonnie recent
-
-Obviously between our present observation point and the Epoch of Reionization lies the entire intervening Universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).
-Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.
-As a first step, areas of the sky where these signals are minimal (for example, outside the galactic plane and away from strong point sources) are targeted (REF FOR THIS).
-More importantly and most fortuitously however is the fact that all of these foreground signals are smooth spectrum sources whereas the expected spectrum of the EOR is expected to be rough since it is made up of non-ionized regions which are randomly distributed over a wide range of redshifts.
-This fact allows us to try and isolate foregrounds from the EOR, as will be discussed.
-
-This section will provide a brief overview of  the theoretical underpinning of how foregrounds contaminate the measurement (the so-called ``wedge''), discuss the various techniques used to make the measurement and finally provide a brief discussion on calibration issues and techniques.
-
-\subsection{The ``Wedge''}
-The past three years have seen deep EoR observations with PAPER, the MWA, 
-the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2013}), and LOFAR.
-PAPER and the MWA have produced progressively deeper limits
-\citep{dillon_et_al2015,parsons_etal2014,2015ApJ...809...61A}, with PAPER
-yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
-(Fig. \ref{fig:limits}).
-The inherent challenge is to simultaneously meet stringent sensitivity requirements 
-while suppressing foregrounds $\sim5$ orders of magnitude
-brighter than the 21\,cm signal, which has been addressed along a number of approaches.  
-To address this challenge,  all aspects of the experimental process must be refined and improved, spanning 
-calibration ({\em e.g.} \cite{zheng_et_al2014, jacobs_et_al2016, barry_et_al2016}), handling foreground contamination
-({\em e.g.} \cite{moore_et_al2013,moore_et_al2016,thyagarajan_et_al2015a,pober_et_al2016}),
-and the interferometer design itself ({\em e.g.} \cite{parsons_et_al2012a,dillon_parsons2016}).
-
-Perhaps the most important advance informing HERA's design is a
-refined understanding of how smooth-spectrum foregrounds interact
-with instrument chromaticity to produce a characteristic ``wedge" of
-foreground leakage in Fourier space (see Fig.~\ref{fig:wedge}), 
-outside of which the 21\,cm signal dominates (the ``EoR window'').  The coordinates of the wedge are discussed below.
-Through theoretical and observational work
-\citep{datta_etal2010,morales_et_al2012,parsons_et_al2012b,vedantham_2012,thyagarajan_et_al2013,hazelton_et_al2013,pober_etal2013b,liu_et_al2014a,liu_et_al2014b},
-we have learned how the boundary between the wedge and  the EoR window is determined by the separation between antennas,
-signal reflections within antennas, and the angular response of the antenna beam.  Deep integrations also show us
-that, to the limits of current sensitivity, foreground emission is absent outside of the wedge; it can only 
-appear there through instrumental leakage \citep{parsons_etal2014,2015,ApJ...809...61A,moore_et_al2016,kohn_et_al2016}.
-
-Though the power spectrum measurement is over a volume, it is instructive to split the spatial wavenumbers into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kvpr$ is determined by the antenna baseline and $\kpar$ by the bandwidth.
-This is useful since it allows us to split out the chromatic response of the interferometer and isolate a phase space where foreground sources ({\em i.e.} everything {\em not} the EOR) contaminate the signal of interest from where they don't.
-As mentioned above, this contaminated phase space may be shown [REF] to be a wedge-shaped region in $\kpr$:$\kpar$ space:
-\begin{equation}
-|\kpar| \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
-\end{equation}
-where $c$ is the speed of light and $S$ is a parameter accounting for offsets related to the combined spectral smoothness of the foregrounds and the antenna response (see Fig. \ref{fig:wedge}).  
-A key issue to measuring the EOR power spectrum is to understand and minimize such effects.
-For an analytical derivation, see \cite{zahn_etal2012,vedantham_2012,liu_et_al2014b},  in simulation see \cite{datta_etal2010,hazelton_et_al2013} and for observations
-\cite{pober_etal2013b,2015arXiv150601026P,parsons_etal2014,2015arXiv150206016A}.
-Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).
-In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_o=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.
-
-\begin{figure}[t]
-\centerline{
-\includegraphics[height=3in]{plots/wedgeWithBG.png}
-}
-\caption{Foregrounds are a primary challenge facing 21\,cm cosmology experiments. 
-HERA leverages a ``wedge'' in the spherically-averaged ($\mathbf{k}$ is broken into $k_\|$ and $k_\perp$) power spectrum (center panel; \citealt{dillon_et_al2015}).
-Smooth-spectrum foregrounds (right panel) $\sim$5 orders of magnitude brighter than fiducial EoR models (left panel; \citealt{mesinger_et_al2011}) create the ``wedge'' when they interact with the interferometer's chromatic response.
-By avoiding foregrounds, PAPER has placed limits within an order
-of magnitude (in mK) of these models \citep{2015ApJ...809...61A} and show the ``EoR Window'' to be foreground-free.
-HERA's dish and configuration optimize wedge/window isolation and direct sensitivity to low-$k_\perp$ modes where EoR is brightest.
-%v2
-Cartoon of the wedge.
-The gray rectangle represents the smooth-spectrum foregrounds as they would be if the instrument had a perfect achromatic response.
-The pink wedge area shows the chromatic response of the interferometer, where additional power is thrown up from the gray region to higher k-values with increasing baseline.
-The additional horizontal pink line illustrates the affect that other systematics, such as reflections on cables, can have in contaminating the measurement.
-The white region shown the 'eor window', where the measurement is not contaminated by foregrounds or systematics.}
-\label{fig:wedge}
-\end{figure}
-
-The techniques to measure the power spectrum may be broken down into two principle techniques (delay-space and map-making), plus additional hydrid methods.
-These are briefly summarized below, with an emphasis on the delay-spectrum technique as the initial approach taken by HERA.
-Finally, we provide a short discussion on calibration in the concept of this design.
-
-\subsection{Delay-Spectrum Approach}
-\label{sec:delayapproach}
-Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement of the EOR spatial power spectrum.
-There are now numerous papers in the literature on the technique of measuring the power spectrum using an interferometer and the attendant issues [[[list of citations]]].
-Below, a high-level overview based on  \cite{parsons_et_al2012b} will be presented.
-The so-called delay-spectrum approach leverages the interferometer measurement to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
-Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum, which contrasts to the map-making techniques briefly described in the following section.
-
-We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment:
-\begin{equation}
-{\hat P}(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} \right]
-\label{eq:Pk_baseline}
-\end{equation}
-where $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes (so, relating wavenumber to physical volume at a given redshift), $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.
-The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.
-Note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
-Rather than ${\hat P}(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.
-
-Since the thermal noise per visibility baseline may be expressed as
-\begin{equation}
-V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
-\label{eq:sensitivity_per_baseline}
-\end{equation}
-where $\tau$ is the integration time (see {\em e.g.} TMS),
-to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq.
-\ref{eq:Pk_baseline}.
-Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
-\begin{equation}
-\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
-\label{eq:sensitivity}
-\end{equation}
-where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement 
-(see \cite{parsons_et_al2012b}).
-Fig.
-\ref{fig:boost} plots this ``boost'' factor for various telescopes under nominal observing campaigns using the delay-spectrum technique.
-The large improvement for PAPER and HERA is the fact that the baselines are redundantly placed on appropriate spatial scales, so that a more limited number of unique baselines get sampled many times.
-
-\begin{figure}[h!]
-\centerline{
-\includegraphics[width=0.5\textwidth]{plots/boost.png}
-}
-\caption{Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) for representative observing campaigns for various instruments.
-ADD NEW HERA CONFIGURATION AND LOFAR AND PUT IN 'HOW TO MAKE THIS PLOT'}
-\label{fig:boost}
-\end{figure}
-
-The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by both the antenna baseline and bandwidth, as discussed above.
-The $\kvpr$ components are proportional to the visibilities and $\kpar$ are proportional to the Fourier transform of the frequency response.
-The Fourier transform of a frequency spectrum is a delay, hence the name ``delay-spectrum'' for this technique.
-Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.
-For HERA, wavenumbers are dominated by the bandwidth, not the baseline.
-
-As an aside, note the parallels to imaging with an interferometer.
-For imaging, one takes a 2-D spatial Fourier transform of the visibilities, and produces images as a function of frequency.
-Here, one is taking a 1-D Fourier transform along the other axis (the bandpass) and produces a 3-D spatial power spectrum.
-In imaging, point sources are transformed to a single pixel, whereas for the delay-spectrum, smooth sources are transformed to a single channel (but then get forced into the wedge due to the inherent chromatic response of an interferometer).
-For imaging, we are concerned with ripples in the passband due to standing waves, whereas for the delay-spectrum we are concerned with delayed versions of the signal due to multiple reflections.
-
-Figure \ref{fig:kperf} is a rather busy plot showing many of the dependencies on wavenumber of bandwidth, configuration and the cosmos.
-It shows the perpendicular, parallel and total resolution wavenumber as a function of redshift for various bandwidths and baselines.
-The term {\em resolution} wavenumber is used, since it is a lower limit for that parameter since larger wavenumbers may be obtained by channelizing.
-The redshift range is appropriate for the extended frequency range that is the goal for HERA, however $z$=6-13 is the primary region for EOR studies.
-
-In Fig.
-\ref{fig:kperf}, blue lines denote $\kpar$ values, black lines denote $|\kvpr|$, and green lines denote total magnitude ($|{\kvec}|$).
-The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.
-Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.
-Also shown just below the blue $\kpar$ curves are redshift ranges for that bandwidth centered at $z$=6,12, and 18 (they appear as horizontal ``error'' bars).
-Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer really appropriate, so some evolutionary model would have to be assumed (this is why the line is dashed and only the $z$=12 range is shown).
-Also in Fig.
-\ref{fig:kperf}, the lime-green shaded region indicates a plausible region in which HERA could operate and the lighter square within that region is the area that is being targeted in the first instance.
-
-\begin{figure}[h!]
-\centerline{
-\includegraphics[width=.4\textwidth]{plots/kperf.png} 
-}
-\caption{\small $\kvec, \kpar, \kperp$ resolution wavenumbers with various bandwidths and baselines as a function of redshift.
-See text for discussion.
-The full redshift range is the HERA extended goal.
-The initial HERA delay-spectrum power spectrum goal is the light square region in the upper right.}
-\label{fig:kperf}
-\end{figure}
-
-\subsection{Map-Making Approach}
-\label{sec:mapapproach}
-In contrast to the delay-spectrum approach, map-making approaches combine baselines to build up information before squaring and calculating the power spectrum.
-The conceptually easiest approach is to first make an image cube of the sky as a function of angular position and frequency, and then to Fourier transform, square, and bin the cube to determine the power spectrum.
-One advantage of such an approach is that sky images can be interesting in their own right, both for dealing with measurement systematics and for accessing the non-Gaussian observational signatures (such as ionized bubble structures) that are missed by variance statistics such as the power spectrum.
-The principal disadvantage of a map-making approach is its stringent calibration and forward-modeling requirements for propagating models of bright foregrounds through instrument systematics.
-In contrast, the delay-spectrum allows one to stay considerably closer to the format of raw visibility measurements, thus allowing systematic effects to be sequestered to localized regions of parameter space where they can much more easily be filtered out.
-
-This approach is being pursued primarily via FHD.
-
-\subsection{Hybrid Approaches}
-\label{sec:hybridapproach}
-
-While they were presented separately above, the delay-spectrum and map-making approaches need not be viewed as mutually exclusive.
-For example, mapmaking need not be limited to the real space image basis; pipelines that combine multiple baselines into estimates of various Fourier amplitudes of the sky are formally also mapmaking pipelines.
-Like the delay-spectrum approach, such pipelines (see, e.g., \citealt{trott_et_al2016}) avoid the image domain entirely.
-This helps to prevent artifacts that may introduced by imaging algorithms.
-
-More generally, it is possible to express a map-making algorithm as a linear operator that acts on visibility data to produce a compressed dataset.
-Squaring this to obtain an estimate of the power spectrum then results in a weighted quadratic combination of visibilities, where one multiplies the visibility from every baseline with that from every other baseline.
-These visibility product pairs are subsequently normalized to form their own individual estimates of the power spectrum before all the normalized pairs are summed together to form a final power spectrum.
-Measuring a power spectrum in this way combines the best aspects of the mapmaking and delay-spectrum approaches.
-Power spectra are formed directly by the cross-multiplication of visibility data, preserving the delay-spectrum approach's strategy of staying close to quantities measured by an interferometer; on the other hand, by multiplying together the data from every possible combination of baselines, one retains all the information captured by mapmaking.
-Indeed, it can be shown \citep{liu_et_al2014a} that in the limit of infinitely fine Fourier space bins, the hybrid method that we have just described is mathematically equivalent to the mapmaking approach.
-In practice, the hybrid method comes with the added benefit of allowing particular baseline pairs that are suspected of being affected by instrumental systematics to be more surgically downweighted.
-
-\subsection{Calibration}
-\label{sec:calibration}
 

### 2016-04-07 manuscript.tex
update to sensitivity figure

@@
 Fig.
-\ref{fig:boost} plots this ``boost'' factor for various telescopes under nominal observing campaigns using the delay-spectrum technique.
-The large improvement for PAPER and HERA is the fact that the baselines are redundantly placed on appropriate spatial scales, so that a more limited number of unique baselines get sampled many times.
+\ref{fig:boost} plots this ``boost'' factor for various configuration types relative to an array in a 19-element hexagonal configuration.
+They are labelled for the type of configuration (hex, grid, imaging) and the number of antennas (from 37 to 350).
+Note that the hex is based on the configuration used for HERA, grid for that of PAPER, Imaging-128 for MWA and Imaging-48 for LOFAR.
+The plot shows that for a fixed number using the delay-spectrum approach, redundant arrays (hex/grid) provide about an order of magnitude improvement over imaging arrays.
+Since the baselines go as order $N^2$, that corresponds to that for about 1/3 the number of elements you get the same performance (as verified by the Hex-37/Imaging-128 lines).
+This advantage does not extend to the other approaches discussed below.
+
+Another approach to improving sensitivity is to make the collecting area per element larger.
 
@@
 \centerline{
-\includegraphics[width=0.5\textwidth]{plots/boost.png}
+\includegraphics[width=0.5\textwidth]{plots/boost.pdf}
+\includegraphics[width=0.5\textwidth]{plots/sensratio.pdf}
 }
-\caption{Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) for representative observing campaigns for various instruments.
-ADD NEW HERA CONFIGURATION AND LOFAR AND PUT IN 'HOW TO MAKE THIS PLOT'}
+\caption{Left: Representative boost factors ($\mathcal{N}$ in Eq.
+\ref{eq:sensitivity}) normalized to a 19-element hexagonal configuration spaced at 14.6m.
+For a given element count, one can get about an order of
+magnitude improvement in delay-spectrum sensitivity.
+Alternatively, for the same sensitivity one can build roughly 1/3 the antennas.
+Right:  Power spectrum sensitivities for various arrays normalized to the
+19-element hex.  Here the element size is included, which is shown as an appropriately sized circle.}
 \label{fig:boost}

### 2016-04-07 plots/boost.py

@@ -1,0 +1,76 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+h127=np.load('h127drift_mod_0.158.npz')
+h19=np.load('h19.drift_mod_0.158.npz')
+h350=np.load('h350drift_mod_0.158.npz')
+h37=np.load('h37drift_mod_0.158.npz')
+lofar=np.load('lofar_coredrift_mod_0.158.npz')
+mwa=np.load('mwa128drift_mod_0.158.npz')
+paper=np.load('paper128drift_mod_0.158.npz')
+h331=np.load('h331drift_mod_0.158.npz')
+
+D_lofar = 30.75
+D_hera = 14.0
+D_mwa = 2.32*2.
+D_paper = 1.18*2.
+
+#plt.figure('T_errs')
+#plt.semilogy(h127['ks'],h127['T_errs'],label='h127')
+#plt.semilogy(h19['ks'],h19['T_errs'],label='h19')
+#plt.semilogy(h37['ks'],h37['T_errs'],label='h37')
+#plt.semilogy(h350['ks'],h350['T_errs'],label='h350')
+#plt.semilogy(lofar['ks'],lofar['T_errs'],label='lofar')
+#plt.semilogy(mwa['ks'],mwa['T_errs'],label='mwa')
+#plt.semilogy(paper['ks'],paper['T_errs'],label='paper')
+
+def plotPwr(p=2,title='-',ylabel='-'):
+    N37_19    = (D_hera/D_hera)**p*(h19['T_errs']/h37['T_errs'])
+    N127_19   = (D_hera/D_hera)**p*(h19['T_errs']/h127['T_errs'])
+    N331_19   = (D_hera/D_hera)**p*(h19['T_errs']/h331['T_errs'])
+    N350_19   = (D_hera/D_hera)**p*(h19['T_errs']/h350['T_errs'])
+    Nmwa_19   = (D_hera/D_mwa)**p*(h19['T_errs']/mwa['T_errs'])
+    Nlofar_19 = (D_hera/D_lofar)**p*(h19['T_errs']/lofar['T_errs'])
+    Npaper_19 = (D_hera/D_paper)**p*(h19['T_errs']/paper['T_errs'])
+    
+    clr = ['b','r','g','k','c','m','g']
+    yshift = [1.35,0.8,0.8,0.76,1.12,1.17,0.72]
+    if p==2:
[…]

### 2016-04-08 manuscript.tex
much less drafty.  Generally ok through sec 3 (but needs cal and map-making input).  Working through 4+

@@
 understanding the astrophysics and fundamental cosmology of our early universe.
-Taking advantage of our new mastery of bright foreground systematics, HERA's
+Taking advantage of our new understanding of bright foreground systematics, HERA's
 purpose-built radio
@@
 
-HERA brings to bear both the sensitivity and the precision to deliver 
-its key science results on the basis of proven techniques.
+HERA brings to bear both the sensitivity and the precision to deliver its key science results on the basis of proven techniques.
 These science results include
@@
 This has resulted in a close-packed array of transit 14-meter dishes (Fig.
-\ref{fig:HERApictures}), with the first 19-element array completed and construction underway for 37 elements, when the first science observing will begin.
-The first phase, extending to August 2016, will deploy 37 elements that will test the HERA element design for performance and manufacturability at location.  These 37 elements will more 
-than double the sensitivity of the 128-element PAPER array.  The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal 
-paths along with , which should provide a very robust detection of the EOR signal.  Finally, extending into 2020, HERA will build out to 350 elements to further EOR science as a function 
+\ref{fig:HERApictures}), with the first 19-element array completed and construction underway for 37 elements with additional funding from Cambridge University.
+
+The current 19 elements are testing the HERA element design for performance and manufacturability at location.
+The 37 elements, to be completed in 2016, will provide nearly an order of magnitude more
+sensitivity than the 128-element PAPER array and provide a significant chance at detecting the EoR power spectrum signal.
+The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER \citep{ali_et_al2015}, which should provide a very robust detection of the EOR signal and allow some characterization.
+Finally, extending to 2020, HERA will build out to 350 elements to further EOR science as a function
 of redshift and spatial scale, potentially producing the first images of the EOR.
@@
 Section \ref{sec:eormeas} will present the techniques used to make the measurement, focussing on the delay-spectrum approach.
-Section \ref{sec:requirements} will present the resulting high-level requirements and section will present the system description \ref{sec:design} as the bulk of the paper.
-Section \ref{sec:status} will provide a status and section \ref{sec:conclusion} will conclude by summarizing and providing additional context.
+Section \ref{sec:existing} will provide a brief summary of the existing measured limits.
+Section \ref{sec:requirements} will present the high-level requirements and Section \ref{sec:design} will present the system description..
+Section \ref{sec:status} will present a brief status and outline the deployement timetable and Section \ref{sec:conclusion} will conclude by summarizing and providing additional context.
 
@@
 \noindent The {\it Cosmic Dawn} of our universe is one of the last unexplored
-frontiers in cosmic history, which is summarized in the cartoon in Fig. \ref{fig:cosmos}. During this period, the 
-intergalactic medium (IGM) encodes a panoply of information about astrophysical and cosmological phenomena. 
-This period saw the evolution of the universe from a very homogeneous ``gas'' to the formation of the first stars and galaxies.
+frontiers in cosmic history, which is summarized in Fig.
+\ref{fig:cosmos}.
+On the left is the Big Bang, then the hot, young universe slowly cools until it begins to gravitationally contract to the first structures that relit the universe.
+The Cosmic Dawn saw the evolution of the universe from a very homogeneous ``gas'' to the formation of the first stars and galaxies en route to the structure of galaxies
+of our current time.
+During this period, encoded in the structure of the 
+intergalactic medium (IGM) lies a panoply of information about the underlying astrophysical and cosmological phenomena governing cosmic evolution.
 Ultimately this early population of gravitationally condensed material produced sufficiently energetic flux to reionize the IGM
-from its previous neutral state in a period called the Epoch of Reionization
-Its evolution depends on the cosmic density field, the relative velocities of
+from its previous neutral state in a period called the Epoch of Reionization.
+This period is shown in Fig \ref{fig:cosmos} as the rapid transition from the big bubble structures to structures that begin to resemble the denizens of our current universe.
+
+The evolution of these structures depends on the cosmic density field, the relative velocities of
 baryons and dark matter, and the sizes and clustering of the first galaxies.
-But it also depends on the constituents of those galaxies---so-called Population
+But it also depends on the constituents of those first galaxies---so-called Population
 III stars (stars formed very early with little to no elements heavier than helium), later generation stars, stellar remnants, X-ray binaries,
@@
 culminating in the Epoch of Reionization, 
-led the most recent US National Academies astronomy decadal survey entitled {\em New Worlds, New Horizons} (NWNH) to highlight it as one of the top three ``priority science objectives'' for
-the decade.
+led the most recent US National Academies astronomy decadal survey entitled {\em New Worlds, New Horizons} to highlight it as one of the top three ``priority science objectives'' for
+the decade \citep{NWNH}.
 
@@
 	\includegraphics[width=0.7\textwidth]{plots/cosmicEvo.png}
-	\caption{Rendering of cosmic evolution from just after the Big Bang to today (credit CCCCC).
-The blue line shows the frequency of the red-shifted hydrogen line (rest frequency 1420 MHz) as a function of redshift (lower horizontal axis) and cosmic time (upper horizontal axis).}
+	\caption{Rendering of cosmic evolution from just after the Big Bang to today (background image credit Loeb/{\em Scientific American}).
+The blue line shows the frequency of the red-shifted hydrogen line (rest frequency 1420 MHz) as a function of redshift (lower horizontal axis) and cosmic time (upper horizontal axis).
+The white dashed horizontal lines are the HERA EoR band and the blue are the extended frequency goal.
+Determining where the structure in the background really falls is the
+goal of HERA.} 
 	\label{fig:cosmos}
@@
 Exploring the interplay of galaxies and large-scale structure during the EOR
-requires complementary observational approaches. CMB measurements 
-provide initial conditions for structure formation, and the Thomson
+requires complementary observational approaches. Measurements of the Cosmic Microwave Background (CMB; the escaping radiation permeating the universe after the universe became
+transparent to its radiation by the recombination of the protons and electrons about 400,000 years after the Big Bang)  provide initial conditions for structure formation, and the Thomson
 scattering of CMB photons constrains the integrated column of 
@@
 schenker_et_al2012, treu_et_al2012, dijkstra_et_al2014} is uncertain and
-highly model-dependent.
+highly model-dependent.  See Fig.~\ref{fig:IonHist} for these constraints on the hydrogen neutral fraction as a function of redshift.
+As shown, existing probes are limited in
+their ability to constrain reionization, and will be for the foreseeable future.
+HERA will use
+another complementary probe --- the 21\,cm ``spin-flip" transition of neutral hydrogen
+--- to bring transformative new capabilities in this area.  The next sub-sections outline these goals.
 
@@
 \end{figure}
-
-As shown in Figure~\ref{fig:IonHist},
-existing probes are limited in
-their ability to constrain reionization, and will be for the foreseeable future.
-HERA will use
-another complementary probe --- the 21\,cm ``spin-flip" transition of neutral hydrogen
---- to bring transformative new capabilities in this area.  The next sub-sections outline these goals.
 
@@
 the Murchison Widefield Array (MWA; \citealt{bowman_et_al2012,tingay_et_al2013}),
-the MIT EoR experiment (MITEoR; \citealt{zheng_et_al2014}) and the Experiment to Detect the Global EoR Step (EDGES; \citealt{bowman_rogers2010}).  
-Efforts by members of the HERA team
+the MIT EoR experiment (MITEoR; \citealt{zheng_et_al2014}) and the Experiment to Detect the Global EoR Step (EDGES; \citealt{bowman_rogers2010}) (see Sec. \ref{sec:existing}).
+Recent measurements
 have produced the first astrophysically constraining upper limits on the 21\,cm EoR power spectrum, 
-providing evidence for significant heating in the IGM prior to reionization \citep{parsons_etal2014, 2015ApJ...809...61A,pober_et_al2015}.
+providing evidence for significant heating in the IGM prior to reionization \citep{parsons_etal2014, parsons_et_al2015, ali_et_al2015,pober_et_al2015}.
 However, current experiments cannot expect more than marginal detections of the EoR signal. 
@@
 \caption{\hspace{-1.2mm}: Predicted SNRs of 21\,cm experiments for an EoR model with 50\% ionization at $z=9.5$, with 1080 hours observation, integrated over a $\Delta z$ of $0.8$.
-Foreground avoidance represents an analysis comparable to \cite{2015ApJ...809...61A}, whereas foreground modeling allows signifcantly more $k$ modes of the cosmological signal to be recovered.}
+Foreground avoidance represents an analysis comparable to \cite{ali_et_al2015}, whereas foreground modeling allows signifcantly more $k$ modes of the cosmological signal to be recovered.}
 \small
@@
 
-\startsquarepar\emph{\textbf{Precision Cosmology.}}
+\emph{\textbf{Precision Cosmology.}}
 \label{sec:tau}
-By advancing our understanding of reionization astrophysics, HERA improves CMB constraints on 
+By advancing our understanding of reionization astrophysics, HERA will improve CMB constraints on 
 fundamental cosmological parameters by
 removing the optical depth
-$\tau$ as a nuisance parameter. HERA breaks the degeneracy between the
+$\tau$ as a ``nuisance'' parameter. HERA breaks the degeneracy between the
 constraints on $\tau$ and the sum of the neutrino masses $\sum m_\nu$, which has
@@
 % that their sum is the minimum value of $0.058\,\textrm{eV}$ 
-still allowed by neutrino oscillation \stopsquarepar
+still allowed by neutrino oscillation experiments
+\citep{allison_et_al2015}, making HERA key to our understanding of neutrino
+physics. HERA's estimate of $\tau$ would also break the degeneracy between
+$\tau$ and the amplitude of matter fluctuations (expressed in Fig.
+\ref{fig:sigma8Tau} as $\sigma_8$) that arises when using only
+CMB data. HERA effectively reduces error bars on $\sigma_8$ by more
+than a factor of three \citep{liu_et_al2015}, potentially elucidating
+current tensions between cluster cosmology constraints and those from primary
+CMB anisotropies.
 
@@
 
-\noindent experiments
-\citep{allison_et_al2015}, making HERA key to our understanding of neutrino
-physics. HERA's estimate of $\tau$ would also break the degeneracy between
-$\tau$ and the amplitude of matter fluctuations (expressed in Fig.
-\ref{fig:sigma8Tau} as $\sigma_8$) that arises when using only
-CMB data. HERA effectively reduces error bars on $\sigma_8$ by more
-than a factor of three \citep{liu_et_al2015}, potentially elucidating
-current tensions between cluster cosmology constraints and those from primary
-CMB anisotropies.
-
 \emph{\textbf{First Images of the Reionization Epoch.}}
 \label{sec:imaging}
-In addition to measuring the power spectrum, HERA has sufficient 
+In addition to measuring the power spectrum, if systematics are kept sufficiently low HERA has sufficient 
 sensitivity to directly image the IGM during reionization,
@@
 of HERA places it in a unique position to directly explore calibration techniques
-(e.g. redundant calibration, \citealt{liu_et_al2010}), while retaining a high quality 
+(e.g. redundant calibration, \citealt{liu_et_al2010,zheng_et_al2014}), while retaining a high quality 
 point spread functions for imaging and identifying foregrounds. Leveraging the expertise gained 
@@
 
-With support from the Moore Foundation to develop feeds sensitive down to 50\,MHz, HERA will probe the IGM prior to
-reionization and constrain the sources of heating, obtaining, in the case of
+While the observational and analytical state of the art of pre-reionization 21\,cm science is not as advanced as for the EoR \citep{ewall-wice_et_al2016-EoXLimits}, with support from the Moore Foundation to develop feeds sensitive down to 50\,MHz, HERA will explore the IGM prior to
+reionization with a goal to constrain the sources of heating, obtaining, in the case of
 X--ray heating, percent-level constraints on the efficiency with which star-forming baryons
-produce X-rays \citep{ewall-wice_et_al2015}.
-While the observational and analytical state of the art of pre-reionization 21\,cm science is not as advanced as for the EoR \citep{ewall-wice_et_al2016-EoXLimits}, the risk of lower-frequency observing is borne by the Moore Foundation.
-Lower
+produce X-rays \citep{ewall-wice_et_al2015}. Lower
 frequency observations also test feedback
@@
 
-\noindent With unprecedented sensitivity at this frequency range, HERA can deliver much more than its core 21\,cm science. 
-Some of HERA's broader science impacts derive directly from the public data products HERA delivers.  Other
-areas of impact may require additional hardware or software that HERA can host at no extra risk. 
-In this respect, HERA serves as a ``platform" for additional science programs with external funding.
+\noindent Given its high sensitivity, frequency support, filled aperture and outriggers, HERA is capable of delivering more than the core 21\,cm science discussed above.
+HERA will make the data available to the community after only a short quality assurance time period to help facilitate low-frequency science.
+Additionally, we encourage groups to leverage HERA as a platform upon which other equipment could be deployed. 
 The Moore Foundation's support of HERA feed development and data analysis targeting the pre-reionization science described in \S\ref{sec:EoX}
-is a prime example of how we envision this functioning.
-Below, we list examples of the broader science that HERA can deliver.
+is a prime example of how we envision this functioning.  Transients also provide an avenue for promising research.
+Below, we list examples of the broader science that HERA could impact.
 
@@
 
-HERA is placed in a unique position to enable such cross-correlation: the GOODS-South field---one of the most panchromatically studied regions of the sky, the site of the Hubble UDF, and home to over a third of all known $z>9$
-galaxies---lies in its field of view. HERA's images of the IGM can provide environmental context to galaxy surveys through identification of
+Fortuitiously, the GOODS-South field---one of the most panchromatically studied regions of the sky, the site of the Hubble UDF, and home to over a third of all known $z>9$
+galaxies---lies in HERA's field of view. HERA's images of the IGM can provide environmental context to galaxy surveys through identification of
 ionized bubbles \citep{malloy_lidz2013}, or in a more statistical sense as
@@
 \label{sec:exoplanets}
-HERA could be an powerful tool in the search for bright, highly polarized auroral bursts
+HERA could be a powerful tool in the search for bright, highly polarized auroral bursts
 from exoplanets %due to the electron cyclotron maser instability 
@@
 
-This section will provide a brief overview of  the theoretical underpinning of how foregrounds contaminate the measurement (the so-called ``wedge''), discuss the various techniques used to make the measurement and finally provide a brief discussion on calibration issues and techniques.
-
-\subsection{The ``Wedge''}
 The past three years have seen deep EoR observations with PAPER, the MWA, 
@@
 PAPER and the MWA have produced progressively deeper limits
-\citep{dillon_et_al2015,parsons_etal2014,2015ApJ...809...61A}, with PAPER
+\citep{dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
 yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
@@
 
+As mentioned, a number of programs are underway to make the first detection of the EoR power spectrum using a number of approaches.  
+PAPER adopted a foreground mitigation strategy based largely
+on identifying and filtering out a region of phase space where the strong foregrounds are largely confined (Sec. \ref{sec:delayapproach}).  
+By reducing the need
+for foreground modeling and subtraction, this approach allowed PAPER to switch to a grid-based antenna layout that
+enhanced sensitivity toward fewer Fourier modes and facilitated calibration
+based on the ``redundancy" of different antenna pairs measuring the same sky modes \citep{liu_et_al2010,zheng_et_al2014}.
+Combined, redundancy and delay filtering provide a robust, inexpensive, and demonstrably successful solution 
+to the foreground problem.
+
+Yet PAPER's lack of imaging support and its uneven $uv$ sampling 
+leave it with limited diagnostic capability, particularly for direction-dependent systematics
+such as polarization leakage from Faraday-rotated emission \Mycitep{moore_et_al2013}.
+While concern over such effects has decreased markedly since discovering that
+intrinsic point source polarization is lower than previously thought
+\citep{asad_et_al2015} and that variable Faraday rotation through the
+ionosphere averages down the polarized signal over long observing seasons
+\Mycitep{moore_et_al2016}, direction-dependent beam effects remain an area of interest.
+The MWA's image-based calibration and foreground subtraction strategy provides complementary
+capabilities (Sec. \ref{sec:mapapproach}). Imaging with subtraction, while still under development as a viable foreground strategy,
+can increase sensitivity by recovering more modes of the cosmological signal (see
+Table~\ref{tab:signif}) and help address systematic errors rooted
+in the image domain.  HERA's antenna configuration---shown in Figure \ref{fig:arrayConfig} and discussed in \S\ref{sec:arrayConfig}---emphasizes
+the proven approaches of redundant calibration and delay filtering, while simultaneously
+increasing the extent and density of $uv$ sampling for high-fidelity imaging.
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=.8\textwidth]{plots/current_limits.pdf}
+	\caption{The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, (solid symbols represent analyses led by HERA collaborators) compared to 21cmFAST-generated models at $k=0.2$\,$h$\,Mpc$^{-1}$.
+Analysis is still underway on PAPER and MWA observations that approach their projected full sensitivities;  HERA can deliver sub-$\text{mK}^2$ sensitivities.}
+	\vspace{-10pt}
+	\label{fig:limits}
+	\label{fig:IGMtemperatureConstraints}
+\end{figure}
+
+Despite progress, the fact remains that the 21\,cm EoR signal is intrinsically very faint; making a detection requires a
+large instrumental collecting area and a long, dedicated observing campaign.
+Although published PAPER and MWA results (Fig. \ref{fig:limits}) do not yet include observations 
+at full sensitivity that are still being analyzed (e.g. Fig. \ref{fig:Sensitivities}), 
+it is clear already that these instruments lack the sensitivity
+to make a conclusive detection (see Table~\ref{tab:signif}). 
+HERA addresses this shortcoming with a dish element that delivers a much larger collecting area 
+while retaining the necessary characteristics for both proven and forward-looking foreground removal strategies.
+
+This section will provide a brief overview of  the theoretical underpinning of how foregrounds contaminate the measurement (the so-called ``wedge''), discuss the various techniques used to make the measurement and finally provide a brief discussion on calibration issues and techniques.
+
+\subsection{The ``Wedge''}
 Perhaps the most important advance informing HERA's design is a
@@
 that, to the limits of current sensitivity, foreground emission is absent outside of the wedge; it can only 
-appear there through instrumental leakage \citep{parsons_etal2014,2015,ApJ...809...61A,moore_et_al2016,kohn_et_al2016}.
+appear there through instrumental leakage \citep{parsons_etal2014,ali_et_al2015,moore_et_al2016,kohn_et_al2016}.
 
@@
 By avoiding foregrounds, PAPER has placed limits within an order
-of magnitude (in mK) of these models \citep{2015ApJ...809...61A} and show the ``EoR Window'' to be foreground-free.
+of magnitude (in mK) of these models \citep{ali_et_al2015} and show the ``EoR Window'' to be foreground-free.
 HERA's dish and configuration optimize wedge/window isolation and direct sensitivity to low-$k_\perp$ modes where EoR is brightest.
@@
 (see \cite{parsons_et_al2012b}).
+
 Fig.
-\ref{fig:boost} plots this ``boost'' factor for various configuration types relative to an array in a 19-element hexagonal configuration.
+\ref{fig:boost} (left) plots this ``redundancy boost'' factor for various configuration types relative to an array in a 19-element hexagonal configuration.
 They are labelled for the type of configuration (hex, grid, imaging) and the number of antennas (from 37 to 350).
-Note that the hex is based on the configuration used for HERA, grid for that of PAPER, Imaging-128 for MWA and Imaging-48 for LOFAR.
+Note that the (a) hex is based on the configuration used for HERA (14.6m center-to-center spacing); (b) grid for that of PAPER (4m N-S, 16m E-W spacing); (c) Imaging-128 for MWA (\cite{mwa}); and (d) Imaging-48 for LOFAR (\cite{2013A&A...556A...2V}; note that this assumes the ability to fully correlate the HBA-pairs).
 The plot shows that for a fixed number using the delay-spectrum approach, redundant arrays (hex/grid) provide about an order of magnitude improvement over imaging arrays.
-Since the baselines go as order $N^2$, that corresponds to that for about 1/3 the number of elements you get the same performance (as verified by the Hex-37/Imaging-128 lines).
-This advantage does not extend to the other approaches discussed below.
+Since the baselines go as order $N^2$, this corresponds to using about 1/3 of the number of elements to yield the same performance (as verified by the Hex-37/Imaging-128 lines).
+This advantage does not extend to the other approaches discussed below and imaging configurations would be preferred.
+Note that a filled hexagonal-packed array provides excellent imaging as well, but with potentially limited resolution unless outriggers are includes.
 
 Another approach to improving sensitivity is to make the collecting area per element larger.
+In the limit, this yields a fully filled aperture as mentioned.
+A larger element limits the accessible field-of-view, which however is not a huge liability for a focussed experiment, as long as an appropriate patch of ``cold'' pass passes within its field.
+One method combining both advantages is to use small elements to build up phased array ``tiles'' and then electronically produce many beams, which may get correlated.
+This adds additional complexity and potential systematics, however it is the method employed by both MWA and LOFAR.
+Fig.
+\ref{fig:boost} (right) shows the relative sensitivity of the various array relative to HERA-19.
+The circles near the array name indicate the relative sizes of the element.
+Note that MWA and LOFAR are cross-hatched to indicated phased array tiles.
 
@@
 Alternatively, for the same sensitivity one can build roughly 1/3 the antennas.
-Right:  Power spectrum sensitivities for various arrays normalized to the
-19-element hex.  Here the element size is included, which is shown as an appropriately sized circle.}
+Right:  Power spectrum sensitivities for various arrays normalized to HERA-19.
+Here the element size is included, which is shown as an appropriately sized circle.}
 \label{fig:boost}
@@
 
-The power spectrum uses the full magnitude of the {\bf k}-vector, which is set by both the antenna baseline and bandwidth, as discussed above.
+The power spectrum statistic uses the full magnitude of the {\bf k}-vector, which is set by both the antenna baseline and bandwidth, as discussed above.
 The $\kvpr$ components are proportional to the visibilities and $\kpar$ are proportional to the Fourier transform of the frequency response.
@@
 This technique has been pioneered with the MIT EoR experiment, and has resulted in a package called omnical \citep{zheng_et_al2014}.
-
-\section{Existing Measurements}
-\label{sec:existing}
-As mentioned earlier, a number of programs are underway to make the first detection of the EoR power spectrum using the above techniques.  
-PAPER adopted a foreground mitigation strategy based largely
-on filtering out the wedge in frequency domain (so-called ``delay filtering"; \citealt{parsons_et_al2012b}).  
-By reducing the need
-for image-domain foreground modeling, this approach allowed PAPER to switch to a grid-based antenna layout that
-enhanced sensitivity toward fewer Fourier modes and facilitated calibration
-based on the ``redundancy" of different antenna pairs measuring the same sky modes \citep{zheng_et_al2014}.
-Combined, redundancy and delay filtering provide a robust, inexpensive, and demonstrably successful solution 
-to the foreground problem.  
-Yet PAPER's lack of imaging support and its uneven $uv$ sampling 
-leave it with limited diagnostic capability, particularly for direction-dependent systematics
-such as polarization leakage from Faraday-rotated emission \Mycitep{moore_et_al2013}.
-While concern over such effects has decreased markedly since discovering that
-intrinsic point source polarization is lower than previously thought
-\citep{asad_et_al2015} and that variable Faraday rotation through the
-ionosphere averages down the polarized signal over long observing seasons
-\Mycitep{moore_et_al2016}, direction-dependent beam effects remain an area of interest.
-The MWA's image-based calibration and foreground subtraction strategy provides complementary
-capabilities. Imaging with subtraction, while still under development as a viable foreground strategy,
-can increase sensitivity by recovering more modes of the cosmological signal (see
-Table~\ref{tab:signif}) and help address systematic errors rooted
-in the image domain.  HERA's antenna configuration---shown in Figure \ref{fig:arrayConfig} and discussed in \S\ref{sec:arrayConfig}---emphasizes
-the proven approaches of redundant calibration and delay filtering, while simultaneously
-increasing the extent and density of $uv$ sampling for high-fidelity imaging.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=.8\textwidth]{plots/current_limits.pdf}
-	\caption{The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, (solid symbols represent analyses led by HERA collaborators) compared to 21cmFAST-generated models at $k=0.2$\,$h$\,Mpc$^{-1}$.
-Analysis is still underway on PAPER and MWA observations that approach their projected full sensitivities;  HERA can deliver sub-$\text{mK}^2$ sensitivities.}
-	\vspace{-10pt}
-	\label{fig:limits}
-	\label{fig:IGMtemperatureConstraints}
-\end{figure}
-
-Despite progress, the fact remains that the 21\,cm EoR signal is intrinsically very faint; making a detection requires a
-large instrumental collecting area and a long, dedicated observing campaign.
-Although published PAPER and MWA results (Fig. \ref{fig:limits}) do not yet include observations 
-at full sensitivity that are still being analyzed (e.g. Fig. \ref{fig:Sensitivities}), 
-it is clear already that these instruments lack the sensitivity
-to make a conclusive detection (see Table~\ref{tab:signif}). 
-HERA addresses this shortcoming with a dish element that delivers a much larger collecting area 
-while retaining the necessary characteristics for both proven and forward-looking foreground removal strategies.
 
@@
 relatively agnostic with respect to observing parameters, making it a safe and robust first step in data processing
-\citep{zheng_et_al2014,parsons_etal2014,2015ApJ...809...61A}.
+\citep{zheng_et_al2014,parsons_etal2014,ali_et_al2015}.
 The Internet link between the Karoo and data centers in the U.S.\ has been extensively tested, with assistance from SKA-SA, TENET

### 2016-04-08 manuscript.tex
update to wedge figure/caption

@@
 \centerline{
-\includegraphics[height=3in]{plots/wedgeWithBG.png}
+\includegraphics[height=3in]{plots/wedges.pdf}
 }
-\caption{Foregrounds are a primary challenge facing 21\,cm cosmology experiments. 
-HERA leverages a ``wedge'' in the spherically-averaged ($\mathbf{k}$ is broken into $k_\|$ and $k_\perp$) power spectrum (center panel; \citealt{dillon_et_al2015}).
-Smooth-spectrum foregrounds (right panel) $\sim$5 orders of magnitude brighter than fiducial EoR models (left panel; \citealt{mesinger_et_al2011}) create the ``wedge'' when they interact with the interferometer's chromatic response.
-By avoiding foregrounds, PAPER has placed limits within an order
-of magnitude (in mK) of these models \citep{ali_et_al2015} and show the ``EoR Window'' to be foreground-free.
-HERA's dish and configuration optimize wedge/window isolation and direct sensitivity to low-$k_\perp$ modes where EoR is brightest.
-%v2
-Cartoon of the wedge.
-The gray rectangle represents the smooth-spectrum foregrounds as they would be if the instrument had a perfect achromatic response.
-The pink wedge area shows the chromatic response of the interferometer, where additional power is thrown up from the gray region to higher k-values with increasing baseline.
-The additional horizontal pink line illustrates the affect that other systematics, such as reflections on cables, can have in contaminating the measurement.
-The white region shown the 'eor window', where the measurement is not contaminated by foregrounds or systematics.}
+\caption{The EoR ``window'' and foreground ``wedge'' are shown in this plot, which splits the wavenumber into $\kvpr$ (horizontal) and $\kpar$ (vertical) components.
+Smooth-spectrum foregrounds with an intrinsic spectrum indicated by the gray bar along the bottom create the ``wedge'' when they interact with the interferometer's chromatic response, shown with a foreground image from the MWA.
+Systematics can extend the wedge to the upper left.
+The EoR window (shown with a simulation
+from \cite{mesinger_et_al2011}) to the upper left should be free from foregrounds.}
 \label{fig:wedge}

### 2016-04-08 manuscript.tex
minor edits through section 2.1

@@
 With the first generation 21\,cm telescopes approaching the limits of their sensitivity, HERA continues the experiment by bringing significantly more sensitivity to bear
-while continuing to optimize its design along the delay-spectrum approach.
+while continuing to optimize its design along the delay-spectrum approach to detect the power spectrum of the EoR (discussed in Sec. \ref{sec:delayapproach}).
+Key science results include
+\begin{itemize}[noitemsep,nolistsep]
+\item measuring the 21\,cm power spectrum with high significance (\S\ref{sec:EoRPowerSpectra}),
+\item characterizing parametrized models of reionization and the first ionizing sources (\S\ref{sec:EoRPowerSpectra}),
+\item improving cosmic microwave background (CMB) constraints on primordial density fluctuations and neutrino masses by removing the optical depth degeneracy (\S\ref{sec:tau}), and
+\item directly imaging large scale reionization structure (\S\ref{sec:imaging}).
+\end{itemize}
 
@@
 
-HERA brings to bear both the sensitivity and the precision to deliver its key science results on the basis of proven techniques.
-These science results include
-\begin{itemize}[noitemsep,nolistsep]
-\item measuring the 21\,cm power spectrum with high significance (\S\ref{sec:EoRPowerSpectra}),
-\item characterizing parametrized models of reionization and the first ionizing sources (\S\ref{sec:EoRPowerSpectra}),
-\item improving cosmic microwave background (CMB) constraints on primordial density fluctuations and neutrino masses by removing the optical depth degeneracy (\S\ref{sec:tau}), and
-\item directly imaging large scale reionization structure (\S\ref{sec:imaging}).
-\end{itemize}
-At the same time, HERA continues the development of new techniques that extend and improve its capabilities.
+At the same time, HERA continues the development of new techniques to extend and improve its capabilities.
 Examples include improving foreground subtraction to expand the information content 
-in power spectral and imaging measurements and, with support from the Moore Foundation,
+in power spectral and imaging measurements and, with support from the Gordon Moore Foundation [[FOOTNOTE?]],
 extending HERA's performance at lower frequencies to probe IGM heating in the pre-reionization epoch (\S\ref{sec:EoX}).
@@
 350 14-m parabolic dishes (320 in a dense core $+$ 30 outriggers) in the South African Karoo Radio
-Astronomy Reserve.
+Astronomy Reserve (see Fig. \ref{fig:HERApictures} (top) for a rendering).
 HERA has been funded under the US National Science Foundation's Mid-Scale Innovations Program to begin building elements optimized for robust power spectrum detection in a 
-staged deployment.
-The key is to produce inexpensive collecting area optimized to detect the EOR.
-This has resulted in a close-packed array of transit 14-meter dishes (Fig.
-\ref{fig:HERApictures}), with the first 19-element array completed and construction underway for 37 elements with additional funding from Cambridge University.
-
-The current 19 elements are testing the HERA element design for performance and manufacturability at location.
+staged deployment.  The key is to produce inexpensive collecting area optimized to detect the EOR.  
+The first 19-element array is complete and, with additional funding from Cambridge University, construction underway on the next 18 elements, for a total of 37.
+The South African National Research Foundation SKA-SA group (SKA-SA) is a key partner in the construction and science.
+
+The current 19 elements are testing the HERA element design for performance and manufacturability on location.
 The 37 elements, to be completed in 2016, will provide nearly an order of magnitude more
@@
 Section \ref{sec:eormeas} will present the techniques used to make the measurement, focussing on the delay-spectrum approach.
-Section \ref{sec:existing} will provide a brief summary of the existing measured limits.
 Section \ref{sec:requirements} will present the high-level requirements and Section \ref{sec:design} will present the system description..
@@
 \noindent The {\it Cosmic Dawn} of our universe is one of the last unexplored
-frontiers in cosmic history, which is summarized in Fig.
-\ref{fig:cosmos}.
-On the left is the Big Bang, then the hot, young universe slowly cools until it begins to gravitationally contract to the first structures that relit the universe.
-The Cosmic Dawn saw the evolution of the universe from a very homogeneous ``gas'' to the formation of the first stars and galaxies en route to the structure of galaxies
+frontiers in cosmic history.
+This history is summarized in Fig.
+\ref{fig:cosmos}, starting with the Big Bang on the left.
+The hot, young universe slowly cools until it begins to gravitationally contract to the first structures that relit the universe.
+This Cosmic Dawn saw the evolution of the universe from a very homogeneous ``gas'' to the formation of the first stars and galaxies en route to the structure of galaxies
 of our current time.
@@
 Through power-spectral measurements of the 21\,cm line of hydrogen in the primordial IGM,
-HERA directly constrains the topology and evolution of reionization, 
+HERA will be able to direct constrain the topology and evolution of reionization, 
 opening a unique window into the complex astrophysical interplay between the 
@@
 with 1080 hours of integration (black)
-compared with various heating and reionization models (colored).}
+compared with various heating and reionization models (colored).  Note that only predicted results using the delay-spectrum approach are shown.}
 	\label{fig:Sensitivities}
@@
 the Murchison Widefield Array (MWA; \citealt{bowman_et_al2012,tingay_et_al2013}),
-the MIT EoR experiment (MITEoR; \citealt{zheng_et_al2014}) and the Experiment to Detect the Global EoR Step (EDGES; \citealt{bowman_rogers2010}) (see Sec. \ref{sec:existing}).
+the MIT EoR experiment (MITEoR; \citealt{zheng_et_al2014}) and the Experiment to Detect the Global EoR Step (EDGES; \citealt{bowman_rogers2010}).
 Recent measurements
@@
 high-significance detections of virtually any realistic reionization scenario, even using only currently demonstrated foreground
-mitigation strategies (see Table~\ref{tab:signif}). The development of enhanced foreground modeling and subtraction further improve the achievable constraints.
+mitigation strategies (see Table~\ref{tab:signif}). The development of enhanced foreground modeling and subtraction further improve the achievable results.
 
@@
 faces large
-uncertainties, 
+uncertainties [[[change this?]]], 
 HERA's tight constraints on these parameters are comparable
@@
 They are labelled for the type of configuration (hex, grid, imaging) and the number of antennas (from 37 to 350).
-Note that the (a) hex is based on the configuration used for HERA (14.6m center-to-center spacing); (b) grid for that of PAPER (4m N-S, 16m E-W spacing); (c) Imaging-128 for MWA (\cite{mwa}); and (d) Imaging-48 for LOFAR (\cite{2013A&A...556A...2V}; note that this assumes the ability to fully correlate the HBA-pairs).
+Note that the (a) hex is based on the configuration used for HERA (14.6m center-to-center spacing); (b) grid for that of PAPER (4m N-S, 16m E-W spacing); (c) Imaging-128 for MWA \citep{tingay_et_al2013}; and (d) Imaging-48 for LOFAR \citep{2013A&A...556A...2V}; note that this assumes the ability to fully correlate the HBA-pairs).
 The plot shows that for a fixed number using the delay-spectrum approach, redundant arrays (hex/grid) provide about an order of magnitude improvement over imaging arrays.

### 2016-04-08 manuscript.tex
a few typos corrected

@@
 Figure~\ref{fig:paramConstraints} shows the results of a Markov Chain Monte
-Carlo (MCMC) pipeline for fitting models to 21\,cm power spectrum data \citep{greig_and_mesinger2015}, which we have conservatively limited to the range $8 < z < 10$ (although in practice a much broader bandwidth will be available; see Table \ref{tab:BasicParameters}).
+Carlo (MCMC) pipeline for fitting models to 21\,cm power spectrum data \citep{greig_and_mesinger2015}, which we have conservatively limited to the range $8 < z < 10$ (although in practice a much broader bandwidth will be available; see Sec.
+\ref{sec:freqs}).
 Based on the excursion-set formalism of
@@
 The bandwidth
-therefore equates to a cosmic distance along the site of the telescope, with the frequency determining the cosmic age.
+therefore equates to a cosmic distance along the line-of-site of the telescope, with the frequency determining the cosmic age.
 An observation of an area of sky over a given bandwidth is therefore providing an average
@@
 \subsection{Frequency and Bandwidth}
+\label{sec:freqs}
 Figure \ref{fig:cosmos} indicates the frequency range requirement to probe the expected timescale of the epoch of reionization using the 21cm line of hydrogen as the probe.  

### 2016-04-09 manuscript.tex
mostly tweaked the wedge discussion

@@
 \label{sec:eormeas}
-Due to the expansion of the Universe over cosmological time, we can identify and measure the early Universe via the red-shift of spectral lines.
-The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire Universe back almost to the period of recombination.
+Due to the expansion of the universe over cosmological time, we can identify and measure the early Universe via the red-shift of spectral lines.
+The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire universe back almost to the period of recombination.
 The bandwidth
@@
 
-Obviously between our present observation point and the Epoch of Reionization lies the entire intervening Universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).
+Obviously between our present observation point and the Epoch of Reionization lies the entire intervening universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).
 Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.
@@
 
-Though the power spectrum measurement is over a volume, it is instructive to split the spatial wavenumbers into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $\kvpr$ is determined by the antenna baseline and $\kpar$ by the bandwidth.
-This is useful since it allows us to split out the chromatic response of the interferometer and isolate a phase space where foreground sources ({\em i.e.} everything {\em not} the EOR) contaminate the signal of interest from where they don't.
-As mentioned above, this contaminated phase space may be shown [REF] to be a wedge-shaped region in $\kpr$:$\kpar$ space:
+Though the power spectrum measurement is over a volume, it is instructive to split the spatial wavenumbers into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $|\kvpr|\equiv \kpr = 2\pi b/\lambda X$ is determined by the antenna baseline ($b$) and $\kpar=2\pi/Y\Delta\nu$ by the bandwidth ($\Delta\nu$).
+Here $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes respectively (so, relating wavenumber to physical volume at a given redshift).
+This is useful since it allows us to split out the chromatic response of the interferometer visibility measurement and isolate a phase space where foreground sources ({\em i.e.} everything {\em not} the EOR) contaminate the signal of interest from where they don't.
+
+In these variables, the ``contaminated'' phase space is a wedge-shaped region in $\kpr$:$\kpar$ space such that the ratio of the time-delay across a given baseline ($b/c$,  which impacts the chromatic visilbility) and the delay associated with a given bandwidth ($1/\Delta\nu$) is less than some parameter determined by the details of the system, which we denote as $1/\beta$.
+Substituting $\kpr$ and $\kpar$ in for $b$ and $\Delta\nu$ in this ratio, we find that this wedge is bounded by
 \begin{equation}
-|\kpar| \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
+\kpar \le \beta\frac{X\lambda}{Yc}\kpr + \frac{S}{Y}
 \end{equation}
-where $c$ is the speed of light and $S$ is a parameter accounting for offsets related to the combined spectral smoothness of the foregrounds and the antenna response (see Fig. \ref{fig:wedge}).  
-A key issue to measuring the EOR power spectrum is to understand and minimize such effects.
-For an analytical derivation, see \cite{zahn_etal2012,vedantham_2012,liu_et_al2014b},  in simulation see \cite{datta_etal2010,hazelton_et_al2013} and for observations
+where we've included an offset parameterized by $S$ accounting for affects related to the combined spectral smoothness of the foregrounds and the antenna response (see Fig. \ref{fig:wedge}).
+
+If this $\beta$ factor is determined by the chromaticity of the longest baseline, we see that $\beta=1$ and we call this line the ``horizon line'' ({\em i.e.}, the delay of a source at the horizon).
+Systematics will push that line such as to close the EOR ``window'' to the upper left ($\beta > 1$).
+If we can completely control systematics and constrain the affects of the foreground to {\em e.g.} the main beam, the line will move to decrease the size of the ``wedge'' ($\beta<1$).
+Therefore, a key issue to measuring the EOR power spectrum is to understand and minimize systematics.
+For further discussion and an analytical derivation, see \cite{zahn_etal2012,vedantham_2012,liu_et_al2014b},  in simulation see \cite{datta_etal2010,hazelton_et_al2013} and for observations
 \cite{pober_etal2013b,2015arXiv150601026P,parsons_etal2014,2015arXiv150206016A}.
+
 Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).
 In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_o=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.
+Also note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
 
@@
 \end{equation}
-where $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes (so, relating wavenumber to physical volume at a given redshift), $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.
+where $X$ and $Y$ were introduced above, $\Omega_b$ is the integrated beam response, $B$ is the effective bandwidth, $\lambda$ is the observation wavelength, and $k_B$ is Boltzmann's constant.
 The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.
-Note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
 Rather than ${\hat P}(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.

### 2016-04-09 manuscript.tex
beginning to reform req and system

@@
 The EoR window (shown with a simulation
-from \cite{mesinger_et_al2011}) to the upper left should be free from foregrounds.}
+from \citealt{mesinger_et_al2011}) to the upper left should be free from foregrounds.}
 \label{fig:wedge}
@@
 \end{equation}
-where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement 
-(see \cite{parsons_et_al2012b}).
+where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement (see \citealt{parsons_et_al2012b}).
+Dependencies on redshift and wavenumber have been included.
 
@@
 They are labelled for the type of configuration (hex, grid, imaging) and the number of antennas (from 37 to 350).
-Note that the (a) hex is based on the configuration used for HERA (14.6m center-to-center spacing); (b) grid for that of PAPER (4m N-S, 16m E-W spacing); (c) Imaging-128 for MWA \citep{tingay_et_al2013}; and (d) Imaging-48 for LOFAR \citep{2013A&A...556A...2V}; note that this assumes the ability to fully correlate the HBA-pairs).
-The plot shows that for a fixed number using the delay-spectrum approach, redundant arrays (hex/grid) provide about an order of magnitude improvement over imaging arrays.
+Note that (a) {\em hex} is based on the configuration used for HERA (14.6m center-to-center spacing); (b) {\em grid} for that of PAPER (4m N-S, 16m E-W spacing); (c) {\em imaging-128} for MWA \citep{tingay_et_al2013}; and (d) {\em imaging-48} for LOFAR (\citealt{2013A&A...556A...2V}; note that this assumes the ability to fully correlate the HBA-pairs).
+The plot shows that for a fixed number of elements using the delay-spectrum approach, redundant arrays (hex/grid) provide about an order of magnitude improvement over imaging arrays.
 Since the baselines go as order $N^2$, this corresponds to using about 1/3 of the number of elements to yield the same performance (as verified by the Hex-37/Imaging-128 lines).
 This advantage does not extend to the other approaches discussed below and imaging configurations would be preferred.
-Note that a filled hexagonal-packed array provides excellent imaging as well, but with potentially limited resolution unless outriggers are includes.
+Note that a filled hexagonal-packed array provides excellent imaging as well, but with potentially limited resolution unless outriggers are included.
 
@@
 In the limit, this yields a fully filled aperture as mentioned.
-A larger element limits the accessible field-of-view, which however is not a huge liability for a focussed experiment, as long as an appropriate patch of ``cold'' pass passes within its field.
+A larger element limits the accessible field-of-view, which however is not a huge liability for a focussed experiment, as long as an appropriate patch of ``cold'' sky passes within its field-of-view.
 One method combining both advantages is to use small elements to build up phased array ``tiles'' and then electronically produce many beams, which may get correlated.
@@
 Fig.
-\ref{fig:boost} (right) shows the relative sensitivity of the various array relative to HERA-19.
+\ref{fig:boost} (right) shows the relative sensitivity of the various array relative to HERA-19, calculated for a redshift corresponding to the peak signal at a hydrogen reionization fraction of 1/2.
 The circles near the array name indicate the relative sizes of the element.
@@
 In Fig.
-\ref{fig:kperf}, blue lines denote $\kpar$ values, black lines denote $|\kvpr|$, and green lines denote total magnitude ($|{\kvec}|$).
+\ref{fig:kperf}, blue lines denote $\kpar$ values, black lines denote $\kpr$, and green lines denote total magnitude ($|{\kvec}|$).
 The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.
@@
 Also shown just below the blue $\kpar$ curves are redshift ranges for that bandwidth centered at $z$=6,12, and 18 (they appear as horizontal ``error'' bars).
-Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer really appropriate, so some evolutionary model would have to be assumed (this is why the line is dashed and only the $z$=12 range is shown).
+Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer appropriate, so some evolutionary model would have to be assumed (this is why the line is dashed and only the $z$=12 range is shown).
 Also in Fig.
@@
 
-\subsection{Antenna Design and Configuration}
-The primary new feature of HERA over previous generation experiments (in this context, primarily PAPER) is the use of large elements to increase the sensitivity.
+\subsection{General Antenna Design and Configuration}
+The primary new feature of HERA over previous generation experiments (in this context, primarily PAPER) is the use of large elements to increase the sensitivity, as indicated in Fig.
+\ref{fig:boost}.
 The cost-performance optimization will
@@
 \citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016}.  Note that to pursue map-making and hybrid techniques, additional antenna outriggers are valuable additions, so the design
-should be able to accommodate ~1km baselines.  In referring to the EOR sensitivity of the array however, the number of antennas just in the core is sometimes used.
+should be able to accommodate $\approx$1km baselines.  Note that in referring to the EOR sensitivity of the array, often the number of antennas just in the core are
+indicated -- that is for EOR HERA-320 and HERA-350 are often used interchangeably.
 
@@
 The fall-back position for the low-frequencies is a serial deployment of a scaled version of the HERA feed or to potentially build additional elements specifically for the low-frequencies.
-The 100 MHz bandwidth is also the minimum fully digitized and correlated bandwidth.
 
@@
 
+\subsection{Delay Response}
+\label{sec:delayspec}
+Need to include initial 60 at 60 and Nithya etc etc etc
+
 \subsection{Sensitivity Optimization}
 \label{sec:cost}
-As an instrument to characterize the power spectrum over its evolution, the specification on sensitivity is to make at least a nominal detection over the redshift region of support, or $z$=6-13, along with a very robust detection at the peak.
+As an instrument to characterize the power spectrum over the evolution of the EoR, the specification on sensitivity is to make at least a nominal detection over the redshift region of support, or $z$=6-13, along with a very robust detection at the peak.
 As seen in Fig.
@@
 \begin{equation}
+\label{eq:ND}
 N_{\text core} = (320*14)/D
@@
 where $N_{\text core}$ and $D$ are the number of antennas in the core and their diameter respectively.
-Note that the number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} Pober) over a range of values for the HERA configuration.
-This dependency differs from Mellema and analytical epressions, mostly likely due to the assumed configuration.
+Note that the linear number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} Pober) over a range of values for the HERA configuration for wavenumbers greater than 0.15.
+This dependency differs from \cite{2013ExA....36..235M} and analytical expressions, mostly likely due to the assumed configuration.
 Optimization is 
@@
 
-To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, where the total number was constrained to keep constant sensitivity at the minimum discussed in the previous paragraph.
+To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, with the total number set by Eq.
+\ref{eq:ND}.
 This included the full system, except the post-processing (that is, everything after the archived correlator output).
@@
 which becomes very flat (colored purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.
+
+OTHER LINES IN FIG 12.
+
 \begin{figure}[t]
 \centerline{
-\includegraphics[width=7.5cm]{plots/costfigNew.png} 
+\includegraphics[width=0.6\textwidth]{plots/costfig.pdf} 
 }
-\caption{\small Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).
+\caption{\small The background coloring with colorbar shows the relative cost/performance at a fixed performance as a function of element diameter.
+Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).
 The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.
@@
 
-\noindent As described in the previous section, we have applied 
+As described in the previous section, we have applied 
 critical insights from first generation 21\,cm EoR experiments to define
 the requirements for HERA---an instrument that ensures foregrounds remain bounded
-within the wedge while \emph{delivering the sensitivity for
+within the wedge while delivering the sensitivity for
 high-significance detections of the 21\,cm reionization power spectrum with
-established foreground filtering techniques} 
+established foreground filtering techniques
 \citep{pober_et_al2014,greig_and_mesinger2015}.
@@
 and that HERA is always producing high quality science.
+
+For details of the first generation HERA system (the existing and commissioned PAPER signal path), see \cite{PAPERSYS}.
+This paper will discuss the new system to be deployed beginning and 2016 and tested in conjunction with the previous system.
+Fig \ref{fig:overallBlockDiagram} provides a system block diagram of the new version.
 
@@
 
-\subsection{Antenna Element}
+\subsection{Antenna}
 \label{sec:antenna}
@@
 
-The construction materials are readily available standard construction materials, making for a very cost-effective design.
+\subsubsection{Element Construction}
+To minimize cost and aid in construction in a remote location, the construction materials are readily available standard construction materials, making for a very cost-effective design.
 The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.
@@
 Note that the perimeter poles can be stayed with guy lines if needed.
-A close-up of one antenna is shown in Fig \ref{fig:elementpic}.
 The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.
-
-The center of the antenna is a cast concrete donut-shaped hub with 12 angled sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.
-The twelve spars are made of 60mm PVC pipe, which are stressed along three support points to approximate a parabolic shape.
+Two of the three posts are shared with adjacent antennas.
+
+The center of the antenna is a cast concrete donut-shaped hub with 12 sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.
+The twelve surface spars are made of 60mm PVC pipe, which are stressed along three support points to approximate a parabolic shape.
 Note that an ideal moment-loaded beam actually attains a true parabola.
 Note that the three support points (one at the hub, one about 3 meters out radially and one at the rim-end) are at the correct height and angle for the underlying parabola.
+The PVC essentially acts as a spatial filter.
 
@@
 Fig \ref{fig:elementplan} shows this surface scheme and the panelization (left) and the offsets (in mm) of that scheme from an ideal parabola.
-The panel that is ``left out'' is the location of a door to allow access into the hub (via a small removable bridge).
+The panel that is ``left out'' is the location of a door to allow access into the hub (via a small removable door and bridge).
 The additional intermediate spars also make the panel sizes much more manageable and better secured.
@@
 
+\subsubsection{Feed}
 The feed is to be deployed along the two separate phases.
@@
 Cones were ruled out since it was found they introduce additional frequency structure with no real improvement in performance and a range of cylinders were evaluated.
-The phase 1 design is documented in a memo \cite{feedmemo}, while additional studies are on-going at NRAO, University of Cambridge and Stellenbosch University.
+The phase 1 design is documented in a memo \cite{feedmemo}, while additional studies are on-going at NRAO and Cambridge University.
 The chosen feed design is shown in Figure X.
 
-Additionally, to lessen cross-coupling, a screen fence around the perimeter is also under investigation.
-Rather than being vertical, it will have a slight angle and potentially some edge treatment to try and lessen diffraction.
-Tests of the antennas are discussed in a number of papers CITE.
-Figure Y shows beam patterns.
-
-As mentioned, the feed is supported by lines from the three poles via a rigid support to which the feed cage and dipole attach.
+%Additionally, to lessen cross-coupling, a screen fence around the perimeter is also under investigation.  
+%Rather than being vertical, it will have a slight angle and potentially some edge treatment to try and lessen diffraction.  
+%Tests of the antennas are discussed in a number of papers CITE.  Figure Y shows beam patterns.
+
+The feed is supported by lines from the three poles via a rigid support to which the feed cage and dipole attach.
 Another line goes down centrally to the hub to hold the feed rigidly in place at the correct height.
@@
 The low noise amplifier/balun integrates directly to the base of the feed dipole in the same configuration as per PAPER (REF).
-The 150m length of 75 $\Omega$ cable also runs down centrally and then over to the post-amplifier module, which is housed near the RFI-shielded container containing the digital electronics.
-
-\noindent The novel design of HERA's antenna element (Figs. \ref{fig:HERApictures} and \ref{fig:orbcommexptandbeammap}), now extensively field-tested with MSIP seed funding, 
+The 150m length of 75 $\Omega$ cable used in the initial design also runs down centrally and then over to the post-amplifier module, which is housed near the RFI-shielded container containing the digital electronics.
+This will be replaced by the 35m cable/node design.
+
+The novel design of HERA's antenna element (Figs. \ref{fig:HERApictures} and \ref{fig:orbcommexptandbeammap}), now extensively field-tested with MSIP seed funding, 
 is one of the critical advances that enables HERA to achieve its science
@@
 and 2 in Green Bank, we are confident in the ability of our trained field teams to construct HERA
-elements to specification.
+elements to specification.  CAMBRIDGE
 
@@
 Resulting calibration errors range from $\sim5\%$ (in the core) to $\sim10\%$ (for outriggers)
-of the residual fractional noise per antenna after averaging.
+of the residual fractional noise per antenna after averaging.  DILLON REF
 
@@
 
+\subsection{Processing}
+\label{sec:processing}
 {\bf Karoo Array Processing Building (KAPB)}.

### 2016-04-09 plots/costfig.py

@@ -1,0 +1,148 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy import interpolate
+import sys, os, os.path
+basedir = os.getenv('UBDIR')
+sys.path.append(os.path.join(basedir,'Code/Lib'))
+import fitlib
+c = 3.0E8
+
+def d2tau(d,fd=0.32):
+    return 2.0*fd*d/c*1E9  # m to ns
+def tau2d(tau,fd=0.32):
+    return c*(tau/1E9)/(2*fd)  # ns to m
+
+xlh = [6.5,26]
+d331 = 5.551
+delayLimSpec = 60.0
+delay2LimDiam = tau2d(delayLimSpec/2.0)
+delayLimPlt = np.array([[xlh[0],delayLimSpec],[xlh[1],delayLimSpec]])
+delayDiamPlt = np.array([[delay2LimDiam,0.0],[delay2LimDiam,delayLimSpec]])
+
+d = np.loadtxt('costfig_summary_newscale.dat')
+cp = d[:,2]/d331
+diam = d[:,1]
+delay = d2tau(d[:,1])
+
+###First version
+fg1 = plt.figure(1)
+ax1 = fg1.add_subplot(111)
+ax2 = ax1.twinx()
+ax2.plot(d[:,1],delay,linestyle='--',color='r')
+ax2.plot(d[:,1],delay*2,linestyle='--',color='r')
+ax2.plot(d[:,1],delay*3,linestyle='--',color='r')
+ax2.plot(delayLimPlt[:,0],delayLimPlt[:,1],linestyle='--',color='k')
+ax2.plot(delayDiamPlt[:,0],delayDiamPlt[:,1],linestyle=':',color='k')
+ax2.text(19,d2tau(19.0+1.0),'1',fontsize=14)
+ax2.text(17,d2tau(2.0*17.0+1.0),'2',fontsize=14)
+ax2.text(14,d2tau(3.0*14.0+1.0),'3',fontsize=14)
+ax2.set_ylabel('Delay [ns] (dashed)',fontsize=14)
+ax2.set_ylim( (0,120) )
+
+ax1.plot(d[:,1],cp,color='b',lw=3)
+ax1.plot(d[:,1],cp,color='g',marker='o')
+ax1.plot([14.0],1.0,'o',markersize=8,color='r')
+ax1.set_ylabel('Cost/Performance (solid)',fontsize=14)
[…]

### 2016-04-09 manuscript.tex
fixed up section 4

@@
 \label{sec:delayspec}
-Need to include initial 60 at 60 and Nithya etc etc etc
+Early analysis \citep{elementmemo} indicated that to trace the EoR over a wide range of redshift, internal reflections should be attenuated by about 60 dB from the initial incoming wave by about 60 ns, or a VSWR $<$ 1.002 for frequencies $<$ 17 MHz.
+This provided the basis for the initial design of the element.
+For the contemplated feeds for a primary focus antenna, it was found that a focal length/diameter radio ($f/D$ of approximately 0.32 was optimal \citep{elementmemo}.
+We may therefore estimate the attenuation as a function of delay and diameter and assign this a ``cost function''.
+The vertical axis of Fig.
+\ref{fig:costfig} shows a very steep cost function at 60 ns.
+The curves labelled `1', `2' and `3' are the round-trip travel time for 1, 2 and 3 reflections.
+
+To get 60 dB at 60 ns we may interpret those lines in one of two ways:  (1) set the spec on how much attentuation per reflection is needed and (2) at a fixed attenuation/reflection, what diameter we need.
+For 1 round-trip, for the assumed $f/D$ over the diameters of interest, all delays are less than 60 ns, so it provides no constraint.
+For
+3 round-trips, all diameters exceed 60 ns, so the attenuation should be less than 60 dB for that path-length.
+The case for 2 round-trips provides a constraint on attenuation and diameter, which is discussed at length in \cite{elementmemo}.
+To estimate, it was assumed that the main reflection time constant was due to the standing waves between the feed and vertex of the parabola and that each reflection redirected about -15 dB of the total power in each reflection for additional standing wave power, which leads to a diameter of about 14m.
+
+As a key specification, this has been a focus of study with our prototypes and a series of papers based on analysis and measurements have been produced to provide much more rigor than this simple early analysis \citep{nithya, ewall-wice, patra}.
+The simple specification set above more appropriately becomes a series of curves for attenuation and delay at different wavenumbers for realistic foreground models.
+Fig \ref{fig:nithadelay} taken from \cite{nitha} shows the results at $k$ = 0.2 $h$/Mpc for the adopted design.
+The analysis shows that for $k>0.15 h/$Mpc, HERA should be able to detect the EoR.
+The analysis also shows which baselines produce the best information.
 
@@
 therefore done for this sensitivity against cost for various diameters and total element numbers, while trying to minimize multiple reflections on long delays.
-Figure \ref{fig:costfig} summarizes this optimization, which
-is discussed below.
-
+
+The cost function in Fig. \ref{fig:costfig} is shown as the colored background and color bar, with the delay-spec vertically and the cost at a fixed sensitivity horizontally. 
 To determine the optimal diameter, full system costings were done on a range of diameter sizes from 5m - 25m, with the total number set by Eq.
@@
 This included the full system, except the post-processing (that is, everything after the archived correlator output).
-Figure \ref{fig:costfig} shows the resulting normalized system cost/performance-curve as the horizontal shading relative to the canonical HERA design 
-which becomes very flat (colored purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.
-
-OTHER LINES IN FIG 12.
+The resulting normalized system cost/performance-curve  becomes very flat (colored purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.
 
@@
 }
-\caption{\small The background coloring with colorbar shows the relative cost/performance at a fixed performance as a function of element diameter.
-Costing model for a fixed sensitivity system as a function of diameter (thick blue line and left axis) as well as the round-trip delay for 1,2 and 3 feed-vertex trips (dashed red lines and right axis).
-The 60 ns spec is indicated by the horizontal dash line, and the two-trip intersection at about 14-meter is the vertical dashed line.
-SEE TEXT...
+\caption{\small The background coloring with colorbar shows the relative cost/performance for delay and cost at a fixed performance as a function of element diameter.
+See text and \cite{elementmemo} for details.
+A diameter of 14-m (vertical dashed line) is near the cost minimum and consistent with assumption for the delay specification.
 \label{fig:costfig}}

### 2016-04-10 manuscript.tex
some references updates

@@
 
-As a key specification, this has been a focus of study with our prototypes and a series of papers based on analysis and measurements have been produced to provide much more rigor than this simple early analysis \citep{nithya, ewall-wice, patra}.
+As a key specification, this has been a focus of study with our prototypes and a series of papers based on analysis and measurements have been produced to provide much more rigor than this simple early analysis \citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016,patra_et_al0216}.
 The simple specification set above more appropriately becomes a series of curves for attenuation and delay at different wavenumbers for realistic foreground models.
-Fig \ref{fig:nithadelay} taken from \cite{nitha} shows the results at $k$ = 0.2 $h$/Mpc for the adopted design.
+Fig \ref{fig:nithadelay} taken from \cite{thyagarajan_et_al2016} shows the results at $k$ = 0.2 $h$/Mpc for the adopted design.
 The analysis shows that for $k>0.15 h/$Mpc, HERA should be able to detect the EoR.

### 2016-04-11 manuscript.tex
going through sec 5

@@
 This redundancy may also be exploited to calibrate the array as well, if the baselines are sufficiently redundant.
-This technique has been pioneered with the MIT EoR experiment, and has resulted in a package called omnical \citep{zheng_et_al2014}.
+This technique has been pioneered with the MIT EoR experiment \citep{zheng_et_al2014}, and has resulted in a package called OMNICAL developed for PAPER and HERA .
+\cite{ali_et_al2015} provides a description of this technique as it applies to calibrating the PAPER array, which is mostly situated on a grid with many redundant baselines.
+This section provides a high-level overview of calibration as it applies to HERA.
+
+Given that the technique requires redundant baselines
 
@@
 This included the full system, except the post-processing (that is, everything after the archived correlator output).
-The resulting normalized system cost/performance-curve  becomes very flat (colored purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.
+The resulting normalized system cost/performance-curve  becomes very flat (colored dark blue/purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.
 
@@
 \section{System Design} \label{sec:design}
-
-As described in the previous section, we have applied 
-critical insights from first generation 21\,cm EoR experiments to define
-the requirements for HERA---an instrument that ensures foregrounds remain bounded
+As described in the previous section, the critical insights from the first generation 21\,cm EoR experiments have been applied to define
+the requirements for HERA---an instrument designed to ensure that foregrounds remain bounded
 within the wedge while delivering the sensitivity for
@@
 
-For details of the first generation HERA system (the existing and commissioned PAPER signal path), see \cite{PAPERSYS}.
+Note that in addition to the 19 elements in South Africa (and an initial prototype near Berkeley, CA), a pair of elements at the Green Bank Observatory in West Virginia have been used extensively in the validation testing, the results of which are highlighted in the earlier referenced material and discussed below.
+Also, three elements are currently under construction at the Mullard Radio Astronomyl Observatory outside Cambridge, UK for additional feed and element tests.
+
+For details of the first generation HERA system (the existing and commissioned PAPER signal path), see \cite{parsons_et_al2010}.
 This paper will discuss the new system to be deployed beginning and 2016 and tested in conjunction with the previous system.
-Fig \ref{fig:overallBlockDiagram} provides a system block diagram of the new version.
+Fig \ref{fig:overallBlockDiagram} provides a system block diagram of the new HERA architecture.
 
@@
 The second item constrains the diameter and element count, as well as the focal length over diameter ratio ($f/D$), based on a cost function and maximizing sensitivity per element.
-And the third items constrains the construction materials and methods and the operational model.
-This led to a fixed transit element with a diameter of 14 meters.
-
-\subsubsection{Element Construction}
-To minimize cost and aid in construction in a remote location, the construction materials are readily available standard construction materials, making for a very cost-effective design.
-The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.
-Given the close-packed design, each pole (except for the perimeter) are shared between three antennas.
-This both reduces the number of poles, but also makes for a balanced load.
-Note that the perimeter poles can be stayed with guy lines if needed.
-The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.
-Two of the three posts are shared with adjacent antennas.
-
-The center of the antenna is a cast concrete donut-shaped hub with 12 sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.
-The twelve surface spars are made of 60mm PVC pipe, which are stressed along three support points to approximate a parabolic shape.
-Note that an ideal moment-loaded beam actually attains a true parabola.
-Note that the three support points (one at the hub, one about 3 meters out radially and one at the rim-end) are at the correct height and angle for the underlying parabola.
-The PVC essentially acts as a spatial filter.
-
-The shape between the spars is actually a one-dimensional parabola rather than a two-dimensional paraboloid, so the dish surface may more properly be called a ``facetted parabola''.
-If only twelve spars were used for the entire surface, simple Ruze losses would be quite high (about 10\% at 150 MHz).
-Therefore, at about 2.1 meters another horizontal member is placed, which launches another parabolic spar.
-The simple Ruze loss is then less than 1\%, which is much more acceptable.
-Fig \ref{fig:elementplan} shows this surface scheme and the panelization (left) and the offsets (in mm) of that scheme from an ideal parabola.
-The panel that is ``left out'' is the location of a door to allow access into the hub (via a small removable door and bridge).
-The additional intermediate spars also make the panel sizes much more manageable and better secured.
-The panel arrangement is also set by a standard width of mesh roll of 1.22m.
-
-\begin{figure}[h!]
-\centerline{
-\includegraphics[height=6cm]{plots/antPlan.png}
-\includegraphics[height=6cm]{plots/offset.png} 
-}
-\caption{\small Element.
-\label{fig:elementplan}}
-\end{figure}
-
-\subsubsection{Feed}
-The feed is to be deployed along the two separate phases.
-Initially, the feed reuses the dipole portion of the PAPER element, however in a new optimized cylindrical feed-cage configuration.
-The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on dish), standing waves (blockage, {\em e.g.} Baars), and frequency response.
-The design parameters were size of backplane, height of mast, height of skirt, and cone versus cylinder.
-Cones were ruled out since it was found they introduce additional frequency structure with no real improvement in performance and a range of cylinders were evaluated.
-The phase 1 design is documented in a memo \cite{feedmemo}, while additional studies are on-going at NRAO and Cambridge University.
-The chosen feed design is shown in Figure X.
-
-%Additionally, to lessen cross-coupling, a screen fence around the perimeter is also under investigation.  
-%Rather than being vertical, it will have a slight angle and potentially some edge treatment to try and lessen diffraction.  
-%Tests of the antennas are discussed in a number of papers CITE.  Figure Y shows beam patterns.
-
-The feed is supported by lines from the three poles via a rigid support to which the feed cage and dipole attach.
-Another line goes down centrally to the hub to hold the feed rigidly in place at the correct height.
-
-The low noise amplifier/balun integrates directly to the base of the feed dipole in the same configuration as per PAPER (REF).
-The 150m length of 75 $\Omega$ cable used in the initial design also runs down centrally and then over to the post-amplifier module, which is housed near the RFI-shielded container containing the digital electronics.
-This will be replaced by the 35m cable/node design.
-
-The novel design of HERA's antenna element (Figs. \ref{fig:HERApictures} and \ref{fig:orbcommexptandbeammap}), now extensively field-tested with MSIP seed funding, 
-is one of the critical advances that enables HERA to achieve its science
-goals cost-effectively.  The 14-m fixed zenith-pointing parabolic dish strikes an optimal
-balance between sensitivity and systematics.  The large collecting area of a
+And the third item constrains the construction materials and methods and the operational model.
+This led to a fixed transit element with a diameter of 14 meters to strike an optimal balance between sensitivity and systematics.
+
+The large collecting area of a
 HERA element yields nearly 5 times the sensitivity of an MWA tile and more
@@
 
-With observing wavelengths of 1.5 to 3 m, meeting the required accuracy
-in placing the components of HERA's dish is straightforward.
-Dish centers are fixed
-by a concrete hub placed with final surveyed accuracy of $\sim$10 cm.  These
-hubs constrain radial PVC spars, tensioned into approximate parabolas against a
-rim supported by utility poles.  The resulting faceted paraboloid has
-a measured RMS surface accuracy of 3 cm, which is well within the 0.1$\lambda$ tolerance for
-coherent phasing.  Feed placement relative to the dish is the most stringent requirement, with
-a tolerance of 5 cm.  This placement is ensured by spring-tensioned Kevlar lines attached to
-three utility poles at the dish rim and a 165-lb tensioning down to the concrete hub with
-a fixed length line that ensures a correct focal height.  With 19 dishes constructed in South Africa
-and 2 in Green Bank, we are confident in the ability of our trained field teams to construct HERA
-elements to specification.  CAMBRIDGE
+\subsubsection{Element Construction}
+To minimize cost and aid in construction in a remote location, the construction materials are readily available standard construction materials, making for a very cost-effective design.
+The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.
+Given the close-packed design, each pole (except for the perimeter) are shared between three antennas.
+This both reduces the number of poles, but also makes for a balanced load.
+Note that the perimeter poles can be stayed with guy lines if needed.
+The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.
+Two of the three posts are shared with adjacent antennas.
+
+The center of the antenna is a cast concrete donut-shaped hub with 12 sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.
+The twelve surface spars are made of 60mm PVC pipe, which are stressed along three support points to approximate a parabolic shape.
+Note that an ideal moment-loaded beam actually attains a true parabola.
+Note that the three support points (one at the hub, one about 3 meters out radially and one at the rim-end) are at the correct height and angle for the underlying parabola.
+The PVC essentially acts as a spatial filter.
+
+The shape between the spars is actually a one-dimensional parabola rather than a two-dimensional paraboloid, so the dish surface may more properly be called a ``facetted parabola''.
+If only twelve spars were used for the entire surface, simple Ruze losses would be quite high (about 10\% at 150 MHz).
+Therefore, at about 2.4 meters another horizontal member is placed, which launches another parabolic spar.
+The simple Ruze loss is then less than 1\%, which is much more acceptable.
+Fig \ref{fig:elementplan} shows this surface scheme and the panelization (left) and the offsets (in mm) of that scheme from an ideal parabola.
+The panel that is ``left out'' is the location of a door to allow access into the hub (via a small removable door and bridge).
+The additional intermediate spars also make the panel sizes much more manageable and better secured.
+The panel arrangement is also set by a standard width of mesh roll of 1.22m.
+
+\begin{figure}[h!]
+\centerline{
+\includegraphics[height=6cm]{plots/antPlan.png}
+\includegraphics[height=6cm]{plots/offset.png} 
+}
+\caption{\small Element.
+\label{fig:elementplan}}
+\end{figure}
+
+\subsubsection{Feed}
+The feed is to be deployed along the two separate phases.
+Initially, the feed reuses the dipole portion of the PAPER element, however in a new optimized cylindrical feed-cage configuration.
+The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on the dish), standing waves (blockage size), and frequency response.
+The original PAPER sleeved dipole is held on a PVC mast within the cylindrical structure.
+The design parameters were size of backplane, height of mast, and height of cylinder.
+A range of conical structures were looked at but were ruled out since it was found they introduce additional frequency structure with no real improvement in performance.
+The phase 1 design is documented in a memo \cite{feedmemo}, which are currently under test on the 19 elements in South Africa.
+Additional studies are on-going at NRAO and Cambridge University.
+
+%Additionally, to lessen cross-coupling, a screen fence around the perimeter is also under investigation.  
+%Rather than being vertical, it will have a slight angle and potentially some edge treatment to try and lessen diffraction.  
+%Tests of the antennas are discussed in a number of papers CITE.  Figure Y shows beam patterns.
+
+The feed is supported by spring-tensioned aramid fiber lines from the three poles via a rigid support to which the feed cage and dipole attach.
+The tensioning line goes down centrally to the hub to hold the feed rigidly in place at the correct height.
+This also stabilizes the feed in position and allows for strain relief on the coaxial cables.
+
+The low noise amplifier/balun integrates directly to the base of the feed dipole in the same configuration as per PAPER (REF).
+The 150m length of 75 $\Omega$ cable used in the initial design also runs down centrally and then over to the post-amplifier module, which is housed near the RFI-shielded container containing the digital electronics.
+This will be replaced by the 35m cable/node design.
 

### 2016-04-11 manuscript.tex
Initial light edits to the science sections

@@
 Section \ref{sec:science} will give a high-level overview of the science motivating HERA's construction, including primary and secondary science objectives.
-Section \ref{sec:eormeas} will present the techniques used to make the measurement, focussing on the delay-spectrum approach.
+Section \ref{sec:eormeas} will present the techniques used to make the measurement.
 Section \ref{sec:requirements} will present the high-level requirements and Section \ref{sec:design} will present the system description..
@@
 \ref{fig:cosmos}, starting with the Big Bang on the left.
-The hot, young universe slowly cools until it begins to gravitationally contract to the first structures that relit the universe.
-This Cosmic Dawn saw the evolution of the universe from a very homogeneous ``gas'' to the formation of the first stars and galaxies en route to the structure of galaxies
-of our current time.
-During this period, encoded in the structure of the 
-intergalactic medium (IGM) lies a panoply of information about the underlying astrophysical and cosmological phenomena governing cosmic evolution.
-Ultimately this early population of gravitationally condensed material produced sufficiently energetic flux to reionize the IGM
-from its previous neutral state in a period called the Epoch of Reionization.
+The hot, young universe expands and cools slowly in the background while gravitational instability causes primordial density fluctuations to grow.
+The Cosmic Dawn represents a specific epoch in this growth, where the first stars and galaxies formed and illuminated the Universe en route to forming the astronomical structures we see today.
+Ultimately, this early population of gravitationally condensed material produced sufficiently energetic flux to reionize the intergalactic medium (IGM) from its previous neutral state in a period called the Epoch of Reionization.
 This period is shown in Fig \ref{fig:cosmos} as the rapid transition from the big bubble structures to structures that begin to resemble the denizens of our current universe.
+The structure of the
+intergalactic medium (IGM) thus contains a panoply of information about the underlying astrophysical and cosmological phenomena governing cosmic evolution.
 
@@
 the detailed evolution of the IGM 
-is only loosely constrained \citep{haiman_holder2003,mortonson_hu2008,mesinger_et_al2012}.
+is only loosely constrained \citep{haiman_holder2003,mortonson_hu2008,zahn_etal2012,mesinger_et_al2012}.
 Lyman-$\alpha$ absorption features in quasar and $\gamma$-ray burst spectra give 
@@
 first luminous objects and their environments.
-Using cosmological redshift, HERA can associate 
-the signal at each observing frequency with emission time (or distance) to determine both the time evolution
+The spectral nature of 21\,cm cosmology means that 
+the signal at each observing frequency can be associated with an emission time (or distance) to determine both the time evolution
 and three-dimensional spatial structure of ionization in the IGM.

### 2016-04-11 manuscript.tex
workign through 5

--- 2016-04-11 manuscript.tex
@@
 \subsection{The ``Wedge''}
+\label{sec:wedge}
 Perhaps the most important advance informing HERA's design is a
@@
 Fig.
-\ref{fig:boost} (right) shows the relative sensitivity of the various array relative to HERA-19, calculated for a redshift corresponding to the peak signal at a hydrogen reionization fraction of 1/2.
+\ref{fig:boost} (right) shows the relative sensitivity of the various array relative to HERA-19, calculated for a redshift corresponding to the peak signal at a hydrogen reionization fraction of 0.5.
 The circles near the array name indicate the relative sizes of the element.
@@
 
-Note that in addition to the 19 elements in South Africa (and an initial prototype near Berkeley, CA), a pair of elements at the Green Bank Observatory in West Virginia have been used extensively in the validation testing, the results of which are highlighted in the earlier referenced material and discussed below.
-Also, three elements are currently under construction at the Mullard Radio Astronomyl Observatory outside Cambridge, UK for additional feed and element tests.
+Note that in addition to the 19 elements in South Africa (and an initial prototype near Berkeley, CA), a pair of elements at the Green Bank Observatory in West Virginia have been used extensively in the validation testing, discussed below.
+Also, three elements are currently under construction at the Mullard Radio Astronomy Observatory outside Cambridge, UK for additional feed and element tests.
 
@@
 relates directly to the leakage of smooth-spectrum foregrounds into regions of Fourier space 
-used to measure reionization.
+used to measure reionization, as discuss in Sec. \ref{sec:wedge}.
 To facilitate foreground filtering, HERA's antenna element is designed to suppress reflections at long time delays.
 
+%%%%%%%%%%FIXUP HERE
+\begin{figure}[h!]
+	\centering
+    \includegraphics[width=0.49\textwidth]{plots/s11_compare_msip.pdf}
+	%\includegraphics[width=0.49\textwidth]{plots/compare_kernels_msip_subband.pdf}
+\includegraphics[width=0.5\textwidth]{plots/ps1d_with_delay_kernel.pdf}
+	\vspace{-25pt}
+	\caption{Left: Electromagnetic simulations (black; \citealt{ewallwice_et_al2016}) of the response of the HERA dish as a function of time delay, $\tau$,
+       agree with field measurements (grey; \citealt{patra_et_al2016}). 
+Right: Modeled foreground power spectra with (green) and without (blue) the antenna chromaticism over a 10\,MHz subband centered at $z=7.5$.
+The spectral smoothness of the dish enables foreground isolation below the level of a fiducial reionization signal (black) for most $k_{\parallel}$ modes, yielding percent-level constraints on model parameters.}
+	\label{fig:reflectometry}
+	\vspace{-10pt}
+\end{figure}
+
+As shown in Fig.~\ref{fig:reflectometry}, HERA's element meets this requirement in
+reflectometry tests, performing comparably to the PAPER element.  The
+relatively short (4.5\,m) focal height of the parabolic dish 
+minimizes the timescale of feed-dish reflections while maintaining the feed's
+illumination of the dish.  HERA's feed design is based on the PAPER sleeved
+dipole element, inverted above a backplane, with a skirting cylinder to
+improve symmetry in the polarized beam response patterns.  
+%%%%%%%%%%%
+
+Figure
+\ref{fig:orbcommexptandbeammap} shows the beam pattern measured at 137 MHz with a beam mapping
+system using the ORBCOMM satellite network \Mycitep{neben_et_al2016}.  Results
+indicate an effective per-element collecting area of 93\,m$^2$, compared with the
+theoretical maximum Airy response of 155 m$^2$.  The measured primary beam is
+consistent with simulations at the 0.1\% to 0.5\% level, with a full width at
+half maximum of $\sim$$10^\circ$, and a first sidelobe at -20 dB \citep{ewallwice_et_al2016,neben_et_al2016,patra_et_al2016,thyagarajan_et_al2016}.
+Feed-to-feed coupling between adjacent antennas has been measured to be below -50 dB, indicating
+that mutual coupling will not be problematic.
+%%%%%%%%%%%%
+SHOW CALCULATED BEAM PATTERNS
 \begin{figure}[h!]
@@
 
-\begin{figure}[h!]
-	\centering
-    \includegraphics[width=0.49\textwidth]{plots/s11_compare_msip.pdf}
-	%\includegraphics[width=0.49\textwidth]{plots/compare_kernels_msip_subband.pdf}
-\includegraphics[width=0.5\textwidth]{plots/ps1d_with_delay_kernel.pdf}
-	\vspace{-25pt}
-	\caption{Left: Electromagnetic simulations (black; \citealt{ewallwice_et_al2016}) of the response of the HERA dish as a function of time delay, $\tau$,
-       agree with field measurements (grey; \citealt{patra_et_al2016}). 
-Right: Modeled foreground power spectra with (green) and without (blue) the antenna chromaticism over a 10\,MHz subband centered at $z=7.5$.
-The spectral smoothness of the dish enables foreground isolation below the level of a fiducial reionization signal (black) for most $k_{\parallel}$ modes, yielding percent-level constraints on model parameters.}
-	\label{fig:reflectometry}
-	\vspace{-10pt}
-\end{figure}
-
-As shown in Fig.~\ref{fig:reflectometry}, HERA's element meets this requirement in
-reflectometry tests, performing comparably to the PAPER element.  The
-relatively short (4.5\,m) focal height of the parabolic dish 
-minimizes the timescale of feed-dish reflections while maintaining the feed's
-illumination of the dish.  HERA's feed design is based on the PAPER sleeved
-dipole element, inverted above a backplane, with a skirting cylinder to
-improve symmetry in the polarized beam response patterns.  Figure
-\ref{fig:orbcommexptandbeammap} shows the beam pattern measured at 137 MHz with a beam mapping
-system using the ORBCOMM satellite network \Mycitep{neben_et_al2016}.  Results
-indicate an effective per-element collecting area of 93\,m$^2$, compared with the
-theoretical maximum Airy response of 155 m$^2$.  The measured primary beam is
-consistent with simulations at the 0.1\% to 0.5\% level, with a full width at
-half maximum of $\sim$$10^\circ$, and a first sidelobe at -20 dB \citep{ewallwice_et_al2016,neben_et_al2016,patra_et_al2016,thyagarajan_et_al2016}.
-Feed-to-feed coupling between adjacent antennas has been measured to be below -50 dB, indicating
-that mutual coupling will not be problematic.
-
 \subsubsection{Element Construction}
@@
 The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.
-Two of the three posts are shared with adjacent antennas.
-
-The center of the antenna is a cast concrete donut-shaped hub with 12 sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.
+Two of the three intermediate posts are shared with adjacent antennas.
+
+The center of the antenna is a cast-in-place concrete donut-shaped hub with 12 sleeves for PVC spars which provide the support for the surface and 12 horizontal spars which allow for additional support.
 The twelve surface spars are made of 60mm PVC pipe, which are stressed along three support points to approximate a parabolic shape.
 Note that an ideal moment-loaded beam actually attains a true parabola.
-Note that the three support points (one at the hub, one about 3 meters out radially and one at the rim-end) are at the correct height and angle for the underlying parabola.
-The PVC essentially acts as a spatial filter.
-
-The shape between the spars is actually a one-dimensional parabola rather than a two-dimensional paraboloid, so the dish surface may more properly be called a ``facetted parabola''.
+The three support points (one at the hub, one about 3 meters out radially and one at the rim-end) are at the correct height and angle for the underlying parabola and the PVC essentially acts as a spatial filter.
+
+The shape between the spars is actually a parabola rather than a tparaboloid, so the dish surface may more properly be called a ``faceted parabola''.
 If only twelve spars were used for the entire surface, simple Ruze losses would be quite high (about 10\% at 150 MHz).
@@
 \centerline{
-\includegraphics[height=6cm]{plots/antPlan.png}
-\includegraphics[height=6cm]{plots/offset.png} 
+\includegraphics[height=6cm]{plots/antPlan.pdf}
+\includegraphics[height=6cm]{plots/offset.pdf} 
 }
-\caption{\small Element.
+\caption{\small Element panel plan (left) and residual from a true paraboloid (right).
 \label{fig:elementplan}}
@@
 \subsubsection{Feed}
-The feed is to be deployed along the two separate phases.
-Initially, the feed reuses the dipole portion of the PAPER element, however in a new optimized cylindrical feed-cage configuration.
+The feed is to be deployed along two separate phases.
+Initially, the feed reuses the original PAPER sleeved dipole, however in a new optimized cylindrical feed-cage configuration.
 The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on the dish), standing waves (blockage size), and frequency response.
-The original PAPER sleeved dipole is held on a PVC mast within the cylindrical structure.
 The design parameters were size of backplane, height of mast, and height of cylinder.
 A range of conical structures were looked at but were ruled out since it was found they introduce additional frequency structure with no real improvement in performance.
-The phase 1 design is documented in a memo \cite{feedmemo}, which are currently under test on the 19 elements in South Africa.
+The phase 1 design documented in a memo \cite{feedmemo} is currently under test on the 19 elements in South Africa.
 Additional studies are on-going at NRAO and Cambridge University.
@@
 
-The low noise amplifier/balun integrates directly to the base of the feed dipole in the same configuration as per PAPER (REF).
+The low noise amplifier/balun integrates directly to the base of the feed dipole in the same configuration as per PAPER.
 The 150m length of 75 $\Omega$ cable used in the initial design also runs down centrally and then over to the post-amplifier module, which is housed near the RFI-shielded container containing the digital electronics.
-This will be replaced by the 35m cable/node design.
+This will be replaced by the 35m cable/node design, to be discussed.
 
@@
 
-\noindent HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments 
+As an experiment, HERA has optimized its configuration for the delay-spectrum approach to measure the EoR power spectrum.
+This means that it should be on short redundant baselines.
+Measurements conducted with PAPER showed that baselines of about 15m allow \cite{dillon_parsons2016}.
+
+HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments 
 to cover the $uv$-plane with sub-element sampling density (see Fig.~\ref{fig:arrayConfig}). The dense core maximizes 
@@
 Resulting calibration errors range from $\sim5\%$ (in the core) to $\sim10\%$ (for outriggers)
-of the residual fractional noise per antenna after averaging.  DILLON REF
+of the residual fractional noise per antenna after averaging.
 

### 2016-04-12 manuscript.tex
updated feed/ant section with new figures

@@
 	\centering
-	\includegraphics[width=.8\textwidth]{plots/current_limits.pdf}
+	\includegraphics[width=.8\textwidth]{plots/summary_results.png}
 	\caption{The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, (solid symbols represent analyses led by HERA collaborators) compared to 21cmFAST-generated models at $k=0.2$\,$h$\,Mpc$^{-1}$.
@@
 that mutual coupling will not be problematic.
-%%%%%%%%%%%%
-SHOW CALCULATED BEAM PATTERNS
+
 \begin{figure}[h!]
@@
 A range of conical structures were looked at but were ruled out since it was found they introduce additional frequency structure with no real improvement in performance.
-The phase 1 design documented in a memo \cite{feedmemo} is currently under test on the 19 elements in South Africa.
+The phase 1 design process and outcome is documented in  \cite{feedmemo} and is currently under test on the 19 elements in South Africa.
 Additional studies are on-going at NRAO and Cambridge University.
+
+The first phase design uses a 172cm diameter backplane, with a 36cm cylinder height.
+The dipole is held at 36cm from the backplane and the back of the feed is rigged at 4.9m above the vertex of the dish.
+The 10 dB input match for the feed itself spans from about 105 - 200 MHz.
+The resultant calculated antenna performance parameters are shown in Fig.
+\ref{fig:ant_params} and beam-patterns for 110 MHz, 155 MHz and 190 MHz are shown in Fig.
+\ref{fig:beampatterns}.
+
+\begin{figure}[h!]
+\centerline{
+\includegraphics[height=0.6\textwidth]{plots/ant_params.pdf}
+}
+\caption{\small Antenna parameters}
+\label{fig:ant_params}
+\end{figure}
+
+\begin{figure}[h!]
+\centerline{
+\includegraphics[height=0.3\textwidth]{plots/pat_110.pdf}
+\includegraphics[height=0.3\textwidth]{plots/pat_155.pdf}
+\includegraphics[height=0.3\textwidth]{plots/pat_190.pdf}
+}
+\caption{\small Antenna patterns at 110 MHz, 155 MHz and 190 MHz.}
+\label{fig:ant_params}
+\end{figure}
 

### 2016-04-12 manuscript.tex
minor updates to feed

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{graphicx}
@@
 \renewcommand{\deg}{^\circ}
+\newcommand{\integral}{\int\limits}
 
@@
 
-%%%%%%%%%%FIXUP HERE
+%%%%%%%%%%%%%%%
+FIXUPHERE
 \begin{figure}[h!]
@@
 improve symmetry in the polarized beam response patterns.  
+TOHERE
 %%%%%%%%%%%
@@
 Initially, the feed reuses the original PAPER sleeved dipole, however in a new optimized cylindrical feed-cage configuration.
-The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on the dish), standing waves (blockage size), and frequency response.
+The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on the dish), standing waves (blockage size), frequency response and polarization match.
 The design parameters were size of backplane, height of mast, and height of cylinder.
 A range of conical structures were looked at but were ruled out since it was found they introduce additional frequency structure with no real improvement in performance.
+The polarization figure-of-merit shown in Fig.
+\ref{fig:ant_params} is a simplified variant of the fractional power leakage as defined in \cite{Moore_et_al2015}:
+\begin{equation}
+\xi = \frac{\integral_{-\pi}^{\pi}\left[\sqrt{G(\theta,\phi=0)} - \sqrt{G(\theta,\phi=90)}\right]^2\sin\theta d\theta}
+               {\integral_{-\pi}^{\pi}\left[\sqrt{G(\theta,\phi=0)} + \sqrt{G(\theta,\phi=90)}\right]^2\sin\theta d\theta}
+\end{equation}
+where $G$ is the beam pattern.
 The phase 1 design process and outcome is documented in  \cite{feedmemo} and is currently under test on the 19 elements in South Africa.
@@
 \caption{\small Antenna patterns at 110 MHz, 155 MHz and 190 MHz.}
-\label{fig:ant_params}
+\label{fig:beampatterns}
 \end{figure}

### 2016-04-12 manuscript.tex
update to delay/array

@@
 Imaging does remain an ultimate goal to fully understand the process, however we will likely need a greater understanding of the signal characteristics and the systematics of our telescopes to achieve this more difficult goal.
-MANY PAPERS, see Nithya and Jonnie recent
 
@@
 Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.
-As a first step, areas of the sky where these signals are minimal (for example, outside the galactic plane and away from strong point sources) are targeted (REF FOR THIS).
+As a first step, areas of the sky where these signals are minimal (for example, outside the galactic plane and away from strong point sources) are targeted \citep{bernardi_2013}.
 More importantly and most fortuitously however is the fact that all of these foreground signals are smooth spectrum sources whereas the expected spectrum of the EOR is expected to be rough since it is made up of non-ionized regions which are randomly distributed over a wide range of redshifts.
@@
 
-As mentioned, a number of programs are underway to make the first detection of the EoR power spectrum using a number of approaches.  
-PAPER adopted a foreground mitigation strategy based largely
-on identifying and filtering out a region of phase space where the strong foregrounds are largely confined (Sec. \ref{sec:delayapproach}).  
+A promising approach to make the first detection of the EoR power spectrum 
+is a foreground mitigation strategy based largely
+on identifying and filtering out a region of phase space where the strong foregrounds are largely confined (Sec. \ref{sec:delayapproach}).  This is the approach
+pioneered by PAPER.
 By reducing the need
@@
 Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement of the EOR spatial power spectrum.
-There are now numerous papers in the literature on the technique of measuring the power spectrum using an interferometer and the attendant issues [[[list of citations]]].
+There are now numerous papers in the literature on the technique of measuring the power spectrum using an interferometer and the attendant issues as addressed in nearly all of the cited references.
 Below, a high-level overview based on  \cite{parsons_et_al2012b} will be presented.
@@
 The simple specification set above more appropriately becomes a series of curves for attenuation and delay at different wavenumbers for realistic foreground models.
-Fig \ref{fig:nithadelay} taken from \cite{thyagarajan_et_al2016} shows the results at $k$ = 0.2 $h$/Mpc for the adopted design.
+Fig \ref{fig:delayspec}, adapted from \cite{thyagarajan_et_al2016}, shows the results at $k$ = 0.2 $h$/Mpc for the adopted design.
 The analysis shows that for $k>0.15 h/$Mpc, HERA should be able to detect the EoR.
-The analysis also shows which baselines produce the best information.
+We also see that the initial simple ``60@60'' specification was an overly restrictive, however was a good initial goal.
+The analysis also indicates which baselines produce the best information on different scales.
+
+\begin{figure}[h!]
+	\centering
+    \includegraphics[width=0.49\textwidth]{plots/delayspecplot.pdf}
+     \includegraphics[width=0.5\textwidth]{plots/ps1d_with_delay_kernel.pdf}
+	\vspace{-25pt}
+	\caption{Left:  Antenna delay attenuation specification with realistic foregrounds from \cite{thyagarajan_et_al2016} with measurements from Green Bank prototype \citep{patra_et_al2016}.
+Also shown is the initial aggressive ``60@60'' specification.
+       agree with field measurements (grey; \citealt{patra_et_al2016}). 
+Right: Modeled foreground power spectra with (green) and without (blue) the antenna chromaticism over a 10\,MHz subband centered at $z=7.5$.
+The spectral smoothness of the dish enables foreground isolation below the level of a fiducial reionization signal (black) for most $k_{\parallel}$ modes, yielding percent-level constraints on model parameters.}
+	\label{fig:delayspec}
+	\vspace{-10pt}
+\end{figure}
 
@@
 and system architecture (see Fig.~\ref{fig:overallBlockDiagram}).
-This architecture
-directly inherits from the PAPER and MWA experiments; HERA begins by reusing the
+This architecture directly inherits from the PAPER and MWA experiments; HERA begins by reusing the
 analog, digital, and real-time processing systems deployed for PAPER-128.  As HERA develops, this
@@
 
-The large collecting area of a
-HERA element yields nearly 5 times the sensitivity of an MWA tile and more
+The large collecting area of a HERA element yields nearly 5 times the sensitivity of an MWA tile and more
 than 20 times that of a PAPER element, but it does so without substantially
@@
 
-%%%%%%%%%%%%%%%
-FIXUPHERE
-\begin{figure}[h!]
-	\centering
-    \includegraphics[width=0.49\textwidth]{plots/s11_compare_msip.pdf}
-	%\includegraphics[width=0.49\textwidth]{plots/compare_kernels_msip_subband.pdf}
-\includegraphics[width=0.5\textwidth]{plots/ps1d_with_delay_kernel.pdf}
-	\vspace{-25pt}
-	\caption{Left: Electromagnetic simulations (black; \citealt{ewallwice_et_al2016}) of the response of the HERA dish as a function of time delay, $\tau$,
-       agree with field measurements (grey; \citealt{patra_et_al2016}). 
-Right: Modeled foreground power spectra with (green) and without (blue) the antenna chromaticism over a 10\,MHz subband centered at $z=7.5$.
-The spectral smoothness of the dish enables foreground isolation below the level of a fiducial reionization signal (black) for most $k_{\parallel}$ modes, yielding percent-level constraints on model parameters.}
-	\label{fig:reflectometry}
-	\vspace{-10pt}
-\end{figure}
-
-As shown in Fig.~\ref{fig:reflectometry}, HERA's element meets this requirement in
+As shown in Fig.~\ref{fig:delayspec}, HERA's element meets this requirement in
 reflectometry tests, performing comparably to the PAPER element.  The
@@
 illumination of the dish.  HERA's feed design is based on the PAPER sleeved
-dipole element, inverted above a backplane, with a skirting cylinder to
+dipole element incorporating a cylindrical backplane to
 improve symmetry in the polarized beam response patterns.
-TOHERE
-%%%%%%%%%%%
-
-Figure
-\ref{fig:orbcommexptandbeammap} shows the beam pattern measured at 137 MHz with a beam mapping
+
+Figure \ref{fig:orbcommexptandbeammap} shows the beam pattern measured at 137 MHz with a beam mapping
 system using the ORBCOMM satellite network \Mycitep{neben_et_al2016}.  Results
@@
 This means that it should be on short redundant baselines.
-Measurements conducted with PAPER showed that baselines of about 15m allow \cite{dillon_parsons2016}.
+Measurements conducted with PAPER showed that baselines of about 15m allow foreground reduction with real systematics and provide a proof of concept for this scale as well.
+The ability to fully calibrate the array is also a key consideration.
+The details of the process and adopted configuration may be found in \cite{dillon_parsons2016}.
+This section will briefly present the configuration.
 

### 2016-04-12 manuscript.tex
still in sec 5

@@
 
-\noindent The {\it Cosmic Dawn} of our universe is one of the last unexplored
+The {\it Cosmic Dawn} of our universe is one of the last unexplored
 frontiers in cosmic history.
@@
 minimize signal reflections over long delays and deliver significant forward gain relative to their horizon response.
-Tests with prototype HERA antennas (Figs. \ref{fig:orbcommexptandbeammap} and \ref{fig:reflectometry}, discussed in \S\ref{sec:antenna})
+Tests with prototype HERA antennas (Figs.  \ref{fig:delayspec} and \ref{fig:orbcommexptandbeammap}, discussed in \S\ref{sec:antenna})
 indicate that a moderately large parabolic dish with a short focal height can meet these requirements
@@
 \small
+\Caption{0.0in}{0.99}{-0.1in}{HERA-350 design parameters and their observational consequences. Angular scales computed at 150\,MHz.}
+\label{tab:BasicParameters}
 \begin{center}
@@
 \end{tabular}
-\Caption{-0.1in}{0.99}{-0.4in}{HERA-350 design parameters and their observational consequences. Angular scales computed at 150\,MHz.}
-\label{tab:BasicParameters}
-\vspace{.2in}
 \end{center}
@@
 dipole element incorporating a cylindrical backplane to
-improve symmetry in the polarized beam response patterns.
+improve symmetry in the polarized beam response patterns (Sec. \ref{sec:feed}, which also presents HFSS calculations of the full feed/element system).
 
@@
 
-The shape between the spars is actually a parabola rather than a tparaboloid, so the dish surface may more properly be called a ``faceted parabola''.
+The shape between the spars is actually a parabola rather than a paraboloid, so the dish surface may more properly be called a ``faceted parabola''.
 If only twelve spars were used for the entire surface, simple Ruze losses would be quite high (about 10\% at 150 MHz).
@@
 \subsubsection{Feed}
+\label{sec:feed}
 The feed is to be deployed along two separate phases.
@@
 \subsection{Analog Signal Path}
-
-\noindent HERA's analog signal path, which operates from 50--250 MHz, emphasizes spectral smoothness and robustness.
-With receiver temperatures of 50--100 K readily achievable with ambient temperature electronics, 
-sky noise dominates the system temperature of telescopes operating below 300 MHz.  As a result,
-HERA's low-noise amplifiers (LNAs) are inexpensive, passively cooled components integrated into the
+HERA's analog signal path from 70--250 MHz with the full performance EoR band of 100-200 MHz, as shown in Fig.
+\ref{fig:cosmos} and discussed in Sec.
+\ref{sec:freqs}, emphasizes spectral smoothness and robustness.
+Although the low-frequency sky has an intrinsically bright signal level, in the desired ``cold patches'' the sky temperature is only about 30 K at the upper frequency end, so relatively low receiver temperatures are desired.
+Note that at the low end of the band (70 MHz), the sky temperature is about 1000 K and across the EoR band it ranges from about 100-400 K.
+With receiver temperatures of 30--100 K readily achievable with ambient temperature electronics however
+HERA's low-noise amplifiers (LNAs) may be inexpensive, passively cooled components integrated into the
 antenna feed.
-HERA's signal path emphasizes careful impedance matching between each component to avoid signal reflections that can induce foreground leakage.
+
+As part of the delay specification,
+HERA's signal path emphasizes careful impedance matching between each component to avoid signal reflections that can worsen foreground leakage.
 In particular, impedance matching between free space (377$\Omega$) and coaxial cable (50$\Omega$) is undertaken
@@
 to limit the total analog path length.
-
-While building to 128 elements during the first two project years, HERA reuses PAPER's tested and characterized 
-LNAs, cables, and post-amplifier modules, operating
-from 100--200 MHz.  In the third project year, these components are replaced by the 50--250 MHz, length-constrained signal
-path described above, supporting HERA's pre-reionization science and extending the lever-arm for modeling and suppressing smooth-spectrum foreground emission.
+This length is set to be 35m, such that the round-trip delay corresponds to about 0.15 $h$/Mpc, which remains in the wavenumbers  contaminated by the wedge.
+The project is also examining the impact of having a range of analog path lengths of $35\pm10$ m.
 

### 2016-04-12 manuscript.tex
incremental sec 5

@@
 
-In order to meet the 35\,m specification for a maximum analog signal path, along with a growing need for system scalability, HERA beyond the first 128 elements will adopt a node-based architecture for amplification, digitization, channelization, and digital
-transmission in the field, building on MWA heritage. This design is merged with PAPER's clean 
+In order to meet the 35\,m specification for a maximum analog signal path length, along with a growing need for system scalability, HERA will adopt a node-based architecture for amplification, digitization, channelization, and digital
+transmission in the field, building on MWA and Allen Telescope Array heritage. This design is merged with PAPER's clean 
 architecture for real-sampling and channelizing the entire analog passband at once, packetizing the data into
-10 Gb Ethernet format, and relying on commercial switches to perform the frequency/antenna corner-turn required for FX correlators
+10 Gb Ethernet format, and relying on commercial switches to perform the frequency/antenna corner-turn required for FX correlators.
+This has led to a new analog-digital-optical board called the Smart Networked Analog-Digital Processor (SNAP), which will become a new modular CASPER device.
+Figure \ref{fig:hardware} shows the procession of boards.
+This section will step through the digital signal path organized by location:  node, container, processing building.
 
@@
 
-\subsection{Processing}
-\label{sec:processing}
 {\bf Karoo Array Processing Building (KAPB)}.
@@
 
-\subsection{Data Management}
+\subsection{Data Processing and Management}
 \label{sec:data}

### 2016-04-13 manuscript.tex
sec 5 and added proto-site section
updated atten spec plot

@@
 
-To get 60 dB at 60 ns we may interpret those lines in one of two ways:  (1) set the spec on how much attentuation per reflection is needed and (2) at a fixed attenuation/reflection, what diameter we need.
+To get 60 dB at 60 ns we may interpret those lines in one of two ways:  (1) set the spec on how much attenuation per reflection is needed and (2) at a fixed attenuation/reflection, what diameter we need.
 For 1 round-trip, for the assumed $f/D$ over the diameters of interest, all delays are less than 60 ns, so it provides no constraint.
@@
 The analysis shows that for $k>0.15 h/$Mpc, HERA should be able to detect the EoR.
-We also see that the initial simple ``60@60'' specification was an overly restrictive, however was a good initial goal.
+We also see that the initial simple ``60@60'' specification was a overly conservative, however it served as a good initial goal.
 The analysis also indicates which baselines produce the best information on different scales.
@@
 
+%\subsection{Site}
+%\label{sec:site}
+HERA is located at the Karoo Astronomy Reserve.
+
 \subsection{Antenna}
 \label{sec:antenna}
-\vspace{-5pt}
 

### 2016-04-13 plots/attendelayspec.py

@@ -1,0 +1,26 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import os.path
+datafile = 'spec_on_foreground_reflected_power_21cmfast_14.6m_150.0_MHz_subband_v2.npz'
+
+npzdata = np.load(datafile)
+tau = npzdata['tau']
+achrm = npzdata['achrmbeam']
+
+bottom = np.zeros(len(tau))
+plt.plot(tau,achrm[0],lw=2,label=r'$k_{||}>0.10$ h/Mpc')
+plt.fill_between(tau,bottom,achrm[0],color='b',alpha='0.25')
+plt.plot(tau,achrm[1],lw=2,label=r'$k_{||}>0.15$ h/Mpc')
+plt.fill_between(tau,bottom,achrm[1],color='g',alpha='0.25')
+plt.plot(tau,achrm[2],lw=2,label=r'$k_{||}>0.20$ h/Mpc')
+plt.fill_between(tau,bottom,achrm[2],color='k',alpha='0.25')
+t60 = [60,60,500]
+a60 = [0,60,60]
+a0 = [0.0,0.0,0.0]
+plt.fill_between(t60,a0,a60,color='k',alpha='0.05')
+plt.plot(t60,a60)
+plt.axis([0,500,80,0])
+plt.xlabel(r'$\tau$ [ns]')
+plt.ylabel('Attenuation [dB]')
+plt.legend()
+plt.savefig('delayspecplot.pdf',dpi=200)

### 2016-04-13 manuscript.tex
Edits to array configuration section

@@
-%XXX: Aaron, do one last check for new typos in my final series of edits.
 \documentclass[preprint,11pt]{aastex} 
@@
 
-%%% ADDED FROM 2014 PROPOSAL
 %\newcommand{\Mycite}[1]{{\bf \cite{#1}}}
@@
 	\centering
-	\includegraphics[height=1.75in]{plots/hera_render.png}
-	\includegraphics[height=1.75in]{plots/hera19H.png}
+	\includegraphics[width=4in]{plots/hera_render.png}
+	\includegraphics[width=4in]{plots/hera19H.png}
 	\caption{Rendering of the 320-element core (left) of the full HERA-350 array and picture of 19 HERA 14-m, zenith-pointing dishes (with PAPER elements in the background) currently deployed in South Africa (right).}
@@
 
-More generally, it is possible to express a map-making algorithm as a linear operator that acts on visibility data to produce a compressed dataset.
+More generally, it is possible to express a map-making algorithm as a linear operator that acts on visibility data to produce a compressed dataset \citep{dillon_et_al2015a}.
 Squaring this to obtain an estimate of the power spectrum then results in a weighted quadratic combination of visibilities, where one multiplies the visibility from every baseline with that from every other baseline.
@@
 \label{sec:calibration}
-As examined above, for maximize sensitivity to a limited number of spatial modes using the delay-spectrum technique, a redundant grid greatly is advantageous.
-This redundancy may also be exploited to calibrate the array as well, if the baselines are sufficiently redundant.
-This technique has been pioneered with the MIT EoR experiment \citep{zheng_et_al2014}, and has resulted in a package called OMNICAL developed for PAPER and HERA .
-\cite{ali_et_al2015} provides a description of this technique as it applies to calibrating the PAPER array, which is mostly situated on a grid with many redundant baselines.
+As examined above, maximizing sensitivity to a limited number of spatial modes for the delay-spectrum technique requires a grid of antennas that simultaneously measures the same baselines using many redundant pairs of antennas.
+This redundancy may also be exploited to calibrate the array as well, as long as the true antenna positions fall close enough to that ideal grid.
+This technique has been pioneered with the MITEoR experiment \citep{zheng_et_al2014}, and has resulted in a package called OMNICAL developed for PAPER and HERA.
+\citet{ali_et_al2015} provides a description of this technique as it applies to calibrating the PAPER array, which is mostly situated on a grid with many redundant baselines.
 This section provides a high-level overview of calibration as it applies to HERA.
@@
 
-As an experiment, HERA has optimized its configuration for the delay-spectrum approach to measure the EoR power spectrum.
-This means that it should be on short redundant baselines.
-Measurements conducted with PAPER showed that baselines of about 15m allow foreground reduction with real systematics and provide a proof of concept for this scale as well.
-The ability to fully calibrate the array is also a key consideration.
-The details of the process and adopted configuration may be found in \cite{dillon_parsons2016}.
-This section will briefly present the configuration.
-
-HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments 
-to cover the $uv$-plane with sub-element sampling density (see Fig.~\ref{fig:arrayConfig}). The dense core maximizes 
-sensitivity on the short baselines for which PAPER's proven foreground filtering strategy works best.
-The core is supplemented by 30 additional outrigger elements 
-to tile the $uv$-plane with instantaneously complete sub-aperture 
-sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz). This sampling 
-strategy suppresses grating lobes in the synthesized beam and provides information for calibrating and correcting 
-direction-dependent antenna responses \citep{dillon_parsons2016}.
-Even with the sub-aperture dithering and long baselines, all 350 HERA elements can be robustly calibrated by taking advantage of 
-its highly redundant configuration \citep{liu_et_al2010,zheng_et_al2014}. 
-Resulting calibration errors range from $\sim5\%$ (in the core) to $\sim10\%$ (for outriggers)
-of the residual fractional noise per antenna after averaging.
+As a focused experiment, HERA's configuration is optimized for the robust foreground-avoidance approach to EoR power spectrum estimation. 
+This requires a densely-packed core of antenna elements, maximizing sensitivity to short baselines---those least contaminated by foreground chromaticity (i.e.\ the wedge). 
+Furthermore, the configuration should be highly redundant, both to enable the delay-spectrum strategy of PAPER \citep{parsons_et_al2012a} and redundant baseline calibration \citep{liu_et_al2010,zheng_et_al2014}.
+On the other hand, it is also desirable to improve the mapmaking ability of the array for calibration and map-based power spectrum estimation with longer baselines and a densely-sampled $uv$-plane \citep{dillon_et_al2015a}.
+This tradeoff was closely examined by \citet{dillon_parsons2016} and the ultimate HERA configuration was drawn from one of the designs they considered.
+
+HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments to cover the $uv$-plane with sub-aperture sampling density (see Fig.~\ref{fig:arrayConfig}).
+Splitting the core into three sections triples the density of instantaneous sampling of the $uv$-plane, improving HERA's mapmaking ability.Furthermore, the core is supplemented by 30 additional outrigger elements to tile the $uv$-plane with instantaneously complete sub-aperture sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz).
+This sub-aperture sampling strategy suppresses grating lobes in the instrument's point spread function and provides information for calibrating and correcting direction-dependent antenna responses \citep{dillon_parsons2016}.
+Even with the sub-aperture dithering and long baselines, the design is sufficiently redundant to take advantage of redundant baseline calibration technique and robust to the failure of individual elements.
+Resulting calibration errors range from $\sim 5\%$ (in the core) to $\sim 10\%$ (for outriggers) of the residual fractional noise per antenna after averaging. 
+The split core only increases the gain errors on the core elements by another $\sim 0.1\%$ from a solid hexagonal configuration.
+HERA's compact and redundant design allows it to fully utilize PAPER's robust approach to calibration and foreground mitigation, but it also provides for excellent imaging capability that can be leveraged to suppress foregrounds and improve access to the 21\,cm reionization signal.
 
@@
 	\vspace{-5pt}
-	\includegraphics[width=1\textwidth,clip]{plots/HERA_Array_Config.eps}
+	\includegraphics[width=1\textwidth,clip]{plots/HERA_Array_Config.pdf}
 	\vspace{-25pt}
 	\caption{HERA's elements are divided between a 320-element, hexagonally-packed core and 30 outriggers (left). This produces instantaneous
-$uv$ coverage at triple the element spacing out to 250$\lambda$ at 150\,MHz (middle). All 350 elements can be redundantly calibrated 
+$uv$ coverage at triple the element packing out to 250$\lambda$ at 150\,MHz, supressing grating lobes in the synthesized beam (middle). All 350 elements can be redundantly calibrated 
 using the \citet{liu_et_al2010} technique, yielding calibration errors that are a small fraction of the residual noise per antenna (right).}

### 2016-04-14 manuscript.tex
Updates to mapmaking, hybrid approaches, and calibration

@@
 PAPER and the MWA have produced progressively deeper limits
-\citep{dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
+\citep{dillon_et_al2014,dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
 yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
-(Fig. \ref{fig:limits}).
+(Fig. \ref{fig:limits}). %XXX: This figure doesn't really say much about spin temperature.
 The inherent challenge is to simultaneously meet stringent sensitivity requirements 
@@
 	\includegraphics[width=.8\textwidth]{plots/summary_results.png}
-	\caption{The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, (solid symbols represent analyses led by HERA collaborators) compared to 21cmFAST-generated models at $k=0.2$\,$h$\,Mpc$^{-1}$.
+	\caption{The current best published $2\sigma$ upper limits on the 21cm power spectrum, $\Delta^2(k)$, compared to a 21cmFAST-generated model at $k=0.2$\,$h$\,Mpc$^{-1}$.
 Analysis is still underway on PAPER and MWA observations that approach their projected full sensitivities;  HERA can deliver sub-$\text{mK}^2$ sensitivities.}
@@
 
-In these variables, the ``contaminated'' phase space is a wedge-shaped region in $\kpr$:$\kpar$ space such that the ratio of the time-delay across a given baseline ($b/c$,  which impacts the chromatic visilbility) and the delay associated with a given bandwidth ($1/\Delta\nu$) is less than some parameter determined by the details of the system, which we denote as $1/\beta$.
+In these variables, the ``contaminated'' phase space is a wedge-shaped region in $\kpr$:$\kpar$ space such that the ratio of the time-delay across a given baseline ($b/c$,  which impacts the chromatic visibility) and the delay associated with a given bandwidth ($1/\Delta\nu$) is less than some parameter determined by the details of the system, which we denote as $1/\beta$.
 Substituting $\kpr$ and $\kpar$ in for $b$ and $\Delta\nu$ in this ratio, we find that this wedge is bounded by
@@
 Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).
-In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_o=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.
+In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_0=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.
 Also note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
@@
 
-\subsection{Map-Making Approach}
+\subsection{Mapmaking Approach}
 \label{sec:mapapproach}
 In contrast to the delay-spectrum approach, map-making approaches combine baselines to build up information before squaring and calculating the power spectrum.
-The conceptually easiest approach is to first make an image cube of the sky as a function of angular position and frequency, and then to Fourier transform, square, and bin the cube to determine the power spectrum.
+The conceptually simpler approach is to first make an image cube of the sky as a function of angular position and frequency, and then to Fourier transform, square, and bin the cube to determine the power spectrum \cite{liu_tegmark2011,dillon_et_al2013a}.
+The image domain is a natural place to combine information from partially-coherent pairs of visibility measurements (either from different baselines or from the same baseline at different times).
+Mapmaking can be lossless data compression step \citep{tegmark1997b}, though keeping track of statistical properties of the maps, when tend to have complex frequency and position dependent point spread functions and noise covariance matrices, can be computationally challenging \citep{dillon_et_al2015a}.
+Existing approaches have had to make a number of approximations, including the approximation that point spread functions do not vary over the field of view and that noise is not correlated between $uv$-cells.
+
+%This approach is being pursued primarily via FHD.
+%XXX: this needs the be expanded.
+
 One advantage of such an approach is that sky images can be interesting in their own right, both for dealing with measurement systematics and for accessing the non-Gaussian observational signatures (such as ionized bubble structures) that are missed by variance statistics such as the power spectrum.
-The principal disadvantage of a map-making approach is its stringent calibration and forward-modeling requirements for propagating models of bright foregrounds through instrument systematics.
-In contrast, the delay-spectrum allows one to stay considerably closer to the format of raw visibility measurements, thus allowing systematic effects to be sequestered to localized regions of parameter space where they can much more easily be filtered out.
-
-This approach is being pursued primarily via FHD.
+Mapmaking also allows for the direct subtraction of bright foregrounds, potentially expanding the EoR window.
+However, accessing these high-order statistics or the modes of the power spectrum inside the wedge using a mapmaking approach requires extreme (and so-far undemonstrated) precision in calibration and forward-modeling of bright foreground models through instrument systematics.
 
@@
 
-More generally, it is possible to express a map-making algorithm as a linear operator that acts on visibility data to produce a compressed dataset \citep{dillon_et_al2015a}.
-Squaring this to obtain an estimate of the power spectrum then results in a weighted quadratic combination of visibilities, where one multiplies the visibility from every baseline with that from every other baseline.
+More generally, it is possible to express a mapmaking algorithm as a linear operator that acts on visibility data to produce a compressed dataset \citep{dillon_et_al2015a}.
+It follows then that squaring this to obtain an estimate of the power spectrum then results in a weighted quadratic combination of visibilities, where one multiplies the visibility from every baseline with that from every other baseline.
 These visibility product pairs are subsequently normalized to form their own individual estimates of the power spectrum before all the normalized pairs are summed together to form a final power spectrum.
 Measuring a power spectrum in this way combines the best aspects of the mapmaking and delay-spectrum approaches.
-Power spectra are formed directly by the cross-multiplication of visibility data, preserving the delay-spectrum approach's strategy of staying close to quantities measured by an interferometer; on the other hand, by multiplying together the data from every possible combination of baselines, one retains all the information captured by mapmaking.
+Power spectra are formed directly by the cross-multiplication of visibility data, preserving the delay-spectrum approach's strategy of staying close to quantities measured by an interferometer; on the other hand, by multiplying together the data from every possible combination of baselines, one retains all the information from partially-coherent visibility pairs captured by mapmaking.
 Indeed, it can be shown \citep{liu_et_al2014a} that in the limit of infinitely fine Fourier space bins, the hybrid method that we have just described is mathematically equivalent to the mapmaking approach.
 In practice, the hybrid method comes with the added benefit of allowing particular baseline pairs that are suspected of being affected by instrumental systematics to be more surgically downweighted.
+
+Fundamentally, the hybrid approach and the optimal mapmaking approach \citep{dillon_et_al2015a} are very similar.
+Whether it makes computational sense to examine all pairs of partially-coherent visibilities or to perform mapmaking as an intermediate data-compression step before power spectrum estimation (despite the complexity of the resultant map statistics) is an open question that depends on a number of factors, including the number of elements in the array, its layout, and its integration time.
 
@@
 As examined above, maximizing sensitivity to a limited number of spatial modes for the delay-spectrum technique requires a grid of antennas that simultaneously measures the same baselines using many redundant pairs of antennas.
-This redundancy may also be exploited to calibrate the array as well, as long as the true antenna positions fall close enough to that ideal grid.
+This redundancy may also be exploited to calibrate the array as well, as long as the true antenna positions fall close enough to that ideal grid \citep{liu_et_al2010}.
 This technique has been pioneered with the MITEoR experiment \citep{zheng_et_al2014}, and has resulted in a package called OMNICAL developed for PAPER and HERA.
@@
 
-Given that the technique requires redundant baselines
+Fundamentally, the problem of calibration is the determination of a single complex, frequency dependent gain for each antenna due to differences in amplifiers, cables, etc.
+Given those gains, the measured visibility measured by antennas $i$ and $j$ at frequency $\nu$ is given by
+\begin{equation}
+V_{ij}^\text{measured}(\nu) = g_i(\nu) g_j^*(\nu) V_{ij}^\text{true}(\nu), 
+\end{equation}
+where each $V_{ij}(\nu)$ is the measured or true visibility and $g_i(\nu)$ is the gain on antenna $i$ at frequency $\nu$.\footnote{This discussion ignores electromagnetic cross-talk between antennas and antenna-to-antenna variation of the primary beam.} With its highly-redundant array configuration (see \S\ref{sec:arrayConfig}), HERA's 350 antennas measure 61,075 visibilities, but only 6,140 unique baselines.
+Given a set of measured visibilities, it is possible estimate both the 6,140 unique true visibilities and the 350 complex gains simultaneously because that system of equations is greatly overdetermined.
+A method for linearizing the system and applying standard minimum-variance linear estimators was developed by \citet{liu_et_al2010} and first applied in \citet{zheng_et_al2014}.
+Unlike traditional calibration of radio interferometers, redundant calibration makes no reference to any sky model, requiring merely that the sky is bright relative to the noise on each antenna (which is always true in the sky-noise dominated regime characteristic of 21\,cm observatories with wide fields of view).
+
+However, at each frequency, degeneracies in the redundant calibration procedure prevent it from solving four numbers per frequency (down from 700 in HERA's case).
+The first is an overall gain---one can multiply each $g_i(\nu)$ by 2 and divide each $V_{ij}^\text{true}$ by 4, for example, and get exactly the same measured visibilities.
+The others are an overall phase and two other phase terms that describe the tip and tilt of the array.
+These degeneracies mean that the output of any redundant baseline calibration also needs a final, ``absolute'' calibration.
+Absolute bandpass calibration can be performed using a sky model, though it remains an open question whether this reduced set of calibration parameters can be further constrained without an explicit sky model.
 

### 2016-04-14 manuscript.tex
Minor tweaks and typo corrections to what I just (re)wrote.

@@
 \label{sec:mapapproach}
-In contrast to the delay-spectrum approach, map-making approaches combine baselines to build up information before squaring and calculating the power spectrum.
+In contrast to the delay-spectrum approach, mapmaking approaches combine baselines to build up information before squaring and calculating the power spectrum.
 The conceptually simpler approach is to first make an image cube of the sky as a function of angular position and frequency, and then to Fourier transform, square, and bin the cube to determine the power spectrum \cite{liu_tegmark2011,dillon_et_al2013a}.
 The image domain is a natural place to combine information from partially-coherent pairs of visibility measurements (either from different baselines or from the same baseline at different times).
-Mapmaking can be lossless data compression step \citep{tegmark1997b}, though keeping track of statistical properties of the maps, when tend to have complex frequency and position dependent point spread functions and noise covariance matrices, can be computationally challenging \citep{dillon_et_al2015a}.
-Existing approaches have had to make a number of approximations, including the approximation that point spread functions do not vary over the field of view and that noise is not correlated between $uv$-cells.
+Mapmaking can be lossless data compression step \citep{tegmark1997b}, though keeping track of statistical properties of the maps, which have complex frequency and position dependent point spread functions and noise covariance matrices, can be computationally challenging \citep{dillon_et_al2015a}.
+Existing approaches have had to make a number of approximations, including that point spread functions do not vary over the field of view and that noise is not correlated between $uv$-cells.
+Additionally, mapmaking is in principle not as vulnerable to polarization leakage as the delay-spectrum approach, since only polarization mismodeling (rather than any kind of polarization asymmetry) can cause leakage from Stokes $Q$ or $U$ to $I$.
 
@@
 This redundancy may also be exploited to calibrate the array as well, as long as the true antenna positions fall close enough to that ideal grid \citep{liu_et_al2010}.
-This technique has been pioneered with the MITEoR experiment \citep{zheng_et_al2014}, and has resulted in a package called OMNICAL developed for PAPER and HERA.
+This technique was pioneered with the MITEoR experiment \citep{zheng_et_al2014}, and has resulted in a package called OMNICAL developed for PAPER and HERA.
 \citet{ali_et_al2015} provides a description of this technique as it applies to calibrating the PAPER array, which is mostly situated on a grid with many redundant baselines.
@@
 
-Fundamentally, the problem of calibration is the determination of a single complex, frequency dependent gain for each antenna due to differences in amplifiers, cables, etc.
-Given those gains, the measured visibility measured by antennas $i$ and $j$ at frequency $\nu$ is given by
+Fundamentally, the problem of calibration is the determination of a complex, frequency-dependent gain for each antenna which arises due to differences in amplifiers, cables, etc.
+Given those gains, the visibility measured by antennas $i$ and $j$ at frequency $\nu$ is given by
 \begin{equation}
@@
 
-However, at each frequency, degeneracies in the redundant calibration procedure prevent it from solving four numbers per frequency (down from 700 in HERA's case).
+However, at each frequency, degeneracies in the redundant calibration procedure prevent it from solving four numbers per frequency for the whole array (down from 700 in HERA's case).
 The first is an overall gain---one can multiply each $g_i(\nu)$ by 2 and divide each $V_{ij}^\text{true}$ by 4, for example, and get exactly the same measured visibilities.
 The others are an overall phase and two other phase terms that describe the tip and tilt of the array.
-These degeneracies mean that the output of any redundant baseline calibration also needs a final, ``absolute'' calibration.
-Absolute bandpass calibration can be performed using a sky model, though it remains an open question whether this reduced set of calibration parameters can be further constrained without an explicit sky model.
+These degeneracies mean that the output of any redundant baseline calibration also need to undergo a final, ``absolute'' calibration.
+Absolute bandpass calibration has thus far been performed for PAPER and MITEoR using a sky model, though it remains an open question whether this reduced set of calibration parameters can be further constrained without an explicit sky model.
 

### 2016-04-14 manuscript.tex
some updated to figures per telecon

--- 2016-04-13 manuscript.tex
@@
+%XXX: Aaron, do one last check for new typos in my final series of edits.
 \documentclass[preprint,11pt]{aastex} 
@@
 
+%%% ADDED FROM 2014 PROPOSAL
 %\newcommand{\Mycite}[1]{{\bf \cite{#1}}}
@@
 	\centering
-	\includegraphics[width=4in]{plots/hera_render.png}
-	\includegraphics[width=4in]{plots/hera19H.png}
+	\includegraphics[height=1.75in]{plots/hera_render_logo.pdf}
+	\includegraphics[height=1.75in]{plots/hera19H.png}
 	\caption{Rendering of the 320-element core (left) of the full HERA-350 array and picture of 19 HERA 14-m, zenith-pointing dishes (with PAPER elements in the background) currently deployed in South Africa (right).}
@@
 \centerline{
-\includegraphics[width=0.5\textwidth]{plots/boost.pdf}
-\includegraphics[width=0.5\textwidth]{plots/sensratio.pdf}
+\includegraphics[height=6.2cm]{plots/boost.pdf}
+\includegraphics[height=6cm]{plots/sensratio.pdf}
 }
@@
 
-More generally, it is possible to express a map-making algorithm as a linear operator that acts on visibility data to produce a compressed dataset \citep{dillon_et_al2015a}.
+More generally, it is possible to express a map-making algorithm as a linear operator that acts on visibility data to produce a compressed dataset.
 Squaring this to obtain an estimate of the power spectrum then results in a weighted quadratic combination of visibilities, where one multiplies the visibility from every baseline with that from every other baseline.
@@
 \label{sec:calibration}
-As examined above, maximizing sensitivity to a limited number of spatial modes for the delay-spectrum technique requires a grid of antennas that simultaneously measures the same baselines using many redundant pairs of antennas.
-This redundancy may also be exploited to calibrate the array as well, as long as the true antenna positions fall close enough to that ideal grid.
-This technique has been pioneered with the MITEoR experiment \citep{zheng_et_al2014}, and has resulted in a package called OMNICAL developed for PAPER and HERA.
-\citet{ali_et_al2015} provides a description of this technique as it applies to calibrating the PAPER array, which is mostly situated on a grid with many redundant baselines.
+As examined above, for maximize sensitivity to a limited number of spatial modes using the delay-spectrum technique, a redundant grid greatly is advantageous.
+This redundancy may also be exploited to calibrate the array as well, if the baselines are sufficiently redundant.
+This technique has been pioneered with the MIT EoR experiment \citep{zheng_et_al2014}, and has resulted in a package called OMNICAL developed for PAPER and HERA .
+\cite{ali_et_al2015} provides a description of this technique as it applies to calibrating the PAPER array, which is mostly situated on a grid with many redundant baselines.
 This section provides a high-level overview of calibration as it applies to HERA.
@@
 
-As a focused experiment, HERA's configuration is optimized for the robust foreground-avoidance approach to EoR power spectrum estimation. 
-This requires a densely-packed core of antenna elements, maximizing sensitivity to short baselines---those least contaminated by foreground chromaticity (i.e.\ the wedge). 
-Furthermore, the configuration should be highly redundant, both to enable the delay-spectrum strategy of PAPER \citep{parsons_et_al2012a} and redundant baseline calibration \citep{liu_et_al2010,zheng_et_al2014}.
-On the other hand, it is also desirable to improve the mapmaking ability of the array for calibration and map-based power spectrum estimation with longer baselines and a densely-sampled $uv$-plane \citep{dillon_et_al2015a}.
-This tradeoff was closely examined by \citet{dillon_parsons2016} and the ultimate HERA configuration was drawn from one of the designs they considered.
-
-HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments to cover the $uv$-plane with sub-aperture sampling density (see Fig.~\ref{fig:arrayConfig}).
-Splitting the core into three sections triples the density of instantaneous sampling of the $uv$-plane, improving HERA's mapmaking ability.Furthermore, the core is supplemented by 30 additional outrigger elements to tile the $uv$-plane with instantaneously complete sub-aperture sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz).
-This sub-aperture sampling strategy suppresses grating lobes in the instrument's point spread function and provides information for calibrating and correcting direction-dependent antenna responses \citep{dillon_parsons2016}.
-Even with the sub-aperture dithering and long baselines, the design is sufficiently redundant to take advantage of redundant baseline calibration technique and robust to the failure of individual elements.
-Resulting calibration errors range from $\sim 5\%$ (in the core) to $\sim 10\%$ (for outriggers) of the residual fractional noise per antenna after averaging. 
-The split core only increases the gain errors on the core elements by another $\sim 0.1\%$ from a solid hexagonal configuration.
-HERA's compact and redundant design allows it to fully utilize PAPER's robust approach to calibration and foreground mitigation, but it also provides for excellent imaging capability that can be leveraged to suppress foregrounds and improve access to the 21\,cm reionization signal.
+As an experiment, HERA has optimized its configuration for the delay-spectrum approach to measure the EoR power spectrum.
+This means that it should be on short redundant baselines.
+Measurements conducted with PAPER showed that baselines of about 15m allow foreground reduction with real systematics and provide a proof of concept for this scale as well.
+The ability to fully calibrate the array is also a key consideration.
+The details of the process and adopted configuration may be found in \cite{dillon_parsons2016}.
+This section will briefly present the configuration.
+
+HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments 
+to cover the $uv$-plane with sub-element sampling density (see Fig.~\ref{fig:arrayConfig}). The dense core maximizes 
+sensitivity on the short baselines for which PAPER's proven foreground filtering strategy works best.
+The core is supplemented by 30 additional outrigger elements 
+to tile the $uv$-plane with instantaneously complete sub-aperture 
+sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz). This sampling 
+strategy suppresses grating lobes in the synthesized beam and provides information for calibrating and correcting 
+direction-dependent antenna responses \citep{dillon_parsons2016}.
+Even with the sub-aperture dithering and long baselines, all 350 HERA elements can be robustly calibrated by taking advantage of 
+its highly redundant configuration \citep{liu_et_al2010,zheng_et_al2014}. 
+Resulting calibration errors range from $\sim5\%$ (in the core) to $\sim10\%$ (for outriggers)
+of the residual fractional noise per antenna after averaging.
 
@@
 	\vspace{-5pt}
-	\includegraphics[width=1\textwidth,clip]{plots/HERA_Array_Config.pdf}
+	\includegraphics[width=1\textwidth,clip]{plots/HERA_Array_Config.eps}
 	\vspace{-25pt}
 	\caption{HERA's elements are divided between a 320-element, hexagonally-packed core and 30 outriggers (left). This produces instantaneous
-$uv$ coverage at triple the element packing out to 250$\lambda$ at 150\,MHz, supressing grating lobes in the synthesized beam (middle). All 350 elements can be redundantly calibrated 
+$uv$ coverage at triple the element spacing out to 250$\lambda$ at 150\,MHz (middle). All 350 elements can be redundantly calibrated 
 using the \citet{liu_et_al2010} technique, yielding calibration errors that are a small fraction of the residual noise per antenna (right).}

### 2016-04-14 plots/attendelayspec.py

@@ -1,6 +1,5 @@
 import numpy as np
 import matplotlib.pyplot as plt
-import os.path
 datafile = 'spec_on_foreground_reflected_power_21cmfast_14.6m_150.0_MHz_subband_v2.npz'
 
 npzdata = np.load(datafile)
@@ -19,8 +18,15 @@ a60 = [0,60,60]
 a0 = [0.0,0.0,0.0]
 plt.fill_between(t60,a0,a60,color='k',alpha='0.05')
 plt.plot(t60,a60)
-plt.axis([0,500,80,0])
+plt.axis([0,500,75,0])
 plt.xlabel(r'$\tau$ [ns]')
 plt.ylabel('Attenuation [dB]')
 plt.legend()
+
+delay = np.load('delay_spectrum.npz')
+#plt.plot(delay['tau'],delay['dpaper'])
+plt.plot(delay['tau'],-1.*delay['dhera'],'k',lw=2)
+#plt.figure('aaa')
+#plt.plot(delay['tau'],delay['dhera'])
+
 plt.savefig('delayspecplot.pdf',dpi=200)

### 2016-04-14 plots/boost.py

@@ -68,7 +68,7 @@ def plotPwr(p=2,title='-',ylabel='-'):
     plt.title(title)
     plt.xlabel('k [h/Mpc]',fontsize=14)
     plt.ylabel(ylabel,fontsize=14)
-    plt.axis(xmin=0.12,xmax=1.5)
+    plt.axis(xmin=0.09,xmax=1.5)
     #plt.legend()
 
 plotPwr(0,'Sensitivity Ratio',r'$\Delta_{19}^2/\Delta_x^2$')

### 2016-04-14 manuscript.tex
more figure updates

--- 2016-04-14 manuscript.tex
@@
 	\centering
-	\includegraphics[width=4in]{plots/hera_render.png}
-	\includegraphics[width=4in]{plots/hera19H.png}
+	\includegraphics[height=4.0in]{plots/hera_render_logo.pdf}
+	\includegraphics[height=4.0in]{plots/hera19H.png}
 	\caption{Rendering of the 320-element core (left) of the full HERA-350 array and picture of 19 HERA 14-m, zenith-pointing dishes (with PAPER elements in the background) currently deployed in South Africa (right).}
@@
 \centerline{
-\includegraphics[width=0.5\textwidth]{plots/boost.pdf}
-\includegraphics[width=0.5\textwidth]{plots/sensratio.pdf}
+\includegraphics[height=6cm]{plots/boost.pdf}
+\includegraphics[height=6cm]{plots/sensratio.pdf}
+\includegraphics[height=5cm]{plots/elementSizes.pdf}
 }
@@
 \centerline{
-\includegraphics[width=.4\textwidth]{plots/kperf.png} 
+\includegraphics[width=.4\textwidth]{plots/kperf.pdf} 
 }
@@
 	\vspace{-25pt}
-	\caption{Left:  Antenna delay attenuation specification with realistic foregrounds from \cite{thyagarajan_et_al2016} with measurements from Green Bank prototype \citep{patra_et_al2016}.
-Also shown is the initial aggressive ``60@60'' specification.
-       agree with field measurements (grey; \citealt{patra_et_al2016}). 
+	\caption{Left:  Antenna delay attenuation specification with realistic foregrounds from \cite{thyagarajan_et_al2016} along with the initial conservative ``60@60'' specification.
+HFSS calculations (black line) show HERA meeting the specification.
+Measurements on the Green Bank prototype are shown in  \citep{patra_et_al2016}.
 Right: Modeled foreground power spectra with (green) and without (blue) the antenna chromaticism over a 10\,MHz subband centered at $z=7.5$.

### 2016-04-14 plots/kperf.py

@@ -9,46 +9,56 @@ n = nedclass.ned(H0=100.0)
 
 def get_dz(z,B,rest=1420.):
     return B*(1+z)**2/rest
+def just_z_stuff(z,restFreq,b):  #just pass one z
+    #X = []  # X Mpc/rad
+    #Xb = [] # scaled version
+    #Xinstr = []  # observed extent in Mpc
+    #Y = []  # Y Mpc/GHz
+    #DA = [] # angular scale term
+    #E = []  # Hubble parameter evolution
+    DAi = n.calcUniverse(z)
+    #DA.append(DAi)
+    Xi = DAi*(1.+z)
+    obsFreq = restFreq/(1.+z)
+    obsWavelength = 0.3/obsFreq
+    Xbi = Xi/((180.0/math.pi)*1.0)  #convert to degrees
+    #X.append(Xi)
+    #Xb.append(Xbi)
+    Xbi = Xi*coeff*obsWavelength/b
+    #Xinstr.append(Xbi)
+    #E.append(n.Ez)
+    Yi = ( n.Ynu )/restFreq
+    #Y.append(Yi)
+    #Y = np.array(Y)
+    #Xinstr = np.array(Xinstr)
+    return Xbi, Yi
 
-def mkPlot(restFreq,BW,b,coeff):
-    X = []  # X Mpc/rad
-    Xb = [] # scaled version
-    Xinstr = []  # observed extent in Mpc
-    Y = []  # Y Mpc/GHz
-    DA = [] # angular scale term
-    E = []  # Hubble parameter evolution
-    #Below are just used to get a power-law fit and then plot it post facto
-    apX = []    # approximated X (post-fit numbers put back in)
-    apY = []    # approximated Y ( " )
-    z2fit = []  # range of z's to fit for approximation
-    X2fit = []  #     "    X's   "
-    Y2fit = []  #     "    Y's   "
-
+def mkPlot(restFreq,BW,b,coeff,z=None):
     # Loop over red-shift
-    z = np.arange(0.1,25.0,.1)
-    for zi in z:
-        DAi = n.calcUniverse(zi)
-        DA.append(DAi)
-        Xi = DAi*(1.+zi)
-        obsFreq = restFreq/(1.+zi)
-        obsWavelength = 0.3/obsFreq
[…]

### 2016-04-14 plots/kperf.py
last tweak to kperf.pdf

@@ -76,6 +76,8 @@ plt.semilogy(zhera,khera2,'w')
 plt.fill_between(zhera,khera1,khera2,color='w',alpha='0.65')
 
 #first part (k_tot)
+zstart = 5.0
+zstop = 25.0
 BW_list = [0.02,.002]
 D_list = [14.6,25.3]
 #txtloc = [[9,.02,-4],[7,.32,-5]]  #wide
@@ -83,15 +85,20 @@ txtloc = [[9.3,.02,-7],[6.0,.34,-10]]  #narrow
 lsc = ['g-','g-']
 tclr = ['g','g']
 for i,BW in enumerate(BW_list):
-    z,kperp,kper,ktot = mkPlot(restFreq,BW,D_list[i],coeff)
+    z = [zstart]
+    while z[-1]<zstop:
+        newz = z[-1] + get_dz(z[-1],BW,restFreq)
+        z.append(newz)
+    z,kperp,kper,ktot = mkPlot(restFreq,BW,D_list[i],coeff,z)
     if i==0:
+        zsave = z
         savek = ktot
     show_label='%.0f MHz, %.1fm' % (1000.*BW,D_list[i])
     lbl = r'${\bf |k|}$ @%s' % (show_label)
     plt.semilogy(z,ktot,lsc[i],label=show_label,lw=2) 
     plt.text(txtloc[i][0],txtloc[i][1],lbl,rotation=txtloc[i][2],color=tclr[i],fontsize=16)
-ktottop = 10.*np.ones(len(z))
-plt.fill_between(z,savek,ktottop,color='g',alpha='0.25')
+ktottop = 10.*np.ones(len(zsave))
+plt.fill_between(zsave,savek,ktottop,color='g',alpha='0.25')
 plt.fill_between(zhera,khera1,khera2,color='w',alpha='0.65')  ###AND REDO LINE 63
 
 #second part  (k_perp)
@@ -112,7 +119,7 @@ for i,D in enumerate(D_list):
 D = 14.6
 BW_list = [0.1,0.01,.001]
 lsc = ['b--','b-','b-']
-txtloc = [[6.0,4.6e-3,-9],[6,.07,-11],[15.3,.475,-5]]
+txtloc = [[7.6,3.5e-3,0],[6,.07,-11],[15.3,.475,-5]]
 BW_show = ['100MHz','10MHz', '1MHz']
 zstart = 5.0
 zstop = 25.0

### 2016-04-14 manuscript.tex
sauls sec 1 comments

@@
 \section{Introduction}
-
-\noindent The Hydrogen Epoch of Reionization Array (HERA) is a staged experiment to use the redshifted 21\,cm line
+The Hydrogen Epoch of Reionization Array (HERA) is a staged experiment to use the redshifted 21\,cm line
 of neutral hydrogen to characterize our Cosmic Dawn, from
@@
 understanding the astrophysics and fundamental cosmology of our early universe.
-Taking advantage of our new understanding of bright foreground systematics, HERA's
+Taking advantage of a new understanding of bright foreground systematics, HERA's
 purpose-built radio
 interferometer is optimized to deliver high signal-to-noise measurements of
-redshifted 21\,cm emission.
-With the first generation 21\,cm telescopes approaching the limits of their sensitivity, HERA continues the experiment by bringing significantly more sensitivity to bear
+redshifted 21\,cm emission to detect and characterize the Epoch of Reionization (EOR).
+
+Because of these key science questions that may be probed by highly redshifted hydrogen, there has been a resurgence in low frequency arrays  -- most with a primary
+objective of measuring the EOR.
+These include the Precision Array Probing the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}), the Murchison Widefield Array (MWA; \citealt{tingay_et_al2013}), the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V}), and the Long Wavelength Array (LWA; \citealt{2009IEEEP..97.1421E}), as well as systems for existing dish arrays like the Very Large Array (VLA; \citealt{2013AAS...22132804K}) and the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2011}).
+With the first generation 21\,cm telescopes approaching the limits of their sensitivity however, HERA continues the experiment by bringing significantly more sensitivity to bear
 while continuing to optimize its design along the delay-spectrum approach to detect the power spectrum of the EoR (discussed in Sec. \ref{sec:delayapproach}).
-Key science results include
-\begin{itemize}[noitemsep,nolistsep]
-\item measuring the 21\,cm power spectrum with high significance (\S\ref{sec:EoRPowerSpectra}),
-\item characterizing parametrized models of reionization and the first ionizing sources (\S\ref{sec:EoRPowerSpectra}),
-\item improving cosmic microwave background (CMB) constraints on primordial density fluctuations and neutrino masses by removing the optical depth degeneracy (\S\ref{sec:tau}), and
-\item directly imaging large scale reionization structure (\S\ref{sec:imaging}).
-\end{itemize}
+At the same time, HERA continues the development of new techniques to extend and improve its capabilities.
+Examples include improving foreground subtraction to expand the information content 
+in power spectral and imaging measurements and
+extending HERA's performance at lower frequencies to probe IGM heating in the pre-reionization epoch (\S\ref{sec:EoX}).
 
@@
 	\centering
-	\includegraphics[height=4.0in]{plots/hera_render_logo.pdf}
-	\includegraphics[height=4.0in]{plots/hera19H.png}
+	\includegraphics[height=2.0in]{plots/hera_render_logo.pdf}
+	\includegraphics[height=2.0in]{plots/hera19H.png}
 	\caption{Rendering of the 320-element core (left) of the full HERA-350 array and picture of 19 HERA 14-m, zenith-pointing dishes (with PAPER elements in the background) currently deployed in South Africa (right).}
@@
 \end{figure}
-
-At the same time, HERA continues the development of new techniques to extend and improve its capabilities.
-Examples include improving foreground subtraction to expand the information content 
-in power spectral and imaging measurements and, with support from the Gordon Moore Foundation [[FOOTNOTE?]],
-extending HERA's performance at lower frequencies to probe IGM heating in the pre-reionization epoch (\S\ref{sec:EoX}).
-Beyond these directly supported applications, HERA's public data products and instrumentation platform enable other high-impact science, including 
-\begin{itemize}[noitemsep,nolistsep]
-\item 21\,cm cross-correlation (e.g. with JWST, WFIRST, CO, CII, and Ly$-\alpha$; \S\ref{sec:crossCorr}),
-\item searching for auroral radio bursts from exoplanetary magnetospheres (\S\ref{sec:exoplanets}), and
-\item performing triggered, low-frequency follow up and characterization of fast radio bursts (\S\ref{sec:FRBs}).
-\end{itemize}
 
@@
 The 37 elements, to be completed in 2016, will provide nearly an order of magnitude more
-sensitivity than the 128-element PAPER array and provide a significant chance at detecting the EoR power spectrum signal.
+sensitivity than current instruments and provide a significant chance at detecting the EoR power spectrum signal.
 The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER \citep{ali_et_al2015}, which should provide a very robust detection of the EOR signal and allow some characterization.
@@
 Section \ref{sec:science} will give a high-level overview of the science motivating HERA's construction, including primary and secondary science objectives.
+Key science results include
+measuring the 21\,cm power spectrum with high significance (\S\ref{sec:EoRPowerSpectra});
+characterizing parametrized models of reionization and the first ionizing sources (\S\ref{sec:EoRPowerSpectra});
+improving cosmic microwave background (CMB) constraints on primordial density fluctuations and neutrino masses by removing the optical depth degeneracy (\S\ref{sec:tau}); and
+directly imaging large scale reionization structure (\S\ref{sec:imaging}).
+Additionally, beyond these directly supported applications, HERA's public data products and instrumentation platform enable other high-impact science, including 
+21\,cm cross-correlation (e.g. with JWST, WFIRST, CO, CII, and Ly$-\alpha$; \S\ref{sec:crossCorr});
+searching for auroral radio bursts from exoplanetary magnetospheres (\S\ref{sec:exoplanets}); and
+performing triggered, low-frequency follow up and characterization of fast radio bursts (\S\ref{sec:FRBs}).
+
 Section \ref{sec:eormeas} will present the techniques used to make the measurement.
@@
 \includegraphics[height=6cm]{plots/sensratio.pdf}
-\includegraphics[height=5cm]{plots/elementSizes.pdf}
+\includegraphics[height=4cm]{plots/elementSizes.pdf}
 }

### 2016-04-16 manuscript.tex
updated figures

@@
 \ref{fig:cosmos}, starting with the Big Bang on the left.
-The hot, young universe expands and cools slowly in the background while gravitational instability causes primordial density fluctuations to grow.
+The hot, young universe expands and cools slowly in the background while gravitational instability around concentrations of dark matter causes primordial density fluctuations to grow.
 The Cosmic Dawn represents a specific epoch in this growth, where the first stars and galaxies formed and illuminated the Universe en route to forming the astronomical structures we see today.
 Ultimately, this early population of gravitationally condensed material produced sufficiently energetic flux to reionize the intergalactic medium (IGM) from its previous neutral state in a period called the Epoch of Reionization.
-This period is shown in Fig \ref{fig:cosmos} as the rapid transition from the big bubble structures to structures that begin to resemble the denizens of our current universe.
+This period is shown in Fig \ref{fig:cosmos} as the rapid transition from separated large reionized ``bubbles'' to the merged reionized state and to structures that begin to resemble the denizens of our current universe.
 The structure of the
-intergalactic medium (IGM) thus contains a panoply of information about the underlying astrophysical and cosmological phenomena governing cosmic evolution.
-
-The evolution of these structures depends on the cosmic density field, the relative velocities of
-baryons and dark matter, and the sizes and clustering of the first galaxies.
+IGM thus contains a panoply of information about the underlying astrophysical and cosmological phenomena governing cosmic evolution.
+
+The evolution of the cosmic structure depends on the local and average cosmic density, the relative velocities of
+baryons and dark matter, and the sizes and clustering of the first galaxies to form.
 But it also depends on the constituents of those first galaxies---so-called Population
@@
 Exploring the interplay of galaxies and large-scale structure during the EOR
-requires complementary observational approaches. Measurements of the Cosmic Microwave Background (CMB; the escaping radiation permeating the universe after the universe became
-transparent to its radiation by the recombination of the protons and electrons about 400,000 years after the Big Bang)  provide initial conditions for structure formation, and the Thomson
-scattering of CMB photons constrains the integrated column of 
-ionized gas, but even with kinetic Sunyaev-Zel'dovich measurements, 
-the detailed evolution of the IGM 
+requires complementary observational approaches. Measurements of the Cosmic Microwave Background 
+(CMB; the escaping radiation permeating the universe after the universe became
+transparent to its radiation by the recombination of the protons and electrons about 400,000 years after the Big Bang)  by COBE, WMAP and Planck provide initial conditions for structure formation.
+Thomson
+scattering of CMB photons by the ionized particles constrains the integrated column of 
+ionized gas and kinetic Sunyaev-Zel'dovich measurements constrain the power spectrum of cosmic structures.
+But even with these measurements, the detailed evolution of the IGM 
 is only loosely constrained \citep{haiman_holder2003,mortonson_hu2008,zahn_etal2012,mesinger_et_al2012}.
@@
 saturate at low neutral fractions
-($x_{\rm HI} \la 10^{-4}$). Measurements of galaxy populations in
+($x_{\rm HI} \la 10^{-4}$), where $x_{\rm HI}$ is the ratio of neutral hydrogen to the total baryonic density.
+
+Measurements of galaxy populations in
 deep {\it Hubble Space Telescope} observations 
@@
 
-\noindent HERA's top science goal is to transform our understanding of the first stars, galaxies, 
+HERA's primary science goal is to transform our understanding of the first stars, galaxies, 
 and black holes, and their role in driving reionization. 
@@
 Figure~\ref{fig:Sensitivities} compares telescope sensitivities as a function of redshift to models of 
-the evolving, spherically averaged 21\,cm EoR power spectrum $\Delta^2 (k) \equiv k^3 P(k) / 2 \pi^2$ (discussed later).  In contrast to current experiments,
-HERA's optimized design and collecting area make it capable of
-high-significance detections of virtually any realistic reionization scenario, even using only currently demonstrated foreground
-mitigation strategies (see Table~\ref{tab:signif}). The development of enhanced foreground modeling and subtraction further improve the achievable results.
+the evolving, spherically averaged 21\,cm EoR power spectrum, characterized by the dimensionless power spectrum parameter $\Delta^2 (k) \equiv k^3 P(k) / 2 \pi^2$.
+The expected performance of HERA relative to current and planned telescopes to detect the peak of reionization (as well as the total collecting area is shown in 
+Table~\ref{tab:signif}.
+
+%In contrast to current experiments,
+%HERA's optimized design and collecting area make it capable of
+%high-significance detections of virtually any realistic reionization scenario, even using only currently demonstrated foreground
+%mitigation strategies (see Table~\ref{tab:signif}). The development of enhanced foreground modeling and subtraction further improve the achievable results.
 

### 2016-04-16 plots/attendelayspec.py

@@ -25,7 +25,7 @@ plt.legend()
 
 delay = np.load('delay_spectrum.npz')
 #plt.plot(delay['tau'],delay['dpaper'])
-plt.plot(delay['tau'],-1.*delay['dhera'],'k',lw=2)
+plt.plot(delay['tau_hera'],-1.*delay['dhera'],'k',lw=2)
 #plt.figure('aaa')
 #plt.plot(delay['tau'],delay['dhera'])
 

### 2016-04-16 plots/kperf.py

@@ -78,13 +78,15 @@ plt.fill_between(zhera,khera1,khera2,color='w',alpha='0.65')
 #first part (k_tot)
 zstart = 5.0
 zstop = 25.0
-BW_list = [0.02,.002]
-D_list = [14.6,25.3]
+BW_list = [0.02,0.01,.002]
+D_list = [14.6,25.3,25.3]
 #txtloc = [[9,.02,-4],[7,.32,-5]]  #wide
-txtloc = [[9.3,.02,-7],[6.0,.34,-10]]  #narrow
-lsc = ['g-','g-']
-tclr = ['g','g']
-for i,BW in enumerate(BW_list):
+txtloc = [[9.3,.02,-7],[10.3,.02,-7],[6.0,.34,-10]]  #narrow
+lsc = ['g-','r-','g-']
+tclr = ['g','r','g']
+use_i = [0,2]
+for i in use_i:
+    BW = BW_list[i]
     z = [zstart]
     while z[-1]<zstop:
         newz = z[-1] + get_dz(z[-1],BW,restFreq)
@@ -158,7 +160,7 @@ if draw_z_bins:
 
 
 plt.xlabel('Redshift [z]',fontsize=16)
-plt.ylabel('k(res) [h/Mpc]',fontsize=16)
+plt.ylabel('k [h/Mpc]',fontsize=16)
 #plt.legend(loc='upper right')
 plt.grid()
 plt.axis([restFreq/0.220 - 1.0,restFreq/0.065-1.0,2e-3,1])

### 2016-04-18 manuscript.tex
rest of sauls edits

@@
 the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V, yatawatta_et_al2013}),
-faces large
-uncertainties [[[change this?]]], 
-HERA's tight constraints on these parameters are comparable
-to what could be achieved with the SKA.  Such results would be valuable for
-informing the design and science case of the low-frequency SKA, highlighting areas where the SKA has unique capability.
+provides some ability to constrain these parameters,
+HERA's constraints are significantly more precise and are comparable
+to what could be achieved with the SKA. 
 Additionally, HERA's constraints enable principal component parameterizations of the
@@
 \label{sec:imaging}
-In addition to measuring the power spectrum, if systematics are kept sufficiently low HERA has sufficient 
-sensitivity to directly image the IGM during reionization,
-and will observe a 1440 deg$^2$ stripe, comparable to future WFIRST
+In addition to measuring the power spectrum, there is the potential for HERA  
+ to directly image the IGM during reionization
+over the 1440 deg$^2$ stripe that transits overhead, which is comparable to future WFIRST
 large area near-IR surveys.
- After 100 hours on a
-single field (achievable in 200 nights), HERA reaches a surface brightness 
+ After 100 hours on a single field (achievable in 200 nights), HERA reaches a surface brightness 
 sensitivity of 50 $\mu$Jy/beam (synthesized beam FWHM $\sim$ 24$'$) compared 
@@
 (e.g. redundant calibration, \citealt{liu_et_al2010,zheng_et_al2014}), while retaining a high quality 
-point spread functions for imaging and identifying foregrounds. Leveraging the expertise gained 
-through both precursor strategies, HERA will yield exceptionally high dynamic range
-images.
+point spread functions for imaging and identifying foregrounds.
 
@@
 
-While the observational and analytical state of the art of pre-reionization 21\,cm science is not as advanced as for the EoR \citep{ewall-wice_et_al2016-EoXLimits}, with support from the Moore Foundation to develop feeds sensitive down to 50\,MHz, HERA will explore the IGM prior to
+While the observational and analytical state of the art of pre-reionization 21\,cm science is not as advanced as for the EoR \citep{ewall-wice_et_al2016-EoXLimits}, 
+with the development of feeds sensitive down to 50\,MHz, HERA will explore the IGM prior to
 reionization with a goal to constrain the sources of heating, obtaining, in the case of
@@
 
-\noindent Given its high sensitivity, frequency support, filled aperture and outriggers, HERA is capable of delivering more than the core 21\,cm science discussed above.
+Given its high sensitivity, frequency support, filled aperture and outriggers, HERA is capable of delivering more than the core 21\,cm science discussed above.
 HERA will make the data available to the community after only a short quality assurance time period to help facilitate low-frequency science.
@@
 galaxies---lies in HERA's field of view. HERA's images of the IGM can provide environmental context to galaxy surveys through identification of
-ionized bubbles \citep{malloy_lidz2013}, or in a more statistical sense as
+ionized bubbles ({\em e.g.} \citealt{malloy_lidz2013}), or in a more statistical sense as
 described in \cite{beardsley_et_al2015}.
@@
 Fast Radio Bursts (FRB) are millisecond-long radio flashes whose origin has remained a great enigma ever since their discovery \citep{2007Sci...318..777L}. 
-If equipped with a suitable backend, HERA could be triggered by nearby, higher-frequency telescopes for FRB followup, saving baseband data and thus full sensitivity to all dispersion measures.
+HERA could be triggered by nearby, higher-frequency telescopes for FRB followup, saving baseband data and thus full sensitivity to all dispersion measures.
+This would require additional hardware, but the architecture allows for such additions.
 Bursts like that discovered by \citet{masui_et_al2015}, the lowest frequency
@@
 
-As the distance to the EOR is great (about 13 billion light-years) the signal is quite weak.
+As the distance to the EOR is great (about 13 billion light-years) the signal is weak.
 However, the signal also subtends the entire sky, so initial measurements of the EOR strive to measure a statistical power spectrum of the signal over the sky since the nature of the reionization process should have a specific spatial signature.
@@
 calibration ({\em e.g.} \cite{zheng_et_al2014, jacobs_et_al2016, barry_et_al2016}), handling foreground contamination
-({\em e.g.} \cite{moore_et_al2013,moore_et_al2016,thyagarajan_et_al2015a,pober_et_al2016}),
+({\em e.g.} \cite{moore_et_al2013,thyagarajan_et_al2015a,pober_et_al2016}),
 and the interferometer design itself ({\em e.g.} \cite{parsons_et_al2012a,dillon_parsons2016}).
@@
 leave it with limited diagnostic capability, particularly for direction-dependent systematics
-such as polarization leakage from Faraday-rotated emission \Mycitep{moore_et_al2013}.
+such as polarization leakage from Faraday-rotated emission \Mycitep{moore_et_al2013, nunhokee_etal_2016}.
 While concern over such effects has decreased markedly since discovering that
@@
 ionosphere averages down the polarized signal over long observing seasons
-\Mycitep{moore_et_al2016}, direction-dependent beam effects remain an area of interest.
+\Mycitep{moore_et_al2016,aguirre_etal_2016}, direction-dependent beam effects remain an area of interest.
 The MWA's image-based calibration and foreground subtraction strategy provides complementary
@@
 foreground leakage in Fourier space (see Fig.~\ref{fig:wedge}), 
-outside of which the 21\,cm signal dominates (the ``EoR window'').  The coordinates of the wedge are discussed below.
+outside of which the 21\,cm signal dominates (the ``EOR window'').  The coordinates of the wedge are discussed below.
 Through theoretical and observational work
@@
 In these variables, the ``contaminated'' phase space is a wedge-shaped region in $\kpr$:$\kpar$ space such that the ratio of the time-delay across a given baseline ($b/c$,  which impacts the chromatic visibility) and the delay associated with a given bandwidth ($1/\Delta\nu$) is less than some parameter determined by the details of the system, which we denote as $1/\beta$.
-Substituting $\kpr$ and $\kpar$ in for $b$ and $\Delta\nu$ in this ratio, we find that this wedge is bounded by
+Substituting $\kpr$ and $\kpar$ in for $b$ and $\Delta\nu$ in this ratio, this wedge is bounded by
 \begin{equation}
@@
 \end{equation}
-where we've included an offset parameterized by $S$ accounting for affects related to the combined spectral smoothness of the foregrounds and the antenna response (see Fig. \ref{fig:wedge}).
-
-If this $\beta$ factor is determined by the chromaticity of the longest baseline, we see that $\beta=1$ and we call this line the ``horizon line'' ({\em i.e.}, the delay of a source at the horizon).
+where an offset parameterized by $S$ accounting for affects related to the combined spectral smoothness of the foregrounds and the antenna response has been included(see Fig. \ref{fig:wedge}).
+
+If this $\beta$ factor is determined by the chromaticity of the longest baseline, we see that $\beta=1$ and this line is referred to as the ``horizon line'' ({\em i.e.}, the delay of a source at the horizon).
 Systematics will push that line such as to close the EOR ``window'' to the upper left ($\beta > 1$).
-If we can completely control systematics and constrain the affects of the foreground to {\em e.g.} the main beam, the line will move to decrease the size of the ``wedge'' ($\beta<1$).
+If we can completely control systematics and constrain the affects of the foreground to {\em e.g.} the main beam, the line will move to decrease the size of the wedge ($\beta<1$).
 Therefore, a key issue to measuring the EOR power spectrum is to understand and minimize systematics.
@@
 
-The techniques to measure the power spectrum may be broken down into two principle techniques (delay-space and map-making), plus additional hydrid methods.
+The techniques to measure the power spectrum may be broken down into two principle techniques -- delay-space (\ref{sec:delayapproach}) and map-making (\ref{sec:mapapproach}) -- plus additional hydrid methods.
 These are briefly summarized below, with an emphasis on the delay-spectrum technique as the initial approach taken by HERA.
@@
 Given that the response of an interferometer natively measures the power in Fourier modes of the sky within its beam, we see that it is a natural instrument to use for this measurement of the EOR spatial power spectrum.
-There are now numerous papers in the literature on the technique of measuring the power spectrum using an interferometer and the attendant issues as addressed in nearly all of the cited references.
-Below, a high-level overview based on  \cite{parsons_et_al2012b} will be presented.
-The so-called delay-spectrum approach leverages the interferometer measurement to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
+The delay-spectrum approach leverages the interferometer measurement to optimize sensitivity to the desired modes while rejecting modes contaminated by the foreground power in the wedge.
 Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum, which contrasts to the map-making techniques briefly described in the following section.
@@
 \end{equation}
-where $\tau$ is the integration time (see {\em e.g.} TMS),
+where $\tau$ is the integration time (see {\em e.g.} \citealt{thompson_et_al2001}),
 to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq.
@@
 
-Figure \ref{fig:kperf} is a rather busy plot showing many of the dependencies on wavenumber of bandwidth, configuration and the cosmos.
+Figure \ref{fig:kperf} shows many of the dependencies on wavenumber of bandwidth, configuration and the cosmos.
 It shows the perpendicular, parallel and total resolution wavenumber as a function of redshift for various bandwidths and baselines.
@@
 Also in Fig.
-\ref{fig:kperf}, the lime-green shaded region indicates a plausible region in which HERA could operate and the lighter square within that region is the area that is being targeted in the first instance.
+\ref{fig:kperf}, the lime-green shaded region indicates a plausible region in which HERA could operate and the lighter square within that region is the area that is being targeted in the first phase.
 
@@
 }
-\caption{\small $\kvec, \kpar, \kperp$ resolution wavenumbers with various bandwidths and baselines as a function of redshift.
+\caption{\small $|\kvec |, \kpar, \kperp$ wavenumber plots at various bandwidths ($k_{||}$; blue), baselines ($\kperp$; black), and total ($|\kvec |$, green) as a function of redshift.
+The stepped profiles for $k_{||}$ and $|\kvec |$ are the line-of-site redshift ranges for that bandwidth (too small to see at 1 MHz).
 See text for discussion.
 The full redshift range is the HERA extended goal.
-The initial HERA delay-spectrum power spectrum goal is the light square region in the upper right.}
+The initial HERA delay-spectrum power spectrum goal is the light square region in the upper left.}
 \label{fig:kperf}
@@
 Whether it makes computational sense to examine all pairs of partially-coherent visibilities or to perform mapmaking as an intermediate data-compression step before power spectrum estimation (despite the complexity of the resultant map statistics) is an open question that depends on a number of factors, including the number of elements in the array, its layout, and its integration time.
+These approaches need further development, and can be incorporated when sufficiently implemented.
 
@@
 The cost-performance optimization will
-be discussed below, however we see that to address the central challenges of the delay-spectrum approach in the context of the wedge, to open the widest possible window for EoR measurements, 
+be discussed below, however we see that to address the central challenges of the delay-spectrum approach in the context of the wedge, to open the widest possible window for EOR measurements, 
 HERA must use close-packed antennas that
@@
 This provided the basis for the initial design of the element.
-For the contemplated feeds for a primary focus antenna, it was found that a focal length/diameter radio ($f/D$ of approximately 0.32 was optimal \citep{elementmemo}.
+For the contemplated feeds for a primary focus antenna, it was found that a focal length/diameter radio ($f/D$ of approximately 0.32 was optimal.
 We may therefore estimate the attenuation as a function of delay and diameter and assign this a ``cost function''.
@@
 
-To get 60 dB at 60 ns we may interpret those lines in one of two ways:  (1) set the spec on how much attenuation per reflection is needed and (2) at a fixed attenuation/reflection, what diameter we need.
+To obtain 60 dB at 60 ns we may interpret those lines in one of two ways:  (1) set the specification on how much attenuation per reflection is needed and (2) at a fixed attenuation/reflection, what diameter is required.
 For 1 round-trip, for the assumed $f/D$ over the diameters of interest, all delays are less than 60 ns, so it provides no constraint.
@@
 For details of the first generation HERA system (the existing and commissioned PAPER signal path), see \cite{parsons_et_al2010}.
-This paper will discuss the new system to be deployed beginning and 2016 and tested in conjunction with the previous system.
+This section will discuss the new system to be deployed beginning and 2016 and tested in conjunction with the previous system.
 Fig \ref{fig:overallBlockDiagram} provides a system block diagram of the new HERA architecture.
@@
 This real-time pipeline, based on a similar system for PAPER, is mature, computationally manageable, and 
-relatively agnostic with respect to observing parameters, making it a safe and robust first step in data processing
-\citep{zheng_et_al2014,parsons_etal2014,ali_et_al2015}.
+agnostic with respect to observing parameters, making it a safe and robust first step in data processing.  For a discussion
+of the PAPER pipeline, see
+\cite{zheng_et_al2014,parsons_etal2014,ali_et_al2015}.
 The Internet link between the Karoo and data centers in the U.S.\ has been extensively tested, with assistance from SKA-SA, TENET
@@
 
-With support from this grant, NRAO hosts the archiving, indexing, and processing of HERA data products. 
+NRAO hosts the archiving, indexing, and processing of HERA data products. 
 Project data and public data products are served via the existing 
@@
 NRAO manages the dynamic acquisition of computing from
-external HPC and cloud-computing facilities, including the Amazon spot market and XSEDE. 
+external HPC and cloud-computing facilities, including the Amazon spot market (\url{aws.amazon.com/ec2/spot}) and XSEDE (\url{www.xsede.com}) . 
 Combined with legacy computing resources at UPenn and MIT, this plan leverages NRAO's
@@
 and \url{github.com/Nithyanandan/PRISim}.}, revision controlled, and unit-tested.  
-These include Omnical, a complete package for redundant baseline calibration;
+These include OMNICAL, a complete package for redundant baseline calibration;
 Astronomical Interferometry in Python (AIPY), a set of
@@
 (PRISim), a package for accurately simulating wide-field interferometric observations;
-21cmFAST, a fast, semi-numerical 21~cm signal simulator,
-and 21cmSense, a tool for forecasting power spectrum sensitivity.
+21cmFAST \citep{mesinger_et_al2011}, a fast, semi-numerical 21~cm signal simulator,
+and 21cmSense \citep{pober_et_al2014}, a tool for forecasting power spectrum sensitivity.
 Other project code is aggregated and revision

### 2016-04-18 manuscript.tex
editing and restructure of sec 3

@@
 foreground leakage in Fourier space (see Fig.~\ref{fig:wedge}), 
-outside of which the 21\,cm signal dominates (the ``EOR window'').  The coordinates of the wedge are discussed below.
+outside of which the 21\,cm signal dominates (the ``EOR window'').  
 Through theoretical and observational work
@@
 
-Though the power spectrum measurement is over a volume, it is instructive to split the spatial wavenumbers into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $|\kvpr|\equiv \kpr = 2\pi b/\lambda X$ is determined by the antenna baseline ($b$) and $\kpar=2\pi/Y\Delta\nu$ by the bandwidth ($\Delta\nu$).
+The power spectrum measurement provides the spatial correlations across the sky, characterized by the magnitude of the wavenumber, ${\bf k}$.
+Though the full magnitude of the {\bf k}-vector is used, it is instructive to split it into two components,  $\kvec =  \kvpr + \kpar{\hat {\bf z}}$ where $|\kvpr|\equiv \kpr = 2\pi b/\lambda X$ is determined by the antenna baseline ($b$) and $\kpar=2\pi/(YB)$ by the bandwidth ($B$).
 Here $X$ and $Y$ are cosmological parameters relating angular size and spectral frequency to cosmic volumes respectively (so, relating wavenumber to physical volume at a given redshift).
@@
 
-In these variables, the ``contaminated'' phase space is a wedge-shaped region in $\kpr$:$\kpar$ space such that the ratio of the time-delay across a given baseline ($b/c$,  which impacts the chromatic visibility) and the delay associated with a given bandwidth ($1/\Delta\nu$) is less than some parameter determined by the details of the system, which we denote as $1/\beta$.
+The $\kvpr$ components are directly proportional to the baselines and $\kpar$ are proportional to the Fourier transform of the frequency response.
+The Fourier transform of a frequency spectrum is a delay, hence the name ``delay-spectrum'' for the technique that uses this approach.
+Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.
+For HERA, wavenumbers are dominated by the bandwidth, not the baseline.
+
+Figure \ref{fig:kperf} shows many of the dependencies on wavenumber by the bandwidth, configuration and cosmos by plotting the perpendicular wavenumber ($\kvpr$; black lines), parallel wavenumber ($\kpar$; blue lines) and total wavenumber ($k$; green lines) as a function of redshift for various bandwidths and baselines.
+The redshift range is appropriate for the extended frequency range that is the goal for HERA, however $z$=6-13 is the primary region for EOR studies.
+The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.
+Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.
+Wavenumbers with an assumed bandwidth ($k$ and $\kpar$ are shown as stepped profiles, with the redshift steps equal to the bandwidth at that redshift.
+Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer appropriate, which is why the line is dashed.
+Also in Fig. \ref{fig:kperf}, the lime-green shaded area indicates the initial region in which HERA intends to detect and characterize the EOR using the delay-spectrum technique.
+
+\begin{figure}[h!]
+\centerline{
+\includegraphics[width=.4\textwidth]{plots/kperf.pdf} 
+}
+\caption{\small $|\kvec |, \kpar, \kperp$ wavenumber plots at various bandwidths ($k_{||}$; blue), baselines ($\kperp$; black), and total ($|\kvec |$, green) as a function of redshift.
+The stepped profiles for $k_{||}$ and $|\kvec |$ are the line-of-site redshift ranges for that bandwidth (too small to see at 1 MHz).
+See text for discussion.
+The full redshift range is the HERA extended goal.
+The initial HERA delay-spectrum power spectrum goal is the light green square region in the upper left.}
+\label{fig:kperf}
+\end{figure}
+
+In these variables, the ``contaminated'' phase space is a wedge-shaped region in $\kpr$:$\kpar$ space such that the ratio of the time-delay across a given baseline ($b/c$,  which impacts the chromatic visibility) and the delay associated with a given bandwidth ($1/B$) is less than some parameter determined by the details of the system, which we denote as $1/\beta$.
 Substituting $\kpr$ and $\kpar$ in for $b$ and $\Delta\nu$ in this ratio, this wedge is bounded by
@@
 
-The techniques to measure the power spectrum may be broken down into two principle techniques -- delay-space (\ref{sec:delayapproach}) and map-making (\ref{sec:mapapproach}) -- plus additional hydrid methods.
+The techniques to measure the power spectrum may be broken down into two principle techniques -- delay-space (\S\ref{sec:delayapproach}) and map-making (\S\ref{sec:mapapproach}) -- plus additional hydrid methods.
 These are briefly summarized below, with an emphasis on the delay-spectrum technique as the initial approach taken by HERA.
@@
 Note that (a) {\em hex} is based on the configuration used for HERA (14.6m center-to-center spacing); (b) {\em grid} for that of PAPER (4m N-S, 16m E-W spacing); (c) {\em imaging-128} for MWA \citep{tingay_et_al2013}; and (d) {\em imaging-48} for LOFAR (\citealt{2013A&A...556A...2V}; note that this assumes the ability to fully correlate the HBA-pairs).
-The plot shows that for a fixed number of elements using the delay-spectrum approach, redundant arrays (hex/grid) provide about an order of magnitude improvement over imaging arrays.
+The plot shows that for a fixed number of elements using the delay-spectrum approach, redundant arrays (hex and grid) provide about an order of magnitude improvement over imaging arrays.
 Since the baselines go as order $N^2$, this corresponds to using about 1/3 of the number of elements to yield the same performance (as verified by the Hex-37/Imaging-128 lines).
@@
 Another approach to improving sensitivity is to make the collecting area per element larger.
-In the limit, this yields a fully filled aperture as mentioned.
-A larger element limits the accessible field-of-view, which however is not a huge liability for a focussed experiment, as long as an appropriate patch of ``cold'' sky passes within its field-of-view.
-One method combining both advantages is to use small elements to build up phased array ``tiles'' and then electronically produce many beams, which may get correlated.
-This adds additional complexity and potential systematics, however it is the method employed by both MWA and LOFAR.
+In the limit, this yields a fully filled aperture.
+A larger element does limit the accessible field-of-view, but this is not a huge liability for a focussed experiment as long as an appropriate patch of ``cold'' sky passes within its field-of-view.
 Fig.
-\ref{fig:boost} (right) shows the relative sensitivity of the various array relative to HERA-19, calculated for a redshift corresponding to the peak signal at a hydrogen reionization fraction of 0.5.
-The circles near the array name indicate the relative sizes of the element.
-Note that MWA and LOFAR are cross-hatched to indicated phased array tiles.
+\ref{fig:boost} (right) shows the relative sensitivity incorporating configuration and element size of various arrays relative to HERA-19, calculated for a redshift corresponding to the peak signal at a hydrogen reionization fraction of 0.5.
+The circles to the right indicate the relative sizes of the element.
+Note that MWA and LOFAR are cross-hatched to indicate phased array tiles, which uses small antennas to phase up a full element response.
 
@@
 
-The power spectrum statistic uses the full magnitude of the {\bf k}-vector, which is set by both the antenna baseline and bandwidth, as discussed above.
-The $\kvpr$ components are proportional to the visibilities and $\kpar$ are proportional to the Fourier transform of the frequency response.
-The Fourier transform of a frequency spectrum is a delay, hence the name ``delay-spectrum'' for this technique.
-Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.
-For HERA, wavenumbers are dominated by the bandwidth, not the baseline.
-
 As an aside, note the parallels to imaging with an interferometer.
@@
 
-Figure \ref{fig:kperf} shows many of the dependencies on wavenumber of bandwidth, configuration and the cosmos.
-It shows the perpendicular, parallel and total resolution wavenumber as a function of redshift for various bandwidths and baselines.
-The term {\em resolution} wavenumber is used, since it is a lower limit for that parameter since larger wavenumbers may be obtained by channelizing.
-The redshift range is appropriate for the extended frequency range that is the goal for HERA, however $z$=6-13 is the primary region for EOR studies.
-
-In Fig.
-\ref{fig:kperf}, blue lines denote $\kpar$ values, black lines denote $\kpr$, and green lines denote total magnitude ($|{\kvec}|$).
-The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.
-Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.
-Also shown just below the blue $\kpar$ curves are redshift ranges for that bandwidth centered at $z$=6,12, and 18 (they appear as horizontal ``error'' bars).
-Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer appropriate, so some evolutionary model would have to be assumed (this is why the line is dashed and only the $z$=12 range is shown).
-Also in Fig.
-\ref{fig:kperf}, the lime-green shaded region indicates a plausible region in which HERA could operate and the lighter square within that region is the area that is being targeted in the first phase.
-
-\begin{figure}[h!]
-\centerline{
-\includegraphics[width=.4\textwidth]{plots/kperf.pdf} 
-}
-\caption{\small $|\kvec |, \kpar, \kperp$ wavenumber plots at various bandwidths ($k_{||}$; blue), baselines ($\kperp$; black), and total ($|\kvec |$, green) as a function of redshift.
-The stepped profiles for $k_{||}$ and $|\kvec |$ are the line-of-site redshift ranges for that bandwidth (too small to see at 1 MHz).
-See text for discussion.
-The full redshift range is the HERA extended goal.
-The initial HERA delay-spectrum power spectrum goal is the light square region in the upper left.}
-\label{fig:kperf}
-\end{figure}
-
 \subsection{Mapmaking Approach}
@@
 In contrast to the delay-spectrum approach, mapmaking approaches combine baselines to build up information before squaring and calculating the power spectrum.
-The conceptually simpler approach is to first make an image cube of the sky as a function of angular position and frequency, and then to Fourier transform, square, and bin the cube to determine the power spectrum \cite{liu_tegmark2011,dillon_et_al2013a}.
+The conceptually simplest approach is to first make an image cube of the sky as a function of angular position and frequency, and then to Fourier transform, square, and bin the cube to determine the power spectrum \citep{liu_tegmark2011,dillon_et_al2013a}.
 The image domain is a natural place to combine information from partially-coherent pairs of visibility measurements (either from different baselines or from the same baseline at different times).
-Mapmaking can be lossless data compression step \citep{tegmark1997b}, though keeping track of statistical properties of the maps, which have complex frequency and position dependent point spread functions and noise covariance matrices, can be computationally challenging \citep{dillon_et_al2015a}.
+Mapmaking can be a lossless data compression step \citep{tegmark1997b}, though keeping track of statistical properties of the maps, which have complex frequency and position dependent point spread functions and noise covariance matrices, can be computationally challenging \citep{dillon_et_al2015a}.
 Existing approaches have had to make a number of approximations, including that point spread functions do not vary over the field of view and that noise is not correlated between $uv$-cells.
@@
 Like the delay-spectrum approach, such pipelines (see, e.g., \citealt{trott_et_al2016}) avoid the image domain entirely.
-This helps to prevent artifacts that may introduced by imaging algorithms.
+This helps to prevent artifacts that may be introduced by imaging algorithms.
 
@@
 Whether it makes computational sense to examine all pairs of partially-coherent visibilities or to perform mapmaking as an intermediate data-compression step before power spectrum estimation (despite the complexity of the resultant map statistics) is an open question that depends on a number of factors, including the number of elements in the array, its layout, and its integration time.
-These approaches need further development, and can be incorporated when sufficiently implemented.
+However, similar to map-making techniques, these approaches need additional precision in calibration and modeling development in order to be implemented with confidence.
 
@@
 
-However, at each frequency, degeneracies in the redundant calibration procedure prevent it from solving four numbers per frequency for the whole array (down from 700 in HERA's case).
+However, at each frequency, degeneracies in the redundant calibration procedure prevent it from solving four numbers per frequency for the whole array (down from 700 in HERA's case if full calibration, as opposed to redundant calibration, were needed).
 The first is an overall gain---one can multiply each $g_i(\nu)$ by 2 and divide each $V_{ij}^\text{true}$ by 4, for example, and get exactly the same measured visibilities.

### 2016-04-18 plots/kperf.py

@@ -100,8 +100,8 @@ for i in use_i:
     plt.semilogy(z,ktot,lsc[i],label=show_label,lw=2) 
     plt.text(txtloc[i][0],txtloc[i][1],lbl,rotation=txtloc[i][2],color=tclr[i],fontsize=16)
 ktottop = 10.*np.ones(len(zsave))
-plt.fill_between(zsave,savek,ktottop,color='g',alpha='0.25')
-plt.fill_between(zhera,khera1,khera2,color='w',alpha='0.65')  ###AND REDO LINE 63
+#plt.fill_between(zsave,savek,ktottop,color='g',alpha='0.25')
+plt.fill_between(zhera,khera1,khera2,color='g',alpha='0.25')  ###AND REDO LINE 63
 
 #second part  (k_perp)
 BW = 0.01

### 2016-04-19 manuscript.tex
scrub of sec 4

@@
 \subsection{General Antenna Design and Configuration}
-The primary new feature of HERA over previous generation experiments (in this context, primarily PAPER) is the use of large elements to increase the sensitivity, as indicated in Fig.
+The primary new feature of HERA over previous generation experiments (in this context primarily PAPER) is the use of large elements to increase the sensitivity, as indicated in Fig.
 \ref{fig:boost}.
 The cost-performance optimization will
-be discussed below, however we see that to address the central challenges of the delay-spectrum approach in the context of the wedge, to open the widest possible window for EOR measurements, 
-HERA must use close-packed antennas that
+be discussed below, however we see that to address the central challenges of the delay-spectrum approach in the context of the wedge.
+HERA should use close-packed antennas that
 minimize signal reflections over long delays and deliver significant forward gain relative to their horizon response.
@@
 Extending the science to include the Dark Ages remains a goal so that efforts are on-going to design a feed to increase the lower limit down to at least 70 MHz without compromising the performance in the above requirement range.
-The fall-back position for the low-frequencies is a serial deployment of a scaled version of the HERA feed or to potentially build additional elements specifically for the low-frequencies.
-
-The scientific requirement on channel bandwidth is to allow access to k-modes of 1 h/Mpc [ref], which requires about 256 channels over the 100 MHz bandwidth.
+The fall-back position for low frequencies is a serial deployment of a scaled version of the HERA feed or to potentially build additional elements specifically for low frequencies.
+
+The scientific requirement on channel bandwidth is to allow access to k-modes of $\sim1$ h/Mpc, which requires about 256 channels over the 100 MHz bandwidth.
 However, in order to handle radio frequency interference as well as to allow the bandpass to be characterized, a specification of 1024 channels has been chosen.
@@
 \label{sec:delayspec}
-Early analysis \citep{elementmemo} indicated that to trace the EoR over a wide range of redshift, internal reflections should be attenuated by about 60 dB from the initial incoming wave by about 60 ns, or a VSWR $<$ 1.002 for frequencies $<$ 17 MHz.
+Early analysis \citep{elementmemo} indicated that to trace the EOR over a wide range of redshift, internal reflections should be attenuated by about 60 dB from the initial incoming wave by about 60 ns, or a VSWR $<$ 1.002 for frequencies $<$ 17 MHz.
 This provided the basis for the initial design of the element.
-For the contemplated feeds for a primary focus antenna, it was found that a focal length/diameter radio ($f/D$ of approximately 0.32 was optimal.
+For the contemplated feeds on a primary focus antenna, it was found that a focal length/diameter radio ($f/D$) of approximately 0.32 was optimal.
 We may therefore estimate the attenuation as a function of delay and diameter and assign this a ``cost function''.
-The vertical axis of Fig.
-\ref{fig:costfig} shows a very steep cost function at 60 ns.
-The curves labelled `1', `2' and `3' are the round-trip travel time for 1, 2 and 3 reflections.
-
-To obtain 60 dB at 60 ns we may interpret those lines in one of two ways:  (1) set the specification on how much attenuation per reflection is needed and (2) at a fixed attenuation/reflection, what diameter is required.
-For 1 round-trip, for the assumed $f/D$ over the diameters of interest, all delays are less than 60 ns, so it provides no constraint.
-For
-3 round-trips, all diameters exceed 60 ns, so the attenuation should be less than 60 dB for that path-length.
-The case for 2 round-trips provides a constraint on attenuation and diameter, which is discussed at length in \cite{elementmemo}.
-To estimate, it was assumed that the main reflection time constant was due to the standing waves between the feed and vertex of the parabola and that each reflection redirected about -15 dB of the total power in each reflection for additional standing wave power, which leads to a diameter of about 14m.
-
-As a key specification, this has been a focus of study with our prototypes and a series of papers based on analysis and measurements have been produced to provide much more rigor than this simple early analysis \citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016,patra_et_al0216}.
+
+Figure \ref{fig:costfig} summarizes the adopted cost function.  
+The vertical axis shows a very steep delay ``cost'' at 60 ns and the horizontal axis is the cost/performance model, as described in the following section.  
+The color coding is the product of the two assigned costs, with red showing a high cost normalized to 350 14m elements (clipped at a ratio of 1.25) and purple being the minimum.
+The curves labelled `1', `2' and `3' are the round-trip travel times for 1, 2 and 3 reflections between the feed and the vertex.
+To obtain 60 dB by 60 ns we may interpret those lines in one of two ways:  
+(a) at a fixed attenuation/reflection, what diameter is required, and
+(b) at a fixed diameter how much attenuation per reflection can be allowed.
+For 1 round-trip, for the assumed $f/D$ over the diameters of interest, from Fig.
+\ref{fig:costfig} all delays are less than 60 ns, so this provides no constraint.
+If we want to be able to handle 2 round-trips worth of delay, the diameter should be less than 14.2m, which is near the cost minimum.
+If we assume 14.2m, then we can estimate that each reflection must average a return loss better than 15 dB, an aggressive target.
+And finally, if we need to allow for 3 round-trips, our antenna diameter would have to be less than 9.4m, where the cost is quickly rising.
+To handle the delay requirement therefore, the diameter should be less than 14.2m, or the return loss specification becomes very aggressive.
+
+As a key specification, this has been a focus of study with a series of prototypes and analyses based on modeling, and measurements have been conducted to provide much more rigor than this early specification \citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016,patra_et_al0216}.
 The simple specification set above more appropriately becomes a series of curves for attenuation and delay at different wavenumbers for realistic foreground models.
-Fig \ref{fig:delayspec}, adapted from \cite{thyagarajan_et_al2016}, shows the results at $k$ = 0.2 $h$/Mpc for the adopted design.
-The analysis shows that for $k>0.15 h/$Mpc, HERA should be able to detect the EoR.
+Fig \ref{fig:delayspec}, adapted from \cite{thyagarajan_et_al2016}, shows the results at $k$ = 0.10, 0.15, and 0.20 $h$/Mpc for the adopted design.
+The analysis shows that
+the dish design for should should allow for EOR detections for $k>0.10 h/$Mpc.
 We also see that the initial simple ``60@60'' specification was a overly conservative, however it served as a good initial goal.
-The analysis also indicates which baselines produce the best information on different scales.
 
@@
     \includegraphics[width=0.49\textwidth]{plots/delayspecplot.pdf}
-     \includegraphics[width=0.5\textwidth]{plots/ps1d_with_delay_kernel.pdf}
-	\vspace{-25pt}
-	\caption{Left:  Antenna delay attenuation specification with realistic foregrounds from \cite{thyagarajan_et_al2016} along with the initial conservative ``60@60'' specification.
-HFSS calculations (black line) show HERA meeting the specification.
+     %\includegraphics[width=0.5\textwidth]{plots/ps1d_with_delay_kernel.pdf}
+	%\caption{Left:  
+	\caption{Antenna delay attenuation specification with realistic foregrounds from \cite{thyagarajan_et_al2016} along with the initial conservative ``60@60'' specification.
+HFSS calculations (black line) show the HERA dish meeting the specification.
 Measurements on the Green Bank prototype are shown in  \citep{patra_et_al2016}.
-Right: Modeled foreground power spectra with (green) and without (blue) the antenna chromaticism over a 10\,MHz subband centered at $z=7.5$.
+}
+%Right: Modeled foreground power spectra with (green) and without (blue) the antenna chromaticism over a 10\,MHz subband centered at $z=7.5$.
 The spectral smoothness of the dish enables foreground isolation below the level of a fiducial reionization signal (black) for most $k_{\parallel}$ modes, yielding percent-level constraints on model parameters.}
@@
 \label{sec:cost}
-As an instrument to characterize the power spectrum over the evolution of the EoR, the specification on sensitivity is to make at least a nominal detection over the redshift region of support, or $z$=6-13, along with a very robust detection at the peak.
+As an instrument to characterize the power spectrum over the evolution of the EOR, the specification on sensitivity is to make at least a nominal detection over the redshift region of support ($z$=6-13) along with a very robust detection at the peak.
 As seen in Fig.
-\ref{fig:Sensitivities} and Table \ref{tab:signif}, an array comprising 320 14-m core antennas (note, labelled HERA-350, which includes the outriggers) has sensitivity to a
+\ref{fig:Sensitivities} and Table \ref{tab:signif}, an array comprising 320 14-m core antennas (labelled HERA-350, which includes the outriggers) has sensitivity to a
 fiducial reionization model across that redshift range.  We therefore adopt the delay-spectrum power spectrum sensitivity of 320 14-m close-packed antennas as the minimum sensitivity needed.  
@@
 where $N_{\text core}$ and $D$ are the number of antennas in the core and their diameter respectively.
-Note that the linear number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} Pober) over a range of values for the HERA configuration for wavenumbers greater than 0.15.
+Note that the linear number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} 21cmSense) over a range of values for the HERA configuration for wavenumbers greater than 0.15.
 This dependency differs from \cite{2013ExA....36..235M} and analytical expressions, mostly likely due to the assumed configuration.

### 2016-04-19 manuscript.tex
scrub through 5.4

@@
 \section{System Design} \label{sec:design}
-As described in the previous section, the critical insights from the first generation 21\,cm EoR experiments have been applied to define
+As described in the previous section, the critical insights from the first generation 21\,cm EOR experiments have been applied to define
 the requirements for HERA---an instrument designed to ensure that foregrounds remain bounded
@@
 and system architecture (see Fig.~\ref{fig:overallBlockDiagram}).
-This architecture directly inherits from the PAPER and MWA experiments; HERA begins by reusing the
-analog, digital, and real-time processing systems deployed for PAPER-128.  As HERA develops, this
+This architecture directly inherits from the PAPER and MWA experiments.  HERA begins by reusing the
+analog, digital, and real-time processing systems deployed for PAPER-128.  
+This allowed for immediate observing with the new elements with a well-characterized system.
+As HERA develops, this
 architecture is incrementally upgraded to improve performance and add features while simultaneously
@@
 
-Note that in addition to the 19 elements in South Africa (and an initial prototype near Berkeley, CA), a pair of elements at the Green Bank Observatory in West Virginia have been used extensively in the validation testing, discussed below.
+Note that in addition to the 19 elements currently observing in South Africa (and an initial prototype near Berkeley, CA), a pair of elements at the Green Bank Observatory in West Virginia have been used extensively in validation testing.
 Also, three elements are currently under construction at the Mullard Radio Astronomy Observatory outside Cambridge, UK for additional feed and element tests.
+This allows for an independent testing platform as the group at Cavendish continues to investigate both a broad-band feed, as well as ways to improve the current analog system for an improved match and noise performance.
 
@@
 
-%\subsection{Site}
-%\label{sec:site}
+\subsection{Site}
+\label{sec:site}
 HERA is located at the Karoo Astronomy Reserve.
@@
 And the third item constrains the construction materials and methods and the operational model.
-This led to a fixed transit element with a diameter of 14 meters to strike an optimal balance between sensitivity and systematics.
+This led to a fixed transit element with a diameter of 14 meters to strike an optimal balance between sensitivity and systematics, as discussed in the previous section.
 
@@
 degrading our ability to isolate and remove foreground emission on the basis of
-spectral smoothness (Fig. \ref{fig:reflectometry}, right panel).  As shown in
+spectral smoothness.  As shown in
 \Mycitet{parsons_et_al2012b}, the amplitude and timescale of signal reflections
@@
 If only twelve spars were used for the entire surface, simple Ruze losses would be quite high (about 10\% at 150 MHz).
-Therefore, at about 2.4 meters another horizontal member is placed, which launches another parabolic spar.
-The simple Ruze loss is then less than 1\%, which is much more acceptable.
+Therefore, at about 2.4 meters another horizontal member is placed, which launches another parabolic spar leading to a Ruze loss less than 1\%.
 Fig \ref{fig:elementplan} shows this surface scheme and the panelization (left) and the offsets (in mm) of that scheme from an ideal parabola.
@@
                {\integral_{-\pi}^{\pi}\left[\sqrt{G(\theta,\phi=0)} + \sqrt{G(\theta,\phi=90)}\right]^2\sin\theta d\theta}
+     \label{eq:polarmismatch}
 \end{equation}
@@
 The 10 dB input match for the feed itself spans from about 105 - 200 MHz.
-The resultant calculated antenna performance parameters are shown in Fig.
+The resultant HFSS-calculated antenna performance parameters are shown in Fig.
 \ref{fig:ant_params} and beam-patterns for 110 MHz, 155 MHz and 190 MHz are shown in Fig.
@@
 }
-\caption{\small Antenna parameters}
+\caption{\small Antenna parameters from the feed optimization analysis.
+All calculations were done using HFSS on the final design as described.
+{\bf Upper left:}  fraction of feed power illuminating the primary (main beam efficiency).
+{\bf Upper right:}  polarization mismatch as defined in Eq.
+\ref{eq:polarmismatch}.
+{\bf Lower left:}  total gain (blue) contrasted with a fixed efficiency of 78\%.
+{\bf Lower right:}  first side-lobe level relative to the main beam peak.}
 \label{fig:ant_params}
@@
 }
-\caption{\small Antenna patterns at 110 MHz, 155 MHz and 190 MHz.}
+\caption{\small Antenna patterns at 110 MHz, 155 MHz and 190 MHz calculated with HFSS and the adopted design.
+Red curves are E-plane, green are H-plane at blue at 45$^\circ$.
+Inset is the full 3D pattern.}
 \label{fig:beampatterns}
@@
 As a focused experiment, HERA's configuration is optimized for the robust foreground-avoidance approach to EoR power spectrum estimation. 
-This requires a densely-packed core of antenna elements, maximizing sensitivity to short baselines---those least contaminated by foreground chromaticity (i.e.\ the wedge). 
+This requires a densely-packed core of antenna elements, maximizing sensitivity to short baselines---those least contaminated by foreground chromaticity ({\em i.e} the wedge). 
 Furthermore, the configuration should be highly redundant, both to enable the delay-spectrum strategy of PAPER \citep{parsons_et_al2012a} and redundant baseline calibration \citep{liu_et_al2010,zheng_et_al2014}.
@@
 \subsection{Analog Signal Path}
-HERA's analog signal path from 70--250 MHz with the full performance EoR band of 100-200 MHz, as shown in Fig.
+HERA's analog signal path from 70--250 MHz with the full performance EOR band of 100-200 MHz, as shown in Fig.
 \ref{fig:cosmos} and discussed in Sec.
@@
 Although the low-frequency sky has an intrinsically bright signal level, in the desired ``cold patches'' the sky temperature is only about 30 K at the upper frequency end, so relatively low receiver temperatures are desired.
-Note that at the low end of the band (70 MHz), the sky temperature is about 1000 K and across the EoR band it ranges from about 100-400 K.
+Note that at the low end of the band (70 MHz), the sky temperature is about 1000 K and across the EOR band it ranges from about 100-400 K.
 With receiver temperatures of 30--100 K readily achievable with ambient temperature electronics however
@@
 As part of the delay specification,
-HERA's signal path emphasizes careful impedance matching between each component to avoid signal reflections that can worsen foreground leakage.
-In particular, impedance matching between free space (377$\Omega$) and coaxial cable (50$\Omega$) is undertaken
-in the feed/balun at the front end of the signal chain, where the time constants of inevitable reflections are
-minimized.  For a similar reason, HERA houses analog-to-digital converters (ADCs) in nodes near the antenna elements
+HERA's signal path emphasizes careful impedance matching between each component to minimize signal reflections that can worsen foreground leakage for the delays of interest.
+This means that the emphasis is to reduce the delay-value of a reflection, which allows for a design degree of freedom.
+One such design element is to
+house the analog-to-digital converters (ADCs) in nodes near the antenna elements
 to limit the total analog path length.
-This length is set to be 35m, such that the round-trip delay corresponds to about 0.15 $h$/Mpc, which remains in the wavenumbers  contaminated by the wedge.
-The project is also examining the impact of having a range of analog path lengths of $35\pm10$ m.
+This length is set to be 35m, such that the round-trip delay corresponds to about 0.15 $h$/Mpc, which remains in the wavenumbers  near the wedge.
+The project is also examining the impact of having a range of varying analog path lengths.
 
@@
 
-\noindent Although correlator development has historically been one of the most 
+Although correlator development has historically been one of the most 
 expensive aspects of building a large radio interferometer, this is no longer the case.
-CASPER \Mycitep{parsons_et_al2006}
-open-sourced the development of digital signal processing engines for astronomy and
+The Collaboration for Astronomy Signal Processing and Electronics Research (CASPER) is an \Mycitep{parsons_et_al2006}
+open-source program to further the development of digital signal processing engines for astronomy and
 now has world-wide participation,
@@
 HERA efforts continue this incremental development cycle, using a packet-switched
-correlator architecture \Mycitep{parsons_et_al2008} that both PAPER and LEDA have
+correlator architecture \Mycitep{parsons_et_al2008} that both PAPER and LEDA \citep{2015JAI.....450003K} have
 applied to systems employing FPGAs and GPUs \citep{clark_et_al2011}.

### 2016-04-19 manuscript.tex
incorporate eloys comments

@@
 
+The feed design will proceed in three phases:
+\begin{itemize}
+\item {\em Phase 1:}  use existing PAPER dipole with new ground screen (\S\ref{sec:feed}) and existing PAPER signal path,
+\item {\em Phase 2:}  updated ground screen, matching network and analog signal path (\S\ref{sec:analog})
+\item {\em Phase 3:}  broad-band or separate low-frequency for full spectral coverage.
+\end{itemize}
+Models and measurements discussed below are for the current Phase 1 system.
+
 As shown in Fig.~\ref{fig:delayspec}, HERA's element meets this requirement in
@@
 half maximum of $\sim$$10^\circ$, and a first sidelobe at -20 dB \citep{ewallwice_et_al2016,neben_et_al2016,patra_et_al2016,thyagarajan_et_al2016}.
-Feed-to-feed coupling between adjacent antennas has been measured to be below -50 dB, indicating
-that mutual coupling will not be problematic.
+Feed-to-feed coupling between the two adjacent antennas in Green Bank has been measured to be below -50 dB, providing some confidence that mutual coupling will be at a 
+manageable level.  Studies of elements in a bigger array and potential screening approaches are also underway.
 
@@
 Initially, the feed reuses the original PAPER sleeved dipole, however in a new optimized cylindrical feed-cage configuration.
-The optimization was based on main beam efficiency, cross-coupling (integrated feed pattern on the dish), standing waves (blockage size), frequency response and polarization match.
+As discussed in \cite{feedmemo}, the optimization was based on (a) main beam efficiency, (b) cross-coupling (integrated feed pattern on the dish), (c) standing waves (blockage size), (d) frequency response and (e) polarization match.
 The design parameters were size of backplane, height of mast, and height of cylinder.
@@
 \subsection{Analog Signal Path}
+\label{sec:analog}
 HERA's analog signal path from 70--250 MHz with the full performance EOR band of 100-200 MHz, as shown in Fig.
@@
 One such design element is to
-house the analog-to-digital converters (ADCs) in nodes near the antenna elements
+house the analog-to-digital converters (ADCs) in ``nodes'' near the antenna elements
 to limit the total analog path length.
@@
 The project is also examining the impact of having a range of varying analog path lengths.
+
+The analog signal path comprises the front-end module (FEM) at the feed, the post-amplfier module (PAM) at field-deployed nodes, and the $\sim 35m$ coaxial cable in between.
+Also included are the monitor and control aspects to control and measure its state.
+
+{\bf Front-End Module: }
+
+{\bf Post-Amplifier Module: }
 
@@
 five generations of hardware.
-On a modest budget, PAPER applied CASPER technology to develop and deploy new correlators
+PAPER has applied CASPER technology to develop and deploy new correlators
 annually for five years running, each quadrupling the computational capacity of its predecessor.
+
 Led at UC Berkeley's Radio Astronomy Lab (RAL),
@@
 
-In order to meet the 35\,m specification for a maximum analog signal path length, along with a growing need for system scalability, HERA will adopt a node-based architecture for amplification, digitization, channelization, and digital
+In order to meet the 35\,m specification for a maximum analog signal path length, along with a growing need for system scalability, HERA will 
+adopt a node-based architecture for amplification, digitization, channelization, and digital
 transmission in the field, building on MWA and Allen Telescope Array heritage. This design is merged with PAPER's clean 
@@
 This has led to a new analog-digital-optical board called the Smart Networked Analog-Digital Processor (SNAP), which will become a new modular CASPER device.  
-Figure \ref{fig:hardware} shows the procession of boards.
+Figure \ref{fig:hardware} shows the series of primary CASPER processing boards, with SNAP on the far right.
 This section will step through the digital signal path organized by location:  node, container, processing building.

### 2016-04-20 manuscript.tex
incorporate (most of) Jonnies comments

@@
 
-The Hydrogen Epoch of Reionization Array (HERA http://reionization.org) is a staged experiment that uses the unique properties of the 21-cm line from neutral hydrogen to probe the Epoch of Reionization (EOR).
+The Hydrogen Epoch of Reionization Array (HERA) is a staged experiment that uses the unique properties of the 21\,cm line from neutral hydrogen to probe the Epoch of Reionization (EOR).
 During this epoch, roughly 0.3 - 1 billion years after the Big Bang, the first galaxies and black holes heated and reionized the early Universe.
 Direct observation of the large scale structure of reionization and its evolution with time will have a profound impact on our understanding of the birth of the first galaxies and black holes, their influence on the intergalactic medium (IGM), and cosmology.
-This paper will provide an overview of the project and describe the design of the system.
+This paper will provide an overview of the project and describe the design of the system, moving from a science overview to requirements, design and status.
 
 \section{Introduction}
-The Hydrogen Epoch of Reionization Array (HERA) is a staged experiment to use the redshifted 21\,cm line
+The Hydrogen Epoch of Reionization Array (HERA; http://reionization.org) is a staged experiment to use the redshifted 21\,cm line
 of neutral hydrogen to characterize our Cosmic Dawn, from
@@
 
-Because of these key science questions that may be probed by highly redshifted hydrogen, there has been a resurgence in low frequency arrays  -- most with a primary
+Because of these key science questions that may be probed by highly-redshifted hydrogen emission, there has been a resurgence in low frequency arrays  -- most with a primary
 objective of measuring the EOR.
@@
 With the first generation 21\,cm telescopes approaching the limits of their sensitivity however, HERA continues the experiment by bringing significantly more sensitivity to bear
-while continuing to optimize its design along the delay-spectrum approach to detect the power spectrum of the EoR (discussed in Sec. \ref{sec:delayapproach}).
+while continuing to optimize its design along the delay-spectrum approach to detect the power spectrum of the EoR (discussed in \S\ref{sec:delayapproach}).
 At the same time, HERA continues the development of new techniques to extend and improve its capabilities.
@@
 
-The current 19 elements are testing the HERA element design for performance and manufacturability on location.
-The 37 elements, to be completed in 2016, will provide nearly an order of magnitude more
-sensitivity than current instruments and provide a significant chance at detecting the EoR power spectrum signal.
-The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER \citep{ali_et_al2015}, which should provide a very robust detection of the EOR signal and allow some characterization.
-Finally, extending to 2020, HERA will build out to 350 elements to further EOR science as a function
-of redshift and spatial scale, potentially producing the first images of the EOR.
+The current 19 elements are testing the HERA element design for performance and manufacturability on location.  The 37 elements, to be completed in 2016, will provide a factor of about 5 more 
+sensitivity than PAPER-128 and provide a significant chance at detecting the EoR power spectrum signal.
+The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER,
+%\citep{ali_et_al2015}
+which should provide a very robust detection of the EOR signal and allow some characterization.  Finally, extending to 2020, HERA will build out to 350 elements to further EOR science as a function 
+of redshift and spatial scale, potentially producing the first images of the EOR.  These estimates will be quantified and shown in this paper.
 
@@
 The blue line shows the frequency of the red-shifted hydrogen line (rest frequency 1420 MHz) as a function of redshift (lower horizontal axis) and cosmic time (upper horizontal axis).
-The white dashed horizontal lines are the HERA EoR band and the blue are the extended frequency goal.
+The white dashed horizontal lines are the HERA EOR band and the blue are the extended frequency goal.
 Determining where the structure in the background really falls is the
@@
 requires complementary observational approaches. Measurements of the Cosmic Microwave Background 
-(CMB; the escaping radiation permeating the universe after the universe became
-transparent to its radiation by the recombination of the protons and electrons about 400,000 years after the Big Bang)  by COBE, WMAP and Planck provide initial conditions for structure formation.
+(CMB; the photons permeating the universe after becoming
+transparent to its own radiation by the recombination of the protons and electrons about 400,000 years after the Big Bang)  by COBE, WMAP and Planck provide initial conditions for structure formation.
 Thomson
 scattering of CMB photons by the ionized particles constrains the integrated column of 
-ionized gas and kinetic Sunyaev-Zel'dovich measurements constrain the power spectrum of cosmic structures.
+ionized gas and kinetic Sunyaev-Zel'dovich measurements constrain the duration of the ``patchy'' phase of cosmic structures.
 But even with these measurements, the detailed evolution of the IGM 
@@
 saturate at low neutral fractions
-($x_{\rm HI} \la 10^{-4}$), where $x_{\rm HI}$ is the ratio of neutral hydrogen to the total baryonic density.
+($x_{\rm HI} \ga 10^{-4}$), where $x_{\rm HI}$ is the fraction of hydrogen in its neutral state.
 
@@
 with 1080 hours of integration (black)
-compared with various heating and reionization models (colored).  Note that only predicted results using the delay-spectrum approach are shown.}
+compared with various heating and reionization models (colored).  Sensitivity analysis is per footnote in Table \ref{sec:signif}}
 	\label{fig:Sensitivities}
@@
 The expected performance of HERA relative to current and planned telescopes to detect the peak of reionization (as well as the total collecting area is shown in 
-Table~\ref{tab:signif}.
+Table~\ref{tab:signif}).
 
@@
 
-\begin{SCtable}
-\caption{\hspace{-1.2mm}: Predicted SNRs of 21\,cm experiments for an EoR model with 50\% ionization at $z=9.5$, with 1080 hours observation, integrated over a $\Delta z$ of $0.8$.
-Foreground avoidance represents an analysis comparable to \cite{ali_et_al2015}, whereas foreground modeling allows signifcantly more $k$ modes of the cosmological signal to be recovered.}
+\begin{table}
+\caption{\small Predicted SNRs of 21\,cm experiments for an EoR model with 50\% ionization at $z=9.5$, with 1080 hours observation, integrated over a $\Delta z$ of $0.8^*$.
+}
 \small
@@
 \end{tabular}
-%\Caption{-0.1in}{0.99}{-0.1in}
-\hspace{-0.1in}
+\caption*{%\tiny{\justify
+$^*$Calculations done via \url{21cmSense} \url{www.github.com/} \citep{2013AJ....145...65P,2014ApJ...782...66P}
+Foreground avoidance represents an analysis comparable to \cite{ali_et_al2015}, whereas foreground modeling allows significantly more $k$ modes of the cosmological signal to be recovered.}
 \label{tab:signif}
-\end{SCtable}
+\end{table}
 
@@
 Figure~\ref{fig:paramConstraints} shows the results of a Markov Chain Monte
-Carlo (MCMC) pipeline for fitting models to 21\,cm power spectrum data \citep{greig_and_mesinger2015}, which we have conservatively limited to the range $8 < z < 10$ (although in practice a much broader bandwidth will be available; see Sec.
-\ref{sec:freqs}).
+Carlo (MCMC) pipeline for fitting models to 21\,cm power spectrum data \citep{greig_and_mesinger2015}, which we have conservatively limited to the range $8 < z < 10$ (although in practice a much broader bandwidth will be available; see \S\ref{sec:freqs}).
 Based on the excursion-set formalism of
@@
 \label{sec:eormeas}
-Due to the expansion of the universe over cosmological time, we can identify and measure the early Universe via the red-shift of spectral lines.
+Due to the expansion of the universe over cosmological time, we can identify and measure the early Universe via the redshift of spectral lines.
 The hydrogen hyperfine transition at a rest frequency of 1420 MHz is a key spectral line due to the ubiquity of hydrogen and, being a ``forbidden'' transition, the optical depth lets us see through the entire universe back almost to the period of recombination.
@@
 
-The past three years have seen deep EoR observations with PAPER, the MWA, 
-the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2013}), and LOFAR.
+The past three years have seen deep EOR observations with PAPER, the MWA, 
+the Giant Metrewave Radio Telescope (GMRT), and LOFAR.
 PAPER and the MWA have produced progressively deeper limits
@@
 To address this challenge,  all aspects of the experimental process must be refined and improved, spanning 
-calibration ({\em e.g.} \cite{zheng_et_al2014, jacobs_et_al2016, barry_et_al2016}), handling foreground contamination
+calibration ({\em e.g.} \cite{zheng_et_al2014, jacobs_et_al2016, 2016arXiv160300607B}), handling foreground contamination
 ({\em e.g.} \cite{moore_et_al2013,thyagarajan_et_al2015a,pober_et_al2016}),
@@
 is a foreground mitigation strategy based largely
-on identifying and filtering out a region of phase space where the strong foregrounds are largely confined (Sec. \ref{sec:delayapproach}).  This is the approach
+on identifying and filtering out a region of parameter space where the strong foregrounds are largely confined (\S\ref{sec:delayapproach}).  This is the approach
 pioneered by PAPER.
@@
 The MWA's image-based calibration and foreground subtraction strategy provides complementary
-capabilities (Sec. \ref{sec:mapapproach}). Imaging with subtraction, while still under development as a viable foreground strategy,
+capabilities (\S\ref{sec:mapapproach}). Imaging with subtraction, while still under development as a viable foreground strategy,
 can increase sensitivity by recovering more modes of the cosmological signal (see
@@
 
-The $\kvpr$ components are directly proportional to the baselines and $\kpar$ are proportional to the Fourier transform of the frequency response.
+The $\kvpr$ components are directly proportional to the baselines and $\kpar$ are proportional to the Fourier transform of the frequency response \footnote{Note that this equivalency is an approximation, which is very good for the short baselines and small bandwidths generally used here.
+For a more complete discussion see \cite{liu_et_al2014a,liu_et_al2014b}}.
 The Fourier transform of a frequency spectrum is a delay, hence the name ``delay-spectrum'' for the technique that uses this approach.
@@
 
-Figure \ref{fig:kperf} shows many of the dependencies on wavenumber by the bandwidth, configuration and cosmos by plotting the perpendicular wavenumber ($\kvpr$; black lines), parallel wavenumber ($\kpar$; blue lines) and total wavenumber ($k$; green lines) as a function of redshift for various bandwidths and baselines.
+Figure \ref{fig:kperf} shows many of the dependencies on wavenumber by the bandwidth, configuration and cosmology by plotting the perpendicular wavenumber ($\kvpr$; black lines), parallel wavenumber ($\kpar$; blue lines) and total wavenumber ($k$; green lines) as a function of redshift for various bandwidths and baselines.
 The redshift range is appropriate for the extended frequency range that is the goal for HERA, however $z$=6-13 is the primary region for EOR studies.
@@
 \caption{\small $|\kvec |, \kpar, \kperp$ wavenumber plots at various bandwidths ($k_{||}$; blue), baselines ($\kperp$; black), and total ($|\kvec |$, green) as a function of redshift.
-The stepped profiles for $k_{||}$ and $|\kvec |$ are the line-of-site redshift ranges for that bandwidth (too small to see at 1 MHz).
+The stepped profiles for $k_{||}$ and $|\kvec |$ are the line-of-sight redshift ranges for that bandwidth (too small to see at 1 MHz).
 See text for discussion.
@@
 
-In these variables, the ``contaminated'' phase space is a wedge-shaped region in $\kpr$:$\kpar$ space such that the ratio of the time-delay across a given baseline ($b/c$,  which impacts the chromatic visibility) and the delay associated with a given bandwidth ($1/B$) is less than some parameter determined by the details of the system, which we denote as $1/\beta$.
+In these variables, the ``contaminated'' phase space is a wedge-shaped region in $\kpr, \kpar$ space such that the ratio of the time-delay across a given baseline ($b/c$,  which impacts the chromatic visibility) and the delay associated with a given bandwidth ($1/B$) is less than some parameter determined by the details of the system, which we denote as $1/\beta$.
 Substituting $\kpr$ and $\kpar$ in for $b$ and $\Delta\nu$ in this ratio, this wedge is bounded by
@@
 \end{equation}
-where an offset parameterized by $S$ accounting for affects related to the combined spectral smoothness of the foregrounds and the antenna response has been included(see Fig. \ref{fig:wedge}).
+where an offset parameterized by $S$ accounting for affects related to the combined spectral smoothness of the foregrounds and the antenna response has been included (see Fig. \ref{fig:wedge}).
 
@@
 In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_0=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.
-Also note that at the red-shifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
+Also note that at the redshifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
 
@@
 Other than potentially handling overlapping bins in {\em uv} space, the delay-spectrum approach does not combine baselines before squaring and calculating the power spectrum, which contrasts to the map-making techniques briefly described in the following section.
-
-We can essentially relate the sky power spectrum mode ${\hat P}(\kvec)$ to an interferometer baseline visibility ${\tilde V}(\bvec)$, which is the equivalent of the (complex) fringe pattern in a double slit experiment:
+For more details, see \cite{parsons_et_al2012b}.
+
+We can essentially relate the sky power spectrum $P(\kvec)$ to the Fourier transform along the frequency axis (the delay-transform) of an interferometer baseline visibility, denoted ${\tilde V}_b(\tau)$:
 \begin{equation}
-{\hat P}(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2(\bvec)}{\Omega_b B/ \lambda^4} \right]
+P(\kvec) \approx\frac{X^2Y}{4k_B^2}   \left[\frac{{\tilde V}^2_b(\tau)}{\Omega_b B/ \lambda^4} \right]
 \label{eq:Pk_baseline}
@@
 The terms in square brackets are instrumental terms, as opposed to the constants and cosmological parameters.
-Rather than ${\hat P}(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}{\hat P}(\kvec)$.
+Rather than $P(\kvec)$, the literature generally works with a volume-normalized parameter given by $\Delta^2(k) = \frac{k^3}{2\pi^2}P(\kvec)$.
 
@@
 \begin{equation}
-V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2B\tau}}\right)\Omega_b B
+V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2Bt}}\right)\Omega_b B
 \label{eq:sensitivity_per_baseline}
 \end{equation}
-where $\tau$ is the integration time (see {\em e.g.} \citealt{thompson_et_al2001}),
+where $t$ is the integration time (see {\em e.g.} \citealt{thompson_et_al2001}),
 to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq.
@@
 They are labelled for the type of configuration (hex, grid, imaging) and the number of antennas (from 37 to 350).
-Note that (a) {\em hex} is based on the configuration used for HERA (14.6m center-to-center spacing); (b) {\em grid} for that of PAPER (4m N-S, 16m E-W spacing); (c) {\em imaging-128} for MWA \citep{tingay_et_al2013}; and (d) {\em imaging-48} for LOFAR (\citealt{2013A&A...556A...2V}; note that this assumes the ability to fully correlate the HBA-pairs).
+Note that (a) {\em hex} is based on the configuration used for HERA (14.6m center-to-center spacing); (b) {\em grid} for that of PAPER (4m N-S, 16m E-W spacing); (c) {\em imaging-128} for MWA \citep{tingay_et_al2013}; and (d) {\em imaging-48} for LOFAR (\citealt{2013A&A...556A...2V}; note that this assumes the ability to fully correlate the HBA-pairs in the Netherlands core).
 The plot shows that for a fixed number of elements using the delay-spectrum approach, redundant arrays (hex and grid) provide about an order of magnitude improvement over imaging arrays.
 Since the baselines go as order $N^2$, this corresponds to using about 1/3 of the number of elements to yield the same performance (as verified by the Hex-37/Imaging-128 lines).
-This advantage does not extend to the other approaches discussed below and imaging configurations would be preferred.
+This specific dependency does not extend to the other approaches discussed below and imaging configurations would likely be preferred.
 Note that a filled hexagonal-packed array provides excellent imaging as well, but with potentially limited resolution unless outriggers are included.
@@
 Another approach to improving sensitivity is to make the collecting area per element larger.
-In the limit, this yields a fully filled aperture.
-A larger element does limit the accessible field-of-view, but this is not a huge liability for a focussed experiment as long as an appropriate patch of ``cold'' sky passes within its field-of-view.
+A larger element does limit the accessible field-of-view, but this is not a huge liability for a focussed experiment as long as an appropriate patch of ``cold'' sky passes within its field-of-view and a large enough piece of sky is surveyed to overcome sample variance.
+These conditions hold true for HERA.
 Fig.
-\ref{fig:boost} (right) shows the relative sensitivity incorporating configuration and element size of various arrays relative to HERA-19, calculated for a redshift corresponding to the peak signal at a hydrogen reionization fraction of 0.5.
+\ref{fig:boost} (right) shows the relative sensitivity incorporating configuration and element size of various arrays relative to HERA-19 using \url{21cmSense}, calculated for a redshift corresponding to the peak signal at a hydrogen reionization fraction of 0.5 (in this case, z=8).
 The circles to the right indicate the relative sizes of the element.
@@
 Here, one is taking a 1-D Fourier transform along the other axis (the bandpass) and produces a 3-D spatial power spectrum.
-In imaging, point sources are transformed to a single pixel, whereas for the delay-spectrum, smooth sources are transformed to a single channel (but then get forced into the wedge due to the inherent chromatic response of an interferometer).
+In imaging, point sources are transformed to a single pixel, whereas for the delay-spectrum, smooth sources are transformed to a single channel.
 For imaging, we are concerned with ripples in the passband due to standing waves, whereas for the delay-spectrum we are concerned with delayed versions of the signal due to multiple reflections.
+Note that these are equivalent views of the same phenomena.
 
@@
 This redundancy may also be exploited to calibrate the array as well, as long as the true antenna positions fall close enough to that ideal grid \citep{liu_et_al2010}.
-This technique was pioneered with the MITEoR experiment \citep{zheng_et_al2014}, and has resulted in a package called OMNICAL developed for PAPER and HERA.
+This technique was pioneered with the MITEoR experiment \citep{zheng_et_al2014}, and has resulted in a package called OMNICAL developed for PAPER and HERA (\url{https://github.com/jeffzhen/omnical}).
 \citet{ali_et_al2015} provides a description of this technique as it applies to calibrating the PAPER array, which is mostly situated on a grid with many redundant baselines.
@@
 where $N_{\text core}$ and $D$ are the number of antennas in the core and their diameter respectively.
-Note that the linear number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} 21cmSense) over a range of values for the HERA configuration for wavenumbers greater than 0.15.
+Note that the linear number-diameter dependence was derived by running multiple sensitivity codes ({\em e.g.} \url{21cmSense}) over a range of values for the HERA configuration for wavenumbers greater than 0.15.
 This dependency differs from \cite{2013ExA....36..235M} and analytical expressions, mostly likely due to the assumed configuration.
@@
 relates directly to the leakage of smooth-spectrum foregrounds into regions of Fourier space 
-used to measure reionization, as discuss in Sec. \ref{sec:wedge}.
+used to measure reionization, as discuss in \S\ref{sec:wedge}.
 To facilitate foreground filtering, HERA's antenna element is designed to suppress reflections at long time delays.
@@
 dipole element incorporating a cylindrical backplane to
-improve symmetry in the polarized beam response patterns (Sec. \ref{sec:feed}, which also presents HFSS calculations of the full feed/element system).
+improve symmetry in the polarized beam response patterns (\S\ref{sec:feed}, which also presents HFSS calculations of the full feed/element system).
 
@@
 HERA's analog signal path from 70--250 MHz with the full performance EOR band of 100-200 MHz, as shown in Fig.
-\ref{fig:cosmos} and discussed in Sec.
-\ref{sec:freqs}, emphasizes spectral smoothness and robustness.
+\ref{fig:cosmos} and discussed in \S\ref{sec:freqs}, emphasizes spectral smoothness and robustness.
 Although the low-frequency sky has an intrinsically bright signal level, in the desired ``cold patches'' the sky temperature is only about 30 K at the upper frequency end, so relatively low receiver temperatures are desired.
@@
 Although correlator development has historically been one of the most 
-expensive aspects of building a large radio interferometer, this is no longer the case.
-The Collaboration for Astronomy Signal Processing and Electronics Research (CASPER) is an \Mycitep{parsons_et_al2006}
-open-source program to further the development of digital signal processing engines for astronomy and
+expensive aspects of building a large radio interferometer, this is no longer generally the case.
+The Collaboration for Astronomy Signal Processing and Electronics Research (CASPER) (\citealt{parsons_et_al2006}, \url{casper.berkeley.edu})
+is an open-source community to further the development of digital signal processing engines for astronomy and
 now has world-wide participation,
@@
 annually for five years running, each quadrupling the computational capacity of its predecessor.
-
 Led at UC Berkeley's Radio Astronomy Lab (RAL),
@@
 In order to meet the 35\,m specification for a maximum analog signal path length, along with a growing need for system scalability, HERA will 
-adopt a node-based architecture for amplification, digitization, channelization, and digital
-transmission in the field, building on MWA and Allen Telescope Array heritage. This design is merged with PAPER's clean 
+adopt a field-deployed node-based architecture for amplification, digitization, channelization, and digital
+transmission, building on MWA and Allen Telescope Array heritage. This design is merged with PAPER's clean 
 architecture for real-sampling and channelizing the entire analog passband at once, packetizing the data into
 10 Gb Ethernet format, and relying on commercial switches to perform the frequency/antenna corner-turn required for FX correlators.  
+This section will step through the digital signal path organized by location:  node, container, and processing building.
+
+\vspace{.2in}
+\noindent{\bf Node}. 
+HERA-350 employs RFI-tight node enclosures that each contain the final gain (PAM) and digitization stages for
+signals from 12 antennas, along with power supplies, cooling, sensors, and a small server for monitor/control. 
 This has led to a new analog-digital-optical board called the Smart Networked Analog-Digital Processor (SNAP), which will become a new modular CASPER device.  
 Figure \ref{fig:hardware} shows the series of primary CASPER processing boards, with SNAP on the far right.
-This section will step through the digital signal path organized by location:  node, container, processing building.
-
-{\bf Node}. HERA-350 employs RFI-tight node enclosures that each contain the final gain and digitization stages for
-signals from 12 antennas, along with power supplies, cooling, sensors, and a small server for monitor/control.  
-As part of previous HERA support,
-a new Smart Network ADC Processor (SNAP; Fig.~\ref{fig:hardware}) board has been incorporated 
-into the Collaboration for Astronomy Signal Processing and Electronics Research (CASPER) suite of hardware and firmware. This inexpensive board was co-designed by UC Berkeley and NRAO to be
+This inexpensive board was co-designed by UC Berkeley and NRAO to be
 both the digitizer and F-engine in HERA's FX correlator architecture.
-Each SNAP board 
-digitizes and channelizes a 0--250 MHz band for 6 input signals (3 antennas, dual-polarization).
+Each SNAP board digitizes and channelizes a 0--250 MHz band for 6 input signals (3 antennas, dual-polarization).
 A 200-MHz band of selectable channels is transmitted over optical fiber
@@
 tests have successfully demonstrated robust transmission from the SNAP board, through 10 km of fiber optic cable, 
-into a 10 GbE switch using inexpensive, commercially available optical transceivers.
-Activities under this proposal include 
-testing and extending FPGA firmware,
-integrating and testing all components in the node subsystem, and providing a monitor/control
-access interface.
+into a 10 GbE switch using inexpensive, commercially available optical transceivers.  For HERA, SNAP is housed in
+a field-deployable RFI-tight box -- for other applications it may be combined in a chassis.
+
+SNAP is an inexpensive and flexible ``RF in, 10 GbE out'' device which uses industry-standard SFP+ modules for output.
+These modules allow for copper or optical fiber output.  Additionally, for optical fiber as in the HERA system, the modules come in different colors
+which may then be multiplexed onto a single fiber within the node (labeled FMUX in Fig. \ref{fig:overallBlockDiagram}).
+The multiplexer module output is a single Quad SFP+ (QSFP+) connector which may be plugged directly into a switch
+as a demultiplexer.   These industry standards allow for inexpensive and very reliable data transport systems.
 
@@
 
-{\bf Central Container}.
+The node enclosure itself is an RFI-tight housing that can handle the signal paths for up to 12 antennas.
+The enclosures are placed throughout the array to minimize the total number constrained by the 35m maximum cable length.
+Note that the delay specification is not critical for the outriggers, so the cable-length to those may be as long as practical.
+This leads to 34 nodes, or an average of just over 10 antennas/node.
+The node uses forced-air, earth-coupled cooling similar to that implemented for the Allen Telescope Array \citep{2009IEEEP..97.1438W}.
+An AC-synchronous motor continually blows air through 30m of underground 150mm PVC pipe, which vents through the enclosure (in the bottom and out the top).
+The underground pipe is at least 1m deep, so the emerging air is very thermally stable on seasonal time constants.
+The PAM is the lowest component since it is the most sensitive and also generates the least heat, with the SNAP boards, monitor/control and power supplies above.
+Temperature, air-flow and power will be monitored at the node, as well as control to turn off/power cycle power to the different types of devices.
+
+\vspace{.2in}
+\noindent{\bf Central Container}.
 HERA's central container houses two significant subsystems adjacent to the array.  The first is a timing subsystem
@@
 
-{\bf Karoo Array Processing Building (KAPB)}.
-The KAPB houses the switch and CPUs
-that 
-complete the HERA correlator system.  A fiber optic bundle enters the KAPB and patches
+\vspace{.2in}
+\noindent{\bf Karoo Array Processing Building (KAPB)}.
+Approximately 10km (by fiber path) from the array central container is the KAPB -- a purpose-built facility for use
+in radio astronomy in the remote desert.
+The KAPB houses the switch and Graphical Processor Units (GPUs) that serve as the correlator cross-multiplier (the ``X-engine'').
+ A fiber optic bundle enters the KAPB and patches
 into local fiber optic cables. These cables terminate in optical transceivers which plug into a 240-port 10 GbE switch.
@@
 21cmFAST \citep{mesinger_et_al2011}, a fast, semi-numerical 21~cm signal simulator,
-and 21cmSense \citep{pober_et_al2014}, a tool for forecasting power spectrum sensitivity.
+and \url{https://github.com/jpober/21cmSense} \citep{pober_et_al2014}, a tool for forecasting power spectrum sensitivity.
 Other project code is aggregated and revision

### 2016-04-20 manuscript.tex
first quickpass through sec 6

@@
 \noindent On-site construction will proceed in stages, starting from the
-initial 37 elements that will be in place by project start.  These will be used
+initial 19 element currently in place.    Elements 20-37
+are currently under construction, to be completed in 2016.  
+ These will be used
 in conjunction with the thoroughly characterized extant PAPER signal path and
-processing hardware for a very low risk initiation of scientific observations
-from the very beginning.  Additionally, the 19 HERA elements currently in
-place will also have had an extensive period of commissioning
-and characterization.  As elements 38 to 128 are installed in year 1, they can
+processing hardware for a very low risk initiation of scientific observations.
+As elements 38 to 128 are installed in 2017, they can
 immediately be placed within the array and be used for observing.
 
-During this period, infrastructure for the new node-based system is installed
+During this period, infrastructure for the new node-based system will be installed
 and tested with the first elements beyond 128.  After the HERA-128 observing
-season, the full array is transitioned to use HERA's new hardware infrastructure.  In this same time frame, the existing PAPER processing
-container is moved to the edge of the array to house the timing sub-system
+season, the full array will be transitioned to use HERA's new hardware infrastructure.  
+In this same time frame, the existing PAPER processing
+container will be moved to the edge of the array to house the timing sub-system
 and fiber optic splice cabinet while the correlator will move to the KAPB.
-Also, the data processing activity at Penn is transitioned to NRAO.
-Summarizing, the hardware deployment phases are in four categories:
-\begin{enumerate}
-\item Element construction:  linear process (years 1, 2 \& 3)
-\item Infrastructure:  (a) node power and fiber reticulation (years 1 \& 2), (b) container relocation (year 1), (c) node installation (years 2 \& 3)
-\item Signal path:  convert to node-based system after 128 elements (years 2 \& 3)
-\item Processing:  (a) data processor/server to NRAO (year 1), (b) X-engine \& real-time processor/storage to KAPB (year 2)
-\end{enumerate}
+ The overall plan
+is shown in Figure \ref{fig:timeline}.
 
@@
 
-The optimal EoR observing window is from September to April, with power spectrum limit results appearing about one year later.
-Concurrent technique development and deployment will be on-going.
-Additional information is in the project management plan.
+The optimal EoR observing window is from September to April, with power spectrum limit results appearing about one year later.  Concurrent technique development and deployment will be on-going.
 \begin{itemize}[leftmargin=0.7in]
-\item[Year 1:] Observe with H37.  Real-time data pipeline, delay-space power spectrum (DSPS) pipeline, FHD pipeline.  21\,cm framework for incorporating with other probes. Construct H128.
-\item[Year 2:] Observe with H128.
+\item[2016/17:] Observe with H37.  Real-time data pipeline, delay-space power spectrum (DSPS) pipeline, FHD pipeline.  21\,cm framework for incorporating with other probes. Construct H128.
+\item[2017/18:] Observe with H128.
 Real-time calibration pipeline.
@@
 {\em Data products:  power spectrum, Stokes I maps.}
-\item[Year 3:] Observe with H240.
+\item[2018/19:] Observe with H240.
 Foreground-filtered imaging pipeline.
@@
 {\em Data products:  power spectrum, Stokes IQUV maps, foreground image cube.}
-\item[Year 4:] Observe with H350.  EoR parameter estimates.  H240 results.  {\em Data products:  power spectrum, global sky model IQUV, snapshots, foreground-filtered image cube.}
-%\item[Year 5:] Observe with H350.  H350 results.  {\em Data products:  power spectrum, global sky model IQUV, snapshots, foreground-filtered image cube.}
+\item[2019/20:] Observe with H350.  EOR parameter estimates.  H240 results.  {\em Data products:  power spectrum, global sky model IQUV, snapshots, foreground-filtered image cube.}
+\item[2020+:] Observe with H350.  H350 results.  {\em Data products:  power spectrum, global sky model IQUV, snapshots, foreground-filtered image cube.}
 \end{itemize}

### 2016-04-20 manuscript.tex
Added citations for ionization history figure

@@
     \includegraphics[width=0.51\textwidth,clip]{plots/ionHist.pdf}
-  \caption{Combining direct constraints on $x_\text{HI}$, the hydrogen neutral fraction, as a function of redshift (black points) with \emph{Planck} priors yields
-an inferred $95\%$ confidence region (gray).  HERA constraints with (dark red) and without (pale red) 
-a conservative $25\%$ modeling error in the $21\,\textrm{cm}$ power spectrum can dramatically narrow this confidence region.}
+  \caption{Combining direct constraints on $x_\text{HI}$, the hydrogen neutral fraction, as a function of redshift (black points) with \emph{Planck} priors \citep{planck_et_al2015} yields
+an inferred $95\%$ confidence region (gray; Greig \& Mesinger, in prep.).  HERA constraints with (dark red) and without (pale red) 
+a conservative $25\%$ modeling error in the $21\,\textrm{cm}$ power spectrum can dramatically narrow this confidence region \citep{liu_parsons2015}.
+Included for reference are constraints from the fraction of Lyman-alpha emitting galaxies \citep{pentericci_et_al2014,schenker_et_al2014}, quasar near-zone studies \citep{mortlock_et_al2011,bolton_et_al2011,schroeder_et_al2013}, Lyman-alpha galaxy clustering \citep{mcquinn_et_al2007_lya_clustering,ouchi_et_al2010}, and counts of dark Lyman-alpha pixels \citep{mcgreer_et_al2015}.}
 	\label{fig:IonHist}

### 2016-04-20 manuscript.tex
commit before pulling
Edited EoR window figure and caption

@@
 	\centering
-	\includegraphics[width=0.7\textwidth]{plots/cosmicEvo.png}
+	\includegraphics[width=0.7\textwidth]{plots/cosmicEvo.pdf}
 	\caption{Rendering of cosmic evolution from just after the Big Bang to today (background image credit Loeb/{\em Scientific American}).
@@
 The white dashed horizontal lines are the HERA EOR band and the blue are the extended frequency goal.
-Determining where the structure in the background really falls is the
-goal of HERA.} 
+%Determining where the structure in the background really falls is the goal of HERA.
+CMB observations observe the afterglow of the Big Bang (far left) and Baryonic Acoustic Oscillation (BAO) surveys proposed 
+target $z\approx$0.8-2.5.  Limited surveys span back to about $z\approx$7.} 
 	\label{fig:cosmos}
@@
 
-\begin{figure}[t]
-\centerline{
-\includegraphics[height=3in]{plots/wedges.pdf}
-}
-\caption{The EoR ``window'' and foreground ``wedge'' are shown in this plot, which splits the wavenumber into $\kvpr$ (horizontal) and $\kpar$ (vertical) components.
-Smooth-spectrum foregrounds with an intrinsic spectrum indicated by the gray bar along the bottom create the ``wedge'' when they interact with the interferometer's chromatic response, shown with a foreground image from the MWA.
+%\begin{figure}[t]
+%\centerline{
+%\includegraphics[height=3in]{plots/wedges.pdf}
+%}
+%\caption{The EoR ``window'' and foreground ``wedge'' are shown in this plot, which splits the wavenumber into $\kvpr$ (horizontal) and $\kpar$ (vertical) components.
+%Smooth-spectrum foregrounds with an intrinsic spectrum indicated by the gray bar along the bottom create the ``wedge'' when they interact with the interferometer's chromatic response, shown with a foreground image from the MWA.
 Systematics can extend the wedge to the upper left.
 The EoR window (shown with a simulation
-from \citealt{mesinger_et_al2011}) to the upper left should be free from foregrounds.}
-\label{fig:wedge}
+%from \citealt{mesinger_et_al2011}) to the upper left should be free from foregrounds.}
+%\label{fig:wedge}
+%\end{figure}
+
+\begin{figure}[t]
+	\centering
+%	\vspace{-10pt}
+	\includegraphics[width=1\textwidth,clip]{plots/Josh_Window_Data_Cartoon_v3.pdf}
+	\vspace{-10pt}
+	\caption{Foregrounds are a primary challenge facing 21\,cm cosmology experiments. 
+HERA leverages a ``wedge'' in the cylindrically-averaged ($\mathbf{k}$ is broken into $k_\|$ and $k_\perp$) power spectrum (center panel).
+Smooth-spectrum foregrounds (right panel) $\sim$5 orders of magnitude brighter than fiducial EoR models (left panel; \citealt{mesinger_et_al2011}) create the ``wedge'' when they interact with the interferometer's chromatic response.
+Flat spectrum foregrounds cannot contaminate the EoR window beyond the horizon wedge, though intrinsic or imperfectly calibrated chromaticity can push it slightly beyond the horizon.
+By avoiding foregrounds, PAPER has placed limits within an order
+of magnitude (in mK) of these models \citep{ali_et_al2015} and show the ``EoR Window'' to be foreground-free.
+HERA's dish and compact configuration optimize wedge/window isolation and direct sensitivity to low-$k_\perp$ modes where the EoR is brightest relative to the foreground wedge.
+}	\label{fig:wedge}
+%\vspace{-10pt}
 \end{figure}

### 2016-04-20 manuscript.tex
New abstract

@@
 
-The Hydrogen Epoch of Reionization Array (HERA) is a staged experiment that uses the unique properties of the 21\,cm line from neutral hydrogen to probe the Epoch of Reionization (EOR).
-During this epoch, roughly 0.3 - 1 billion years after the Big Bang, the first galaxies and black holes heated and reionized the early Universe.
-Direct observation of the large scale structure of reionization and its evolution with time will have a profound impact on our understanding of the birth of the first galaxies and black holes, their influence on the intergalactic medium (IGM), and cosmology.
-This paper will provide an overview of the project and describe the design of the system, moving from a science overview to requirements, design and status.
+The Hydrogen Epoch of Reionization Array (HERA) is a staged experiment
+optimized to measure 21 cm emission from the primordial
+intergalactic medium (IGM) throughout cosmic reionization ($z=6-12$), with
+support for exploring earlier stages of our Cosmic Dawn ($z\sim30$).  During these
+epochs, early stars and black holes heat and ionize the IGM, introducing
+fluctuations in 21 cm emission.  HERA aims to characterize the evolution of the 21 cm
+power spectrum in detail to constrain the timing and morphology of
+reionization, the properties of the first galaxies, the evolution of
+large-scale structure, and the early sources of heating in the IGM.  
+The full HERA instrument will be a
+350-element interferometer in South Africa consisting of 14-m parabolic dishes observing from 50 to 250 MHz.
+Currently, 19 dishes have been deployed on site and are being commissioned.
+
+In this paper, we summarize HERA's scientific context and provide forecasts for
+its key science results.  After reviewing the current state of the art in
+foreground mitigation, we use the delay-spectrum technique to
+motivate high-level performance requirements for the HERA instrument.  Next,
+we present the HERA instrument design, along with the subsystem specifications that
+ensure that HERA meets its performance requirements.  Finally, we summarize the schedule
+and status of the project.  We conclude by suggesting that, given the realities of foreground
+contamination, current-generation 21 cm instruments are approaching their sensitivity limits.
+HERA is designed to bring both the
+sensitivity and the precision to deliver its primary science on the basis of
+proven foreground filtering techniques, while developing new subtraction
+techniques to unlock new capabilities.  The result will be a major step toward
+realizing the widely recognized scientific potential of 21 cm cosmology.
 
 \section{Introduction}
+
 The Hydrogen Epoch of Reionization Array (HERA; http://reionization.org) is a staged experiment to use the redshifted 21\,cm line

### 2016-04-20 manuscript.tex
Tweaks and polishing of Introduction

@@
 structure of the primordial IGM as it is heated and reionized, HERA complements
-other probes, adding transformative capabilities for
+probes at other wavelengths, adding transformative capabilities for
 understanding the astrophysics and fundamental cosmology of our early universe.
@@
 
-Because of these key science questions that may be probed by highly-redshifted hydrogen emission, there has been a resurgence in low frequency arrays  -- most with a primary
+The power that observations of highly redshifted hydrogen emission have for answering key science about our early universe
+has motivated a resurgence of interest in low-frequency arrays --- most with the primary
 objective of measuring the EOR.
 These include the Precision Array Probing the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}), the Murchison Widefield Array (MWA; \citealt{tingay_et_al2013}), the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V}), and the Long Wavelength Array (LWA; \citealt{2009IEEEP..97.1421E}), as well as systems for existing dish arrays like the Very Large Array (VLA; \citealt{2013AAS...22132804K}) and the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2011}).
-With the first generation 21\,cm telescopes approaching the limits of their sensitivity however, HERA continues the experiment by bringing significantly more sensitivity to bear
-while continuing to optimize its design along the delay-spectrum approach to detect the power spectrum of the EoR (discussed in \S\ref{sec:delayapproach}).
-At the same time, HERA continues the development of new techniques to extend and improve its capabilities.
-Examples include improving foreground subtraction to expand the information content 
-in power spectral and imaging measurements and
-extending HERA's performance at lower frequencies to probe IGM heating in the pre-reionization epoch (\S\ref{sec:EoX}).
+These experiments all struggle with the
+inherent challenge of simultaneously meeting stringent sensitivity requirements
+while suppressing foregrounds $\sim5$ orders of magnitude
+brighter than the 21\,cm signal.
+HERA improves on its predecessors by bringing significantly more sensitivity to bear
+on the angular and spectral scales where recent work (discussed in Section \ref{sec:delayapproach}) has indicated that
+the power spectrum of the EoR may dominate over foregrounds.
+% XXX ARP I thought this was too early in the paper for the stuff below.  Relocate?
+%At the same time, HERA continues to develop new techniques to extend and improve its capabilities.
+%Examples include improving foreground subtraction to expand the information content 
+%in power spectral and imaging measurements and
+%extending HERA's performance at lower frequencies to probe IGM heating in the pre-reionization epoch (\S\ref{sec:EoX}).
 
@@
 350 14-m parabolic dishes (320 in a dense core $+$ 30 outriggers) in the South African Karoo Radio
-Astronomy Reserve (see Fig. \ref{fig:HERApictures} (top) for a rendering).
-HERA has been funded under the US National Science Foundation's Mid-Scale Innovations Program to begin building elements optimized for robust power spectrum detection in a 
-staged deployment.  The key is to produce inexpensive collecting area optimized to detect the EOR.  
-The first 19-element array is complete and, with additional funding from Cambridge University, construction underway on the next 18 elements, for a total of 37.
-The South African National Research Foundation SKA-SA group (SKA-SA) is a key partner in the construction and science.
-
-The current 19 elements are testing the HERA element design for performance and manufacturability on location.  The 37 elements, to be completed in 2016, will provide a factor of about 5 more 
-sensitivity than PAPER-128 and provide a significant chance at detecting the EoR power spectrum signal.
-The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER,
-%\citep{ali_et_al2015}
-which should provide a very robust detection of the EOR signal and allow some characterization.  Finally, extending to 2020, HERA will build out to 350 elements to further EOR science as a function 
-of redshift and spatial scale, potentially producing the first images of the EOR.  These estimates will be quantified and shown in this paper.
-
-HERA is a collaboration of the following Partner institutions:  
+Astronomy Reserve (see Fig. \ref{fig:HERApictures}, top, for a rendering).
+HERA's antenna element and its compact configuration are
+optimized for robust power spectrum detection,
+delivering the requisite collecting area inexpensively, while carefully controlling foreground systematics.
+The design and construction of this instrument is supported by the HERA collaboration 
+consisting of the following Partner institutions:  
 Arizona State University (Tempe, AZ USA), 
@@
 California State Polytechnic University (Pomona, CA USA).
+The South African National Research Foundation SKA-SA group (SKA-SA) is a key partner in HERA's construction and science.
+
+HERA's first stage of development has been funded under the US National Science Foundation's Mid-Scale Innovations Program,
+which has supported the construction of a 19-element array for testing instrumental performance
+and manufaturability on location.  
+The first 19-element array is complete and, with additional funding from Cambridge University, 
+construction is now underway on the next 18 elements.
+The total 37 elements, to be completed in 2016, will provide a factor of about 5 more 
+sensitivity than PAPER-128 and provide a significant chance at detecting the EoR power spectrum signal.
+The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER,
+%\citep{ali_et_al2015}
+HERA-128 should provide a robust detection of the EOR signal and allow some characterization.  Finally, extending to 2020, HERA will build out to 350 elements to further EOR science as a function 
+of redshift and spatial scale, potentially producing the first images of the EOR.  
+%These estimates will be quantified and shown in this paper.
 
 This paper has the following structure.
-Section \ref{sec:science} will give a high-level overview of the science motivating HERA's construction, including primary and secondary science objectives.
+Section \ref{sec:science} gives a high-level overview of the science motivating HERA's construction, including primary and secondary science objectives.
 Key science results include
@@
 directly imaging large scale reionization structure (\S\ref{sec:imaging}).
-Additionally, beyond these directly supported applications, HERA's public data products and instrumentation platform enable other high-impact science, including 
+Beyond these directly supported applications, HERA's public data products and instrumentation platform enable other high-impact science, including 
 21\,cm cross-correlation (e.g. with JWST, WFIRST, CO, CII, and Ly$-\alpha$; \S\ref{sec:crossCorr});
@@
 
-Section \ref{sec:eormeas} will present the techniques used to make the measurement.
-Section \ref{sec:requirements} will present the high-level requirements and Section \ref{sec:design} will present the system description..
-Section \ref{sec:status} will present a brief status and outline the deployement timetable and Section \ref{sec:conclusion} will conclude by summarizing and providing additional context.
+Section \ref{sec:eormeas} presents the techniques used to make the measurement.
+Section \ref{sec:requirements} presents the high-level requirements and Section \ref{sec:design} presents the system description..
+Section \ref{sec:status} presents a brief status and outline the deployement timetable and Section \ref{sec:conclusion} concludes by summarizing and providing additional context.
 

### 2016-04-20 manuscript.tex
Removed a lot of "our"s

@@
 sensitivity than PAPER-128 and provide a significant chance at detecting the EoR power spectrum signal.
-The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER,
+The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER.
 %\citep{ali_et_al2015}
@@
 their ability to constrain reionization, and will be for the foreseeable future.
-HERA will use
+HERA uses
 another complementary probe --- the 21\,cm ``spin-flip" transition of neutral hydrogen
@@
 
-HERA's 21\,cm measurements can be used in conjunction with our semi-analytic models to constrain the ionization history. 
+HERA's 21\,cm measurements can be used in conjunction with semi-analytic models to constrain the ionization history. 
 The red band in Fig.~\ref{fig:IonHist} shows the forecasted 95\% confidence region derived from HERA data after marginalizing over astrophysical and cosmological parameters.
@@
 still allowed by neutrino oscillation experiments
-\citep{allison_et_al2015}, making HERA key to our understanding of neutrino
+\citep{allison_et_al2015}, making HERA key to understanding neutrino
 physics. HERA's estimate of $\tau$ would also break the degeneracy between
@@
 The goal is therefore to measure a range of aggregate spatial scales on the sky, rather than to image the signal directly.
-Imaging does remain an ultimate goal to fully understand the process, however we will likely need a greater understanding of the signal characteristics and the systematics of our telescopes to achieve this more difficult goal.
+Imaging does remain an ultimate goal to fully understand the process, however we will likely need a greater understanding of the signal characteristics and the systematics to achieve this more difficult goal.
 
@@
 If we assume 14.2m, then we can estimate that each reflection must average a return loss better than 15 dB, an aggressive target.
-And finally, if we need to allow for 3 round-trips, our antenna diameter would have to be less than 9.4m, where the cost is quickly rising.
+And finally, if we need to allow for 3 round-trips, the antenna diameter would have to be less than 9.4m, where the cost is quickly rising.
 To handle the delay requirement therefore, the diameter should be less than 14.2m, or the return loss specification becomes very aggressive.
@@
 than 20 times that of a PAPER element, but it does so without substantially
-degrading our ability to isolate and remove foreground emission on the basis of
+degrading the ability to isolate and remove foreground emission on the basis of
 spectral smoothness.  As shown in
@@
 HERA's data management system archives this on-site in a 1.5 PB storage array, calibrates it in real-time
-on the basis of redundancy, compresses it by a factor of 20, and transfers it to our computing facility at NRAO.
+on the basis of redundancy, compresses it by a factor of 20, and transfers it to a computing facility at NRAO.
 This real-time pipeline, based on a similar system for PAPER, is mature, computationally manageable, and 
@@
 
-\noindent With first-generation instruments, the HERA team has led a revolution in our understanding of how to best perform 21\,cm cosmology measurements.
+\noindent With first-generation instruments, the HERA team has led a revolution in the understanding of how to best perform 21\,cm cosmology measurements.
 In the past three years, we have developed the EoR window paradigm for isolating foreground systematics, implemented novel
@@
 
-HERA is ``the right instrument at the right time.'' With first generation experiments approaching the limits of their sensitivity, it is now critical to build upon our new knowledge and experience to advance beyond them.
+HERA is ``the right instrument at the right time.'' With first generation experiments approaching the limits of their sensitivity, it is now critical to build upon new knowledge and experience to advance beyond them.
 HERA's large collecting area and optimized design enable precise constraints of EoR astrophysics and a broad range of high-impact secondary science, all before SKA-low becomes operational.

### 2016-04-21 manuscript.tex
Jack's pass on Digital Section.

@@
 \label{sec:digital}
-
-Although correlator development has historically been one of the most 
-expensive aspects of building a large radio interferometer, this is no longer generally the case.
-The Collaboration for Astronomy Signal Processing and Electronics Research (CASPER) (\citealt{parsons_et_al2006}, \url{casper.berkeley.edu})
-is an open-source community to further the development of digital signal processing engines for astronomy and
-now has world-wide participation,
-with over 500 members at 73 institutions, and 
-five generations of hardware.
-PAPER has applied CASPER technology to develop and deploy new correlators
-annually for five years running, each quadrupling the computational capacity of its predecessor.
-Led at UC Berkeley's Radio Astronomy Lab (RAL),
-HERA efforts continue this incremental development cycle, using a packet-switched
-correlator architecture \Mycitep{parsons_et_al2008} that both PAPER and LEDA \citep{2015JAI.....450003K} have
-applied to systems employing FPGAs and GPUs \citep{clark_et_al2011}.
-
-In order to meet the 35\,m specification for a maximum analog signal path length, along with a growing need for system scalability, HERA will 
-adopt a field-deployed node-based architecture for amplification, digitization, channelization, and digital
-transmission, building on MWA and Allen Telescope Array heritage. This design is merged with PAPER's clean 
-architecture for real-sampling and channelizing the entire analog passband at once, packetizing the data into
-10 Gb Ethernet format, and relying on commercial switches to perform the frequency/antenna corner-turn required for FX correlators.  
-This section will step through the digital signal path organized by location:  node, container, and processing building.
+%The computational cost of a correlator scales with the square of the number of antennas in an array.
+Although in the past, correlator development has been one of the most expensive and complex aspects of building a large array or radio interferometers, this is no longer generally the case.
+The Collaboration for Astronomy Signal Processing and Electronics Research (CASPER) (\citealt{parsons_et_al2006}, \url{casper.berkeley.edu}) is a community of astronomers and engineers who work to reduce the cost and complexity of radio astronomy signal processing systems through the development of open-source, general-purpose hardware and software.
+CASPER currently has several hundred members at 73 institutions, and has developed six generations of FPGA-based signal processing hardware, shown in Fig. \ref{fig:hardware}.
+PAPER has applied CASPER technology to develop and deploy new correlators annually for five years running, each quadrupling the computational capacity of its predecessor.
+Key to the upgradability of the PAPER correlator is the use of modular processing engines, and industry-standard digital interconnect based on off-the-shelf Ethernet switches \Mycitep{parsons_et_al2008} to perform the antenna/frequency data transpose required by FX correlators.
+
+HERA will maintain both PAPER's well-proven digital system architecture along with the simple scheme of real-sampling and channelizing the entire analog passband at once.
+However, in order to meet the 35\,m specification for maximum analog signal path length, as well as ensure future scalability of the system, HERA-350 will adopt an architecture of field-deployed amplification, digitization, and channelization nodes, building on MWA and Allen Telescope Array heritage.
+Digital data streams from multiple nodes will be converge on a container by the HERA array, from which they will be routed to a central processor building where correlation and further processing will take place.
+This section will step through the digital signal path, shown in Fig. \ref{fig:overallBlockDiagram}, organized by location:  node, container, and processing building.
+
+%Led at UC Berkeley's Radio Astronomy Lab (RAL), HERA efforts continue this development cycle, using a packet-switched correlator architecture \Mycitep{parsons_et_al2008} that both PAPER and LEDA \citep{2015JAI.....450003K} have
+%applied to systems employing FPGAs and GPUs \citep{clark_et_al2011}.
 
@@
 \noindent{\bf Node}. 
-HERA-350 employs RFI-tight node enclosures that each contain the final gain (PAM) and digitization stages for
-signals from 12 antennas, along with power supplies, cooling, sensors, and a small server for monitor/control. 
-This has led to a new analog-digital-optical board called the Smart Networked Analog-Digital Processor (SNAP), which will become a new modular CASPER device.  
-Figure \ref{fig:hardware} shows the series of primary CASPER processing boards, with SNAP on the far right.
-This inexpensive board was co-designed by UC Berkeley and NRAO to be
-both the digitizer and F-engine in HERA's FX correlator architecture.
-Each SNAP board digitizes and channelizes a 0--250 MHz band for 6 input signals (3 antennas, dual-polarization).
-A 200-MHz band of selectable channels is transmitted over optical fiber
-to a central container (see below), and on to the correlator.  Laboratory
-tests have successfully demonstrated robust transmission from the SNAP board, through 10 km of fiber optic cable, 
-into a 10 GbE switch using inexpensive, commercially available optical transceivers.  For HERA, SNAP is housed in
-a field-deployable RFI-tight box -- for other applications it may be combined in a chassis.
-
-SNAP is an inexpensive and flexible ``RF in, 10 GbE out'' device which uses industry-standard SFP+ modules for output.
-These modules allow for copper or optical fiber output.  Additionally, for optical fiber as in the HERA system, the modules come in different colors
-which may then be multiplexed onto a single fiber within the node (labeled FMUX in Fig. \ref{fig:overallBlockDiagram}).
-The multiplexer module output is a single Quad SFP+ (QSFP+) connector which may be plugged directly into a switch
-as a demultiplexer.   These industry standards allow for inexpensive and very reliable data transport systems.
+HERA-350 employs RFI-tight node enclosures that each contain the final gain (PAM) and digitization stages for signals from 12 antennas, along with power supplies, cooling, sensors, and a small server for monitor/control.
+This has led to the development of a new CASPER platform based around a Xilinx Kintex 7 FPGA, which incorporates on-board ADCs: the Smart Networked Analog-Digital Processor (SNAP, shown in Fig.
+\ref{fig:hardware}).
+SNAP is an inexpensive and flexible ``Analog in, 10 GbE out'' device which uses a pair of industry-standard SFP+ modules for output over either copper or fiber cables.
+Co-designed by UC Berkeley and NRAO, the SNAP board will serve as both the digitizer and F-engine in HERA's FX correlator architecture.
+Each SNAP board digitizes and channelizes a 0--250 MHz band for 6 input signals (3 antennas, dual-polarization), with a complete node containing 4 SNAP boards.
+
+Following digitization and channelization, a 200-MHz band of runtime-selectable channels is output as a UDP stream over optical fiber.
+Coarse Wavelength Division Multiplexing (CWDM) technology allows the 10 Gb/s Ethernet streams from the four SNAP boards in a node to share a single fiber, which is routed to a central container, and on to the correlator.
+In this configuration, the fiber from each node is input directly into an Ethernet switch using a single QSFP+ 40GBASE-LR4 transciever, which provides built-in multiplexing and demultiplexing capabilities and may be operated as 4 independent transceivers.
+Laboratory tests have successfully demonstrated robust transmission from the SNAP board, through CWDM multiplexers and 10 km single mode fiber optic cable, into a 10 GbE switch using inexpensive, commercially available optical transceivers.
 
@@
 Six generations of CASPER digital signal processing (left to right) culminating in the SNAP board (right, along with the long-haul fiber-link test setup), which was developed and tested under the previous HERA MSIP award.
-By preserving its toolflow, signal processing libraries, and interface code between hardware generations,
-CASPER combines with a scalable correlator architecture to enable the PAPER correlator to be 
-easily upgraded for HERA \Mycitep{parsons_et_al2006,parsons_et_al2008}.
+By preserving its design tools, signal processing libraries, and interface code between hardware generations, CASPER hardware and a modular architecture have enabled the PAPER correlator to be easily upgraded for HERA \Mycitep{parsons_et_al2006,parsons_et_al2008}.
 %
@@
 
+While some applications may utilise SNAPs in multiple-board chasses, for HERA, SNAP is housed in a field-deployable RFI-tight box.
 The node enclosure itself is an RFI-tight housing that can handle the signal paths for up to 12 antennas.
-The enclosures are placed throughout the array to minimize the total number constrained by the 35m maximum cable length.
-Note that the delay specification is not critical for the outriggers, so the cable-length to those may be as long as practical.
-This leads to 34 nodes, or an average of just over 10 antennas/node.
+The enclosures are placed throughout the array to minimize the total number constrained by the 35\,m maximum cable length. 
+For outrigger antennas the delay specification is not critical, so the cable-length to those may be as long as practical.
+In total, including outriggers, the HERA-350 design demands 34 nodes, or an average of just over 10 antennas/node.
+
 The node uses forced-air, earth-coupled cooling similar to that implemented for the Allen Telescope Array \citep{2009IEEEP..97.1438W}.
-An AC-synchronous motor continually blows air through 30m of underground 150mm PVC pipe, which vents through the enclosure (in the bottom and out the top).
-The underground pipe is at least 1m deep, so the emerging air is very thermally stable on seasonal time constants.
-The PAM is the lowest component since it is the most sensitive and also generates the least heat, with the SNAP boards, monitor/control and power supplies above.
-Temperature, air-flow and power will be monitored at the node, as well as control to turn off/power cycle power to the different types of devices.
+An AC-synchronous motor continually blows air through 30\,m of underground 150\,mm PVC pipe, which vents through the enclosure from bottom to top.
+The underground pipe is at least 1\,m deep, so that the emerging air is very thermally stable on seasonal time constants.
+The PAM is the lowest component since it is the most sensitive and also generates the least heat.
+SNAP boards, monitor and control hardware, and power supplies will be positioned above the PAMs.
+Remote monitoring capabilities will include node temperature, air-flow and power, with remote control of power to the node's various active components.
 
@@
 \noindent{\bf Central Container}.
-HERA's central container houses two significant subsystems adjacent to the array.  The first is a timing subsystem
-that maintains a GPS-disciplined oscillator and distributes timing
-signals (the sampling clock and 1 pulse-per-second synchronization) to the nodes.  The second
-subsystem is a passive fiber optic patch panel that couples
-the optical network from the nodes into the 192-filament optical fiber bundle 
-that connects to the KAPB.
+HERA's central container houses two significant subsystems adjacent to the array.
+The first is a timing subsystem that maintains a GPS-disciplined oscillator and distributes timing signals to the nodes.
+These signals comprise a sampling clock (or other frequency reference from which the sampling clock may be derived) and 1 pulse-per-second (PPS) synchronization pulses.
+The timing subsystem will distribute clock and synchronisation signals over fiber and, depending on cost and performance, will utilise either industry-standard distribution solutions, or the White Rabbit protocol \Mycitep{Moreira_WR}.
+
+The second subsystem is a passive fiber optic patch panel that couples the optical network from the nodes into the 192-filament optical fiber bundle that is routed to the KAPB.
 
@@
 \noindent{\bf Karoo Array Processing Building (KAPB)}.
-Approximately 10km (by fiber path) from the array central container is the KAPB -- a purpose-built facility for use
-in radio astronomy in the remote desert.
-The KAPB houses the switch and Graphical Processor Units (GPUs) that serve as the correlator cross-multiplier (the ``X-engine'').
- A fiber optic bundle enters the KAPB and patches
-into local fiber optic cables. These cables terminate in optical transceivers which plug into a 240-port 10 GbE switch.
-Such switches, while large, are readily available commercially today.  Also connected to
-this switch are 30 servers, each hosting two dual-GPU graphics cards and two dual
-10 GbE network interface cards, which implement the cross-multiplication (X-Engine) component
-of the correlator during observations.  This estimate for the number of X-Engine servers
-is extrapolated from current GPU servers deployed on PAPER, assuming no improvement in bus
-speeds for transferring data into the GPU cores, but assuming that the computational
-capacity of GPU cores doubles once according to Moore's Law prior to purchasing
-these servers in year three of the project.
-Output data from the correlator are written to the data storage system described
-in the following section.
+Approximately 10\,km (by fiber path) from the array central container is the KAPB -- a purpose-built facility to house radio astronomy computing resources in the remote desert.
+A fiber optic bundle from the HERA container enters the KAPB and is patched into local fiber cables, which terminate in QSFP+ optical transceivers.
+These QSFP+ transceivers are input to a 64-port QSFP+ 10/40 GbEthernet switch -- such devices are readily available on the commercial market today, at relatively low cost.
+The Ethernet switch forms the core of the HERA correlator's data interconnect, and provides the antenna-frequency data transpose which allows cross-multiplication of antenna signals to be easily parallelised by frequency over many compute nodes performing cross-multiplication (``X-Engines'').
+
+These compute nodes are anticipated to be 30 x86 servers, each hosting a pair of Nvidia Graphics Processing Units (GPUs) and two dual 10GbE network interface cards.
+This estimate assumes an identical configuration to the current PAPER X-Engine hardware, and conservatively assumes that the computational capacity of GPU accelerator cards will double once prior to purchasing these servers in year three of the project.
+No improvement in data transfer speed from CPU to host is assumed.
+
+Other X-Engine realizations are under active evaluation, including a solution involving AMD GPUs and the highly-optimised cross-multiplication kernels developed for the CHIME Pathfinder array \Mycitep{Denman_Chime_X}, and FPGA-based implementations involving both CASPER, and commercially available, platforms.
+
+Output data from the correlator are written to the data storage system described in the following section.
 

### 2016-04-21 manuscript.tex
included jonnies 2nd set of comments

@@
 \title{Hydrogen Epoch of Reionization Array (HERA)}
-\author{A.U.Thor}
+\author{David R.
+DeBoer, Aaron R.
+Parsons, Judd D.
+Bowman, Jonathan C.
+Pober, Jacqueline N.
+Hewitt, Chris L.
+Carilli, Gianni Bernardi, Anita Loots, Andrei Messenger, Paul Alexander, Steve Furlanetto, James E.
+Aguirre, Miguel F.
+Morales, Adam Beardsley, Daniel C Jacobs, Nithyanandan Thyagarajan, Zak, Aaron Ewall-Wice, Abraham R.
+Neben, Max Tegmark, Haoxuan Zheng, Richard Bradley, Matthys Maree, Kathryn Rosie, Randall Fritz, Nathan Mathison, Cresshim Malgas, Raddwine Sell, Samantha Pieterse, Lourence Malan, Austin Julius, MacCalvin Kariseb, Craig Smith, Jasper Grobbelaar, Telalo Lekalake, Eunice Matsetela, Angelo Syce, Bradley Grieg, Eloy de Lera Acedo, Nicolas Fagnoni, Nima Razavi-Ghods, Zaki S.
+Ali, Carina Cheng, Josh Dillon, Jack Hickish Adrian Liu, David MacMahon, Nipanjana Patra, Saul A.
+Kohn, Bryna Hazelton, Peter Williams}
 
@@
 The cost-performance optimization will
-be discussed below, however we see that to address the central challenges of the delay-spectrum approach in the context of the wedge.
+be discussed below, however we see that to address the central challenges of the delay-spectrum approach in the context of the wedge
 HERA should use close-packed antennas that
@@
 \label{sec:delayspec}
-Early analysis \citep{elementmemo} indicated that to trace the EOR over a wide range of redshift, internal reflections should be attenuated by about 60 dB from the initial incoming wave by about 60 ns, or a VSWR $<$ 1.002 for frequencies $<$ 17 MHz.
+Early analysis \citep{elementmemo} indicated that to trace the EOR over a wide range of redshift, internal reflections should be attenuated by about 60 dB from the initial incoming wave by about 60 ns, or a voltage standing wave ratio (VSWR) $<$ 1.002 for frequencies $<$ 17 MHz.
+This was a conservative value based on estimates of power in spectrally smooth foreground sources from \cite{santos_et_al2005,2008MNRAS.385.2166A, deoliveira2008, jelic_et_al2008, bernardi_et_al2010} and experience with PAPER dealing with foreground systematics.
 This provided the basis for the initial design of the element.
@@
 The color coding is the product of the two assigned costs, with red showing a high cost normalized to 350 14m elements (clipped at a ratio of 1.25) and purple being the minimum.
-The curves labelled `1', `2' and `3' are the round-trip travel times for 1, 2 and 3 reflections between the feed and the vertex.
+The black lines labelled `1', `2' and `3' are the round-trip travel times for 1, 2 and 3 reflections between the feed and the vertex.
 To obtain 60 dB by 60 ns we may interpret those lines in one of two ways:  
@@
 
+\begin{figure}[t]
+\centerline{
+\includegraphics[width=0.6\textwidth]{plots/costfig.pdf} 
+}
+\caption{\small The background coloring with colorbar shows the relative cost/performance for delay and cost at a fixed performance as a function of element diameter.
+See \S\ref{sec:delayspec}, \S\ref{sec:cost} and \cite{elementmemo} for details.
+A diameter of 14-m (vertical dashed line) is near the cost minimum and consistent with assumption for the delay specification.
+\label{fig:costfig}}
+\end{figure}
+
 As a key specification, this has been a focus of study with a series of prototypes and analyses based on modeling, and measurements have been conducted to provide much more rigor than this early specification \citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016,patra_et_al0216}.
 The simple specification set above more appropriately becomes a series of curves for attenuation and delay at different wavenumbers for realistic foreground models.
-Fig \ref{fig:delayspec}, adapted from \cite{thyagarajan_et_al2016}, shows the results at $k$ = 0.10, 0.15, and 0.20 $h$/Mpc for the adopted design.
+Figure \ref{fig:delayspec}, adapted from \cite{thyagarajan_et_al2016}, shows the results at $k$ = 0.10, 0.15, and 0.20 $h$/Mpc for the adopted design.
 The analysis shows that
-the dish design for should should allow for EOR detections for $k>0.10 h/$Mpc.
-We also see that the initial simple ``60@60'' specification was a overly conservative, however it served as a good initial goal.
+the dish design for should allow for EOR detections for $k>0.10 h/$Mpc.
+We also see that the initial simple ``60@60'' specification was a overly conservative (shown as the lightly shaded area bounded by the teal line), however it served as a good initial goal.
 
@@
 	%\caption{Left:  
-	\caption{Antenna delay attenuation specification with realistic foregrounds from \cite{thyagarajan_et_al2016} along with the initial conservative ``60@60'' specification.
-HFSS calculations (black line) show the HERA dish meeting the specification.
-Measurements on the Green Bank prototype are shown in  \citep{patra_et_al2016}.
+	\caption{Antenna delay attenuation specification with realistic foregrounds from \cite{thyagarajan_et_al2016}.  
+	The blue, green and red lines show the maximum attenuation needed at a given delay for the foregrounds to be below the expected EOR signal level for different $\kpar$.
+	The teal line is the ``60@60'' initial conservative specification.
+	 HFSS calculations (black line) for the expected attenuation show the HERA dish meeting the more sophisticated specification. 
+	 %Measurements on the Green Bank prototype are shown in  \citep{patra_et_al2016}.
 }
@@
 The resulting normalized system cost/performance-curve  becomes very flat (colored dark blue/purple) for diameters greater between about 12 and15 m, which is consistent with the chosen value of 14 meters based on other system considerations.
-
-\begin{figure}[t]
-\centerline{
-\includegraphics[width=0.6\textwidth]{plots/costfig.pdf} 
-}
-\caption{\small The background coloring with colorbar shows the relative cost/performance for delay and cost at a fixed performance as a function of element diameter.
-See text and \cite{elementmemo} for details.
-A diameter of 14-m (vertical dashed line) is near the cost minimum and consistent with assumption for the delay specification.
-\label{fig:costfig}}
-\end{figure}
 
@@
 onto a 10 km fiber bundle connecting to the Karoo Array Processing Building, where signals are cross-multiplied
-in the X processor.  After correlation, visibilities are stored by the Librarian, compressed and redundantly calibrated
+in the X processor.  After correlation, visibilities are stored by the Librarian, pre-processed and redundantly calibrated
 by the Real-Time Processor, and transmitted over the network to clusters for storage and analysis.  Final products are
@@
 \item minimize costs, and
-\item the experiment has a limited lifetime of about five years.
+\item design for a limited lifetime of about five years.
 \end{itemize}
@@
 relates directly to the leakage of smooth-spectrum foregrounds into regions of Fourier space 
-used to measure reionization, as discuss in \S\ref{sec:wedge}.
+used to measure reionization, as discussed in \S\ref{sec:wedge}.
 To facilitate foreground filtering, HERA's antenna element is designed to suppress reflections at long time delays.
@@
 
-As shown in Fig.~\ref{fig:delayspec}, HERA's element meets this requirement in
-reflectometry tests, performing comparably to the PAPER element.  The
-relatively short (4.5\,m) focal height of the parabolic dish 
-minimizes the timescale of feed-dish reflections while maintaining the feed's
-illumination of the dish.  HERA's feed design is based on the PAPER sleeved
-dipole element incorporating a cylindrical backplane to
-improve symmetry in the polarized beam response patterns (\S\ref{sec:feed}, which also presents HFSS calculations of the full feed/element system).
+%As shown in Fig.~\ref{fig:delayspec}, HERA's element meets this requirement in
+%reflectometry tests, performing comparably to the PAPER element.  The
+%relatively short (4.5\,m) focal height of the parabolic dish 
+%minimizes the timescale of feed-dish reflections while maintaining the feed's
+%illumination of the dish.  HERA's feed design is based on the PAPER sleeved
+%dipole element incorporating a cylindrical backplane to
+%improve symmetry in the polarized beam response patterns (\S\ref{sec:feed}, which also presents HFSS calculations of the full feed/element system).
 
@@
 	\begin{minipage}{2.05in}
-	\caption{Left: The first of three prototype dishes at NRAO--Green Bank, used for measuring beam frequency structure with reflectometry and the beam pattern at 137\,MHz by comparing satellite signals to the reference dipole in the foreground \citep{neben_et_al2016}.
+	\caption{Left: The first of two prototype dishes at NRAO--Green Bank, used for measuring beam frequency structure with reflectometry and the beam pattern at 137\,MHz by comparing satellite signals to the reference dipole in the foreground \citep{neben_et_al2016}.
 Right: The measured EW power pattern plotted with dashed lines marking zenith angles of 20$^\circ$, 40$^\circ$, 60$^\circ$, 80$^\circ$.}
@@
 This requires a densely-packed core of antenna elements, maximizing sensitivity to short baselines---those least contaminated by foreground chromaticity ({\em i.e} the wedge). 
-Furthermore, the configuration should be highly redundant, both to enable the delay-spectrum strategy of PAPER \citep{parsons_et_al2012a} and redundant baseline calibration \citep{liu_et_al2010,zheng_et_al2014}.
+Furthermore, the configuration should be highly redundant, both to increase sensitivity for the delay-spectrum strategy of PAPER \citep{parsons_et_al2012a} and provide for redundant baseline calibration \citep{liu_et_al2010,zheng_et_al2014}.
 On the other hand, it is also desirable to improve the mapmaking ability of the array for calibration and map-based power spectrum estimation with longer baselines and a densely-sampled $uv$-plane \citep{dillon_et_al2015a}.
@@
 HERA's 320 core elements are arranged in a compact hexagonal grid, split into three displaced segments to cover the $uv$-plane with sub-aperture sampling density (see Fig.~\ref{fig:arrayConfig}).
-Splitting the core into three sections triples the density of instantaneous sampling of the $uv$-plane, improving HERA's mapmaking ability.Furthermore, the core is supplemented by 30 additional outrigger elements to tile the $uv$-plane with instantaneously complete sub-aperture sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz).
+Splitting the core into three sections triples the density of instantaneous sampling of the $uv$-plane, improving HERA's mapmaking ability.
+Furthermore, the core is supplemented by 30 additional outrigger elements to tile the $uv$-plane with instantaneously complete sub-aperture sampling out to 250$\lambda$ and complete aperture-scale sampling out to 350$\lambda$ (at 150\,MHz).
 This sub-aperture sampling strategy suppresses grating lobes in the instrument's point spread function and provides information for calibrating and correcting direction-dependent antenna responses \citep{dillon_parsons2016}.
@@
 Although in the past, correlator development has been one of the most expensive and complex aspects of building a large array or radio interferometers, this is no longer generally the case.
-The Collaboration for Astronomy Signal Processing and Electronics Research (CASPER) (\citealt{parsons_et_al2006}, \url{casper.berkeley.edu}) is a community of astronomers and engineers who work to reduce the cost and complexity of radio astronomy signal processing systems through the development of open-source, general-purpose hardware and software.
+The Collaboration for Astronomy Signal Processing and Electronics Research (CASPER; \citealt{parsons_et_al2006}, \url{casper.berkeley.edu}) is a community of astronomers and engineers who work to reduce the cost and complexity of radio astronomy signal processing systems through the development of open-source, general-purpose hardware and software.
 CASPER currently has several hundred members at 73 institutions, and has developed six generations of FPGA-based signal processing hardware, shown in Fig. \ref{fig:hardware}.
@@
 \caption{
-Six generations of CASPER digital signal processing (left to right) culminating in the SNAP board (right, along with the long-haul fiber-link test setup), which was developed and tested under the previous HERA MSIP award.
+Six generations of CASPER digital signal processing (left to right) culminating in the SNAP board (right, along with the long-haul fiber-link test setup).
 By preserving its design tools, signal processing libraries, and interface code between hardware generations, CASPER hardware and a modular architecture have enabled the PAPER correlator to be easily upgraded for HERA \Mycitep{parsons_et_al2006,parsons_et_al2008}.
@@
 NRAO manages the dynamic acquisition of computing from
-external HPC and cloud-computing facilities, including the Amazon spot market (\url{aws.amazon.com/ec2/spot}) and XSEDE (\url{www.xsede.com}) . 
+external HPC and cloud-computing facilities, including the Amazon spot market (\url{aws.amazon.com/ec2/spot}) and XSEDE (\url{www.xsede.com}). 
 Combined with legacy computing resources at UPenn and MIT, this plan leverages NRAO's
@@
 \url{github.com/MiguelFMorales/FHD}, \url{github.com/MiguelFMorales/eppsilon}, \url{github.com/jpober/21cmSense},
-and \url{github.com/Nithyanandan/PRISim}.}, revision controlled, and unit-tested.  
+\url{github.com/Nithyanandan/PRISim}, and \url{https://github.com/BradGreig/21CMMC}.}, revision controlled, and unit-tested.  
 These include OMNICAL, a complete package for redundant baseline calibration;

### 2016-04-22 manuscript.tex
first pass at Nima and Eloy

@@
 \title{Hydrogen Epoch of Reionization Array (HERA)}
-\author{David R.
-DeBoer, Aaron R.
-Parsons, Judd D.
-Bowman, Jonathan C.
-Pober, Jacqueline N.
-Hewitt, Chris L.
-Carilli, Gianni Bernardi, Anita Loots, Andrei Messenger, Paul Alexander, Steve Furlanetto, James E.
-Aguirre, Miguel F.
-Morales, Adam Beardsley, Daniel C Jacobs, Nithyanandan Thyagarajan, Zak, Aaron Ewall-Wice, Abraham R.
-Neben, Max Tegmark, Haoxuan Zheng, Richard Bradley, Matthys Maree, Kathryn Rosie, Randall Fritz, Nathan Mathison, Cresshim Malgas, Raddwine Sell, Samantha Pieterse, Lourence Malan, Austin Julius, MacCalvin Kariseb, Craig Smith, Jasper Grobbelaar, Telalo Lekalake, Eunice Matsetela, Angelo Syce, Bradley Grieg, Eloy de Lera Acedo, Nicolas Fagnoni, Nima Razavi-Ghods, Zaki S.
-Ali, Carina Cheng, Josh Dillon, Jack Hickish Adrian Liu, David MacMahon, Nipanjana Patra, Saul A.
-Kohn, Bryna Hazelton, Peter Williams}
+\author{David DeBoer, Aaron Parsons, Judd Bowman, Jonathan Pober, Jacqueline Hewitt, Chris Carilli, Gianni Bernardi, Anita Loots, Andrei Messenger, Paul Alexander, Steve Furlanetto, James Aguirre, Miguel Morales, Zaki Ali, Adam Beardsley, Richard Bradley, Carina Cheng, Eloy de Lera Acedo, Josh Dillon, Aaron Ewall-Wice, Nicolas Fagnoni, Randall Fritz, Bradley Grieg, Jasper Grobbelaar, Bryna Hazelton, Daniel Jacobs, Austin Julius, MacCalvin Kariseb, Saul Kohn, Telalo Lekalake, David MacMahon, Lourence Malan, Cresshim Malgas, Matthys Maree, Nathan Mathison, Eunice Matsetela, Abraham Neben, Nipanjana Patra, Samantha Pieterse, Nima Razavi-Ghods, Kathryn Rosie, Raddwine Sell, Craig Smith, Angelo Syce, Max Tegmark, Nithyanandan Thyagarajan, Peter Williams, Zak, Haoxuan Zheng}
 
@@
 contamination, current-generation 21 cm instruments are approaching their sensitivity limits.
-HERA is designed to bring both the
-sensitivity and the precision to deliver its primary science on the basis of
+HERA is designed to bring both the sensitivity and the precision to deliver its primary science on the basis of
 proven foreground filtering techniques, while developing new subtraction
@@
 realizing the widely recognized scientific potential of 21 cm cosmology.
+
+\tableofcontents
 
@@
 analog, digital, and real-time processing systems deployed for PAPER-128.  
-This allowed for immediate observing with the new elements with a well-characterized system.
+This allows for immediate observing with the new elements with a well-characterized system.
 As HERA develops, this
@@
 This allows for an independent testing platform as the group at Cavendish continues to investigate both a broad-band feed, as well as ways to improve the current analog system for an improved match and noise performance.
+
+We are furthermore analyzing the effect of coupling with numerical simulations where we can quantify and understand the effects of mutual coupling, which so far only looks substantial in the far out side-lobes.
+This will allow us optimized the feed design for reduced coupling as well as evaluate the need of ``skirts'' for cross-coupling reduction or other further dish optimizations.
+This work will benefit from the existence of several prototype systems including the 3-dish HERA system at Lord's Bridge, Cambridge.
 
@@
 To facilitate foreground filtering, HERA's antenna element is designed to suppress reflections at long time delays.
-
-The feed design will proceed in three phases:
-\begin{itemize}
-\item {\em Phase 1:}  use existing PAPER dipole with new ground screen (\S\ref{sec:feed}) and existing PAPER signal path,
-\item {\em Phase 2:}  updated ground screen, matching network and analog signal path (\S\ref{sec:analog})
-\item {\em Phase 3:}  broad-band or separate low-frequency for full spectral coverage.
-\end{itemize}
-Models and measurements discussed below are for the current Phase 1 system.
+Models and measurements discussed below are for the initial phase system.
 
@@
 \label{sec:feed}
-The feed is to be deployed along two separate phases.
-Initially, the feed reuses the original PAPER sleeved dipole, however in a new optimized cylindrical feed-cage configuration.
+The feed design will proceed in three phases:
+\begin{itemize}
+\item {\em Phase 1:}  use existing PAPER dipole with new ground screen (\S\ref{sec:feed}) and existing PAPER signal path,
+\item {\em Phase 2:}  updated ground screen, matching network and analog signal path (\S\ref{sec:analog})
+\item {\em Phase 3:}  broad-band or separate low-frequency for full spectral coverage.
+\end{itemize}
+
+\noindent
+{\bf Phase 1:}  Initially, the feed reuses the original PAPER sleeved dipole, however in a new optimized cylindrical feed-cage configuration.
 As discussed in \cite{feedmemo}, the optimization was based on (a) main beam efficiency, (b) cross-coupling (integrated feed pattern on the dish), (c) standing waves (blockage size), (d) frequency response and (e) polarization match.
@@
 
+\noindent
+{\bf Phase 3:} 
+A wideband feed is being designed for the next phase of HERA (see Fig.
+\ref{fig:futureFeed}).
+This feed will have wide bandwidth (50-250 MHz) and improved performance compare to the current PAPER dipoles.
+This wider band will open the window of Cosmic Dawn and the latest stages of the Epoch of Re-ionization to HERA.
+This feed is based on a modified TEM horn concept where the feed?s beam is optimized for the appropriate illumination of the HERA dish to maximize sensitivity.
+This is to maximize the effective aperture while minimizing the system temperature (receiver noise, ohmic losses, spill-over, etc.).
+This feed will as well replace the current PAPER dipoles and will therefore be mechanically compatible with the current dishes being deployed in South Africa.
+
+\begin{figure}[h!]
+\centerline{
+\includegraphics[height=0.3\textwidth]{plots/futureFeed.png}
+}
+\caption{\small Future HERA feed showing its directive beam pointed towards the dish.}
+\label{fig:futureFeed}
+\end{figure}
+
+The proposed feed belongs to the family of travelling wave end-fire antennas.
+This type of antennas, as opposed to standing wave antennas such as dipoles, conducts the wave through the structure until it is radiated.
+These antennas therefore tend to show good ultra band behavior in the time domain, which is key for HERA.
+TEM horn antennas also show ultra wide band impedance patterns for relatively small sizes as well as frequency stable circular beam patterns.
+Polarization purity tends to be well behaved in these antennas, being worst in the diagonal planes.
+Our current work is relying on numerical simulations for the optimization of the feed (see Fig.
+2).
+
+Furthermore, as shown in \cite{2015ExA....39..567D}, room temperature ultra low receiver noise (<35K including matching noise) above ~100 MHz can be achieved with COTS transistors if a proper feeding mechanism and matching are designed.
+Despite sky noise is dominant for most of the EoR band, good matching is always critical in wide band systems, and specifically in the HERA case it is important to ensure good power match as well as noise match in order to reduce the effects of unwanted reflections [2].
+We envisage having such a feeding mechanism in the new feed where the first stage LNA will be connected directly to the feeding point of the feed antenna.
+
 \subsection{Array Configuration}
@@
 
+\noindent
 {\bf Front-End Module: }
-
+The Front-End Module consists of an LNA as the very first stage in the chain for each polarization.
+This component for the existing feed design will be a differential low noise amplifier which consists of two sets of LNAs feeding each dipole arm followed by a balun (or hybrid coupler).
+The output from this point is then all single ended and mostly matched to 50?
+impedance.
+Whilst it is desirable to be sky noise dominated in the HERA band, meaning for the first stage, the best achievable noise match must be obtained, it must not come at a too great cost to the power match since reducing the ripple across the band is absolutely essential.
+This trade-off will be observed carefully to ensure the FEM is both low noise and well matched.
+This stage is followed by carefully matched reflection-less filters and further amplifications of the signal.
+In order to ensure very low reflection on the line a large matching attenuator will be used at each end of the cable after the bias-Tee circuit.
+
+The FEM will also include a number of useful signal conditioning operations.
+One of these will be to have a phase switch for each polarization as early as possible in the signal chain.
+The switch will offer very low phase imbalance and will be controlled by differential lines.
+The Walsh functions (orthogonal phase switching signals) will be generated in the SNAP board and used to control all FEMs in the field.
+
+Another useful feature of the FEM will be an integrated ?Dicke?
+switching radiometer similar to the EDGES experiment \cite{2012RaSc...47.0K06R}.
+The circuit provides the capability of switching between the output of the antenna, a calibrated noise source and a 50?
+ambient load.
+The antenna measures only a fraction of the sky brightness temperature, determined by the matching between the antenna and receiver.
+Three spectrometer measurements (antenna, ambient load and noise source) as well as additional lab measurements of the s-parameters and noise-parameters of the FEM prior to deployment then allow good calibration of the output spectrum.
+An on-board temperature sensor near the noise source will also aid with the calibration of the receiver chain.
+The FEM will be housed in a rugged 5x5x10 cm unit which is water and dust resistant.
+
+\noindent
 {\bf Post-Amplifier Module: }
+The Front-End Module consists of an LNA as the very first stage in the chain for each polarization.
+This component for the existing feed design will be a differential low noise amplifier which consists of two sets of LNAs feeding each dipole arm followed by a balun (or hybrid coupler).
+The output from this point is then all single ended and mostly matched to 50?
+impedance.
+Whilst it is desirable to be sky noise dominated in the HERA band, meaning for the first stage, the best achievable noise match must be obtained, it must not come at a too great cost to the power match since reducing the ripple across the band is absolutely essential.
+This trade-off will be observed carefully to ensure the FEM is both low noise and well matched.
+This stage is followed by carefully matched reflection-less filters and further amplifications of the signal.
+In order to ensure very low reflection on the line a large matching attenuator will be used at each end of the cable after the bias-Tee circuit.
+
+The FEM will also include a number of useful signal conditioning operations.
+One of these will be to have a phase switch for each polarization as early as possible in the signal chain.
+The switch will offer very low phase imbalance and will be controlled by differential lines.
+The Walsh functions (orthogonal phase switching signals) will be generated in the SNAP board and used to control all FEMs in the field.
+
+Another useful feature of the FEM will be an integrated ?Dicke?
+switching radiometer similar to the EDGES experiment [1].
+The circuit provides the capability of switching between the output of the antenna, a calibrated noise source and a 50?
+ambient load.
+The antenna measures only a fraction of the sky brightness temperature, determined by the matching between the antenna and receiver.
+Three spectrometer measurements (antenna, ambient load and noise source) as well as additional lab measurements of the s-parameters and noise-parameters of the FEM prior to deployment then allow good calibration of the output spectrum.
+An on-board temperature sensor near the noise source will also aid with the calibration of the receiver chain.
+The FEM will be housed in a rugged 5x5x10 cm unit which is water and dust resistant.
 

### 2016-04-24 manuscript.tex
Fix a few typos.

@@
 Section \ref{sec:eormeas} presents the techniques used to make the measurement.
-Section \ref{sec:requirements} presents the high-level requirements and Section \ref{sec:design} presents the system description..
+Section \ref{sec:requirements} presents the high-level requirements and Section \ref{sec:design} presents the system description.
 Section \ref{sec:status} presents a brief status and outline the deployement timetable and Section \ref{sec:conclusion} concludes by summarizing and providing additional context.
@@
 	\caption{Rendering of cosmic evolution from just after the Big Bang to today (background image credit Loeb/{\em Scientific American}).
-The blue line shows the frequency of the red-shifted hydrogen line (rest frequency 1420 MHz) as a function of redshift (lower horizontal axis) and cosmic time (upper horizontal axis).
-The white dashed horizontal lines are the HERA EOR band and the blue are the extended frequency goal.
+The labels show the redshift and the frequency of the red-shifted hydrogen line (rest frequency 1420 MHz) at different ages of the Universe.
+The solid white lines bracket the HERA EOR band and the dashed ones bracket the extended frequency goal.
 %Determining where the structure in the background really falls is the goal of HERA.
 CMB observations observe the afterglow of the Big Bang (far left) and Baryonic Acoustic Oscillation (BAO) surveys proposed 
-target $z\approx$0.8-2.5.  Limited surveys span back to about $z\approx$7.} 
+target $z\approx0.8$-2.5.  Limited surveys span back to about $z\approx7$.}
 	\label{fig:cosmos}
@@
 \caption*{%\tiny{\justify
-$^*$Calculations done via \url{21cmSense} \url{www.github.com/} \citep{2013AJ....145...65P,2014ApJ...782...66P}
+$^*$Calculations done via {\tt 21cmSense}  \citep[\url{www.github.com/jpober/21cmSense};][]{2013AJ....145...65P,2014ApJ...782...66P}.
 Foreground avoidance represents an analysis comparable to \cite{ali_et_al2015}, whereas foreground modeling allows significantly more $k$ modes of the cosmological signal to be recovered.}
@@
 The bandwidth
-therefore equates to a cosmic distance along the line-of-site of the telescope, with the frequency determining the cosmic age.
+therefore equates to a cosmic distance along the line-of-sight of the telescope, with the frequency determining the cosmic age.
 An observation of an area of sky over a given bandwidth is therefore providing an average
@@
 Obviously between our present observation point and the Epoch of Reionization lies the entire intervening universe, which has an intrinsically much brighter signal (up to 6 orders of magnitude or more).
+% Almost everywhere else says 5 orders of magnitude, not 6; intended?
 Primarily, this power is due to diffuse Galactic synchrotron radiation, supernova remnants and extragalactic radio sources.

### 2016-04-24 manuscript.tex
Polish the paragraph on exoplanetary radio bursts.

--- 2016-04-22 manuscript.tex
@@
 \label{sec:exoplanets}
-HERA could be a powerful tool in the search for bright, highly polarized auroral bursts
-from exoplanets %due to the electron cyclotron maser instability 
-\citep{treumann2006,hallinan_et_al2015}. These
-distinctive, coherent bursts occur periodically at the planetary
-rotation rate, which cannot otherwise be measured. 
-A burst from a terrestrial planet could point toward
-habitability, since it implies a powerful magnetic field
-protecting the atmosphere and perhaps the biosphere from energetic stellar wind
-particles \citep{tarter_et_al2007}. HERA's sensitivity, large FoV, long campaigns, and precise calibration are all well-suited
-for discovering exoplanetary radio bursts from Jupiter-like planets out to 25~pc.
-Many well studied exoplanetary systems within that distance are in the HERA stripe, including Fomalhaut, Gl 667 C, Gl 433, HD 147513, and Gl 317.
+HERA could revolutionize the study of extrasolar magnetospheres if it detects bright, highly polarized exoplanetary auroral radio bursts \citep{zarka2011}.
+Observers have tried to detect these distinctive events since even before the discovery of the first exoplanet \citep{winglee_et_al1986} but have thus far not succeeded.
+The bursts occur at the electron cyclotron frequency and are highly beamed, allowing remote sensing of planetary rotation periods and magnetic field strengths, two quantities that are otherwise only indirectly measurable.
+Since magnetic fields may play a key role in protecting atmospheres and/or biospheres from energetic stellar wind particles \citep[e.g.,][]{tarter_et_al2007}, studies of planetary magnetism may be vital to evaluating habitability.
+HERA's sensitivity, large field of view, long campaigns, and precise calibration make it well-suited to the search for radio bursts from Jupiter-like planets out to 25~pc.
+Well-studied exoplanet host stars in the HERA stripe within that distance include Fomalhaut, Gl~317, Gl~433, Gl~667~C, and HD~147513.
 

### 2016-04-25 manuscript.tex
Finish reading through for typos.

--- 2016-04-24 manuscript.tex
@@
 To address this challenge,  all aspects of the experimental process must be refined and improved, spanning 
-calibration ({\em e.g.} \cite{zheng_et_al2014, jacobs_et_al2016, 2016arXiv160300607B}), handling foreground contamination
-({\em e.g.} \cite{moore_et_al2013,thyagarajan_et_al2015a,pober_et_al2016}),
-and the interferometer design itself ({\em e.g.} \cite{parsons_et_al2012a,dillon_parsons2016}).
+calibration \citep[{\em e.g.},][]{zheng_et_al2014, jacobs_et_al2016, 2016arXiv160300607B}, handling foreground contamination
+\citep[{\em e.g.},][]{moore_et_al2013,thyagarajan_et_al2015a,pober_et_al2016},
+and the interferometer design itself \citep[{\em e.g.},][]{parsons_et_al2012a,dillon_parsons2016}.
 
@@
 For a more complete discussion see \cite{liu_et_al2014a,liu_et_al2014b}}.
-The Fourier transform of a frequency spectrum is a delay, hence the name ``delay-spectrum'' for the technique that uses this approach.
+The Fourier transform of a frequency spectrum is a delay spectrum, hence the name ``delay-spectrum'' for the technique that uses this approach.
 Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.
@@
 Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.
-Wavenumbers with an assumed bandwidth ($k$ and $\kpar$ are shown as stepped profiles, with the redshift steps equal to the bandwidth at that redshift.
+Wavenumbers with an assumed bandwidth ($k$ and $\kpar$) are shown as stepped profiles, with the redshift steps equal to the bandwidth at that redshift.
 Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer appropriate, which is why the line is dashed.
@@
 Systematics will push that line such as to close the EOR ``window'' to the upper left ($\beta > 1$).
-If we can completely control systematics and constrain the affects of the foreground to {\em e.g.} the main beam, the line will move to decrease the size of the wedge ($\beta<1$).
+% above: should "close" be "move"?
+If we can completely control systematics and constrain the effects of the foreground to {\em e.g.} the main beam, the line will move to decrease the size of the wedge ($\beta<1$).
 Therefore, a key issue to measuring the EOR power spectrum is to understand and minimize systematics.
-For further discussion and an analytical derivation, see \cite{zahn_etal2012,vedantham_2012,liu_et_al2014b},  in simulation see \cite{datta_etal2010,hazelton_et_al2013} and for observations
-\cite{pober_etal2013b,2015arXiv150601026P,parsons_etal2014,2015arXiv150206016A}.
-
-Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).
-In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_0=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.
+For further discussion and an analytical derivation, see \citet{zahn_etal2012}, \citet{vedantham_2012}, and \citet{liu_et_al2014b}.
+In simulation see \citet{datta_etal2010} and \citet{hazelton_et_al2013}, and for observations
+\citet{pober_etal2013b}, \citet{2015arXiv150601026P}, \citet{parsons_etal2014}, and \citet{2015arXiv150206016A}.
+
+Note that the unit of the spatial wavenumber $k$ is length$^{-1}$, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).
+In order to normalize for updated measurements of the present-day Hubble parameter $H_0$, it is further normalized by a factor $h$, where $H_0=100h$ (km/s)/Mpc, such that the wavenumbers are expressed as $h$/Mpc.
 Also note that at the redshifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.
@@
 
-The techniques to measure the power spectrum may be broken down into two principle techniques -- delay-space (\S\ref{sec:delayapproach}) and map-making (\S\ref{sec:mapapproach}) -- plus additional hydrid methods.
+The techniques to measure the power spectrum may be broken down into two principle techniques -- delay-space (\S\ref{sec:delayapproach}) and map-making (\S\ref{sec:mapapproach}) -- plus additional hybrid methods.
 These are briefly summarized below, with an emphasis on the delay-spectrum technique as the initial approach taken by HERA.
@@
 \begin{equation}
-\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
+\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}(k)}\right]
 \label{eq:sensitivity}
@@
 Right:  Power spectrum sensitivities for various arrays normalized to HERA-19.
-Here the element size is included, which is shown as an appropriately sized circle.}
+The element sizes are indicated at the far right with appropriately sized circles.}
 \label{fig:boost}
@@
 For example, $x_{HI}$ spans almost the entire allowable range of [0,1]
-at $z=8$. 21\,cm reionization experiments place much tighter constraints on ionization, with the
+at $z=8$. Twenty-one centimeter reionization experiments place much tighter constraints on ionization, with the
 red band showing the forecasted 95\% confidence region derived from HERA data,
@@
 
-From these measurements and models, HERA is required to cover a redshift range of $z=$6 to 13, corresponding to a frequency range of 101 - 203 MHz.
-We adopt 100-200 MHz for full performance as the requirement.
+From these measurements and models, HERA is required to cover a redshift range of $z=6$ to 13, corresponding to a frequency range of 101--203 MHz.
+We adopt 100--200 MHz for full performance as the requirement.
 Extending the science to include the Dark Ages remains a goal so that efforts are on-going to design a feed to increase the lower limit down to at least 70 MHz without compromising the performance in the above requirement range.
@@
 Early analysis \citep{elementmemo} indicated that to trace the EOR over a wide range of redshift, internal reflections should be attenuated by about 60 dB from the initial incoming wave by about 60 ns, or a voltage standing wave ratio (VSWR) $<$ 1.002 for frequencies $<$ 17 MHz.
-This was a conservative value based on estimates of power in spectrally smooth foreground sources from \cite{santos_et_al2005,2008MNRAS.385.2166A, deoliveira2008, jelic_et_al2008, bernardi_et_al2010} and experience with PAPER dealing with foreground systematics.
+This was a conservative value based on estimates of power in spectrally smooth foreground sources \citep{santos_et_al2005,2008MNRAS.385.2166A, deoliveira2008, jelic_et_al2008, bernardi_et_al2010} and experience with PAPER dealing with foreground systematics.
 This provided the basis for the initial design of the element.
@@
 The vertical axis shows a very steep delay ``cost'' at 60 ns and the horizontal axis is the cost/performance model, as described in the following section.
+% should this just be "horizontal axis is the element diameter"?
 The color coding is the product of the two assigned costs, with red showing a high cost normalized to 350 14m elements (clipped at a ratio of 1.25) and purple being the minimum.
@@
 See \S\ref{sec:delayspec}, \S\ref{sec:cost} and \cite{elementmemo} for details.
-A diameter of 14-m (vertical dashed line) is near the cost minimum and consistent with assumption for the delay specification.
+A diameter of 14-m (vertical dashed line) is near the cost minimum and consistent with assumptions for the delay specification.
 \label{fig:costfig}}
@@
 the dish design for should allow for EOR detections for $k>0.10 h/$Mpc.
-We also see that the initial simple ``60@60'' specification was a overly conservative (shown as the lightly shaded area bounded by the teal line), however it served as a good initial goal.
+We also see that the initial simple ``60@60'' specification was overly conservative (shown as the lightly shaded area bounded by the teal line), however it served as a good initial goal.
 
@@
 
-As discussed in the previous section, the design of the design principles are three-fold:
+As discussed in the previous section, the goals of the design principles are three-fold:
 \begin{itemize}
@@
 \subsubsection{Element Construction}
-To minimize cost and aid in construction in a remote location, the construction materials are readily available standard construction materials, making for a very cost-effective design.
+To minimize cost and aid in construction in a remote location, the elements are built from readily available standard construction materials, making for a very cost-effective design.
 The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.
-Given the close-packed design, each pole (except for the perimeter) are shared between three antennas.
+Given the close-packed design, each pole (except for the perimeter) is shared between three antennas.
 This both reduces the number of poles, but also makes for a balanced load.
 Note that the perimeter poles can be stayed with guy lines if needed.
-The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.
+The dish rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.
 Two of the three intermediate posts are shared with adjacent antennas.
@@
 \ref{fig:futureFeed}).
-This feed will have wide bandwidth (50-250 MHz) and improved performance compare to the current PAPER dipoles.
+This feed will have wider bandwidth (50-250 MHz) and improved performance compare to the current PAPER dipoles.
 This wider band will open the window of Cosmic Dawn and the latest stages of the Epoch of Re-ionization to HERA.
-This feed is based on a modified TEM horn concept where the feed?s beam is optimized for the appropriate illumination of the HERA dish to maximize sensitivity.
+This feed is based on a modified TEM horn concept where the feed's beam is optimized for the appropriate illumination of the HERA dish to maximize sensitivity.
 This is to maximize the effective aperture while minimizing the system temperature (receiver noise, ohmic losses, spill-over, etc.).
@@
 
-Furthermore, as shown in \cite{2015ExA....39..567D}, room temperature ultra low receiver noise (<35K including matching noise) above ~100 MHz can be achieved with COTS transistors if a proper feeding mechanism and matching are designed.
-Despite sky noise is dominant for most of the EoR band, good matching is always critical in wide band systems, and specifically in the HERA case it is important to ensure good power match as well as noise match in order to reduce the effects of unwanted reflections [2].
+Furthermore, as shown in \cite{2015ExA....39..567D}, room temperature ultra low receiver noise ($<$35K including matching noise) above ~100 MHz can be achieved with COTS transistors if a proper feeding mechanism and matching are designed.
+Despite the fact that sky noise is dominant for most of the EoR band, good matching is always critical in wide band systems, and specifically in the HERA case it is important to ensure good power match as well as noise match in order to reduce the effects of unwanted reflections [2].
+% this "[2]" is probably a citation that has not been properly propagated??
 We envisage having such a feeding mechanism in the new feed where the first stage LNA will be connected directly to the feeding point of the feed antenna.
@@
 \label{sec:analog}
-HERA's analog signal path from 70--250 MHz with the full performance EOR band of 100-200 MHz, as shown in Fig.
+HERA's analog signal path from 70--250 MHz with the full performance EOR band of 100--200 MHz, as shown in Fig.
 \ref{fig:cosmos} and discussed in \S\ref{sec:freqs}, emphasizes spectral smoothness and robustness.
 Although the low-frequency sky has an intrinsically bright signal level, in the desired ``cold patches'' the sky temperature is only about 30 K at the upper frequency end, so relatively low receiver temperatures are desired.
-Note that at the low end of the band (70 MHz), the sky temperature is about 1000 K and across the EOR band it ranges from about 100-400 K.
+Note that at the low end of the band (70 MHz), the sky temperature is about 1000 K and across the EOR band it ranges from about 100--400 K.
 With receiver temperatures of 30--100 K readily achievable with ambient temperature electronics however
@@
 
-The analog signal path comprises the front-end module (FEM) at the feed, the post-amplfier module (PAM) at field-deployed nodes, and the $\sim 35m$ coaxial cable in between.
+The analog signal path comprises the front-end module (FEM) at the feed, the post-amplifier module (PAM) at field-deployed nodes, and the $\sim 35$ m coaxial cable in between.
 Also included are the monitor and control aspects to control and measure its state.
@@
 This component for the existing feed design will be a differential low noise amplifier which consists of two sets of LNAs feeding each dipole arm followed by a balun (or hybrid coupler).
-The output from this point is then all single ended and mostly matched to 50?
-impedance.
+The output from this point is then all single ended and mostly matched to 50$\Omega$ impedance.
 Whilst it is desirable to be sky noise dominated in the HERA band, meaning for the first stage, the best achievable noise match must be obtained, it must not come at a too great cost to the power match since reducing the ripple across the band is absolutely essential.
@@
 
-Another useful feature of the FEM will be an integrated ?Dicke?
-switching radiometer similar to the EDGES experiment \cite{2012RaSc...47.0K06R}.
-The circuit provides the capability of switching between the output of the antenna, a calibrated noise source and a 50?
-ambient load.
+Another useful feature of the FEM will be an integrated ``Dicke'' switching radiometer similar to the EDGES experiment \citep{2012RaSc...47.0K06R}.
+The circuit provides the capability of switching between the output of the antenna, a calibrated noise source and a 50$\Omega$ ambient load.
 The antenna measures only a fraction of the sky brightness temperature, determined by the matching between the antenna and receiver.
@@
 An on-board temperature sensor near the noise source will also aid with the calibration of the receiver chain.
-The FEM will be housed in a rugged 5x5x10 cm unit which is water and dust resistant.
+The FEM will be housed in a rugged 5$\times$5$\times$10 cm unit which is water and dust resistant.
 
@@
 {\bf Post-Amplifier Module: }
-The Front-End Module consists of an LNA as the very first stage in the chain for each polarization.
-This component for the existing feed design will be a differential low noise amplifier which consists of two sets of LNAs feeding each dipole arm followed by a balun (or hybrid coupler).
-The output from this point is then all single ended and mostly matched to 50?
-impedance.
-Whilst it is desirable to be sky noise dominated in the HERA band, meaning for the first stage, the best achievable noise match must be obtained, it must not come at a too great cost to the power match since reducing the ripple across the band is absolutely essential.
-This trade-off will be observed carefully to ensure the FEM is both low noise and well matched.
-This stage is followed by carefully matched reflection-less filters and further amplifications of the signal.
-In order to ensure very low reflection on the line a large matching attenuator will be used at each end of the cable after the bias-Tee circuit.
-
-The FEM will also include a number of useful signal conditioning operations.
-One of these will be to have a phase switch for each polarization as early as possible in the signal chain.
-The switch will offer very low phase imbalance and will be controlled by differential lines.
-The Walsh functions (orthogonal phase switching signals) will be generated in the SNAP board and used to control all FEMs in the field.
-
-Another useful feature of the FEM will be an integrated ?Dicke?
-switching radiometer similar to the EDGES experiment [1].
-The circuit provides the capability of switching between the output of the antenna, a calibrated noise source and a 50?
-ambient load.
-The antenna measures only a fraction of the sky brightness temperature, determined by the matching between the antenna and receiver.
-Three spectrometer measurements (antenna, ambient load and noise source) as well as additional lab measurements of the s-parameters and noise-parameters of the FEM prior to deployment then allow good calibration of the output spectrum.
-An on-board temperature sensor near the noise source will also aid with the calibration of the receiver chain.
-The FEM will be housed in a rugged 5x5x10 cm unit which is water and dust resistant.
+TODO.
+% The text here was a duplicate of the FEM subsection.
 
@@
 
-While some applications may utilise SNAPs in multiple-board chasses, for HERA, SNAP is housed in a field-deployable RFI-tight box.
+While some applications may utilise SNAPs in multiple-board chassis, for HERA, SNAP is housed in a field-deployable RFI-tight box.
 The node enclosure itself is an RFI-tight housing that can handle the signal paths for up to 12 antennas.
@@
 of the PAPER pipeline, see
-\cite{zheng_et_al2014,parsons_etal2014,ali_et_al2015}.
+\citet{zheng_et_al2014}, \citet{parsons_etal2014}, and \citet{ali_et_al2015}.
 The Internet link between the Karoo and data centers in the U.S.\ has been extensively tested, with assistance from SKA-SA, TENET
@@
 21cmFAST \citep{mesinger_et_al2011}, a fast, semi-numerical 21~cm signal simulator,
-and \url{https://github.com/jpober/21cmSense} \citep{pober_et_al2014}, a tool for forecasting power spectrum sensitivity.
+and 21cmSense \citep{pober_et_al2014}, a tool for forecasting power spectrum sensitivity.
 Other project code is aggregated and revision
@@
 \noindent On-site construction will proceed in stages, starting from the
-initial 19 element currently in place.    Elements 20-37
+initial 19 elements currently in place.    Elements 20--37
 are currently under construction, to be completed in 2016.  
@@
 HERA's large collecting area and optimized design enable precise constraints of EoR astrophysics and a broad range of high-impact secondary science, all before SKA-low becomes operational.
-With CHAMP, HERA expands CAMPARE's proven methods and recruitment network to bring more students from underrepresented groups into cutting-edge astronomy research and support them with year-round mentoring.
 By bringing the U.S.\ teams from PAPER, the MWA, MITEoR, and EDGES together with strategic international partners,
